### 2020-04-28 README.Rmd
Initial Commit: Copying File Structure as in asp20model

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

set.seed(1337)
```

# Advanced statistical programming with R

Are you taking the course "Advanced statistical programming with R" in the summer term 2020? Then this page is for you. It explains what you need to do to get the credits for this module. `r emo::ji("wink")`

In this course, you are going work on a programming project in a group of three students. Each group needs to develop an R package, in which they implement a statistical method for the so-called location-scale regression model class, and a small simulation study to evaluate their method and implementation. In September, you need to submit a term paper about your project.

This git repository contains the `asp20model` package for R, which implements the location-scale regression model class and will be the basis for your package.

**Please note:** If you are planning to take this course, **you need to preregister on Stud.IP before March 31.** Follow this link to sign up: https://studip.uni-goettingen.de/dispatch.php/course/details?sem_id=5b8822a839a628e8166648a57f4f1eac.

**Update on the Corona crisis:** Unsurprisingly, the Corona crisis makes some changes to this course necessary. All previously announced dates remain valid, but I am going to offer the introductory session and the Q&A session as YouTube livestreams. See below for the links. The intermediate and final presentations might be replaced with short written reports, depending on the further developments. I am going to make a final decision on the presentations before the Q&A session on May 4.

## Important dates

- Tuesday, April 7, 10:00--15:00, ~~Blauer Turm / MZG 8.163~~: Introductory session, project assignment \
  YouTube livestream: https://youtu.be/-a02j6YTxKw \
  Rocket.Chat channel for questions and comments: https://chat.gwdg.de/channel/asp20
- Monday, May 4, 14:00--16:00, ~~Blauer Turm / MZG 8.163~~: Q&A session, feel free to ask technical, statistical, and organizational questions \
  YouTube livestream: https://youtu.be/vi2BqiDUCFI \
  Rocket.Chat channel for questions and comments: https://chat.gwdg.de/channel/asp20
- Friday, June 5, 8:30--15:00, Blauer Turm / MZG 8.163: Intermediate presentations
- Friday, July 10, 8:30--15:00, Blauer Turm / MZG 8.163: Final presentations
- Tuesday, September 15: Submission deadline for the term papers

## Technical & statistical prerequisites

### The R6 OOP system

The `asp20model` package uses R6 classes, so make sure to get familiar with mutable objects, inheritance, etc. See the chapter on R6 classes in [3], available online: https://adv-r.hadley.nz/r6.html.

### Version control using git & GitLab

The `asp20model` package is hosted on https://gitlab.gwdg.de (the GitLab instance of the computing center of the university), where you are going to develop your R package as well. I created a GitLab group for this course and a GitLab project for each group of students.

**Please note:** To start working on your project, log in to GitLab with your student account once and send me an email. After that, I can add you to the group and the project.

GitLab is an online platform where you can host git repositories, very similar to GitHub, and git is a popular version control system for software development. Systematic, version-controlled software development is an essential component of this course, so if you are not yet familiar with git, please take a look at the git chapter in [4], available online: http://r-pkgs.had.co.nz/git.html.

**Please note:** Your code will only be visible to me, the project supervisors, and your fellow students. Feel free to make mistakes and ask questions!

### Location-scale regression

As mentioned above, we are going to work with the location-scale regression model class in this course. So, what **is** location-scale regression? Let's take a step back and look at the standard linear model first. It is defined as

``` math
y_i = \boldsymbol{x}_i' \boldsymbol{\beta} + \varepsilon_i, \text{ where } \varepsilon_i \overset{i.i.d.}{\sim} \mathcal{N}(0, \sigma^2).
```

From the definition, it follows that

``` math
y_i \overset{ind.}{\sim} \mathcal{N}(\boldsymbol{x}_i' \boldsymbol{\beta}, \sigma^2),
```

and from this representation, the model can easily be extended with a second linear predictor for the standard deviation. For this purpose, we introduce a covariate vector \$`\boldsymbol{z}_i`\$ and a parameter vector \$`\boldsymbol{\gamma}`\$:

``` math
y_i \overset{ind.}{\sim} \mathcal{N}(\boldsymbol{x}_i' \boldsymbol{\beta}, (\exp(\boldsymbol{z}_i' \boldsymbol{\gamma}))^2).
```

What is the \$`\exp`\$ function doing in this formula? Well, one thing that is a little bit tricky about predicting the standard deviation is that the standard deviation needs to be positive, while the linear predictor \$`\boldsymbol{z}_i' \boldsymbol{\gamma}`\$ can become negative for some choices of the parameter vector \$`\boldsymbol{\gamma}`\$. Hence, to ensure that our predictions for the standard deviation are valid, we introduce the \$`\exp`\$ function as a so-called response function for the standard deviation and define \$`\sigma_i = \exp(\boldsymbol{z}_i' \boldsymbol{\gamma})`\$.

Why should we care about location-scale regression? It is not uncommon to observe heteroscedastic data like in the figure below. The standard way to deal with heteroscedastic residuals in a linear model is FGLS estimation, but FGLS does not provide an interpretable description of the variance structure of the data. A location-scale regression model can fill this gap, if we have access to explanatory variables for the variance or the standard deviation of the response variable.

```{r data, fig.width = 11}
n <- 500
x <- runif(n)
y <- x + rnorm(n, sd = exp(-3 + 2 * x))
plot(x, y)
abline(0, 1, lwd = 2)
curve(x + 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
curve(x - 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
```

### Working with the `asp20model` package

First, you need to install the `asp20model` package from GitLab and load it:

```{r installation, eval = 3}
install.packages("devtools")
devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
library(asp20model)
```

Using the data from the plot above, we can set up a location-scale regression model. Note that the first command only sets up the design matrices and the parameter vectors but does not do any kind of inference. All parameters are initialized with a value of 0. We can use the `loglik()` method to obtain the log-likelihood of the model in the initial state (i.e. with all parameters set to 0):

```{r loglik1}
model <- LocationScaleRegression$new(y ~ x, ~ x)
model$beta
model$gamma
model$loglik()
```

Now, let's update the parameter values manually and see how the log-likelihood changes:

```{r loglik2}
model$beta <- c(0.1, 0.1)
model$loglik()
```

For most inference algorithms, we also need the gradient of the log-likelihood with respect to the parameters, which we can obtain with the `grad()` method:

```{r grad}
model$grad_beta()
model$grad_gamma()
```

Finally, the `asp20model` package also comes with a simple [gradient descent algorithm](https://en.wikipedia.org/wiki/Gradient_descent) for maximum likelihood inference. Let's apply it to our model:

```{r graddesc1}
gradient_descent(model)
model$beta
model$gamma
model$grad_beta()
model$grad_gamma()
```

Whoops, that didn't work very well! The algorithm didn't converge, the estimated parameters are far away from the true values we used to simulate the data (\$`\hat{\boldsymbol{\beta}}`\$ should be close to (0, 1), \$`\hat{\boldsymbol{\gamma}}`\$ should be close to (-3, 2)), and the gradient is not close to 0. Let's try again with a smaller step size and more iterations:

```{r graddesc2}
gradient_descent(model, stepsize = 1e-06, maxit = 500000)
model$beta
model$gamma
model$grad_beta()
model$grad_gamma()
```

These results do look better, but they also show how inefficient the gradient descent algorithm is. The algorithm took more than a quarter **million** iterations to converge!

Your task is to do better than me and implement a more efficient inference algorithm for the location-scale regression model class. Use my code for the `gradient_descent()` function as an example. Build on the `LocationScaleRegression` R6 class and its methods and extend them if necessary. Think about inheritance and whether your extensions could be useful for other groups. If yes, feel free to share your code on GitLab and open a pull request in the `asp20model` repository.

## Student projects

### `asp20boot`

- Supervisor: Benjamin Säfken
- Students: Alisa, Jasmin, Katharina
- Tasks: Parameter estimation using Fisher scoring, bootstrap confidence intervals for the parameter estimates
- Readings: https://en.wikipedia.org/wiki/Bootstrapping_(statistics), the `boot` R package

### `asp20cv`

- Supervisor: René-Marcel Kruse
- Students: Jan, Malte, Mattias
- Tasks: Parameter estimation using Fisher scoring, model selection using leave-p-out and k-fold cross validation
- Readings: https://en.wikipedia.org/wiki/Cross-validation_(statistics), the cross validation functions in the `gamlss` R package

### `asp20lasso`

- Supervisor: Maike Hohberg
- Students: Marah, Niklas, Sophie
- Tasks: Parameter estimation, variable selection, and regularization using the least absolute shrinkage and selection operator (LASSO)
- Readings: Section 4.2.3 in [1]

### `asp20boost`

- Supervisor: Thomas Kneib
- Students: Johannes, Levin, Sebastián
- Tasks: Parameter estimation and variable selection using the boosting ensemble learning algorithm
- Readings: Section 4.3 in [1]

### `asp20regPrior`

- Supervisor: Paul Wiemann
- Students: Jinyi, Markus, Sören
- Tasks: Interface for the specification of regularization priors, statistical inference using random walk/Langevin MCMC
- Readings: Section 4.4.2 in [1], https://en.wikipedia.org/wiki/Metropolis-adjusted_Langevin_algorithm

### `asp20ssPrior`

- Supervisor: Manuel Carlan
- Students: Not assigned
- Tasks: Interface for the specification of spike and slab priors, statistical inference using random walk/Langevin MCMC
- Readings: Section 4.4.4 in [1], https://en.wikipedia.org/wiki/Metropolis-adjusted_Langevin_algorithm

### `asp20hMCMC`

- Supervisor: Manuel Carlan
- Students: Carsten, Joanna, Tim
- Tasks: Statistical inference using Hamiltonian MCMC with flat priors
- Readings: "A conceptual introduction to Hamiltonian Monte Carlo" by Michael Betancourt, "MCMC using Hamiltonian dynamics" by Radford Neal

### `asp20iwlsMCMC`

- Supervisor: Isa Marques
- Students: Henrike, Kim, Nils
- Tasks: Statistical inference using iterative weighted least squares (IWLS) MCMC with flat priors
- Readings: Section 5.6.2 in [1], "Sampling from the posterior distribution in generalized linear mixed models" by Dani Gamerman, "BAMLSS: Bayesian additive models for location, scale, and shape (and beyond)" by Nikolaus Umlauf, Nadja Klein, and Achim Zeileis

### `asp20plot`

- Supervisor: Hannes Riebl
- Students: Björn, Friederike, Manuel
- Tasks: Parameter estimation based on weighted/generalized least squares (WLS/GLS), plot functions for model predictions and diagnostics
- Readings: Section 3.4.4 in [1], the plot functions in the `car` and `gamlss` R packages

### `asp20user`

- Supervisor: Hannes Riebl
- Students: Not assigned
- Tasks: Parameter estimation based on weighted/generalized least squares (WLS/GLS), evaluation of the user-friendliness of the `asp20*` packages, implementation of convenience functions, application to the `datasets::airquality` dataset
- Readings: Section 4.1 in [1], Section 16.3 in [3]

<!--
more datasets:
- gamlss.data::TODO
- MASS::Boston
- https://archive.ics.uci.edu/ml/datasets.php
-->

## Requirements to pass the course

To get the credits for this module, you need to...

- ... develop an R package that builds on the `asp20model` package and adds the functionality described above. Your code needs to be comprehensible, well documented, and covered by automated tests.
- ... develop and carry out a small simulation study to show that your method works reliably. Please discuss the scope and design of the simulation study with your supervisor.
- Your development process should be continuous, collaborative, and transparent. The commit history on GitLab should reflect your progress at any time.
- ... give two 20-minute presentations on June 9 and July 14.
- ... submit a term paper by September 15.

You may earn "bonus points" for active exchange with other groups, preferably on GitLab via issues and pull requests. For example, the four groups working on Bayesian inference could develop a common API for the specification of priors. If you need additional features in the `asp20model` package, feel free to open issues and pull requests in this repository.

## Readings

- [1] "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx, especially Section 2.9.1 on "Regression models for location, scale, and shape"
- [2] The tidyverse style guide, available online: https://style.tidyverse.org
- [3] "Advanced R" by Hadley Wickham, available online: https://adv-r.hadley.nz
- [4] "R packages" by Hadley Wickham, available online: http://r-pkgs.had.co.nz

### 2020-04-28 R/model.R

@@ -1,0 +1,211 @@
+#' R6 class for location-scale regression models
+#'
+#' This model class assumes a normally distributed response variable with
+#' one linear predictor for the location (i.e. the mean) and one for the scale
+#' (i.e. the standard deviation). The linear predictors for the location and
+#' the scale are called \eqn{X\beta} and \eqn{Z\gamma} respectively. The scale
+#' uses a log link.
+#'
+#' @field beta A numeric vector with the `beta` parameters.
+#' @field gamma A numeric vector with the `gamma` parameters.
+#'
+#' @importFrom R6 R6Class
+#' @export
+
+LocationScaleRegression <- R6Class(
+  classname = "LocationScaleRegression",
+  public = list(
+    #' @details
+    #' Create a new `LocationScaleRegression` object.
+    #'
+    #' @param location A two-sided formula with the response variable on the
+    #'                 LHS and the predictor for the location (i.e. the mean)
+    #'                 on the RHS.
+    #' @param scale A one-sided formula with the predictor for the scale
+    #'              (i.e. the standard deviation) on the RHS.
+    #' @param data A data frame (or list or environment) in which to evaluate
+    #'             the `location` and `scale` formulas.
+    #' @param ... Passed on to [stats::model.matrix()].
+    #'
+    #' @return
+    #' A `LocationScaleRegression` object.
+    #'
+    #' @examples
+    #' y <- rnorm(30)
+    #' LocationScaleRegression$new(y ~ 1)
+    #'
+    #' @importFrom stats model.matrix
+
+    initialize = function(location,
+                          scale = ~1,
+                          data = environment(location),
+                          ...) {
+      scale <- update(scale, paste(location[[2]], "~ ."))
+      private$y <- eval(location[[2]], data, environment(location))
+      private$X <- model.matrix(location, data, ...)
+      private$Z <- model.matrix(scale, data, ...)
+
+      self$beta <- rep.int(0, ncol(private$X))
+      self$gamma <- rep.int(0, ncol(private$Z))
+
+      invisible(self)
+    },
+
+    #' @details
+    #' Returns the log-likelihood of a `LocationScaleRegression` object
+    #' at the current parameter values.
+    #'
+    #' @return
+    #' A single number.
+    #'
+    #' @examples
+    #' y <- rnorm(30)
[…]

### 2020-04-28 gettingStarted.R

@@ -1,35 +1,0 @@
-
-#First run only
-#install.packages("devtools")
-#devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
-
-
-library(asp20model)
-library(tidyverse)
-
-### Example Data
-
-n <- 500
-x <- runif(n)
-y <- x + rnorm(n, sd = exp(-3 + 2 * x))
-plot(x, y)
-abline(0, 1, lwd = 2)
-curve(x + 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
-curve(x - 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
-
-###
-
-### Build simple linear model
-
-simpleModel <- lm(x ~ y)
-summary(simpleModel)
-plot(x = x , y = y, type = 'p')
-abline(simpleModel)
-
-
-### Plot Resdiuals
-simpleModelRes = resid(simpleModel)
-plot(y, simpleModelRes, ylab="Residuals", xlab="X",  main="Residual Plot") 
-abline(0, 0) 
-#It shows Heteroscedasticity
-#Note: Residuals should be standardized / studentized, P.184

### 2020-04-28 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(asp20model)
+
+test_check("asp20model")

### 2020-04-28 tests/testthat.R
Adjusting Package Name in Description and test-file (formerly asp20model, now asp20boost); adjusting autho in Description

@@ -1,4 +1,4 @@
 library(testthat)
-library(asp20model)
+library(asp20boost)
 
-test_check("asp20model")
+test_check("asp20boost")

### 2020-04-28 R/model.R
Implementing empty boost function and adjusting Namespace-file to be able to use it (adjusting namespace-file not recommended as it is autogenerated by roxygen2)

@@ -209,3 +209,16 @@ gradient_descent <- function(model,
   message("Finishing after ", i, " iterations")
   invisible(model)
 }
+
+
+
+
+
+# boost-function first try -----------------------------------------------------------
+boost_levin <- function(model,
+                        weight_learn = 0.1,
+                        m_stop = 1000,
+                        verbose = FALSE) {
+  print("Boost-Function called. Nothing implemented yet. Sorry, Dude.")
+  invisible(model)
+}

### 2020-04-30 README.Rmd
1st commit. Tinkered arround by deleting what I considered to be "irrelevant" information about other projects. I left LASSO still there: They were asked to regularize while we weren't. Interesting for further disscussion.

@@
 
-### `asp20boot`
-
-- Supervisor: Benjamin Säfken
-- Students: Alisa, Jasmin, Katharina
-- Tasks: Parameter estimation using Fisher scoring, bootstrap confidence intervals for the parameter estimates
-- Readings: https://en.wikipedia.org/wiki/Bootstrapping_(statistics), the `boot` R package
-
-### `asp20cv`
-
-- Supervisor: René-Marcel Kruse
-- Students: Jan, Malte, Mattias
-- Tasks: Parameter estimation using Fisher scoring, model selection using leave-p-out and k-fold cross validation
-- Readings: https://en.wikipedia.org/wiki/Cross-validation_(statistics), the cross validation functions in the `gamlss` R package
-
 ### `asp20lasso`
@@
 - Readings: Section 4.3 in [1]
-
-### `asp20regPrior`
-
-- Supervisor: Paul Wiemann
-- Students: Jinyi, Markus, Sören
-- Tasks: Interface for the specification of regularization priors, statistical inference using random walk/Langevin MCMC
-- Readings: Section 4.4.2 in [1], https://en.wikipedia.org/wiki/Metropolis-adjusted_Langevin_algorithm
-
-### `asp20ssPrior`
-
-- Supervisor: Manuel Carlan
-- Students: Not assigned
-- Tasks: Interface for the specification of spike and slab priors, statistical inference using random walk/Langevin MCMC
-- Readings: Section 4.4.4 in [1], https://en.wikipedia.org/wiki/Metropolis-adjusted_Langevin_algorithm
-
-### `asp20hMCMC`
-
-- Supervisor: Manuel Carlan
-- Students: Carsten, Joanna, Tim
-- Tasks: Statistical inference using Hamiltonian MCMC with flat priors
-- Readings: "A conceptual introduction to Hamiltonian Monte Carlo" by Michael Betancourt, "MCMC using Hamiltonian dynamics" by Radford Neal
-
-### `asp20iwlsMCMC`
-
-- Supervisor: Isa Marques
-- Students: Henrike, Kim, Nils
-- Tasks: Statistical inference using iterative weighted least squares (IWLS) MCMC with flat priors
-- Readings: Section 5.6.2 in [1], "Sampling from the posterior distribution in generalized linear mixed models" by Dani Gamerman, "BAMLSS: Bayesian additive models for location, scale, and shape (and beyond)" by Nikolaus Umlauf, Nadja Klein, and Achim Zeileis
-
-### `asp20plot`
-
-- Supervisor: Hannes Riebl
-- Students: Björn, Friederike, Manuel
-- Tasks: Parameter estimation based on weighted/generalized least squares (WLS/GLS), plot functions for model predictions and diagnostics
-- Readings: Section 3.4.4 in [1], the plot functions in the `car` and `gamlss` R packages
-
-### `asp20user`
-
-- Supervisor: Hannes Riebl
-- Students: Not assigned
-- Tasks: Parameter estimation based on weighted/generalized least squares (WLS/GLS), evaluation of the user-friendliness of the `asp20*` packages, implementation of convenience functions, application to the `datasets::airquality` dataset
-- Readings: Section 4.1 in [1], Section 16.3 in [3]
 

### 2020-04-30 gettingStarted.R
test

@@ -33,3 +33,56 @@ plot(y, simpleModelRes, ylab="Residuals", xlab="X",  main="Residual Plot")
 abline(0, 0) 
 #It shows Heteroscedasticity
 #Note: Residuals should be standardized / studentized, P.184
+#
+LocationScaleRegression <- R6Class("LocationScaleRegression")
+
+#
+#Extend LocationScaleRegression Model 
+#
+gradient_descentBoost <- function(model,
+                             stepsize = 0.001,
+                             maxit = 1000,
+                             abstol = 0.001,
+                             verbose = FALSE) {
+  
+  grad_beta <- model$grad_beta()
+  grad_gamma <- model$grad_gamma()
+  
+  for (i in seq_len(maxit)) {
+    model$beta <- model$beta + stepsize * grad_beta
+    model$gamma <- model$gamma + stepsize * grad_gamma
+    
+    grad_beta <- model$grad_beta()
+    grad_gamma <- model$grad_gamma()
+    
+    if (verbose) {
+      par_msg <- c(model$beta, model$gamma)
+      par_msg <- format(par_msg, trim = TRUE, digits = 3)
+      par_msg <- paste(par_msg, collapse = " ")
+      
+      grad_msg <- c(grad_beta, grad_gamma)
+      grad_msg <- format(grad_msg, trim = TRUE, digits = 3)
+      grad_msg <- paste(grad_msg, collapse = " ")
+      
+      loglik_msg <- format(model$loglik(), digits = 3)
+      
+      message(
+        "Iteration:      ", i, "\n",
+        "Parameters:     ", par_msg, "\n",
+        "Gradient:       ", grad_msg, "\n",
+        "Log-likelihood: ", loglik_msg, "\n",
+        "==============="
+      )
+    }
+    
+    if (all(abs(c(grad_beta, grad_gamma)) <= abstol)) break
+  }
+  
+  message("Finishing after ", i, " iterations")
+  invisible(model)
+}
+
+gradient_descentBoost(model)
+model$grad_gamma()
+model <- LocationScaleRegression$new(y ~ x, ~ x)
+

### 2020-04-30 TO_DO.Rmd

---
title: 'Goals and Tasks'
author: "ASP Boost Team"
date: "30 4 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I propose this file for technical discussion and goal setting and tracking. 
Goal setting could look like this (This is merely a sketch):

## Tasks to achieve till next Q&A session 20.05.04 
### (as set in the last meeting)

* ~~README~~ L. 
* ~~Rmd file for communication~~ S (this file)
* Push a boost function JLS.

Initials could indicate whom belongs which task.

Discussion could look like this (The html file is intended to visualize LaTeX):

## Estimator:

Residuals: $u = y = x\hat{\beta}$
Estimator *coming from the residuals*: $\hat{b} = (X'X)^{-1} X'u$

After inserting $u$ in $\hat{b}$ the base formula results:
$\hat{\beta}^{t}= \hat{\beta}^{t-1} + v(\hat{\beta} - \hat{\beta}^{t-1})$

I am unsure, if the first term inside parenthesis is just $\hat{\beta}$ or rather $\hat{\beta}^{something}$. Please check.

### 2020-04-30 LocationScaleRegressionBoost.R
Initial commit of extended LocationScaleregression R6 Class

@@ -1,0 +1,97 @@
+library(asp20model)
+library(tidyverse)
+
+
+LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
+                                        inherit = LocationScaleRegression,
+                                        
+                                        private = list(
+                                          residScale = numeric(),
+                                          update_gamma = function(value) {
+                                            private$.gamma <- value
+                                            private$fitted$scale <- exp(drop(private$Z %*% self$gamma))
+                                            private$residScale <- private$y - private$fitted$scale
+                                            invisible(self)
+                                          }
+                                        ),
+                                        
+                                        
+                                        active = list(
+                                          lstSqrResid = function()
+                                          {
+                                            
+                                            X_tr_X_inv <- solve(t(private$X) %*% private$X)    
+                                            Proj_M <- X_tr_X_inv %*% t(private$X)
+                                            Proj_M %*% private$resid
+                                          },
+                                          
+                                          lstSqrResidScale = function()
+                                          {
+                                            #
+                                            # How to calculate the factor to change for gamma?
+                                            #
+                                            X_tr_X_inv2 <- solve(t(private$X) %*% private$X)    
+                                            Proj_M2 <- X_tr_X_inv2 %*% t(private$X)
+                                            #Proj_M2 %*% private$residScale
+                                            #(resid / scale)^2 - 1
+                                          }
+                                          
+                                        )
+                                        
+                                        
+)
+
+
+
+
+boostJohannes = function(model,
+                         stepsize = 0.001,
+                         maxit = 10000,
+                         abstol = 0.1,
+                         verbose = FALSE) {
+  grad_beta <- model$grad_beta()
+  grad_gamma <- model$grad_gamma()
+  
+  message("boost js")
+  v <- 0.1
+  
+  for (i in seq_len(maxit)) {
+    model$beta <- model$beta + v*model$lstSqrResid
+    #message(model$beta)
+    grad_beta <- model$grad_beta()
+    ##todo gamma
+    model$gamma <- model$gamma + v*model$lstSqrResid
+    #message(model$loglik())
+    
+  }
+  
+  invisible(model)
+}
+
+##
+##Test
+##
[…]

### 2020-04-30 LocationScaleRegressionBoost.R
fix include R6 libary

@@ -1,6 +1,6 @@
 library(asp20model)
 library(tidyverse)
-
+library(R6)
 
 LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                         inherit = LocationScaleRegression,

### 2020-05-04 TO_DO.Rmd
new tasks clipped

@@
 author: "ASP Boost Team"
-date: "30 4 2020"
+date: "30 April 2020"
 output: html_document
@@
 
-I propose this file for technical discussion and goal setting and tracking. 
-Goal setting could look like this (This is merely a sketch):
+This file is for goal setting and tracking and technical discussion.
 
-## Tasks to achieve till next Q&A session 20.05.04 
+## Tasks until ~ 11 May
+
+* simple boost function for scale (estimate var(yi))
+* read Wickham "R-Packages"
+* try implementing function-testing
+* get used to roxygen2
+
+## Tasks to achieve till next Q&A session 04 May 2020 
 ### (as set in the last meeting)
 
-* ~~README~~ L. 
-* ~~Rmd file for communication~~ S (this file)
-* Push a boost function JLS.
+* L ~~README~~ . 
+* S ~~Rmd file for communication~~  (this file)
+* JLS ~~Push a boost function~~
 
-Initials could indicate whom belongs which task.
+Initials indicate whom belongs which task.
 

### 2020-05-04 README.Rmd
Cleaned up readme, added tasklist to readme

@@
 
-# Advanced statistical programming with R
+# asp20boost
 
-Are you taking the course "Advanced statistical programming with R" in the summer term 2020?
-Then this page is for you.
-It explains what you need to do to get the credits for this module.
-`r emo::ji("wink")`
+* Supervisor: Thomas Kneib
+* Students: Johannes, Levin, Sebastián
+* Tasks: Parameter estimation and variable selection using the boosting ensemble learning algorithm
 
-In this course, you are going work on a programming project in a group of three students.
-Each group needs to develop an R package, in which they implement a statistical method for the so-called location-scale regression model class, and a small simulation study to evaluate their method and implementation.
-In September, you need to submit a term paper about your project.
-
-This git repository contains the `asp20model` package for R, which implements the location-scale regression model class and will be the basis for your package.
-
-**Please note:** If you are planning to take this course, **you need to preregister on Stud.IP before March 31.** Follow this link to sign up: https://studip.uni-goettingen.de/dispatch.php/course/details?sem_id=5b8822a839a628e8166648a57f4f1eac.
-
-**Update on the Corona crisis:** Unsurprisingly, the Corona crisis makes some changes to this course necessary.
-All previously announced dates remain valid, but I am going to offer the introductory session and the Q&A session as YouTube livestreams.
-See below for the links.
-The intermediate and final presentations might be replaced with short written reports, depending on the further developments.
-I am going to make a final decision on the presentations before the Q&A session on May 4.
-
-## Important dates
-
-- Tuesday, April 7, 10:00--15:00, ~~Blauer Turm / MZG 8.163~~: Introductory session, project assignment \
-  YouTube livestream: https://youtu.be/-a02j6YTxKw \
-  Rocket.Chat channel for questions and comments: https://chat.gwdg.de/channel/asp20
-- Monday, May 4, 14:00--16:00, ~~Blauer Turm / MZG 8.163~~: Q&A session, feel free to ask technical, statistical, and organizational questions \
-  YouTube livestream: https://youtu.be/vi2BqiDUCFI \
-  Rocket.Chat channel for questions and comments: https://chat.gwdg.de/channel/asp20
-- Friday, June 5, 8:30--15:00, Blauer Turm / MZG 8.163: Intermediate presentations
-- Friday, July 10, 8:30--15:00, Blauer Turm / MZG 8.163: Final presentations
-- Tuesday, September 15: Submission deadline for the term papers
-
-## Technical & statistical prerequisites
-
-### The R6 OOP system
-
-The `asp20model` package uses R6 classes, so make sure to get familiar with mutable objects, inheritance, etc. See the chapter on R6 classes in [3], available online: https://adv-r.hadley.nz/r6.html.
-
-### Version control using git & GitLab
-
-The `asp20model` package is hosted on https://gitlab.gwdg.de (the GitLab instance of the computing center of the university), where you are going to develop your R package as well.
-I created a GitLab group for this course and a GitLab project for each group of students.
-
-**Please note:** To start working on your project, log in to GitLab with your student account once and send me an email. After that, I can add you to the group and the project.
-
-GitLab is an online platform where you can host git repositories, very similar to GitHub, and git is a popular version control system for software development.
-Systematic, version-controlled software development is an essential component of this course, so if you are not yet familiar with git, please take a look at the git chapter in [4], available online: http://r-pkgs.had.co.nz/git.html.
-
-**Please note:** Your code will only be visible to me, the project supervisors, and your fellow students. Feel free to make mistakes and ask questions!
-
-### Location-scale regression
-
-As mentioned above, we are going to work with the location-scale regression model class in this course.
-So, what **is** location-scale regression?
-Let's take a step back and look at the standard linear model first.
-It is defined as
-
-``` math
-y_i = \boldsymbol{x}_i' \boldsymbol{\beta} + \varepsilon_i, \text{ where } \varepsilon_i \overset{i.i.d.}{\sim} \mathcal{N}(0, \sigma^2).
-```
-
-From the definition, it follows that
-
-``` math
-y_i \overset{ind.}{\sim} \mathcal{N}(\boldsymbol{x}_i' \boldsymbol{\beta}, \sigma^2),
-```
-
-and from this representation, the model can easily be extended with a second linear predictor for the standard deviation.
-For this purpose, we introduce a covariate vector \$`\boldsymbol{z}_i`\$ and a parameter vector \$`\boldsymbol{\gamma}`\$:
-
-``` math
-y_i \overset{ind.}{\sim} \mathcal{N}(\boldsymbol{x}_i' \boldsymbol{\beta}, (\exp(\boldsymbol{z}_i' \boldsymbol{\gamma}))^2).
-```
-
-What is the \$`\exp`\$ function doing in this formula?
-Well, one thing that is a little bit tricky about predicting the standard deviation is that the standard deviation needs to be positive, while the linear predictor \$`\boldsymbol{z}_i' \boldsymbol{\gamma}`\$ can become negative for some choices of the parameter vector \$`\boldsymbol{\gamma}`\$.
-Hence, to ensure that our predictions for the standard deviation are valid, we introduce the \$`\exp`\$ function as a so-called response function for the standard deviation and define \$`\sigma_i = \exp(\boldsymbol{z}_i' \boldsymbol{\gamma})`\$.
-
-Why should we care about location-scale regression?
-It is not uncommon to observe heteroscedastic data like in the figure below.
-The standard way to deal with heteroscedastic residuals in a linear model is FGLS estimation, but FGLS does not provide an interpretable description of the variance structure of the data.
-A location-scale regression model can fill this gap, if we have access to explanatory variables for the variance or the standard deviation of the response variable.
+## The Model
 
@@
 
-### Working with the `asp20model` package
+## Tasks
 
-First, you need to install the `asp20model` package from GitLab and load it:
+### until ~ 11 May
 
-```{r installation, eval = 3}
-install.packages("devtools")
-devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
-library(asp20model)
-```
+* simple boost function for scale (estimate var(yi))
+* read Wickham "R-Packages"
+* try implementing function-testing
+* get used to roxygen2
 
-Using the data from the plot above, we can set up a location-scale regression model.
-Note that the first command only sets up the design matrices and the parameter vectors but does not do any kind of inference.
-All parameters are initialized with a value of 0.
-We can use the `loglik()` method to obtain the log-likelihood of the model in the initial state (i.e.
-with all parameters set to 0):
-
-```{r loglik1}
-model <- LocationScaleRegression$new(y ~ x, ~ x)
-model$beta
-model$gamma
-model$loglik()
-```
-
-Now, let's update the parameter values manually and see how the log-likelihood changes:
-
-```{r loglik2}
-model$beta <- c(0.1, 0.1)
-model$loglik()
-```
-
-For most inference algorithms, we also need the gradient of the log-likelihood with respect to the parameters, which we can obtain with the `grad()` method:
-
-```{r grad}
-model$grad_beta()
-model$grad_gamma()
-```
-
-Finally, the `asp20model` package also comes with a simple [gradient descent algorithm](https://en.wikipedia.org/wiki/Gradient_descent) for maximum likelihood inference. Let's apply it to our model:
-
-```{r graddesc1}
-gradient_descent(model)
-model$beta
-model$gamma
-model$grad_beta()
-model$grad_gamma()
-```
-
-Whoops, that didn't work very well!
-The algorithm didn't converge, the estimated parameters are far away from the true values we used to simulate the data (\$`\hat{\boldsymbol{\beta}}`\$ should be close to (0, 1), \$`\hat{\boldsymbol{\gamma}}`\$ should be close to (-3, 2)), and the gradient is not close to 0.
-Let's try again with a smaller step size and more iterations:
-
-```{r graddesc2}
-gradient_descent(model, stepsize = 1e-06, maxit = 500000)
-model$beta
-model$gamma
-model$grad_beta()
-model$grad_gamma()
-```
-
-These results do look better, but they also show how inefficient the gradient descent algorithm is. The algorithm took more than a quarter **million** iterations to converge!
-
-Your task is to do better than me and implement a more efficient inference algorithm for the location-scale regression model class.
-Use my code for the `gradient_descent()` function as an example.
-Build on the `LocationScaleRegression` R6 class and its methods and extend them if necessary.
-Think about inheritance and whether your extensions could be useful for other groups.
-If yes, feel free to share your code on GitLab and open a pull request in the `asp20model` repository.
-
-## Student projects
-
-### `asp20lasso`
-
-- Supervisor: Maike Hohberg
-- Students: Marah, Niklas, Sophie
-- Tasks: Parameter estimation, variable selection, and regularization using the least absolute shrinkage and selection operator (LASSO)
-- Readings: Section 4.2.3 in [1]
-
-### `asp20boost`
-
-- Supervisor: Thomas Kneib
-- Students: Johannes, Levin, Sebastián
-- Tasks: Parameter estimation and variable selection using the boosting ensemble learning algorithm
-- Readings: Section 4.3 in [1]
+## Literature
+* Section **4.3** in "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx, especially Section 2.9.1 on "Regression models for location, scale, and shape"
 
@@
 -->
-
-## Requirements to pass the course
-
-To get the credits for this module, you need to...
-
-- ... develop an R package that builds on the `asp20model` package and adds the functionality described above. Your code needs to be comprehensible, well documented, and covered by automated tests.
-- ... develop and carry out a small simulation study to show that your method works reliably. Please discuss the scope and design of the simulation study with your supervisor.
-- Your development process should be continuous, collaborative, and transparent. The commit history on GitLab should reflect your progress at any time.
-- ... give two 20-minute presentations on June 9 and July 14.
-- ... submit a term paper by September 15.
-
-You may earn "bonus points" for active exchange with other groups, preferably on GitLab via issues and pull requests.
-For example, the four groups working on Bayesian inference could develop a common API for the specification of priors.
-If you need additional features in the `asp20model` package, feel free to open issues and pull requests in this repository.
-
-## Readings
-
-- [1] "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx, especially Section 2.9.1 on "Regression models for location, scale, and shape"
-- [2] The tidyverse style guide, available online: https://style.tidyverse.org
-- [3] "Advanced R" by Hadley Wickham, available online: https://adv-r.hadley.nz
-- [4] "R packages" by Hadley Wickham, available online: http://r-pkgs.had.co.nz

### 2020-05-04 README.Rmd
clipped task list to readme

@@
 
+## Tasks
+
+### until ~ 11 May
+
+* simple boost function for scale (estimate var(yi))
+* read Wickham "R-Packages"
+* try implementing function-testing
+* get used to roxygen2
+
 ## The Model
@@
 
-## Tasks
-
-### until ~ 11 May
-
-* simple boost function for scale (estimate var(yi))
-* read Wickham "R-Packages"
-* try implementing function-testing
-* get used to roxygen2
-
 ## Literature
-* Section **4.3** in "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx, especially Section 2.9.1 on "Regression models for location, scale, and shape"
+* Section **4.3** on "Boosting Linear Regression Models" in "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx
+* Section **2.9.1** on "Regression models for location, scale, and shape" in ibd.
 

### 2020-05-04 README.Rmd
test

@@
 * Students: Johannes, Levin, Sebastián
-* Tasks: Parameter estimation and variable selection using the boosting ensemble learning algorithm
+* Topic: Parameter estimation and variable selection using the boosting ensemble learning algorithm
 
 ## Tasks 
+ignore this subheader
 

### 2020-05-05 LocationScaleRegressionBoost.R
use deviance for least Suare Residuals of Z

@@ -1,42 +1,37 @@
+#devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de", force=TRUE)
 library(asp20model)
 library(tidyverse)
 library(R6)
 
 LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                         inherit = LocationScaleRegression,
-                                        
                                         private = list(
-                                          residScale = numeric(),
-                                          update_gamma = function(value) {
-                                            private$.gamma <- value
-                                            private$fitted$scale <- exp(drop(private$Z %*% self$gamma))
-                                            private$residScale <- private$y - private$fitted$scale
-                                            invisible(self)
-                                          }
+                                          update_beta = function(value) {
+                                          private$.beta <- value
+                                          self$fitted_location <- drop(private$X %*% self$beta)
+                                          private$.resid <- private$y - self$fitted_location
+                                          #message("update beta \n")
+                                          #message( self$beta)
+                                          invisible(self)
+                                        }
                                         ),
-                                        
-                                        
-                                        active = list(
+                                   active = list(
                                           lstSqrResid = function()
                                           {
                                             
                                             X_tr_X_inv <- solve(t(private$X) %*% private$X)    
                                             Proj_M <- X_tr_X_inv %*% t(private$X)
-                                            Proj_M %*% private$resid
+                                            Proj_M %*% self$resid()
                                           },
-                                          
-                                          lstSqrResidScale = function()
+                                          lstSqrResidZ = function()
                                           {
-                                            #
-                                            # How to calculate the factor to change for gamma?
-                                            #
[…]

### 2020-05-05 LocationScaleRegressionBoost.R
c

@@ -5,16 +5,7 @@ library(R6)
 
 LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                         inherit = LocationScaleRegression,
-                                        private = list(
-                                          update_beta = function(value) {
-                                          private$.beta <- value
-                                          self$fitted_location <- drop(private$X %*% self$beta)
-                                          private$.resid <- private$y - self$fitted_location
-                                          #message("update beta \n")
-                                          #message( self$beta)
-                                          invisible(self)
-                                        }
-                                        ),
+                                       
                                    active = list(
                                           lstSqrResid = function()
                                           {
@@ -104,7 +95,7 @@ model$gamma<-c(0,0)
 model$beta<-c(0,0)
 model$loglik()
 boostJohannes(model, 
-              stepsize = 0.0001, maxit = 40000,
+              stepsize = 0.001, maxit = 40000,
               abstol = 0.001,
               verbose = TRUE)
 

### 2020-05-07 LocationScaleRegressionBoost.R
implement partial derivative for gamma

@@ -5,24 +5,71 @@ library(R6)
 
 LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                         inherit = LocationScaleRegression,
-                                       
-                                   active = list(
+                                        public = list(
+                                          
+                                          X_tr_X_inv = numeric(),
+                                          Proj_M_X = numeric(),
+                                          Z_tr_Z_inv = numeric(),
+                                          Proj_M_Z = numeric(),
+                                          y_upperTube = numeric(),
+                                          y_downTube = numeric(),
+                                          
+                                          initProj_M = function() {
+                                            #𝑃𝑟𝑜𝑗𝑀𝑎𝑡𝑟𝑖𝑥=(𝑋𝑇𝑋)−1𝑋𝑇
+                                            #https://stats.stackexchange.com/questions/154485/least-squares-regression-step-by-step-linear-algebra-computation?noredirect=1&lq=1
+                                            self$X_tr_X_inv <- solve(t(private$X) %*% private$X)    
+                                            self$Proj_M_X <- self$X_tr_X_inv %*% t(private$X)
+                                            
+                                            self$Z_tr_Z_inv <- solve(t(private$Z) %*% private$Z)    
+                                            self$Proj_M_Z <- self$Z_tr_Z_inv %*% t(private$Z)
+                                            
+                                          },
+                                          negloglik = function() {
+                                            location <- self$fitted_location
+                                            scale <- self$fitted_scale
+                                            
+                                            sum(dnorm(private$y, location, scale, log = TRUE))
+                                          }
+                                          
+                                          
+                                          
+                                          
+                                        ),
+                                        active = list(
                                           lstSqrResid = function()
                                           {
                                             
-                                            X_tr_X_inv <- solve(t(private$X) %*% private$X)    
-                                            Proj_M <- X_tr_X_inv %*% t(private$X)
-                                            Proj_M %*% self$resid()
+                                            self$Proj_M_X %*% self$resid()
                                           },
[…]

### 2020-05-07 LocationScaleRegressionBoost.spin.R

@@ -1,137 +1,0 @@
-#devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de", force=TRUE)
-library(asp20model)
-library(tidyverse)
-library(R6)
-
-LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
-                                        inherit = LocationScaleRegression,
-                                       
-                                   active = list(
-                                          lstSqrResid = function()
-                                          {
-                                            
-                                            X_tr_X_inv <- solve(t(private$X) %*% private$X)    
-                                            Proj_M <- X_tr_X_inv %*% t(private$X)
-                                            Proj_M %*% self$resid()
-                                          },
-                                          lstSqrResidZ = function()
-                                          {
-                                            
-                                            X_tr_X_inv <- solve(t(private$X) %*% private$X)    
-                                            Proj_M <- X_tr_X_inv %*% t(private$X)
-                                            
-                                            Proj_M %*% (self$resid("deviance"))
-                                          }
-                                          )
-                                        
-                                        
-)
-
-
-
-
-boostJohannes = function(model,
-                         stepsize = 0.001,
-                         maxit = 1000,
-                         abstol = 0.001,
-                         verbose = TRUE) {
-  grad_beta <- model$grad_beta()
-  grad_gamma <- model$grad_gamma()
-  
-  message("boost js")
-  v <- stepsize
-  
-  
-  #check if location scale model
-  for (i in seq_len(maxit)) {
-    model$beta <- model$beta + v*model$lstSqrResid
-    grad_beta <- model$grad_beta()
-    
-    model$gamma <- model$gamma + 0.00001*model$lstSqrResidZ
-   
-    grad_gamma <- model$grad_gamma()
-    
-    
-    
-    if (verbose) {
-      par_msg <- c(model$beta, model$gamma)
-      par_msg <- format(par_msg, trim = TRUE, digits = 3)
-      par_msg <- paste(par_msg, collapse = " ")
-      
-      grad_msg <- c(grad_beta, grad_gamma)
[…]

### 2020-05-12 README.Rmd
clipped task list
Re-Init Repo to match R Package Guidelines and include  roxygen syntax

@@
 ## Tasks
-ignore this subheader
 
@@
 
-## The Model
+## Literature
+* Section **4.3** on "Boosting Linear Regression Models" in "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx
+* Section **2.9.1** on "Regression models for location, scale, and shape" in ibd.
+
+<!--
+more datasets:
+- gamlss.data::TODO
+- MASS::Boston
+- https://archive.ics.uci.edu/ml/datasets.php
+-->
+
+## Location Scale Regression Model
 
@@
 
-## Literature
-* Section **4.3** on "Boosting Linear Regression Models" in "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx
-* Section **2.9.1** on "Regression models for location, scale, and shape" in ibd.
+## Location Scale Regression Boost Model
 
-<!--
-more datasets:
-- gamlss.data::TODO
-- MASS::Boston
-- https://archive.ics.uci.edu/ml/datasets.php
--->
+```{r}
+library(asp20boost)
+model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+gradient_boost(model, 
+              stepsize = 0.001, maxit = 1000,
+              abstol = 0.001,
+              verbose = FALSE)
+plot(x, y)
+lines(x,model$fitted_location)
+#Todo do create similar graphic
+curve(x + 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)
+curve(x - 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)
+
+```

### 2020-05-12 LocationScaleRegressionBoost.R

@@ -1,200 +1,0 @@
-#devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de", force=TRUE)
-library(asp20model)
-library(tidyverse)
-library(R6)
-
-LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
-                                        inherit = LocationScaleRegression,
-                                        public = list(
-                                          
-                                          X_tr_X_inv = numeric(),
-                                          Proj_M_X = numeric(),
-                                          Z_tr_Z_inv = numeric(),
-                                          Proj_M_Z = numeric(),
-                                          y_upperTube = numeric(),
-                                          y_downTube = numeric(),
-                                          
-                                          initProj_M = function() {
-                                            #𝑃𝑟𝑜𝑗𝑀𝑎𝑡𝑟𝑖𝑥=(𝑋𝑇𝑋)−1𝑋𝑇
-                                            #https://stats.stackexchange.com/questions/154485/least-squares-regression-step-by-step-linear-algebra-computation?noredirect=1&lq=1
-                                            self$X_tr_X_inv <- solve(t(private$X) %*% private$X)    
-                                            self$Proj_M_X <- self$X_tr_X_inv %*% t(private$X)
-                                            
-                                            self$Z_tr_Z_inv <- solve(t(private$Z) %*% private$Z)    
-                                            self$Proj_M_Z <- self$Z_tr_Z_inv %*% t(private$Z)
-                                            
-                                          },
-                                          negloglik = function() {
-                                            location <- self$fitted_location
-                                            scale <- self$fitted_scale
-                                            
-                                            sum(dnorm(private$y, location, scale, log = TRUE))
-                                          }
-                                          
-                                          
-                                          
-                                          
-                                        ),
-                                        active = list(
-                                          lstSqrResid = function()
-                                          {
-                                            
-                                            self$Proj_M_X %*% self$resid()
-                                          },
-                                          
-                                          derivative1 = function()
[…]

### 2020-05-12 R/model.R

@@ -8,150 +8,91 @@
 #'
 #' @field beta A numeric vector with the `beta` parameters.
 #' @field gamma A numeric vector with the `gamma` parameters.
+#' @field fitted_location A numeric vector with the fitted values
+#'                        for the location.
+#' @field fitted_scale A numeric vector with the fitted values
+#'                     for the scale.
 #'
-#' @importFrom R6 R6Class
+#' @import R6
+#' @import asp20model
 #' @export
 
-LocationScaleRegression <- R6Class(
-  classname = "LocationScaleRegression",
-  public = list(
-    #' @details
-    #' Create a new `LocationScaleRegression` object.
-    #'
-    #' @param location A two-sided formula with the response variable on the
-    #'                 LHS and the predictor for the location (i.e. the mean)
-    #'                 on the RHS.
-    #' @param scale A one-sided formula with the predictor for the scale
-    #'              (i.e. the standard deviation) on the RHS.
-    #' @param data A data frame (or list or environment) in which to evaluate
-    #'             the `location` and `scale` formulas.
-    #' @param ... Passed on to [stats::model.matrix()].
-    #'
-    #' @return
-    #' A `LocationScaleRegression` object.
-    #'
-    #' @examples
-    #' y <- rnorm(30)
-    #' LocationScaleRegression$new(y ~ 1)
-    #'
-    #' @importFrom stats model.matrix
+LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
+                                        inherit = LocationScaleRegression,
+                                        public = list(
 
-    initialize = function(location,
-                          scale = ~1,
-                          data = environment(location),
-                          ...) {
-      scale <- update(scale, paste(location[[2]], "~ ."))
-      private$y <- eval(location[[2]], data, environment(location))
-      private$X <- model.matrix(location, data, ...)
-      private$Z <- model.matrix(scale, data, ...)
+                                          X_tr_X_inv = numeric(),
+                                          Proj_M_X = numeric(),
+                                          Z_tr_Z_inv = numeric(),
+                                          Proj_M_Z = numeric(),
 
-      self$beta <- rep.int(0, ncol(private$X))
-      self$gamma <- rep.int(0, ncol(private$Z))
+                                          initialize = function(location,
+                                                                scale = ~1,
+                                                                data = environment(location),
[…]

### 2020-05-12 gettingStarted.R

@@ -1,88 +1,0 @@
-
-#First run only
-#install.packages("devtools")
-#devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
-
-
-library(asp20model)
-library(tidyverse)
-
-### Example Data
-
-n <- 500
-x <- runif(n)
-y <- x + rnorm(n, sd = exp(-3 + 2 * x))
-plot(x, y)
-abline(0, 1, lwd = 2)
-curve(x + 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
-curve(x - 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
-
-###
-
-### Build simple linear model
-
-simpleModel <- lm(x ~ y)
-summary(simpleModel)
-plot(x = x , y = y, type = 'p')
-abline(simpleModel)
-
-
-### Plot Resdiuals
-simpleModelRes = resid(simpleModel)
-plot(y, simpleModelRes, ylab="Residuals", xlab="X",  main="Residual Plot") 
-abline(0, 0) 
-#It shows Heteroscedasticity
-#Note: Residuals should be standardized / studentized, P.184
-#
-LocationScaleRegression <- R6Class("LocationScaleRegression")
-
-#
-#Extend LocationScaleRegression Model 
-#
-gradient_descentBoost <- function(model,
-                             stepsize = 0.001,
-                             maxit = 1000,
-                             abstol = 0.001,
-                             verbose = FALSE) {
-  
-  grad_beta <- model$grad_beta()
-  grad_gamma <- model$grad_gamma()
-  
-  for (i in seq_len(maxit)) {
-    model$beta <- model$beta + stepsize * grad_beta
-    model$gamma <- model$gamma + stepsize * grad_gamma
-    
-    grad_beta <- model$grad_beta()
-    grad_gamma <- model$grad_gamma()
-    
-    if (verbose) {
-      par_msg <- c(model$beta, model$gamma)
-      par_msg <- format(par_msg, trim = TRUE, digits = 3)
-      par_msg <- paste(par_msg, collapse = " ")
-      
-      grad_msg <- c(grad_beta, grad_gamma)
-      grad_msg <- format(grad_msg, trim = TRUE, digits = 3)
-      grad_msg <- paste(grad_msg, collapse = " ")
-      
-      loglik_msg <- format(model$loglik(), digits = 3)
-      
-      message(
[…]

### 2020-05-12 pseudo_tests/call_boost_functions.R

@@ -1,17 +1,0 @@
-### --------------------------------------------------------------
-# Function-Calls -------------------------------------------------
-
-# This file simply tries to asses the boost-function we intend to push.
-
-
-# Load our own package -------------------------------------------
-library(asp20boost)
-
-
-# Construct pseudo-model to pass to boost-function ---------------
-x <- 1
-
-# Call Functions -------------------------------------------------
-boost_levin(x)
-try(boost_sebastian(x))
-try(boost_johannes(x))

### 2020-05-12 tests/testthat/test-test-model.R

@@ -8,7 +8,7 @@ x2 <- runif(n)
 x3 <- runif(n)
 y <- rnorm(n, x1 + x3, exp(-3 + x2 + x3))
 
-model <- LocationScaleRegression$new(y ~ x1 + x3, ~ x2 + x3)
+model <- LocationScaleRegressionBoost$new(y ~ x1 + x3, ~ x2 + x3)
 
 f <- function(x) {
   model <- model$clone()
@@ -30,6 +30,8 @@ test_that("gamma gradient works", {
   expect_equivalent(model$grad_gamma(), grad(f, model$gamma))
 })
 
-test_that("gradient descent does not throw an error", {
-  expect_error(gradient_descent(model, verbose = TRUE), NA)
-})
+
+#test_that("gradient boost does not throw an error", {
+#  expect_error(gradient_boost(model, verbose = TRUE), NA)
+#})
+

### 2020-05-12 README.Rmd
Update to readme

--- 2020-05-04 README.Rmd
@@
 
-## Literature
-* Section **4.3** on "Boosting Linear Regression Models" in "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx
-* Section **2.9.1** on "Regression models for location, scale, and shape" in ibd.
+## Working with Location Scale Regression Boost Model
+First, you need to install the `asp20model` &  `asp20boost` package from GitLab and load it:
 
-<!--
-more datasets:
-- gamlss.data::TODO
-- MASS::Boston
-- https://archive.ics.uci.edu/ml/datasets.php
--->
+```{r installation, eval = 3}
+install.packages("devtools")
+devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
+devtools::install_gitlab("asp20/asp20boost", host = "gitlab.gwdg.de")
+library(asp20boost)
+```
+
+```{r}
+model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+gradient_boost(model, 
+              stepsize = 0.001, maxit = 1000,
+              abstol = 0.001,
+              verbose = FALSE)
+plot(x, y)
+lines(x,model$fitted_location)
+#Todo do create similar graphic
+curve(x + 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)
+curve(x - 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)
+
+```

### 2020-05-12 README.Rmd
small change to readme

--- 2020-05-12 README.Rmd
@@
 
-## Location Scale Regression Boost Model
+## Working with Location Scale Regression Boost Model
+First, you need to install the `asp20model` &  `asp20boost` package from GitLab and load it:
+
+```{r installation, eval = 3}
+install.packages("devtools")
+#This throws an error for asp20boost
+#devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
+#devtools::install_gitlab("asp20/asp20boost", host = "gitlab.gwdg.de")
+
+```
 

### 2020-05-22 R/model.R
update componentwise boosting for Beta

@@ -18,7 +18,7 @@
 #' @export
 
 
-#
+library(R6)
 
 
 LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
@@ -75,42 +75,25 @@ LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                           #' model$<-derivativesZ
                                           derivativesZ = function()
                                           {
-                                            #Calculates current first derivates for Zn of Log Likelehood
+                                            #Calculates current first derivates for Gamma n of Log Likelehood
                                             #Could be further improved by moving the matrix multiplaction to the update gamma function to reduce the calculation time
                                             Z_countCoefficient <- dim(model$getZ)[2]
-                                            coefficients <- numeric()
+                                            derivatives <- numeric()
                                             for(n in 1:dim(private$Z)[2]) {
-                                              coefficients[n]<-sum((self$resid()^2)*private$Z[,n]*exp(-2*(drop(private$Z %*% self$gamma)))-private$Z[,n])
+                                              derivatives[n]<-sum((self$resid()^2)*private$Z[,n]*exp(-2*(drop(private$Z %*% self$gamma)))-private$Z[,n])
                                             }
-                                            coefficients
-
+                                            derivatives
                                           },
-
-                                          lstSqrCovariates = function()
+                                          derivativesB = function()
                                           {
-                                            #Page 219
-
-                                            lstSqrsCovariate <- numeric()
-
-                                            for(j in 1:dim(private$X)[2]) {
-                                              xj_mean <- mean(private$X[,j])
-
-                                              #Compute the resdiuals for the jth covariate
-                                              ui <- y - self$beta[j]*private$X[,j]
-
[…]

### 2020-05-22 README.Rmd
fix componentwise boosting for beta

@@
 
+model$gamma
 ```
 
-```{r}
+```{r fig.width = 11, echo=FALSE}
 library(asp20boost)
@@
 gradient_boost(model, 
-              stepsize = 0.001, maxit = 1000,
+              stepsize = 0.001, maxit = 10000,
               abstol = 0.001,
               verbose = FALSE)
+
 plot(x, y)
 lines(x,model$fitted_location)
-#Todo do create similar graphic
-curve(x + 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)
-curve(x - 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)
+#For a 95% confidence interval, the value c = 1.96 is agood approximation
+curve(x + 1.96 * exp(model$gamma[1]+model$gamma[2]*x), -0.1, 1.1, add = TRUE)
+curve(x - 1.96 * exp(model$gamma[1]+model$gamma[2]*x), -0.1, 1.1, add = TRUE)
 

### 2020-05-22 R/model.R

@@ -88,10 +88,12 @@ LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                           {
                                             #Calculates current first derivates for Beta n of Squared Residuals ()
                                             #Could be further improved by moving the matrix multiplaction to the update beta function to reduce the calculation time
-                                            derivatives <- numeric()
+                                            derivatives<-numeric()
                                             for(n in 1:dim(private$X)[2]) {
-                                              #First derivative of (y-X*B)^2
-                                              derivatives[n]<-sum(2*private$X[,n]*(model$resid()))
+                                              ui <- (self$resid())
+                                              xn <- private$X[,n]
+                                              bjhat <- solve(t(xn)%*%xn)%*%t(xn)%*%ui
+                                              derivatives[n]<-sum((ui-xn%*%bjhat)^2)
                                             }
                                             derivatives
                                           },
@@ -154,19 +156,28 @@ gradient_boost = function(model,
 
 
     #Update Beta with greatest gradient
-    #should be evaluated to use minial Least Squares
-    indexOfBetaUpdate = which.max(abs(tmp_DerviatesB))
-    tmp_newbeta<-model$beta
-    tmp_newbeta[indexOfBetaUpdate] <- model$beta[indexOfBetaUpdate] + v*tmp_DerviatesB[indexOfBetaUpdate]
-    model$beta<-tmp_newbeta
-    tmp_DerviatesB <- model$derivativesB
-    grad_beta <- model$grad_beta()
+    #should be evaluated to check Least Squares Criterion to identify component to update?
+    # indexOfBetaUpdate = which.max(abs(tmp_DerviatesB))
+    # tmp_newbeta<-model$beta
+    # tmp_newbeta[indexOfBetaUpdate] <- model$beta[indexOfBetaUpdate] + v*tmp_DerviatesB[indexOfBetaUpdate]
+    # model$beta<-tmp_newbeta
+    # tmp_DerviatesB <- model$derivativesB
+    # grad_beta <- model$grad_beta()
+    #END OF OLD
 
+    #needs custom step size for Beta, otherwise it would take to long
[…]

### 2020-05-27 Reports/1_REPORT.Rmd
- Basic RMD structure for report. - JSS template for reference, if output desired (untitled). - Literature and packages and Bibtex files for citation. - First knit to pdf and to html

---
title: "1 Report ASP20 Boost"
subtitle: "First Report"
author:
- "Johannes Strauß"
- "Levin Wiebelt"
- "Sebastian Aristizabal"
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: [nothing, nothingness]
abstract: |
  This is the abstract.

  It consists of two paragraphs.
documentclass: report
bibliography: ["citavi_27.05.bib", "packages.bib"]
biblio-style: apalike
geometry: "left=4cm, right=3cm, top=2.5cm, bottom=2.5cm"
lof: FALSE
lot: FALSE
output: 
    pdf_document:
      toc: true
      toc_depth: 2
      latex_engine: xelatex
      keep_tex: yes
      template: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# How to cite and configure the document's appearance

## Write

To cite:

* Parentheses: Implementation builds upon the asp20 model [@R-asp20model]. 
* Inline: It is implemented as R6 class @R-R6.
* Including pages: For the theoretical foundations of our implementation, we rely mostly on  [@Fahrmeir.2013, 219] and [@Hastie.2009, 358].

In the `report` folder there's two `.bib` files. 
These contain all reference we should need. 
To update the `package.bib` write the name of the package in the following chunk and *actively* run it (It should be installed in your terminal).

```{r pack tex, eval=FALSE, include=TRUE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'R6', 'knitr', 'rmarkdown', 'asp20model', 'gamboostLSS','mboost', 'tidyverse'  
), 'packages.bib')
```

To cite write a ``@` followed by `firstname.year` as shown in the `.bib` files for reference. 
See further [cite](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)

As a tip hit always enter after each period for better readability in R Studio.
Like this.

Two enters give me a new paragraph.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

## Appearance using the `YAML` Header:

We can regulate the appearance of the document from here. 
There's a lot flexibility but the learning curve is somehow steep.
I'll link useful resources to learn, but, in principle, the document is *ready for writing*.

Here the corresponding references:

*  [PDF output](https://bookdown.org/yihui/rmarkdown/pdf-document.html)
*  [Pandoc Manuals](https://pandoc.org/MANUAL.html#templates)
*  [TOC](https://bookdown.org/yihui/rmarkdown/pdf-document.html#table-of-contents-1)
*  [Cite with Pandoc](https://github.com/jgm/pandoc-citeproc/blob/master/man/pandoc-citeproc.1.md)

Here those inputs that are easy to configuration for you guys to get an idea:

* With `documentclass` has three base templates: `article`, `report` and `book`. 
* `geometry` let's me change the margins. 
* `lof` and `lot` set to true give me a list of figures and tables respectively.

# 0.Lead Questions to be answered (from Hannes' e-mail):

* Which open questions do you have about the statistical model and methodology?
*  What functionality are you planning to (or did you already) implement? What did you decide to leave out?
* What are the critical aspects of your implementation and your package? Stability? Performance? User-friendliness? How are you going to deal with these potential issues?
    What design designs did you make with respect to your implementation? How are you going to extend the R6 class? What additional functions are you going to provide?

# 1. Description of the problem

We have been working on an implementation for a boosting implementation with gradient descent for location and scale.

# 2. Description of progress

## Johannes is gzus.

# 3. Intended extensions to be implemented. 
# 4. Comments
# 5. References

### 2020-05-27 Reports/1_REPORT.tex
- Basic RMD structure for report. - JSS template for reference, if output desired (untitled). - Literature and packages and Bibtex files for citation. - First knit to pdf and to html

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{report}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={1 Report ASP20 Boost},
  pdfauthor={Johannes Strauß; Levin Wiebelt; Sebastian Aristizabal},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=4cm, right=3cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{1 Report ASP20 Boost}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{First Report}
\author{Johannes Strauß \and Levin Wiebelt \and Sebastian Aristizabal}
\date{27 Mai 2020}

\begin{document}
\maketitle
\begin{abstract}
This is the abstract.

It consists of two paragraphs.
\end{abstract}

{
\setcounter{tocdepth}{1}
\tableofcontents
}
\hypertarget{how-to-cite-and-configure-the-documents-appearance}{%
\chapter{How to cite and configure the document's
appearance}\label{how-to-cite-and-configure-the-documents-appearance}}

\hypertarget{write}{%
\section{Write}\label{write}}

To cite:

\begin{itemize}
\tightlist
\item
  Parentheses: Implementation builds upon the asp20 model (Riebl 2020).
\item
  Inline: It is implemented as R6 class Chang (2019).
\item
  Including pages: For the theoretical foundations of our
  implementation, we rely mostly on (Fahrmeir 2013, 219) and (Hastie,
  Tibshirani, and Friedman 2009, 358).
\end{itemize}

In the \texttt{report} folder there's two \texttt{.bib} files. These
contain all reference we should need. To update the \texttt{package.bib}
write the name of the package in the following chunk and \emph{actively}
run it (It should be installed in your terminal).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# automatically create a bib database for R packages}
\NormalTok{knitr}\OperatorTok{::}\KeywordTok{write_bib}\NormalTok{(}\KeywordTok{c}\NormalTok{(}
  \KeywordTok{.packages}\NormalTok{(), }\StringTok{'R6'}\NormalTok{, }\StringTok{'knitr'}\NormalTok{, }\StringTok{'rmarkdown'}\NormalTok{, }\StringTok{'asp20model'}\NormalTok{, }\StringTok{'gamboostLSS'}\NormalTok{,}\StringTok{'mboost'}\NormalTok{, }\StringTok{'tidyverse'}  
\NormalTok{), }\StringTok{'packages.bib'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

To cite write a `\texttt{@} followed by \texttt{firstname.year} as shown
in the \texttt{.bib} files for reference. See further
\href{https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html}{cite}

As a tip hit always enter after each period for better readability in R
Studio. Like this.

Two enters give me a new paragraph.

\hypertarget{including-plots}{%
\section{Including Plots}\label{including-plots}}

You can also embed plots, for example:

\includegraphics{1_REPORT_files/figure-latex/pressure-1.pdf}

\hypertarget{appearance-using-the-yaml-header}{%
\section{\texorpdfstring{Appearance using the \texttt{YAML}
Header:}{Appearance using the YAML Header:}}\label{appearance-using-the-yaml-header}}

We can regulate the appearance of the document from here. There's a lot
flexibility but the learning curve is somehow steep. I'll link useful
resources to learn, but, in principle, the document is \emph{ready for
writing}.

Here the corresponding references:

\begin{itemize}
\tightlist
\item
  \href{https://bookdown.org/yihui/rmarkdown/pdf-document.html}{PDF
  output}
\item
  \href{https://pandoc.org/MANUAL.html\#templates}{Pandoc Manuals}
\item
  \href{https://bookdown.org/yihui/rmarkdown/pdf-document.html\#table-of-contents-1}{TOC}
\item
  \href{https://github.com/jgm/pandoc-citeproc/blob/master/man/pandoc-citeproc.1.md}{Cite
  with Pandoc}
\end{itemize}

Here those inputs that are easy to configuration for you guys to get an
idea:

\begin{itemize}
\tightlist
\item
  With \texttt{documentclass} has three base templates:
  \texttt{article}, \texttt{report} and \texttt{book}.
\item
  \texttt{geometry} let's me change the margins.
\item
  \texttt{lof} and \texttt{lot} set to true give me a list of figures
  and tables respectively.
\end{itemize}

\hypertarget{lead-questions-to-be-answered-from-hannes-e-mail}{%
\chapter{0.Lead Questions to be answered (from Hannes'
e-mail):}\label{lead-questions-to-be-answered-from-hannes-e-mail}}

\begin{itemize}
\tightlist
\item
  Which open questions do you have about the statistical model and
  methodology?
\item
  What functionality are you planning to (or did you already) implement?
  What did you decide to leave out?
\item
  What are the critical aspects of your implementation and your package?
  Stability? Performance? User-friendliness? How are you going to deal
  with these potential issues? What design designs did you make with
  respect to your implementation? How are you going to extend the R6
  class? What additional functions are you going to provide?
\end{itemize}

\hypertarget{description-of-the-problem}{%
\chapter{1. Description of the
problem}\label{description-of-the-problem}}

We have been working on an implementation for a boosting implementation
with gradient descent for location and scale.

\hypertarget{description-of-progress}{%
\chapter{2. Description of progress}\label{description-of-progress}}

\hypertarget{johannes-is-gzus.}{%
\section{Johannes is gzus.}\label{johannes-is-gzus.}}

\hypertarget{intended-extensions-to-be-implemented.}{%
\chapter{3. Intended extensions to be
implemented.}\label{intended-extensions-to-be-implemented.}}

\hypertarget{comments}{%
\chapter{4. Comments}\label{comments}}

\hypertarget{references}{%
\chapter*{5. References}\label{references}}
\addcontentsline{toc}{chapter}{5. References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-R-R6}{}%
Chang, Winston. 2019. \emph{R6: Encapsulated Classes with Reference
Semantics}. \url{https://CRAN.R-project.org/package=R6}.

\leavevmode\hypertarget{ref-Fahrmeir.2013}{}%
Fahrmeir, Ludwig. 2013. \emph{Regression: Models, Methods and
Applications}. New York: Springer.

\leavevmode\hypertarget{ref-Hastie.2009}{}%
Hastie, Trevor, Robert Tibshirani, and J. H. Friedman. 2009. \emph{The
Elements of Statistical Learning: Data Mining, Inference, and Prediction
/ Trevor Hastie, Robert Tibshirani, Jerome Friedman}. 2nd ed. Springer
Series in Statistics. New York: Springer.

\leavevmode\hypertarget{ref-R-asp20model}{}%
Riebl, Hannes. 2020. \emph{Asp20model: An R6 Class for Location-Scale
Regression Models}.

\end{document}

### 2020-05-27 Reports/test_report.Rmd
- Basic RMD structure for report. - JSS template for reference, if output desired (untitled). - Literature and packages and Bibtex files for citation. - First knit to pdf and to html

---
title: ASP20 Boost
subtitle: jai jai
titlerunning: Short form of title (if too long for head)
authorrunning:
  Short form of author list if too long for running head
thanks: |
  Grants or other notes about the article that should go
  on the front page should be placed here. General
  acknowledgments should be placed at the end of the article.
authors:
  - name: Author 1
    address: Department of YYY, University of XXX
    email: abc@def
  - name: Author 2
    address: Department of ZZZ, University of WWW
    email: djf@wef
keywords:
  - key
  - dictionary
  - word
MSC:
  - MSC code 1
  - MSC code 2    
abstract: |
  The text of your abstract.  150 -- 250 words.
bibliography: citavi_27.05.bib
output: rticles::springer_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'dlstats', 'randomForest', 'tidyverse'  
), 'packages.bib')
```

### 2020-05-27 Reports/test_report.tex
- Basic RMD structure for report. - JSS template for reference, if output desired (untitled). - Literature and packages and Bibtex files for citation. - First knit to pdf and to html

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

\begin{document}

\title{ASP20 Boost \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }
 \subtitle{jai jai}

    \titlerunning{Short form of title (if too long for head)}

\author{  Author 1 \and  Author 2 \and  }

    \authorrunning{ Short form of author list if too long for running head }

\institute{
        Author 1 \at
     Department of YYY, University of XXX \\
     \email{\href{mailto:abc@def}{\nolinkurl{abc@def}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Author 2 \at
     Department of ZZZ, University of WWW \\
     \email{\href{mailto:djf@wef}{\nolinkurl{djf@wef}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
The text of your abstract. 150 -- 250 words.
\\
\keywords{
        key \and
        dictionary \and
        word \and
    }

    \subclass{
                    MSC code 1 \and
                    MSC code 2 \and
            }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\bibliographystyle{spbasic}
\bibliography{citavi\_27.05.bib}

\end{document}

### 2020-05-27 R/model.R
fix componentwise boosting for gamma , move componentwise boosting logic to functions
1c6850

@@ -17,9 +17,8 @@
 #' @import asp20model
 #' @export
 
-
-library(R6)
-
+#Uncomment for testing
+# library(R6)
 
 LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                         inherit = LocationScaleRegression,
@@ -47,63 +46,48 @@ LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                             self$Proj_M_Z <- self$Z_tr_Z_inv %*% t(private$Z)
 
                                           }
-
-
-
-
                                         ),
                                         active = list(
-                                          #' @details
-                                          #' Least Square Residual
-                                          #'
-                                          #' @return
-                                          #' Current gradient of the coefficents
-                                          #'
-                                          #' @examples
-                                          #' model$<-lstSqrResid
-                                          lstSqrResid = function()
-                                          {
-                                            self$Proj_M_X %*% self$resid()
-                                          },
-                                          #' @details
-                                          #' First derivatives of Log Likelehood
-                                          #'
-                                          #' @return
-                                          #' Current first derivates of each Coeffizent
-                                          #'
-                                          #' @examples
-                                          #' model$<-derivativesZ
-                                          derivativesZ = function()
-                                          {
-                                            #Calculates current first derivates for Gamma n of Log Likelehood
-                                            #Could be further improved by moving the matrix multiplaction to the update gamma function to reduce the calculation time
-                                            Z_countCoefficient <- dim(model$getZ)[2]
-                                            derivatives <- numeric()
+                                             bestFittingVariableGamma = function()
+                                            {
+                                            #Fit separate linear models for all covariates of gamma
+                                            loss <- numeric()
+                                            ui <- self$resid("deviance")
[…]

### 2020-05-27 R/model.R
fix typo

@@ -141,7 +141,7 @@ gradient_boost = function(model,
     #needs custom step size for Beta, otherwise it would take to long ? increase with this factor
     stepsize <-10
 
-    model$bet0a<-model$beta + stepsize*v*model$bestFittingVariableBeta
+    model$beta<-model$beta + stepsize*v*model$bestFittingVariableBeta
     grad_beta <- model$grad_beta()
 
 

### 2020-05-27 Reports/1_REPORT.Rmd
Ammend to last commit text: - Included short tutorial to use Rmd for academic wrtiting (with references). - Set the outline for first report. - Set "report" folder including all relevant files for compiling + R Studio Rmd cheatsheets.
- Reformated outline - Expanded the introduction.

@@
 # How to cite and configure the document's appearance 
+If you don't have Latex download latex for R with `tinytex::install_tinytex()`.
 
@@
 To update the `package.bib` write the name of the package in the following chunk and *actively* run it (It should be installed in your terminal).
+The`citavi.bib` file needs to be exported from citavi.
 
@@
 
-# 0.Lead Questions to be answered (from Hannes' e-mail):
+# 1. Introduction:
 
-* Which open questions do you have about the statistical model and methodology?
-*  What functionality are you planning to (or did you already) implement? What did you decide to leave out?
-* What are the critical aspects of your implementation and your package? Stability? Performance? User-friendliness? How are you going to deal with these potential issues?
-    What design designs did you make with respect to your implementation? How are you going to extend the R6 class? What additional functions are you going to provide?
-
-# 1. Description of the problem
-
-We have been working on an implementation for a boosting implementation with gradient descent for location and scale.
+* The developement of this package takes place in the framework of the seminar "Advanced Statisical Programming" -> SS20
+* We expand the R6 class *"LocationScaleRegression"* belonging to the `asp20model` package concieved specifficaly for this seminar [@R-asp20model]
+* Our aim is to develope an multi-faceted implementation of boosting for location and scale including component-wise boosting, use of cross-validation to determine the optimal stopping critera  and useful options for visualization.
 
@@
 
-## Johannes is gzus.
+> Hannes lead question: What are the critical aspects of your implementation and your package? Stability? Performance? User-friendliness? How are you going to deal with these potential issues?
 
-# 3. Intended extensions to be implemented. 
-# 4. Comments
-# 5. References
+> 2. What design designs did you make with respect to your implementation? How are you going to extend the R6 class? What additional functions are you going to provide?
+
+I propose a timeline where the implementation process is described step by step as a list e.g:
+
+1. Primitive implementation of boosting
+2. Simple boosting for location implemented
+3. Adressed the estimation for the variance of gamma. 
+4. Johannes finished everything #lol. 27.05.
+
+## Johannes is *gzus*.
+
+# 3. Description of the problem
+
+> Hannes lead question: Which open questions do you have about the statistical model and methodology?
+
+# 4. Extensions to be implemented.
+
+> Hannes' Lead question: What functionality are you planning to (or did you already) implement? What did you decide to leave out?
+
+# 5. Summary?
+# 6. References

### 2020-05-27 Reports/1_REPORT.tex
- Reformated outline - Expanded the introduction.

@@
 
+If you don't have Latex download latex for R with
+\texttt{tinytex::install\_tinytex()}.
+
 \hypertarget{write}{%
@@
 write the name of the package in the following chunk and \emph{actively}
-run it (It should be installed in your terminal).
+run it (It should be installed in your terminal). The\texttt{citavi.bib}
+file needs to be exported from citavi.
 
@@
 
-\hypertarget{lead-questions-to-be-answered-from-hannes-e-mail}{%
-\chapter{0.Lead Questions to be answered (from Hannes'
-e-mail):}\label{lead-questions-to-be-answered-from-hannes-e-mail}}
+\hypertarget{introduction}{%
+\chapter{1. Introduction:}\label{introduction}}
 
@@
 \item
-  Which open questions do you have about the statistical model and
-  methodology?
+  The developement of this package takes place in the framework of the
+  seminar ``Advanced Statisical Programming'' -\textgreater{} SS20
 \item
-  What functionality are you planning to (or did you already) implement?
-  What did you decide to leave out?
+  We expand the R6 class \emph{``LocationScaleRegression''} belonging to
+  the \texttt{asp20model} package concieved specifficaly for this
+  seminar (Riebl 2020)
 \item
-  What are the critical aspects of your implementation and your package?
-  Stability? Performance? User-friendliness? How are you going to deal
-  with these potential issues? What design designs did you make with
-  respect to your implementation? How are you going to extend the R6
-  class? What additional functions are you going to provide?
+  Our aim is to develope an multi-faceted implementation of boosting for
+  location and scale including component-wise boosting, use of
+  cross-validation to determine the optimal stopping critera and useful
+  options for visualization.
 \end{itemize}
-
-\hypertarget{description-of-the-problem}{%
-\chapter{1. Description of the
-problem}\label{description-of-the-problem}}
-
-We have been working on an implementation for a boosting implementation
-with gradient descent for location and scale.
 
@@
 
+\begin{quote}
+Hannes lead question: What are the critical aspects of your
+implementation and your package? Stability? Performance?
+User-friendliness? How are you going to deal with these potential
+issues?
+\end{quote}
+
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  What design designs did you make with respect to your implementation?
+  How are you going to extend the R6 class? What additional functions
+  are you going to provide?
+\end{enumerate}
+\end{quote}
+
+I propose a timeline where the implementation process is described step
+by step as a list e.g:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Primitive implementation of boosting
+\item
+  Simple boosting for location implemented
+\item
+  Adressed the estimation for the variance of gamma.
+\item
+  Johannes finished everything \#lol. 27.05.
+\end{enumerate}
+
 \hypertarget{johannes-is-gzus.}{%
-\section{Johannes is gzus.}\label{johannes-is-gzus.}}
+\section{\texorpdfstring{Johannes is
+\emph{gzus}.}{Johannes is gzus.}}\label{johannes-is-gzus.}}
 
-\hypertarget{intended-extensions-to-be-implemented.}{%
-\chapter{3. Intended extensions to be
-implemented.}\label{intended-extensions-to-be-implemented.}}
+\hypertarget{description-of-the-problem}{%
+\chapter{3. Description of the
+problem}\label{description-of-the-problem}}
 
-\hypertarget{comments}{%
-\chapter{4. Comments}\label{comments}}
+\begin{quote}
+Hannes lead question: Which open questions do you have about the
+statistical model and methodology?
+\end{quote}
+
+\hypertarget{extensions-to-be-implemented.}{%
+\chapter{4. Extensions to be
+implemented.}\label{extensions-to-be-implemented.}}
+
+\begin{quote}
+Hannes' Lead question: What functionality are you planning to (or did
+you already) implement? What did you decide to leave out?
+\end{quote}
+
+\hypertarget{summary}{%
+\chapter{5. Summary?}\label{summary}}
 
 \hypertarget{references}{%
-\chapter*{5. References}\label{references}}
-\addcontentsline{toc}{chapter}{5. References}
+\chapter*{6. References}\label{references}}
+\addcontentsline{toc}{chapter}{6. References}
 

### 2020-05-27 R/model.R
add comment

@@ -58,7 +58,7 @@ LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
 
                                                u <- ((self$resid()^2)*private$Z[,n]*exp(-2*(drop(private$Z %*% self$gamma)))-private$Z[,n])
 
-                                               bjhat <- solve(t(private$Z[,n])%*%private$Z[,n])%*%t(private$Z[,n])%*% u
+                                                             #Can be further improvied by moving ProjectionMatrix Calculation to initializalisation                                 bjhat <- solve(t(private$Z[,n])%*%private$Z[,n])%*%t(private$Z[,n])%*% u
                                                loss[n]<-sum((ui-(self$resid()/zn%*%bjhat))^2)
                                                }
                                             #determine index of the best-fitting variable
@@ -76,7 +76,8 @@ LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                             ui <- (self$resid())
                                             for(n in 1:dim(private$X)[2]) {
                                               xn <- private$X[,n]
-                                              bjhat <- solve(t(private$X[,n])%*%private$X[,n])%*%t(private$X[,n])%*%ui
+                                              bjhat <- solve(#Can be further improvied by moving ProjectionMatrix Calculation to initializalisation
+                                              t(private$X[,n])%*%private$X[,n])%*%t(private$X[,n])%*%ui
                                               loss[n]<-sum((ui-xn%*%bjhat)^2)
                                             }
                                             #determine index of the best-fitting variable

### 2020-05-27 R/model.R
fix comments

@@ -57,8 +57,8 @@ LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                                zn <- private$Z[,n]
 
                                                u <- ((self$resid()^2)*private$Z[,n]*exp(-2*(drop(private$Z %*% self$gamma)))-private$Z[,n])
-
-                                                             #Can be further improvied by moving ProjectionMatrix Calculation to initializalisation                                 bjhat <- solve(t(private$Z[,n])%*%private$Z[,n])%*%t(private$Z[,n])%*% u
+                                               #Can be further improvied by moving ProjectionMatrix Calculation to initializalisation
+                                               bjhat <- solve(t(private$Z[,n])%*%private$Z[,n])%*%t(private$Z[,n])%*% u
                                                loss[n]<-sum((ui-(self$resid()/zn%*%bjhat))^2)
                                                }
                                             #determine index of the best-fitting variable
@@ -76,8 +76,8 @@ LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                             ui <- (self$resid())
                                             for(n in 1:dim(private$X)[2]) {
                                               xn <- private$X[,n]
-                                              bjhat <- solve(#Can be further improvied by moving ProjectionMatrix Calculation to initializalisation
-                                              t(private$X[,n])%*%private$X[,n])%*%t(private$X[,n])%*%ui
+                                              #Can be further improvied by moving ProjectionMatrix Calculation to initializalisation
+                                              bjhat <- solve(t(private$X[,n])%*%private$X[,n])%*%t(private$X[,n])%*%ui
                                               loss[n]<-sum((ui-xn%*%bjhat)^2)
                                             }
                                             #determine index of the best-fitting variable
@@ -117,9 +117,6 @@ LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
 #' gradient_boost(model)
 #'
 #' @export
-
-
-
 gradient_boost = function(model,
                           stepsize = 0.001,
                           maxit = 1000,

### 2020-05-28 README.Rmd
clipped new tasks, clipped notes report

--- 2020-05-12 README.Rmd
@@
 ## Tasks
+
+### until ~ mid June
+
+* extend documentation of R6-Class
+  + inherit documentation from Hannes
+  + add own extensions
+  + both can be done via roxygen (see Wickham R-PAckages chp5)
+
+* maybe move component-selection mechanism from active field in R6-Class to gradient_boost-function
+  + intuition tells me R6-Class is not the right place to do central calculations of the algorithm
+  + maybe create new helper function component_selection() which gradien_boost() can call
 
@@
 ```{r installation, eval = 3}
-install.packages("devtools")
-devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
-devtools::install_gitlab("asp20/asp20boost", host = "gitlab.gwdg.de")
-library(asp20boost)
+#install.packages("devtools")
+#This throws an error for asp20boost
+#devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
+#devtools::install_gitlab("asp20/asp20boost", host = "gitlab.gwdg.de")
+
+model$gamma
 ```
 
-```{r}
-model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
-gradient_boost(model, 
-              stepsize = 0.001, maxit = 1000,
-              abstol = 0.001,
-              verbose = FALSE)
-plot(x, y)
-lines(x,model$fitted_location)
-#Todo do create similar graphic
-curve(x + 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)
-curve(x - 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)
+```{r fig.width = 11, echo=FALSE}
+# library(asp20boost)
+# model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+# gradient_boost(model, 
+#               stepsize = 0.001, maxit = 10000,
+#               abstol = 0.001,
+#               verbose = FALSE)
+# 
+# 
+# 
+# plot(x, y)
+# lines(x,model$fitted_location)
+# #For a 95% confidence interval, the value c = 1.96 is agood approximation
+# curve(x + 1.96 * exp(model$gamma[1]+model$gamma[2]*x), -0.1, 1.1, add = TRUE)
+# curve(x - 1.96 * exp(model$gamma[1]+model$gamma[2]*x), -0.1, 1.1, add = TRUE)
 

### 2020-05-28 Reports/1_REPORT.Rmd
add my notes 1_report
clipped new tasks, clipped notes report

@@
 
+JS: Performance Needs to be checked and maybe more outputs - messages?
+Starting point of Beta and Gamma?
+
+L:
+
+* user-friendliness: allow for convenient input of model, possibly as data.frame
+* stability: trying to implement automated testing (unit tests)
+* performance: 
+  + custom step size for Beta needed, otherwise slow --> improve
+  + move ProjectionMatrix Calculation to initializalisation to enhance efficiency
+
 > 2. What design designs did you make with respect to your implementation? How are you going to extend the R6 class? What additional functions are you going to provide?
@@
 
-## Johannes is *gzus*.
+## Johannes is *gzus* ^{2}.
+
+L: implementation designs r6-Class "LocScRegBoost"
+
+* moved calculation of central calculation matrices in boosting algorithm to initializer of R6-Class via super-command
+* component-selection implemented inside R6-Class via additional active-fields
+  + mechanism: calculation of n loss-function-values (n = #gamma-parameters = #beta-parameters), then compare to loss-functions of last iteration and choose highest loss-improvement
+  + calculation does NOT work via derivatives as Kneib suggested, but by Hannes' deviance-residuals
 
@@
 
+JS : Can this model be used to mathematically evaluate outliers, for example in the Munich rent data set -> use it to identify cheap apartments
+
+L:
+
+* How does cross-validation work in order to optimize stopping?
+  + This connects to other applications of boosters such as variable selection. Variable selection is especially viable since component-wise boosting is already implemented
+*	Is it possible to optimize the learning rate?
+
 # 4. Extensions to be implemented.
@@
 
+L: Already implemented
+
+* boosting for multidimensional (>2) beta or gamma 
+* component-wise boosting
+
+L: Ideas for further functionality
+
+* better user-friendliness - allow inputs like my_model <- gradient_boost(formula, data = mtcars)
+  + allow for model creation via dataframe-input
+  + this may reduce errors due to unexpected inputs
+* illustrate functions of package with real dataframe (= Vignette)
+*	Visualization of results via connecting to plot team
+*	functionality: predict 
+*	Include model diagnostics
+
 # 5. Summary?

### 2020-05-28 Reports/1_REPORT.tex
clipped new tasks, clipped notes report

@@
 \author{Johannes Strauß \and Levin Wiebelt \and Sebastian Aristizabal}
-\date{27 Mai 2020}
+\date{28 Mai 2020}
 
@@
 
+JS: Performance Needs to be checked and maybe more outputs - messages?
+Starting point of Beta and Gamma?
+
+L:
+
+\begin{itemize}
+\tightlist
+\item
+  user-friendliness: allow for convenient input of model, possibly as
+  data.frame
+\item
+  stability: trying to implement automated testing (unit tests)
+\item
+  performance:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    custom step size for Beta needed, otherwise slow --\textgreater{}
+    improve
+  \item
+    move ProjectionMatrix Calculation to initializalisation to enhance
+    efficiency
+  \end{itemize}
+\end{itemize}
+
 \begin{quote}
@@
 
-\hypertarget{johannes-is-gzus.}{%
-\section{\texorpdfstring{Johannes is
-\emph{gzus}.}{Johannes is gzus.}}\label{johannes-is-gzus.}}
+\hypertarget{johannes-is-gzus-2.}{%
+\section{\texorpdfstring{Johannes is \emph{gzus}
+\^{}\{2\}.}{Johannes is gzus \^{}\{2\}.}}\label{johannes-is-gzus-2.}}
+
+L: implementation designs r6-Class ``LocScRegBoost''
+
+\begin{itemize}
+\tightlist
+\item
+  moved calculation of central calculation matrices in boosting
+  algorithm to initializer of R6-Class via super-command
+\item
+  component-selection implemented inside R6-Class via additional
+  active-fields
+
+  \begin{itemize}
+  \tightlist
+  \item
+    mechanism: calculation of n loss-function-values (n =
+    \#gamma-parameters = \#beta-parameters), then compare to
+    loss-functions of last iteration and choose highest loss-improvement
+  \item
+    calculation does NOT work via derivatives as Kneib suggested, but by
+    Hannes' deviance-residuals
+  \end{itemize}
+\end{itemize}
 
@@
 
+JS : Can this model be used to mathematically evaluate outliers, for
+example in the Munich rent data set -\textgreater{} use it to identify
+cheap apartments
+
+L:
+
+\begin{itemize}
+\tightlist
+\item
+  How does cross-validation work in order to optimize stopping?
+
+  \begin{itemize}
+  \tightlist
+  \item
+    This connects to other applications of boosters such as variable
+    selection. Variable selection is especially viable since
+    component-wise boosting is already implemented
+  \end{itemize}
+\item
+  Is it possible to optimize the learning rate?
+\end{itemize}
+
 \hypertarget{extensions-to-be-implemented.}{%
@@
 \end{quote}
+
+L: Already implemented
+
+\begin{itemize}
+\tightlist
+\item
+  boosting for multidimensional (\textgreater2) beta or gamma
+\item
+  component-wise boosting
+\end{itemize}
+
+L: Ideas for further functionality
+
+\begin{itemize}
+\tightlist
+\item
+  better user-friendliness - allow inputs like my\_model \textless-
+  gradient\_boost(formula, data = mtcars)
+
+  \begin{itemize}
+  \tightlist
+  \item
+    allow for model creation via dataframe-input
+  \item
+    this may reduce errors due to unexpected inputs
+  \end{itemize}
+\item
+  illustrate functions of package with real dataframe (= Vignette)
+\item
+  Visualization of results via connecting to plot team
+\item
+  functionality: predict
+\item
+  Include model diagnostics
+\end{itemize}
 

### 2020-06-03 Reports/1_REPORT.Rmd
update on the report

@@
 If you don't have Latex download latex for R with `tinytex::install_tinytex()`. 
-
+To install pandoc on Mac, install it via Terminal with `brew install pandoc-citeproc`.
 ## Write
@@
 * The developement of this package takes place in the framework of the seminar "Advanced Statisical Programming" -> SS20
-* We expand the R6 class *"LocationScaleRegression"* belonging to the `asp20model` package concieved specifficaly for this seminar [@R-asp20model]
-* Our aim is to develope an multi-faceted implementation of boosting for location and scale including component-wise boosting, use of cross-validation to determine the optimal stopping critera  and useful options for visualization.
+* We extend the R6 class *"LocationScaleRegression"* belonging to the `asp20model` package concieved specifficaly for this seminar [@R-asp20model]
+* Our aim is to develope an multi-faceted implementation of boosting for location and scale regression including component-wise boosting, use of cross-validation to determine the optimal stopping critera  and useful options for visualization.
 
@@
 
-JS: Performance Needs to be checked and maybe more outputs - messages?
-Starting point of Beta and Gamma?
+Our package rely heavily on the asp20model package, any changes, or errors in the code of the package might break the functionality.
+Currently there is no validation of the given input, with respect to heteroscedasticity.
+
+The package also does not provide any auto optimization of the gradient boost step-size, a wrongly picked step-size can lead to a deficient performance, but an already implemented early stopping mechanism can prevent an overfitting.
+An input check and cross-validation of the step-size is on the to-do list.
+Also, the initial starting values for Beta and Gamma can be further improved, see [@Fahrmeir.2013, 219].
+Further optimizations can be made to reduce the computation time, by moving the calculation of the Projection Matrix for individual covariates in the initialization section of the model.
+In regard to User-friendliness of the package, we will try to implement further input validation and try to write up with a comprehensive vignette containing various examples in the next weeks.
 
@@
 > 2. What design designs did you make with respect to your implementation? How are you going to extend the R6 class? What additional functions are you going to provide?
+
+JS:
+During the learning process, we tried several methods to implement the required functionality.
+At first, we started with a simple function in a R-File.
+While looking at other online resources, we agreed on using the asp20model packages as a requirement for our package and inheriting the existing LocationScaleRegression class.
+Our `LocationScaleRegressionBoost` R6 class implements a boosting algorithm, two active fields within the model for calculating the best fitting covariate with respect to Gamma and Beta and has four new variables which are set at the initialization of the model to reduce computation time of the Projection Matrixes during the gradient boosting (subject to change).
+The package provides the gradient boost function for the `LocationScaleRegressionBoost` class, which uses the two model functions to optimize stepwise the Beta and Gamma Vector.
 
@@
 
-* boosting for multidimensional (>2) beta or gamma 
+* boosting for multidimensional (>2) beta and gamma 
 * component-wise boosting
@@
 * illustrate functions of package with real dataframe (= Vignette)
-*	Visualization of results via connecting to plot team
+*	Visualization of results via connecting to plot team `asp20plot`
 *	functionality: predict 
@@
 
+JS:
+check when run gradientboost, if model is LocationScaleRegressionBoost class
+
 # 5. Summary?

### 2020-06-04 Reports/report_1/1_REPORT.Rmd
write first report

---
title: "Report ASP20 Boost"
subtitle: "First Report"
author:
- "Johannes Strauß"
- "Levin Wiebelt"
- "Sebastian Aristizabal"
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: [nothing, nothingness]
documentclass: report
bibliography: ["citavi_27.05.bib", "packages.bib"]
biblio-style: apalike
geometry: "left=4cm, right=3cm, top=2.5cm, bottom=2.5cm"
lof: FALSE
lot: FALSE
output: 
    pdf_document:
      toc: true
      toc_depth: 2
      latex_engine: xelatex
      keep_tex: yes
      template: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# 1. Introduction:

The course "Advanced Statistical Programming with R" consists in implementing a package to address location-scale regression. 
The sample dataset consists of a response y, that's expectation is in linear relationship to a set of predictors xi. 
Its variance, in contrast, is dependent on a different set of linear predictors zi, however transformed by the response function g(x) = exp(x). 
This ensures positiveness of the variances. 
The goal approach is to estimate the effects of both: xi on the expectation of y (location) - this is captured by the estimates beta, and zi on the variance of y (scale) - this is captured by the estimates gamma. Our student group `asp20boost` solves this task applying the concept of boosting.

In part 2 of this report we explain the concept of boosting and elaborate on how it suites to address the problem of location-scale regression. 
In the following part 3 we describe our implementation milestones, whereby part 4 especially elaborates on componentwise boosting. 
In part 5 we present our thoughts on upcoming challenges and critical aspects of the package development.

# 2. Concept: Boosting

The first step in our progress was to understand the theoretical concept of boosting, and which types of problems it is able to solve. In the following the term "booster" refers to a boosting algorithm.

## Location-Booster

To explain the concept and follow our process of understanding we first explain boosting for location parameters. Boosting scale parameters requires another way of thinking about it and constituted a first milestone in our understanding and implementation.
The boosting concept relies on the idea of iterative estimation. 
To estimate the location parameters beta via boosting, one starts with initial estimates, which may be far from optimal. 
However, with estimates and the response vector at hand, residuals can be calculated. 
The boosting-algortihm focuses on those calculated residuals.
The effect of the predictors xi on these residuals is estimated by least squares and the resulting effect is added to the location-estimate beta, adjusted by some learning rate. 
This yields a better fit of the location-model, which results in smaller residuals.
These new residuals are estimated again in the next iteration, yielding smaller effect sizes, and hence a convergence of the boosting algorithm towards the OLS-estimate.
A simple booster stops after a fixed number of iterations is reached.
**(comp. Kneib et al)**

## Scale-Booster

[Loss-function erklären: Boosting can be viewed as lowering iteratively the amount of a loss function, which is in our case the score of the normal distribution]

The concept for boosting scale parameters is the same. However, residuals for the variance estimates can not be calculated, since there is no such thing as "observed" variances for individual response values. The solution is to substitute the calculated residuals with a different term: the derivative of the loss function, which in our case is the log-likelihood of the normal distribution. Instead of estiamting stepwise residuals, one estiamtes hence stepwise the score.

The resulting loss fcuntion then, is defined by sum of squares oa the term:
ui (=deviance-residuals) - inverse predicted scores * working-residuals

We employed the resid-function of the asp20model class, on which our class is built on.
<!-- The working and response residuals are the observed values -->
<!-- minus the fitted values for the location. -->
<!-- The deviance and Pearson residuals are weighted residuals, -->
<!-- divided by the fitted values for the scale. -->

# 3. Implementation

In the current version of the asp20boost-package, the boosting algorithm is already implemented in order to allow componentwise boosting, which will be explained in part 4. 
For good explanation and for following our development process in this part we first describe the implementation of ordinary boosting - which however does not appear in our code anymore.

A simple booster for the location parameters beta works without extending the LocationScaleRegression-Class. 
The current residuals may be extrated with the resid()-method, then estimated. The location parameter is updated. 
This causes the LocationScaleRegression-Class to calculate updated state-dependent objects, such as the loglikelihood, gradients and the residuals. 
This is what's needed to repeat the procedure of estimating residuals and updating location parameters.

The simple booster for gamma consists in the same two repeating steps of first estimating a term, and second updating the scale parameters gamma by adding the estimated effect - adjusted by a learning rate. 
The term to be estimated, however, is not the residual, but the derivative of the loglikelihood of the normal distribution. 
To calculate this, our code makes use of the resid-function(), but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.

[Is the gamma booster dependent on simultaneous running of beta booster?]

Having implemented a boosting algorithm for location and scale, the code-basis for our package was achieved.

# 4. Functionality: Componentwise Boosting

A usefull functionality of boosting is componentwise boosting. 
The idea is to not update a whole parameter vector, but only one entry of it.
The entry chosen for the update at this juncture is the one yielding the best improvement in the sense of lowering the loss function.
The loss function in our case is the loglikelihood of the normally distributed data.
Hence in each iteration only one component of the location-, as well as one component of the scale-parameters are estimated.

The implementation in our package works via the extension of the LocationScaleRegressionClass by two active fields "bestFittingVariableBeta" and "bestFittingVariableGamma". 
These functions partition the design matrix X, respectively Z, into its single columns and then estimate the residuals - respectively the score - seperately for each component, and determine loss functions for a hypothetical update with the respective component. 
Comparing with the old loss function value, the highest loss-improvement may be determined and the respective component is used to update the parameter-vector.

We are in the process of reconsidering the design of this implementation.
Other design possibilities are the following:

* create public fields for the heavily used score-function values
* move the calculation of componentwise losses to an external function
* harmonize boosting and componentwise boosting into one external function, determining the mode of operation by an argument 'componentwise = TRUE'

Another conceptual questions that comes up is if the best loss-improvement may be indicated by the already existing gradients, and hence there is no need for extra calculation.

# 5. Prospects

## Further Functionalities

The most important functionality we intend to implement in our package is variable selection via cross validation. Working out the theoretical concept behind this idea is one of our next milestones.

One further, rather loose, idea is to find ways to optimize the learning rate in our boosting algorithm.

## Stability

We intend to implement further automated unit tests. This will enable us to assess quickly the stability of our code given differing inputs. Departing from this point we intend to reduce proneness to input errors.

## Performance

* A major performance problem in our code is that small learning rates for the beta-booster cause processing time of the code to increase sharply. This remains to be solved.

*A good choice of starting values for beta and gamma may enhance performance. A possible candidate is the mean of the response.

## Further

* Allow user-input of the LocationScaleRegressionBoost-model in form of a dataframe, for example by allowing to pass an optional "data"-argument as known from the lm-call. This may reduce proneness to input-errors.

* Visualize the results of the boosted estimates

* Document the extension of the R6-Class properly and inherit documentation of the LocationScaleRegression-Class.

## Application

[Johannes: Munich Rent Data? - Identify Outliers]

### 2020-06-04 Reports/report_1/1_REPORT.tex
write first report

--- 2020-06-04 Reports/report_1/1_REPORT.Rmd
@@
----
-title: "Report ASP20 Boost"
-subtitle: "First Report"
-author:
-- "Johannes Strauß"
-- "Levin Wiebelt"
-- "Sebastian Aristizabal"
-date: "`r format(Sys.time(), '%d %B %Y')`"
-tags: [nothing, nothingness]
-documentclass: report
-bibliography: ["citavi_27.05.bib", "packages.bib"]
-biblio-style: apalike
-geometry: "left=4cm, right=3cm, top=2.5cm, bottom=2.5cm"
-lof: FALSE
-lot: FALSE
-output: 
-    pdf_document:
-      toc: true
-      toc_depth: 2
-      latex_engine: xelatex
-      keep_tex: yes
-      template: NULL
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{report}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Report ASP20 Boost},
+  pdfauthor={Johannes Strauß; Levin Wiebelt; Sebastian Aristizabal},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[left=4cm, right=3cm, top=2.5cm, bottom=2.5cm]{geometry}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+\title{Report ASP20 Boost}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{First Report}
+\author{Johannes Strauß \and Levin Wiebelt \and Sebastian Aristizabal}
+\date{04 Juni 2020}
 
-```
+\begin{document}
+\maketitle
 
-# 1. Introduction:
+{
+\setcounter{tocdepth}{1}
+\tableofcontents
+}
+\hypertarget{introduction}{%
+\chapter{1. Introduction:}\label{introduction}}
 
-The course "Advanced Statistical Programming with R" consists in implementing a package to address location-scale regression. 
-The sample dataset consists of a response y, that's expectation is in linear relationship to a set of predictors xi. 
-Its variance, in contrast, is dependent on a different set of linear predictors zi, however transformed by the response function g(x) = exp(x). 
-This ensures positiveness of the variances. 
-The goal approach is to estimate the effects of both: xi on the expectation of y (location) - this is captured by the estimates beta, and zi on the variance of y (scale) - this is captured by the estimates gamma.
-Our student group `asp20boost` solves this task applying the concept of boosting.
+The course ``Advanced Statistical Programming with R'' consists in
+implementing a package to address location-scale regression. The sample
+dataset consists of a response y, that's expectation is in linear
+relationship to a set of predictors xi. Its variance, in contrast, is
+dependent on a different set of linear predictors zi, however
+transformed by the response function g(x) = exp(x). This ensures
+positiveness of the variances. The goal approach is to estimate the
+effects of both: xi on the expectation of y (location) - this is
+captured by the estimates beta, and zi on the variance of y (scale) -
+this is captured by the estimates gamma. Our student group
+\texttt{asp20boost} solves this task applying the concept of boosting.
 
-In part 2 of this report we explain the concept of boosting and elaborate on how it suites to address the problem of location-scale regression. 
-In the following part 3 we describe our implementation milestones, whereby part 4 especially elaborates on componentwise boosting. 
-In part 5 we present our thoughts on upcoming challenges and critical aspects of the package development.
+In part 2 of this report we explain the concept of boosting and
+elaborate on how it suites to address the problem of location-scale
+regression. In the following part 3 we describe our implementation
+milestones, whereby part 4 especially elaborates on componentwise
+boosting. In part 5 we present our thoughts on upcoming challenges and
+critical aspects of the package development.
 
-# 2. Concept: Boosting
+\hypertarget{concept-boosting}{%
+\chapter{2. Concept: Boosting}\label{concept-boosting}}
 
-The first step in our progress was to understand the theoretical concept of boosting, and which types of problems it is able to solve.
-In the following the term "booster" refers to a boosting algorithm.
+The first step in our progress was to understand the theoretical concept
+of boosting, and which types of problems it is able to solve. In the
+following the term ``booster'' refers to a boosting algorithm.
 
-## Location-Booster
+\hypertarget{location-booster}{%
+\section{Location-Booster}\label{location-booster}}
 
-To explain the concept and follow our process of understanding we first explain boosting for location parameters.
-Boosting scale parameters requires another way of thinking about it and constituted a first milestone in our understanding and implementation.
-The boosting concept relies on the idea of iterative estimation. 
-To estimate the location parameters beta via boosting, one starts with initial estimates, which may be far from optimal. 
-However, with estimates and the response vector at hand, residuals can be calculated. 
-The boosting-algortihm focuses on those calculated residuals.
-The effect of the predictors xi on these residuals is estimated by least squares and the resulting effect is added to the location-estimate beta, adjusted by some learning rate. 
-This yields a better fit of the location-model, which results in smaller residuals.
-These new residuals are estimated again in the next iteration, yielding smaller effect sizes, and hence a convergence of the boosting algorithm towards the OLS-estimate.
-A simple booster stops after a fixed number of iterations is reached.
-**(comp. Kneib et al)**
+To explain the concept and follow our process of understanding we first
+explain boosting for location parameters. Boosting scale parameters
+requires another way of thinking about it and constituted a first
+milestone in our understanding and implementation. The boosting concept
+relies on the idea of iterative estimation. To estimate the location
+parameters beta via boosting, one starts with initial estimates, which
+may be far from optimal. However, with estimates and the response vector
+at hand, residuals can be calculated. The boosting-algortihm focuses on
+those calculated residuals. The effect of the predictors xi on these
+residuals is estimated by least squares and the resulting effect is
+added to the location-estimate beta, adjusted by some learning rate.
+This yields a better fit of the location-model, which results in smaller
+residuals. These new residuals are estimated again in the next
+iteration, yielding smaller effect sizes, and hence a convergence of the
+boosting algorithm towards the OLS-estimate. A simple booster stops
+after a fixed number of iterations is reached. \textbf{(comp. Kneib et
+al)}
 
-## Scale-Booster
+\hypertarget{scale-booster}{%
+\section{Scale-Booster}\label{scale-booster}}
 
-[Loss-function erklären: Boosting can be viewed as lowering iteratively the amount of a loss function, which is in our case the score of the normal distribution]
+{[}Loss-function erklären: Boosting can be viewed as lowering
+iteratively the amount of a loss function, which is in our case the
+score of the normal distribution{]}
 
-The concept for boosting scale parameters is the same.
-However, residuals for the variance estimates can not be calculated, since there is no such thing as "observed" variances for individual response values.
-The solution is to substitute the calculated residuals with a different term: the derivative of the loss function, which in our case is the log-likelihood of the normal distribution.
-Instead of estiamting stepwise residuals, one estiamtes hence stepwise the score.
+The concept for boosting scale parameters is the same. However,
+residuals for the variance estimates can not be calculated, since there
+is no such thing as ``observed'' variances for individual response
+values. The solution is to substitute the calculated residuals with a
+different term: the derivative of the loss function, which in our case
+is the log-likelihood of the normal distribution. Instead of estiamting
+stepwise residuals, one estiamtes hence stepwise the score.
 
-The resulting loss fcuntion then, is defined by sum of squares oa the term:
-ui (=deviance-residuals) - inverse predicted scores * working-residuals
+The resulting loss fcuntion then, is defined by sum of squares oa the
+term: ui (=deviance-residuals) - inverse predicted scores *
+working-residuals
 
-We employed the resid-function of the asp20model class, on which our class is built on.
-<!-- The working and response residuals are the observed values -->
-<!-- minus the fitted values for the location. -->
-<!-- The deviance and Pearson residuals are weighted residuals, -->
-<!-- divided by the fitted values for the scale. -->
+We employed the resid-function of the asp20model class, on which our
+class is built on.
 
-# 3. Implementation
+\hypertarget{implementation}{%
+\chapter{3. Implementation}\label{implementation}}
 
-In the current version of the asp20boost-package, the boosting algorithm is already implemented in order to allow componentwise boosting, which will be explained in part 4. 
-For good explanation and for following our development process in this part we first describe the implementation of ordinary boosting - which however does not appear in our code anymore.
+In the current version of the asp20boost-package, the boosting algorithm
+is already implemented in order to allow componentwise boosting, which
+will be explained in part 4. For good explanation and for following our
+development process in this part we first describe the implementation of
+ordinary boosting - which however does not appear in our code anymore.
 
-A simple booster for the location parameters beta works without extending the LocationScaleRegression-Class. 
-The current residuals may be extrated with the resid()-method, then estimated. The location parameter is updated. 
-This causes the LocationScaleRegression-Class to calculate updated state-dependent objects, such as the loglikelihood, gradients and the residuals. 
-This is what's needed to repeat the procedure of estimating residuals and updating location parameters.
+A simple booster for the location parameters beta works without
+extending the LocationScaleRegression-Class. The current residuals may
+be extrated with the resid()-method, then estimated. The location
+parameter is updated. This causes the LocationScaleRegression-Class to
+calculate updated state-dependent objects, such as the loglikelihood,
+gradients and the residuals. This is what's needed to repeat the
+procedure of estimating residuals and updating location parameters.
 
-The simple booster for gamma consists in the same two repeating steps of first estimating a term, and second updating the scale parameters gamma by adding the estimated effect - adjusted by a learning rate.
-The term to be estimated, however, is not the residual, but the derivative of the loglikelihood of the normal distribution. 
-To calculate this, our code makes use of the resid-function(), but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
+The simple booster for gamma consists in the same two repeating steps of
+first estimating a term, and second updating the scale parameters gamma
+by adding the estimated effect - adjusted by a learning rate. The term
+to be estimated, however, is not the residual, but the derivative of the
+loglikelihood of the normal distribution. To calculate this, our code
+makes use of the resid-function(), but this time passing the argument
+``deviance'', which results in residuals adjusted by the fitted scale
+estimates.
 
-[Is the gamma booster dependent on simultaneous running of beta booster?]
+{[}Is the gamma booster dependent on simultaneous running of beta
+booster?{]}
 
-Having implemented a boosting algorithm for location and scale, the code-basis for our package was achieved.
+Having implemented a boosting algorithm for location and scale, the
+code-basis for our package was achieved.
 
-# 4. Functionality: Componentwise Boosting
+\hypertarget{functionality-componentwise-boosting}{%
+\chapter{4. Functionality: Componentwise
+Boosting}\label{functionality-componentwise-boosting}}
 
-A usefull functionality of boosting is componentwise boosting. 
-The idea is to not update a whole parameter vector, but only one entry of it.
-The entry chosen for the update at this juncture is the one yielding the best improvement in the sense of lowering the loss function.
-The loss function in our case is the loglikelihood of the normally distributed data.
-Hence in each iteration only one component of the location-, as well as one component of the scale-parameters are estimated.
+A usefull functionality of boosting is componentwise boosting. The idea
+is to not update a whole parameter vector, but only one entry of it. The
+entry chosen for the update at this juncture is the one yielding the
+best improvement in the sense of lowering the loss function. The loss
+function in our case is the loglikelihood of the normally distributed
+data. Hence in each iteration only one component of the location-, as
+well as one component of the scale-parameters are estimated.
 
-The implementation in our package works via the extension of the LocationScaleRegressionClass by two active fields "bestFittingVariableBeta" and "bestFittingVariableGamma". 
-These functions partition the design matrix X, respectively Z, into its single columns and then estimate the residuals - respectively the score - seperately for each component, and determine loss functions for a hypothetical update with the respective component.
-Comparing with the old loss function value, the highest loss-improvement may be determined and the respective component is used to update the parameter-vector.
+The implementation in our package works via the extension of the
+LocationScaleRegressionClass by two active fields
+``bestFittingVariableBeta'' and ``bestFittingVariableGamma''. These
+functions partition the design matrix X, respectively Z, into its single
+columns and then estimate the residuals - respectively the score -
+seperately for each component, and determine loss functions for a
+hypothetical update with the respective component. Comparing with the
+old loss function value, the highest loss-improvement may be determined
+and the respective component is used to update the parameter-vector.
 
-We are in the process of reconsidering the design of this implementation.
-Other design possibilities are the following:
+We are in the process of reconsidering the design of this
+implementation. Other design possibilities are the following:
 
-* create public fields for the heavily used score-function values
-* move the calculation of componentwise losses to an external function
-* harmonize boosting and componentwise boosting into one external function, determining the mode of operation by an argument 'componentwise = TRUE'
+\begin{itemize}
+\tightlist
+\item
+  create public fields for the heavily used score-function values
+\item
+  move the calculation of componentwise losses to an external function
+\item
+  harmonize boosting and componentwise boosting into one external
+  function, determining the mode of operation by an argument
+  `componentwise = TRUE'
+\end{itemize}
 
-Another conceptual questions that comes up is if the best loss-improvement may be indicated by the already existing gradients, and hence there is no need for extra calculation.
+Another conceptual questions that comes up is if the best
+loss-improvement may be indicated by the already existing gradients, and
+hence there is no need for extra calculation.
 
-# 5. Prospects
+\hypertarget{prospects}{%
+\chapter{5. Prospects}\label{prospects}}
 
-## Further Functionalities
+\hypertarget{further-functionalities}{%
+\section{Further Functionalities}\label{further-functionalities}}
 
-The most important functionality we intend to implement in our package is variable selection via cross validation. Working out the theoretical concept behind this idea is one of our next milestones.
+The most important functionality we intend to implement in our package
+is variable selection via cross validation. Working out the theoretical
+concept behind this idea is one of our next milestones.
 
-One further, rather loose, idea is to find ways to optimize the learning rate in our boosting algorithm.
+One further, rather loose, idea is to find ways to optimize the learning
+rate in our boosting algorithm.
 
-## Stability
+\hypertarget{stability}{%
+\section{Stability}\label{stability}}
 
-We intend to implement further automated unit tests.
-This will enable us to assess quickly the stability of our code given differing inputs.
+We intend to implement further automated unit tests. This will enable us
+to assess quickly the stability of our code given differing inputs.
 Departing from this point we intend to reduce proneness to input errors.
 
-## Performance
+\hypertarget{performance}{%
+\section{Performance}\label{performance}}
 
-* A major performance problem in our code is that small learning rates for the beta-booster cause processing time of the code to increase sharply. This remains to be solved.
+\begin{itemize}
+\tightlist
+\item
+  A major performance problem in our code is that small learning rates
+  for the beta-booster cause processing time of the code to increase
+  sharply. This remains to be solved.
+\end{itemize}
 
-*A good choice of starting values for beta and gamma may enhance performance. A possible candidate is the mean of the response.
+*A good choice of starting values for beta and gamma may enhance
+performance. A possible candidate is the mean of the response.
 
-## Further
+\hypertarget{further}{%
+\section{Further}\label{further}}
 
-* Allow user-input of the LocationScaleRegressionBoost-model in form of a dataframe, for example by allowing to pass an optional "data"-argument as known from the lm-call.
-This may reduce proneness to input-errors.
+\begin{itemize}
+\item
+  Allow user-input of the LocationScaleRegressionBoost-model in form of
+  a dataframe, for example by allowing to pass an optional
+  ``data''-argument as known from the lm-call. This may reduce proneness
+  to input-errors.
+\item
+  Visualize the results of the boosted estimates
+\item
+  Document the extension of the R6-Class properly and inherit
+  documentation of the LocationScaleRegression-Class.
+\end{itemize}
 
-* Visualize the results of the boosted estimates
+\hypertarget{application}{%
+\section{Application}\label{application}}
 
-* Document the extension of the R6-Class properly and inherit documentation of the LocationScaleRegression-Class.
+{[}Johannes: Munich Rent Data? - Identify Outliers{]}
 
-## Application
-
-[Johannes: Munich Rent Data? - Identify Outliers]
+\end{document}

### 2020-06-04 Reports/sample_report/1_REPORT.tex
write first report

--- 2020-05-28 Reports/1_REPORT.tex
@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
 
@@
 \author{Johannes Strauß \and Levin Wiebelt \and Sebastian Aristizabal}
-\date{28 Mai 2020}
+\date{03 June 2020}
 
@@
 If you don't have Latex download latex for R with
-\texttt{tinytex::install\_tinytex()}.
-
-\hypertarget{write}{%
-\section{Write}\label{write}}
+\texttt{tinytex::install\_tinytex()}. To install pandoc on Mac, install
+it via Terminal with \texttt{brew\ install\ pandoc-citeproc}. \#\# Write
 
@@
 \begin{Highlighting}[]
-\CommentTok{# automatically create a bib database for R packages}
-\NormalTok{knitr}\OperatorTok{::}\KeywordTok{write_bib}\NormalTok{(}\KeywordTok{c}\NormalTok{(}
-  \KeywordTok{.packages}\NormalTok{(), }\StringTok{'R6'}\NormalTok{, }\StringTok{'knitr'}\NormalTok{, }\StringTok{'rmarkdown'}\NormalTok{, }\StringTok{'asp20model'}\NormalTok{, }\StringTok{'gamboostLSS'}\NormalTok{,}\StringTok{'mboost'}\NormalTok{, }\StringTok{'tidyverse'}
-\NormalTok{), }\StringTok{'packages.bib'}\NormalTok{)}
+\CommentTok{\# automatically create a bib database for R packages}
+\NormalTok{knitr}\OperatorTok{::}\KeywordTok{write\_bib}\NormalTok{(}\KeywordTok{c}\NormalTok{(}
+  \KeywordTok{.packages}\NormalTok{(), }\StringTok{\textquotesingle{}R6\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}knitr\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}rmarkdown\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}asp20model\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}gamboostLSS\textquotesingle{}}\NormalTok{,}\StringTok{\textquotesingle{}mboost\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}tidyverse\textquotesingle{}}
+\NormalTok{), }\StringTok{\textquotesingle{}packages.bib\textquotesingle{}}\NormalTok{)}
 \end{Highlighting}
@@
 \item
-  We expand the R6 class \emph{``LocationScaleRegression''} belonging to
+  We extend the R6 class \emph{``LocationScaleRegression''} belonging to
   the \texttt{asp20model} package concieved specifficaly for this
@@
   Our aim is to develope an multi-faceted implementation of boosting for
-  location and scale including component-wise boosting, use of
-  cross-validation to determine the optimal stopping critera and useful
-  options for visualization.
+  location and scale regression including component-wise boosting, use
+  of cross-validation to determine the optimal stopping critera and
+  useful options for visualization.
 \end{itemize}
@@
 
-JS: Performance Needs to be checked and maybe more outputs - messages?
-Starting point of Beta and Gamma?
+Our package rely heavily on the asp20model package, any changes, or
+errors in the code of the package might break the functionality.
+Currently there is no validation of the given input, with respect to
+heteroscedasticity.
+
+The package also does not provide any auto optimization of the gradient
+boost step-size, a wrongly picked step-size can lead to a deficient
+performance, but an already implemented early stopping mechanism can
+prevent an overfitting. An input check and cross-validation of the
+step-size is on the to-do list. Also, the initial starting values for
+Beta and Gamma can be further improved, see (Fahrmeir 2013, 219).
+Further optimizations can be made to reduce the computation time, by
+moving the calculation of the Projection Matrix for individual
+covariates in the initialization section of the model. In regard to
+User-friendliness of the package, we will try to implement further input
+validation and try to write up with a comprehensive vignette containing
+various examples in the next weeks.
 
@@
 \end{quote}
+
+JS: During the learning process, we tried several methods to implement
+the required functionality. At first, we started with a simple function
+in a R-File. While looking at other online resources, we agreed on using
+the asp20model packages as a requirement for our package and inheriting
+the existing LocationScaleRegression class. Our
+\texttt{LocationScaleRegressionBoost} R6 class implements a boosting
+algorithm, two functions within the model for calculating the best
+fitting covariate with respect to Gamma and Beta and has four new
+variables which are set at the initialization of the model to reduce
+computation time of the Projection Matrixes during the gradient boosting
+(subject to change). The package provides the gradient boost function
+for the \texttt{LocationScaleRegressionBoost} class, which uses the two
+model functions to optimize stepwise the Beta and Gamma Vector.
 
@@
 \item
-  boosting for multidimensional (\textgreater2) beta or gamma
+  boosting for multidimensional (\textgreater2) beta and gamma
 \item
@@
   Visualization of results via connecting to plot team
+  \texttt{asp20plot}
 \item
@@
 \end{itemize}
+
+JS: check when run gradientboost, if model is
+LocationScaleRegressionBoost class
 
@@
 \hypertarget{refs}{}
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-R-R6}{}%
@@
 Regression Models}.
+\end{cslreferences}
 

### 2020-06-04 Reports/report_1/1_REPORT.Rmd
update report_1

@@
 
+**[S: please format all equations and terms - y, xi with a subscripted i, zi (same), ...]**
+
 The course "Advanced Statistical Programming with R" consists in implementing a package to address location-scale regression. 
@@
 
-In part 2 of this report we explain the concept of boosting and elaborate on how it suites to address the problem of location-scale regression. 
-In the following part 3 we describe our implementation milestones, whereby part 4 especially elaborates on componentwise boosting. 
-In part 5 we present our thoughts on upcoming challenges and critical aspects of the package development.
+In *part 2* of this report we explain the concept of boosting and elaborate on how it suites to address the problem of location-scale regression. 
+In the following *part 3* we describe our implementation milestones, whereby *part 4* especially elaborates on componentwise boosting. 
+In *part 5* we present our thoughts on upcoming challenges and critical aspects of the package development.
 
@@
 
+**[S: please format all terms - xi, beta and gamma as greek letters, v also as greek letter ...]**
+**[S: please add reference]**
+
 The first step in our progress was to understand the theoretical concept of boosting, and which types of problems it is able to solve.
-In the following the term "booster" refers to a boosting algorithm.
+To explain the concept and follow our process of understanding we first explain boosting for location parameters.
+Boosting scale parameters requires another way of thinking about it and constituted a first milestone in our understanding and implementation.
+In the following, the term "booster" refers to a boosting algorithm.
 
@@
 
-To explain the concept and follow our process of understanding we first explain boosting for location parameters.
-Boosting scale parameters requires another way of thinking about it and constituted a first milestone in our understanding and implementation.
 The boosting concept relies on the idea of iterative estimation. 
@@
 However, with estimates and the response vector at hand, residuals can be calculated. 
-The boosting-algortihm focuses on those calculated residuals.
-The effect of the predictors xi on these residuals is estimated by least squares and the resulting effect is added to the location-estimate beta, adjusted by some learning rate. 
+The boosting algortihm focuses on those calculated residuals.
+The effect of the predictors xi on these residuals is estimated by least squares and the resulting effect is added to the location-estimate beta, adjusted by some learning rate *v*. 
 This yields a better fit of the location-model, which results in smaller residuals.
@@
 A simple booster stops after a fixed number of iterations is reached.
-**(comp. Kneib et al)**
+**(compare Kneib et al, p217,218)**
 
@@
 
-[Loss-function erklären: Boosting can be viewed as lowering iteratively the amount of a loss function, which is in our case the score of the normal distribution]
+A boosting algorithm for scale parameters **gamma** consists in the same two repeating steps of first estimating a term, and second updating **gamma** by adding the estimated effect - adjusted by a learning rate.
+The term in the estimating step is, however, not the residual.
+For boosting **gamma** this term equals the score function of the model, that is the derivative of the loglikelihood of the normal distribution.
 
-The concept for boosting scale parameters is the same.
+**[Why is that? kurze Erklärung]**
+
+<!-- The concept for boosting scale parameters is the same.
 However, residuals for the variance estimates can not be calculated, since there is no such thing as "observed" variances for individual response values.
@@
 Instead of estiamting stepwise residuals, one estiamtes hence stepwise the score.
+-->
 
-The resulting loss fcuntion then, is defined by sum of squares oa the term:
-ui (=deviance-residuals) - inverse predicted scores * working-residuals
+<!-- The resulting loss fcuntion then, is defined by sum of squares oa the term: -->
+<!-- ui (=deviance-residuals) - inverse predicted scores * working-residuals -->
 
-We employed the resid-function of the asp20model class, on which our class is built on.
+<!-- We employed the resid-function of the asp20model class, on which our class is built on. -->
 <!-- The working and response residuals are the observed values -->
@@
 
-In the current version of the asp20boost-package, the boosting algorithm is already implemented in order to allow componentwise boosting, which will be explained in part 4. 
+The implementation of the boosting algorithm is accomplished by a function *gradient_boost*, which harmonizes with the R6-Class *LocationScaleRegressionBoost*. 
+This class, in turn, is built upon the *LocationScaleRegression* class of the `asp20model`-package. 
+It follows a strict dependency of the `asp20boost`- on the `asp20model`-package.
+
+In the current version of the `asp20boost`-package, the boosting algorithm is already implemented in order to allow componentwise boosting, which will be explained in part 4. 
 For good explanation and for following our development process in this part we first describe the implementation of ordinary boosting - which however does not appear in our code anymore.
 
-A simple booster for the location parameters beta works without extending the LocationScaleRegression-Class. 
-The current residuals may be extrated with the resid()-method, then estimated. The location parameter is updated. 
-This causes the LocationScaleRegression-Class to calculate updated state-dependent objects, such as the loglikelihood, gradients and the residuals. 
-This is what's needed to repeat the procedure of estimating residuals and updating location parameters.
+A simple booster for the location parameters **beta** works without extending the *LocationScaleRegression* class. 
+The current residuals may be extrated with the *resid()*-method, then estimated. The location parameter is updated. 
+This causes the *LocationScaleRegression* class to calculate updated state-dependent objects, such as the loglikelihood, gradients and the residuals. 
+This is what's needed to repeat the procedure of estimating residuals and updating location parameters **Beta**.
 
-The simple booster for gamma consists in the same two repeating steps of first estimating a term, and second updating the scale parameters gamma by adding the estimated effect - adjusted by a learning rate.
-The term to be estimated, however, is not the residual, but the derivative of the loglikelihood of the normal distribution. 
+The simple booster for the scale parameters **gamma**, as already mentioned, consists in the same two repeating steps of first estimating a term, and second updating the scale parameters gamma by adding the estimated effect - adjusted by a learning rate **v**.
+The estimated term here is the derivative of the loglikelihood of the normal distribution.
 To calculate this, our code makes use of the resid-function(), but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
 
-[Is the gamma booster dependent on simultaneous running of beta booster?]
+**[Erklärung weiter ausführen]**
 
@@
 The loss function in our case is the loglikelihood of the normally distributed data.
-Hence in each iteration only one component of the location-, as well as one component of the scale-parameters are estimated.
+Hence in each iteration only one component of the location-, as well as one component of the scale-parameters are updated.
 
-The implementation in our package works via the extension of the LocationScaleRegressionClass by two active fields "bestFittingVariableBeta" and "bestFittingVariableGamma". 
-These functions partition the design matrix X, respectively Z, into its single columns and then estimate the residuals - respectively the score - seperately for each component, and determine loss functions for a hypothetical update with the respective component.
+The implementation in our package works via the extension of the **LocationScaleRegression** class by two active fields *"bestFittingVariableBeta"* and *"bestFittingVariableGamma"*. 
+These functions partition the design matrix *X* - respectively *Z* - into its single columns and then estimate the residuals - respectively the scores - seperately for each component, and determine loss functions for a hypothetical update with the respective component.
 Comparing with the old loss function value, the highest loss-improvement may be determined and the respective component is used to update the parameter-vector.
@@
 
-The most important functionality we intend to implement in our package is variable selection via cross validation. Working out the theoretical concept behind this idea is one of our next milestones.
+The most important functionality we intend to implement in our package is optimizing the stopping point of our booster.
+This optimization of the number of iterations may be done via cross validation.
+Working out the theoretical concept behind this idea is one of our next milestones.
 
@@
 * A major performance problem in our code is that small learning rates for the beta-booster cause processing time of the code to increase sharply. This remains to be solved.
-
 *A good choice of starting values for beta and gamma may enhance performance. A possible candidate is the mean of the response.
@@
 
-* Document the extension of the R6-Class properly and inherit documentation of the LocationScaleRegression-Class.
+* Document the extension of the R6-Class properly and inherit documentation of the LocationScaleRegression-Class applying roxygen2.
 

### 2020-06-04 Reports/report_1/1_REPORT.tex
update report_1

@@
 
+\textbf{{[}S: please format all equations and terms - y, xi with a
+subscripted i, zi (same), \ldots{]}}
+
 The course ``Advanced Statistical Programming with R'' consists in
@@
 
-In part 2 of this report we explain the concept of boosting and
+In \emph{part 2} of this report we explain the concept of boosting and
 elaborate on how it suites to address the problem of location-scale
-regression. In the following part 3 we describe our implementation
-milestones, whereby part 4 especially elaborates on componentwise
-boosting. In part 5 we present our thoughts on upcoming challenges and
-critical aspects of the package development.
+regression. In the following \emph{part 3} we describe our
+implementation milestones, whereby \emph{part 4} especially elaborates
+on componentwise boosting. In \emph{part 5} we present our thoughts on
+upcoming challenges and critical aspects of the package development.
 
@@
 
+\textbf{{[}S: please format all terms - xi, beta and gamma as greek
+letters, v also as greek letter \ldots{]}} \textbf{{[}S: please add
+reference{]}}
+
 The first step in our progress was to understand the theoretical concept
-of boosting, and which types of problems it is able to solve. In the
-following the term ``booster'' refers to a boosting algorithm.
+of boosting, and which types of problems it is able to solve. To explain
+the concept and follow our process of understanding we first explain
+boosting for location parameters. Boosting scale parameters requires
+another way of thinking about it and constituted a first milestone in
+our understanding and implementation. In the following, the term
+``booster'' refers to a boosting algorithm.
 
@@
 
-To explain the concept and follow our process of understanding we first
-explain boosting for location parameters. Boosting scale parameters
-requires another way of thinking about it and constituted a first
-milestone in our understanding and implementation. The boosting concept
-relies on the idea of iterative estimation. To estimate the location
-parameters beta via boosting, one starts with initial estimates, which
-may be far from optimal. However, with estimates and the response vector
-at hand, residuals can be calculated. The boosting-algortihm focuses on
-those calculated residuals. The effect of the predictors xi on these
-residuals is estimated by least squares and the resulting effect is
-added to the location-estimate beta, adjusted by some learning rate.
-This yields a better fit of the location-model, which results in smaller
-residuals. These new residuals are estimated again in the next
-iteration, yielding smaller effect sizes, and hence a convergence of the
-boosting algorithm towards the OLS-estimate. A simple booster stops
-after a fixed number of iterations is reached. \textbf{(comp. Kneib et
-al)}
+The boosting concept relies on the idea of iterative estimation. To
+estimate the location parameters beta via boosting, one starts with
+initial estimates, which may be far from optimal. However, with
+estimates and the response vector at hand, residuals can be calculated.
+The boosting algortihm focuses on those calculated residuals. The effect
+of the predictors xi on these residuals is estimated by least squares
+and the resulting effect is added to the location-estimate beta,
+adjusted by some learning rate \emph{v}. This yields a better fit of the
+location-model, which results in smaller residuals. These new residuals
+are estimated again in the next iteration, yielding smaller effect
+sizes, and hence a convergence of the boosting algorithm towards the
+OLS-estimate. A simple booster stops after a fixed number of iterations
+is reached. \textbf{(compare Kneib et al, p217,218)}
 
@@
 
-{[}Loss-function erklären: Boosting can be viewed as lowering
-iteratively the amount of a loss function, which is in our case the
-score of the normal distribution{]}
+A boosting algorithm for scale parameters \textbf{gamma} consists in the
+same two repeating steps of first estimating a term, and second updating
+\textbf{gamma} by adding the estimated effect - adjusted by a learning
+rate. The term in the estimating step is, however, not the residual. For
+boosting \textbf{gamma} this term equals the score function of the
+model, that is the derivative of the loglikelihood of the normal
+distribution.
 
-The concept for boosting scale parameters is the same. However,
-residuals for the variance estimates can not be calculated, since there
-is no such thing as ``observed'' variances for individual response
-values. The solution is to substitute the calculated residuals with a
-different term: the derivative of the loss function, which in our case
-is the log-likelihood of the normal distribution. Instead of estiamting
-stepwise residuals, one estiamtes hence stepwise the score.
-
-The resulting loss fcuntion then, is defined by sum of squares oa the
-term: ui (=deviance-residuals) - inverse predicted scores *
-working-residuals
-
-We employed the resid-function of the asp20model class, on which our
-class is built on.
+\textbf{{[}Why is that? kurze Erklärung{]}}
 
@@
 
-In the current version of the asp20boost-package, the boosting algorithm
-is already implemented in order to allow componentwise boosting, which
-will be explained in part 4. For good explanation and for following our
-development process in this part we first describe the implementation of
-ordinary boosting - which however does not appear in our code anymore.
+The implementation of the boosting algorithm is accomplished by a
+function \emph{gradient\_boost}, which harmonizes with the R6-Class
+\emph{LocationScaleRegressionBoost}. This class, in turn, is built upon
+the \emph{LocationScaleRegression} class of the
+\texttt{asp20model}-package. It follows a strict dependency of the
+\texttt{asp20boost}- on the \texttt{asp20model}-package.
 
-A simple booster for the location parameters beta works without
-extending the LocationScaleRegression-Class. The current residuals may
-be extrated with the resid()-method, then estimated. The location
-parameter is updated. This causes the LocationScaleRegression-Class to
-calculate updated state-dependent objects, such as the loglikelihood,
-gradients and the residuals. This is what's needed to repeat the
-procedure of estimating residuals and updating location parameters.
+In the current version of the \texttt{asp20boost}-package, the boosting
+algorithm is already implemented in order to allow componentwise
+boosting, which will be explained in part 4. For good explanation and
+for following our development process in this part we first describe the
+implementation of ordinary boosting - which however does not appear in
+our code anymore.
 
-The simple booster for gamma consists in the same two repeating steps of
-first estimating a term, and second updating the scale parameters gamma
-by adding the estimated effect - adjusted by a learning rate. The term
-to be estimated, however, is not the residual, but the derivative of the
-loglikelihood of the normal distribution. To calculate this, our code
-makes use of the resid-function(), but this time passing the argument
-``deviance'', which results in residuals adjusted by the fitted scale
-estimates.
+A simple booster for the location parameters \textbf{beta} works without
+extending the \emph{LocationScaleRegression} class. The current
+residuals may be extrated with the \emph{resid()}-method, then
+estimated. The location parameter is updated. This causes the
+\emph{LocationScaleRegression} class to calculate updated
+state-dependent objects, such as the loglikelihood, gradients and the
+residuals. This is what's needed to repeat the procedure of estimating
+residuals and updating location parameters \textbf{Beta}.
 
-{[}Is the gamma booster dependent on simultaneous running of beta
-booster?{]}
+The simple booster for the scale parameters \textbf{gamma}, as already
+mentioned, consists in the same two repeating steps of first estimating
+a term, and second updating the scale parameters gamma by adding the
+estimated effect - adjusted by a learning rate \textbf{v}. The estimated
+term here is the derivative of the loglikelihood of the normal
+distribution. To calculate this, our code makes use of the
+resid-function(), but this time passing the argument ``deviance'', which
+results in residuals adjusted by the fitted scale estimates.
+
+\textbf{{[}Erklärung weiter ausführen{]}}
 
@@
 data. Hence in each iteration only one component of the location-, as
-well as one component of the scale-parameters are estimated.
+well as one component of the scale-parameters are updated.
 
 The implementation in our package works via the extension of the
-LocationScaleRegressionClass by two active fields
-``bestFittingVariableBeta'' and ``bestFittingVariableGamma''. These
-functions partition the design matrix X, respectively Z, into its single
-columns and then estimate the residuals - respectively the score -
-seperately for each component, and determine loss functions for a
-hypothetical update with the respective component. Comparing with the
-old loss function value, the highest loss-improvement may be determined
-and the respective component is used to update the parameter-vector.
+\textbf{LocationScaleRegression} class by two active fields
+\emph{``bestFittingVariableBeta''} and
+\emph{``bestFittingVariableGamma''}. These functions partition the
+design matrix \emph{X} - respectively \emph{Z} - into its single columns
+and then estimate the residuals - respectively the scores - seperately
+for each component, and determine loss functions for a hypothetical
+update with the respective component. Comparing with the old loss
+function value, the highest loss-improvement may be determined and the
+respective component is used to update the parameter-vector.
 
@@
 The most important functionality we intend to implement in our package
-is variable selection via cross validation. Working out the theoretical
-concept behind this idea is one of our next milestones.
+is optimizing the stopping point of our booster. This optimization of
+the number of iterations may be done via cross validation. Working out
+the theoretical concept behind this idea is one of our next milestones.
 
@@
   for the beta-booster cause processing time of the code to increase
-  sharply. This remains to be solved.
+  sharply. This remains to be solved. *A good choice of starting values
+  for beta and gamma may enhance performance. A possible candidate is
+  the mean of the response.
 \end{itemize}
-
-*A good choice of starting values for beta and gamma may enhance
-performance. A possible candidate is the mean of the response.
 
@@
   Document the extension of the R6-Class properly and inherit
-  documentation of the LocationScaleRegression-Class.
+  documentation of the LocationScaleRegression-Class applying roxygen2.
 \end{itemize}

### 2020-06-04 Reports/report_1/1_REPORT.Rmd
updated report_1

@@
 
-**[Why is that? kurze Erklärung]**
-
-<!-- The concept for boosting scale parameters is the same.
-However, residuals for the variance estimates can not be calculated, since there is no such thing as "observed" variances for individual response values.
-The solution is to substitute the calculated residuals with a different term: the derivative of the loss function, which in our case is the log-likelihood of the normal distribution.
-Instead of estiamting stepwise residuals, one estiamtes hence stepwise the score.
--->
-
-<!-- The resulting loss fcuntion then, is defined by sum of squares oa the term: -->
-<!-- ui (=deviance-residuals) - inverse predicted scores * working-residuals -->
-
-<!-- We employed the resid-function of the asp20model class, on which our class is built on. -->
-<!-- The working and response residuals are the observed values -->
-<!-- minus the fitted values for the location. -->
-<!-- The deviance and Pearson residuals are weighted residuals, -->
-<!-- divided by the fitted values for the scale. -->
-
 # 3. Implementation
@@
 To calculate this, our code makes use of the resid-function(), but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
-
-**[Erklärung weiter ausführen]**
 
@@
 * A major performance problem in our code is that small learning rates for the beta-booster cause processing time of the code to increase sharply. This remains to be solved.
-*A good choice of starting values for beta and gamma may enhance performance. A possible candidate is the mean of the response.
+* A good choice of starting values for beta and gamma may enhance performance. A possible candidate is the mean of the response.
 
-## Further
+## Further Aspects
 

### 2020-06-04 R/model.R
add new comments

@@ -55,17 +55,22 @@ LocationScaleRegressionBoost <- R6Class("LocationScaleRegression",
                                             ui <- self$resid("deviance")
                                             for(n in 1:dim(private$Z)[2]) {
                                                zn <- private$Z[,n]
-
+                                                #u -> first derivation of Log Likelihood - see documentation for further reading
                                                u <- ((self$resid()^2)*private$Z[,n]*exp(-2*(drop(private$Z %*% self$gamma)))-private$Z[,n])
+                                               #bjhat -> Fit separate linear models
+                                               #Kneib book Page 226
+                                               #Johannes: I can't really explain this
                                                #Can be further improvied by moving ProjectionMatrix Calculation to initializalisation
                                                bjhat <- solve(t(private$Z[,n])%*%private$Z[,n])%*%t(private$Z[,n])%*% u
-                                               loss[n]<-sum((ui-(self$resid()/zn%*%bjhat))^2)
+                                               #Calculate new squared error by subtracting the partial deviance of covariate with new gamma (first derivation u) of the total deviance
+                                               #partial deviance -> (self$resid()/zn%*%bjhat)
+                                               loss[n] <- sum((ui-(self$resid()/zn%*%bjhat))^2)
                                                }
-                                            #determine index of the best-fitting variable
+                                            #determine index of the best-fitting variable, the covariate with the greates influance on the deviance will decrease the loss at most
                                             indexOfGammaUpdate = which.min(loss)
                                             #build update vector
                                             updateGamma <- replicate(length(model$gamma), 0)
-                                            #Calculates current first derivates for Gamma (index of the best-fitting variable) of Log Likelehood
+                                            #Calculates current first derivates for Gamma (index of the best-fitting variable) again like u, but as sum, since Gamma covariate is a single
                                             updateGamma[indexOfGammaUpdate] <- sum((self$resid()^2)*private$Z[,indexOfGammaUpdate]*exp(-2*(drop(private$Z %*% self$gamma)))-private$Z[,indexOfGammaUpdate])
                                             updateGamma
                                           },

### 2020-06-05 Reports/report_1/1_REPORT.Rmd
Formatting round 1 -Style -Latex -Language

@@
 title: "Report ASP20 Boost"
-subtitle: "First Report"
 author:
-- "Johannes Strauß"
-- "Levin Wiebelt"
-- "Sebastian Aristizabal"
+- Johannes Strauß
+- Levin Wiebelt
+- Sebastian Aristizabal
 date: "`r format(Sys.time(), '%d %B %Y')`"
-tags: [nothing, nothingness]
+output:
+  html_document:
+    df_print: paged
+    toc: yes
+    toc_depth: '2'
+  pdf_document:
+    keep_tex: yes
+    latex_engine: xelatex
+    template: null
+    toc: yes
+    toc_depth: 2
 documentclass: report
-bibliography: ["citavi_27.05.bib", "packages.bib"]
+geometry: left=4cm, right=3cm, top=2.5cm, bottom=2.5cm
+lof: no
+lot: no
+bibliography:
+- citavi_27.05.bib
+- packages.bib
+subtitle: First Report
+tags:
+- nothing
+- nothingness
 biblio-style: apalike
-geometry: "left=4cm, right=3cm, top=2.5cm, bottom=2.5cm"
-lof: FALSE
-lot: FALSE
-output: 
-    pdf_document:
-      toc: true
-      toc_depth: 2
-      latex_engine: xelatex
-      keep_tex: yes
-      template: NULL
 ---
@@
 
+```{r pack tex, eval=FALSE, include=TRUE}
+# automatically create a bib database for R packages
+knitr::write_bib(c(
+  .packages(), 'R6', 'knitr', 'rmarkdown', 'asp20model', 'gamboostLSS','mboost', 'tidyverse'  
+), 'packages.bib')
+```
+
 # 1. Introduction:
 
-**[S: please format all equations and terms - y, xi with a subscripted i, zi (same), ...]**
-
 The course "Advanced Statistical Programming with R" consists in implementing a package to address location-scale regression. 
-The sample dataset consists of a response y, that's expectation is in linear relationship to a set of predictors xi. 
-Its variance, in contrast, is dependent on a different set of linear predictors zi, however transformed by the response function g(x) = exp(x). 
+The sample dataset consists of a response $y$, i.e, the expectation is in linear relationship to a set of predictors $x_i$. 
+Its variance, in contrast, is dependent on a different set of linear predictors $z_i$, however transformed by the response function $g(x) = exp(x)$. 
 This ensures positiveness of the variances. 
-The goal approach is to estimate the effects of both: xi on the expectation of y (location) - this is captured by the estimates beta, and zi on the variance of y (scale) - this is captured by the estimates gamma.
+The goal approach is to estimate the effects of both: $x_i$ on the expectation of $y$ (location) - this is captured by the estimates $\beta$, and $z_i$ on the variance of $y$ (scale) - this is captured by the estimates $\gamma$.
 Our student group `asp20boost` solves this task applying the concept of boosting.
 
+```{r data, echo=FALSE, fig.width=}
+n <- 500
+x <- runif(n)
+y <- x + rnorm(n, sd = exp(-3 + 2 * x))
+plot(x, y)
+abline(0, 1, lwd = 2)
+curve(x + 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
+curve(x - 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
+```
+
 In *part 2* of this report we explain the concept of boosting and elaborate on how it suites to address the problem of location-scale regression. 
-In the following *part 3* we describe our implementation milestones, whereby *part 4* especially elaborates on componentwise boosting. 
+In *part 3* we describe our implementation milestones, 
+whereby *part 4* especially elaborates on componentwise boosting. 
 In *part 5* we present our thoughts on upcoming challenges and critical aspects of the package development.
@@
 
-**[S: please format all terms - xi, beta and gamma as greek letters, v also as greek letter ...]**
-**[S: please add reference]**
-
-The first step in our progress was to understand the theoretical concept of boosting, and which types of problems it is able to solve.
+The first step in our progress was to understand the theoretical concept of boosting and which types of problems it is able to solve.
 To explain the concept and follow our process of understanding we first explain boosting for location parameters. 
-Boosting scale parameters requires another way of thinking about it and constituted a first milestone in our understanding and implementation.
+Boosting scale parameters required another way of thinking about it and constituted a first milestone in our understanding and implementation.
 In the following, the term "booster" refers to a boosting algorithm.
@@
 
+The following descrition of simple boosting algorithm is mainly based on @Fahrmeir.2013, 217-219.
+
 The boosting concept relies on the idea of iterative estimation. 
-To estimate the location parameters beta via boosting, one starts with initial estimates, which may be far from optimal. 
+To estimate the location parameters $\beta$ via boosting, one starts with initial estimates, which may be far from optimal. 
 However, with estimates and the response vector at hand, residuals can be calculated. 
 The boosting algortihm focuses on those calculated residuals.
-The effect of the predictors xi on these residuals is estimated by least squares and the resulting effect is added to the location-estimate beta, adjusted by some learning rate *v*. 
+The effect of the predictors $x_i$ on these residuals is estimated using least squares and the resulting effect is added to the location-estimate $\beta$ adjusting it  by means of the learning rate $\nu$.
 This yields a better fit of the location-model, which results in smaller residuals.
-These new residuals are estimated again in the next iteration, yielding smaller effect sizes, and hence a convergence of the boosting algorithm towards the OLS-estimate.
+> These new residuals are estimated again in the next iteration, yielding smaller effect sizes, and hence a convergence of the boosting algorithm towards the OLS-estimate.
+
 A simple booster stops after a fixed number of iterations is reached.
-**(compare Kneib et al, p217,218)**
 
@@
 
-A boosting algorithm for scale parameters **gamma** consists in the same two repeating steps of first estimating a term, and second updating **gamma** by adding the estimated effect - adjusted by a learning rate.
+A boosting algorithm for scale parameters $\gamma$ consists in the same two repeating steps of first estimating a term, and second updating $\gamma$ by adding the estimated effect - adjusted by a learning rate.
 The term in the estimating step is, however, not the residual.
-For boosting **gamma** this term equals the score function of the model, that is the derivative of the loglikelihood of the normal distribution.
+For boosting $\gamma$ this term equals the score function of the model i.e  the derivative of the loglikelihood of the normal distribution.
 
@@
 
-The implementation of the boosting algorithm is accomplished by a function *gradient_boost*, which harmonizes with the R6-Class *LocationScaleRegressionBoost*. 
-This class, in turn, is built upon the *LocationScaleRegression* class of the `asp20model`-package. 
-It follows a strict dependency of the `asp20boost`- on the `asp20model`-package.
+To implement a boosting algorithm we constructed the function `gradient_boost` as a part of our R6-Class `LocationScaleRegressionBoost`.
+which  directly builds upon the `LocationScaleRegression` R6 class of the `asp20model`package set at our disposal for this course and this purpose specifically [@R-asp20model].
+?[@R-R6].
 
-In the current version of the `asp20boost`-package, the boosting algorithm is already implemented in order to allow componentwise boosting, which will be explained in part 4. 
-For good explanation and for following our development process in this part we first describe the implementation of ordinary boosting - which however does not appear in our code anymore.
+In the current version of the `asp20boost`-package, the boosting algorithm is already implemented in order to allow for componentwise boosting, which will be explained in part 4.
 
-A simple booster for the location parameters **beta** works without extending the *LocationScaleRegression* class. 
-The current residuals may be extrated with the *resid()*-method, then estimated. The location parameter is updated. 
-This causes the *LocationScaleRegression* class to calculate updated state-dependent objects, such as the loglikelihood, gradients and the residuals. 
-This is what's needed to repeat the procedure of estimating residuals and updating location parameters **Beta**.
+To further provide clarity, accountability and understanding during to our development process we describe briefly our first implementation of an ordinary boosting algorithm.
+This isn't visible in our code anymore.
 
-The simple booster for the scale parameters **gamma**, as already mentioned, consists in the same two repeating steps of first estimating a term, and second updating the scale parameters gamma by adding the estimated effect - adjusted by a learning rate **v**.
+A simple booster for the location parameters $\beta$ works without extending the `asp20model`'s `LocationScaleRegression` class. 
+The current residuals may be extracted with the `resid()` command and then estimated.
+The location parameter is updated. 
+This causes the `LocationScaleRegression` class to calculate updated state-dependent objects, such as the loglikelihood, the gradients and the residuals. 
+This objects are again used to repeatedly  estimate residuals and update the location parameter $\beta$.
+
+The simple booster for the scale parameters $\gamma$, as already mentioned, consists in repeating the same two steps of first, estimating a term and then updating the scale parameters $\gamma$ by adding the estimated effect - adjusted by the learning rate $\nu$.
 The estimated term here is the derivative of the loglikelihood of the normal distribution.
-To calculate this, our code makes use of the resid-function(), but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
-
-Having implemented a boosting algorithm for location and scale, the code-basis for our package was achieved.
+To calculate this, our code makes use of the `resid-function()`, but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
+This simple implementation of a boosting algorithm for location and scale served the code-basis or "skeleton" for our package and can, hence, be though as the file milestone of this project.
 
@@
 
-A usefull functionality of boosting is componentwise boosting. 
-The idea is to not update a whole parameter vector, but only one entry of it.
-The entry chosen for the update at this juncture is the one yielding the best improvement in the sense of lowering the loss function.
-The loss function in our case is the loglikelihood of the normally distributed data.
-Hence in each iteration only one component of the location-, as well as one component of the scale-parameters are updated.
+A usefull extention of our simple boosting algorithm regards *componentwise boosting*. 
+The key idea here is to not update a whole parameter vector, but only one entry of it.
+The entry chosen for the update at this juncture is the one yielding the best improvement in terms of minimization of the loss function.
+The loss function in our case is the loglikelihood function of our normally distributed data.
+Hence, in each iteration only one component of the location- and one component of the scale-parameters are updated.
 
-The implementation in our package works via the extension of the **LocationScaleRegression** class by two active fields *"bestFittingVariableBeta"* and *"bestFittingVariableGamma"*. 
-These functions partition the design matrix *X* - respectively *Z* - into its single columns and then estimate the residuals - respectively the scores - seperately for each component, and determine loss functions for a hypothetical update with the respective component.
-Comparing with the old loss function value, the highest loss-improvement may be determined and the respective component is used to update the parameter-vector.
+We implement this in our package by the extending the `LocationScaleRegression` class to include the two active fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.
+
+These functions partition the design matrix $X$- respectively $Z$ - into its single columns and then estimate the residuals - respectively the scores - seperately for each component, and determine loss functions for a hypothetical update with the respective component.
+Respective to the old loss function value, the highest loss-improvement is determined and the respective component is used to update the parameter-vector.
 
@@
 
-* create public fields for the heavily used score-function values
-* move the calculation of componentwise losses to an external function
-* harmonize boosting and componentwise boosting into one external function, determining the mode of operation by an argument 'componentwise = TRUE'
+* Create public fields for the heavily used score-function values.
+* Move the calculation of componentwise losses to an external function.
+* Harmonize boosting and componentwise boosting into one external function, determining the mode of operation by an argument `componentwise = TRUE`.
 
@@
 
-The most important functionality we intend to implement in our package is optimizing the stopping point of our booster.
-This optimization of the number of iterations may be done via cross validation.
-Working out the theoretical concept behind this idea is one of our next milestones.
-
+The most important functionality we intend to implement in our package is the optimization of the number of iterations via cross validation. 
+To work out the theoretical concept behind this idea is, as well as to implementing it are our current next goals.
 One further, rather loose, idea is to find ways to optimize the learning rate in our boosting algorithm.
@@
 This will enable us to assess quickly the stability of our code given differing inputs.
-Departing from this point we intend to reduce proneness to input errors.
+We also intend to implement measures for proneness reduction to input errors.
 
@@
 
-* A major performance problem in our code is that small learning rates for the beta-booster cause processing time of the code to increase sharply. This remains to be solved.
-* A good choice of starting values for beta and gamma may enhance performance. A possible candidate is the mean of the response.
+* A major performance problem in our code is that small learning rates for the $\beta$-booster lead to a  sharp increase of our code's processing time. This remains to be solved.
+* A good choice of starting values for $\beta$ and $\gamma$ may enhance performance. A possible candidate is the mean of the response.
 
@@
 
-* Document the extension of the R6-Class properly and inherit documentation of the LocationScaleRegression-Class applying roxygen2.
+* Document the extension of the R6-Class properly and inherit documentation of the `LocationScaleRegression-Class` using roxygen2.
 
@@
 [Johannes: Munich Rent Data? - Identify Outliers]
+If you want to cite Kneib to include this copy the following and update the page number:  @Fahrmeir.2013, 217-219.
+
+# 6. References

### 2020-06-05 Reports/report_1/1_REPORT.Rmd
Formatting round 1 -Style -Latex -Language

@@
 output:
-  html_document:
-    df_print: paged
-    toc: yes
-    toc_depth: '2'
   pdf_document:
-    keep_tex: yes
+    keep_tex: no
     latex_engine: xelatex
@@
 geometry: left=4cm, right=3cm, top=2.5cm, bottom=2.5cm
-lof: no
-lot: no
+lof: FALSE
+lot: FALSE
 bibliography:
-- citavi_27.05.bib
+- citavi.bib
 - packages.bib
 subtitle: First Report
-tags:
-- nothing
-- nothingness
 biblio-style: apalike
@@
 
-```{r pack tex, eval=FALSE, include=TRUE}
+```{r pack tex, include=FALSE}
 # automatically create a bib database for R packages
@@
 
-The course "Advanced Statistical Programming with R" consists in implementing a package to address location-scale regression. 
+The course "Advanced Statistical Programming with R" consists in implementing a package to address location-scale regression by building upon an R6 class [@R-R6] given to us as part of the `asp20model` repository [@R-asp20model].
 The sample dataset consists of a response $y$, i.e, the expectation is in linear relationship to a set of predictors $x_i$. 
@@
 
-```{r data, echo=FALSE, fig.width=}
+```{r data, echo=FALSE, cache=TRUE, fig.cap = "A visual example of Heteroskedasticity. Source: asp20model"}
 n <- 500
@@
 In *part 2* of this report we explain the concept of boosting and elaborate on how it suites to address the problem of location-scale regression. 
-In *part 3* we describe our implementation milestones, 
-whereby *part 4* especially elaborates on componentwise boosting. 
+In *part 3* we describe our implementation milestones, whereby *part 4* especially elaborates on componentwise boosting. 
 In *part 5* we present our thoughts on upcoming challenges and critical aspects of the package development.
@@
 
-The following descrition of simple boosting algorithm is mainly based on @Fahrmeir.2013, 217-219.
+The following description of simple boosting algorithm is mainly based on @Fahrmeir.2013, 217-219.
 
@@
 However, with estimates and the response vector at hand, residuals can be calculated. 
-The boosting algortihm focuses on those calculated residuals.
+The boosting algorithm focuses on those calculated residuals.
 The effect of the predictors $x_i$ on these residuals is estimated using least squares and the resulting effect is added to the location-estimate $\beta$ adjusting it  by means of the learning rate $\nu$.
@@
 
+\begin{equation}
+u = y - X\hat{\beta}^{(t-1)}
+\end{equation}
+
 ## Scale-Booster
@@
 The term in the estimating step is, however, not the residual.
-For boosting $\gamma$ this term equals the score function of the model i.e  the derivative of the loglikelihood of the normal distribution.
+For boosting $\gamma$ this term equals the score function of the model i.e  the derivative of the log likelihood of the normal distribution.
+
+\begin{equation}
+ll = \frac{1}{2\pi\exp(Z\Gamma)^2}\times\exp(-\frac{u^2}{2\exp(Z\Gamma)^2})
+\end{equation}
 
@@
 which  directly builds upon the `LocationScaleRegression` R6 class of the `asp20model`package set at our disposal for this course and this purpose specifically [@R-asp20model].
-?[@R-R6].
 
@@
 The location parameter is updated. 
-This causes the `LocationScaleRegression` class to calculate updated state-dependent objects, such as the loglikelihood, the gradients and the residuals. 
+This causes the `LocationScaleRegression` class to calculate updated state-dependent objects, such as the log likelihood, the gradients and the residuals. 
 This objects are again used to repeatedly  estimate residuals and update the location parameter $\beta$.
@@
 The simple booster for the scale parameters $\gamma$, as already mentioned, consists in repeating the same two steps of first, estimating a term and then updating the scale parameters $\gamma$ by adding the estimated effect - adjusted by the learning rate $\nu$.
-The estimated term here is the derivative of the loglikelihood of the normal distribution.
+The estimated term here is the derivative of the log likelihood of the normal distribution.
 To calculate this, our code makes use of the `resid-function()`, but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
-This simple implementation of a boosting algorithm for location and scale served the code-basis or "skeleton" for our package and can, hence, be though as the file milestone of this project.
+This simple implementation of a boosting algorithm for location and scale served the code-basis or "skeleton" for our package and can, hence, be thought as the file milestone of this project.
 
@@
 
-A usefull extention of our simple boosting algorithm regards *componentwise boosting*. 
+A useful extension of our simple boosting algorithm regards *componentwise boosting*. 
 The key idea here is to not update a whole parameter vector, but only one entry of it.
 The entry chosen for the update at this juncture is the one yielding the best improvement in terms of minimization of the loss function.
-The loss function in our case is the loglikelihood function of our normally distributed data.
+> Corrected: Check again. The loss function in our case is the deviance calculated using $\gamma$ and our $\beta$.
+
 Hence, in each iteration only one component of the location- and one component of the scale-parameters are updated.
@@
 
-These functions partition the design matrix $X$- respectively $Z$ - into its single columns and then estimate the residuals - respectively the scores - seperately for each component, and determine loss functions for a hypothetical update with the respective component.
+These functions partition the design matrix $X$- respectively $Z$ - into its single columns and then estimate the residuals - respectively the scores - separately for each component, and determine loss functions for a hypothetical update with the respective component.
 Respective to the old loss function value, the highest loss-improvement is determined and the respective component is used to update the parameter-vector.
@@
 
-* Allow user-input of the LocationScaleRegressionBoost-model in form of a dataframe, for example by allowing to pass an optional "data"-argument as known from the lm-call.
+* Allow user-input of the LocationScaleRegressionBoost-model in form of a data frame, for example by allowing to pass an optional "data"-argument as known from the `lm-call`.
 This may reduce proneness to input-errors.

### 2020-06-05 Reports/sample_report/1_REPORT.tex

@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{graphicx}
+\usepackage{graphicx,grffile}
 \makeatletter
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
-  {\par}
 
@@
 \author{Johannes Strauß \and Levin Wiebelt \and Sebastian Aristizabal}
-\date{03 June 2020}
+\date{05 Juni 2020}
 
@@
 run it (It should be installed in your terminal). The\texttt{citavi.bib}
-file needs to be exported from citavi.
+file needs to be exported from \texttt{citavi}.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# automatically create a bib database for R packages}
-\NormalTok{knitr}\OperatorTok{::}\KeywordTok{write\_bib}\NormalTok{(}\KeywordTok{c}\NormalTok{(}
-  \KeywordTok{.packages}\NormalTok{(), }\StringTok{\textquotesingle{}R6\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}knitr\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}rmarkdown\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}asp20model\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}gamboostLSS\textquotesingle{}}\NormalTok{,}\StringTok{\textquotesingle{}mboost\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}tidyverse\textquotesingle{}}
-\NormalTok{), }\StringTok{\textquotesingle{}packages.bib\textquotesingle{}}\NormalTok{)}
+\CommentTok{# automatically create a bib database for R packages}
+\NormalTok{knitr}\OperatorTok{::}\KeywordTok{write_bib}\NormalTok{(}\KeywordTok{c}\NormalTok{(}
+  \KeywordTok{.packages}\NormalTok{(), }\StringTok{'R6'}\NormalTok{, }\StringTok{'knitr'}\NormalTok{, }\StringTok{'rmarkdown'}\NormalTok{, }\StringTok{'asp20model'}\NormalTok{, }\StringTok{'gamboostLSS'}\NormalTok{,}\StringTok{'mboost'}\NormalTok{, }\StringTok{'tidyverse'}
+\NormalTok{), }\StringTok{'packages.bib'}\NormalTok{)}
 \end{Highlighting}
@@
 \texttt{LocationScaleRegressionBoost} R6 class implements a boosting
-algorithm, two functions within the model for calculating the best
+algorithm, two active fields within the model for calculating the best
 fitting covariate with respect to Gamma and Beta and has four new
@@
 \hypertarget{refs}{}
-\begin{cslreferences}
 \leavevmode\hypertarget{ref-R-R6}{}%
@@
 Regression Models}.
-\end{cslreferences}
 

### 2020-06-05 Reports/report_1/1_REPORT.Rmd
added Derivatives Gamma and  Beta
minor updates to report_1

@@
 This ensures positiveness of the variances. 
-The goal approach is to estimate the effects of both: $x_i$ on the expectation of $y$ (location) - this is captured by the estimates $\beta$, and $z_i$ on the variance of $y$ (scale) - this is captured by the estimates $\gamma$.
+The goal is to estimate the effects of both: $x_i$ on the expectation of $y$ (location) - this is captured by the estimates $\beta$, and $z_i$ on the variance of $y$ (scale) - this is captured by the estimates $\gamma$.
 Our student group `asp20boost` solves this task applying the concept of boosting.
@@
 The boosting algorithm focuses on those calculated residuals.
-The effect of the predictors $x_i$ on these residuals is estimated using least squares and the resulting effect is added to the location-estimate $\beta$ adjusting it  by means of the learning rate $\nu$.
+The effect of the predictors $x_i$ on these residuals is estimated using least squares and the resulting effect is added to the location-estimate $\beta$ - adjusted by the learning rate $\nu$. 
 This yields a better fit of the location-model, which results in smaller residuals.
-> These new residuals are estimated again in the next iteration, yielding smaller effect sizes, and hence a convergence of the boosting algorithm towards the OLS-estimate.
+These new residuals are estimated again in the next iteration, yielding smaller effect sizes, and hence a convergence of the boosting algorithm towards the OLS-estimate.
 
@@
 
-To implement a boosting algorithm we constructed the function `gradient_boost` as a part of our R6-Class `LocationScaleRegressionBoost`.
-which  directly builds upon the `LocationScaleRegression` R6 class of the `asp20model`package set at our disposal for this course and this purpose specifically [@R-asp20model].
+To implement a boosting algorithm we constructed the function `gradient_boost` as a part of our R6-Class `LocationScaleRegressionBoost`, which  directly builds upon the `LocationScaleRegression` R6 class of the `asp20model`package set at our disposal for this course and this purpose specifically [@R-asp20model].
 
@@
 This causes the `LocationScaleRegression` class to calculate updated state-dependent objects, such as the log likelihood, the gradients and the residuals. 
-This objects are again used to repeatedly  estimate residuals and update the location parameter $\beta$.
+These objects are again used to repeatedly  estimate residuals and update the location parameter $\beta$.
 
-The simple booster for the scale parameters $\gamma$, as already mentioned, consists in repeating the same two steps of first, estimating a term and then updating the scale parameters $\gamma$ by adding the estimated effect - adjusted by the learning rate $\nu$.
+The simple booster for the scale parameters $\gamma$, as already mentioned, consists in repeating the same two steps of first estimating a term and then updating the scale parameters $\gamma$ by adding the estimated effect - adjusted by the learning rate $\nu$.
 The estimated term here is the derivative of the log likelihood of the normal distribution.
-To calculate this, our code makes use of the `resid-function()`, but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
-This simple implementation of a boosting algorithm for location and scale served the code-basis or "skeleton" for our package and can, hence, be thought as the file milestone of this project.
+To calculate this, our code makes use of the `resid()`-function, but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
+
+This simple implementation of a boosting algorithm for location and scale served the code-basis or "skeleton" for our package and can be thought of as the first milestone of this project.
 
@@
 The entry chosen for the update at this juncture is the one yielding the best improvement in terms of minimization of the loss function.
-> Corrected: Check again. The loss function in our case is the deviance calculated using $\gamma$ and our $\beta$.
+The loss function in our case is the deviance calculated using $\gamma$ and our $\beta$.
 
@@
 
-We implement this in our package by the extending the `LocationScaleRegression` class to include the two active fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.
+### 1. Beta
+
+\begin{equation} 
+\begin{aligned}
+\beta_{j}^{(0)} = 0, \text{ for} j = 1,..,k. \\
+u_i = y - X\hat{\beta}^{(t-1)} \\
+\hat{b}_j = ((\vec{x}^j)^T \space\vec{x}^j)^{-1}((\vec{x}^j)^T \vec{u} \\
+\end{aligned}
+\end{equation}
+
+### 2. Gamma
+
+\begin{equation} 
+\begin{aligned}
+\gamma_{j}^{(0)} = 0, \text{ for} j = 1,..,k. \\
+u_i = \text{deviance} \\
+\hat{b}_j = U^2Z^2e^{-2(Z\Gamma)} \\
+\end{aligned}
+\end{equation}
+
+### 3. Same procedure.
+
+Fit separate models for all covariates i.e obtain
+
+\begin{equation} 
+\hat{b}_j = ((\vec{x}^j)^T \space\vec{x}^j)^{-1}((\vec{x}^j)^T \vec{u}, \text{ for} \space j = 1,..,k. 
+\end{equation}
+
+And determine the best fitting variable via
+
+\begin{equation} 
+j^* = \arg\min_{j = 0,..,k} \sum_{i=1}^{n}(u_i-x_{ij}\hat{b}_j)^2.
+\end{equation}
+
+### 4. Update Coefficients.
+
+\begin{equation} 
+\begin{aligned}
+\hat{\beta}_{j^*}^{(1)}  = \hat{\beta}_{j^*}^{(0)}+ \nu\hat{b}_{j^*} \\
+\hat{\beta}_{j}^{(1)}  = \hat{\beta}_{j}^{(0)}, j\neq j^*.
+\end{aligned}
+\end{equation}
+
+We implement this in our package by extending the `LocationScaleRegression` class to include the two active fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.
 
@@
 
-## Application
-
-[Johannes: Munich Rent Data? - Identify Outliers]
-If you want to cite Kneib to include this copy the following and update the page number:  @Fahrmeir.2013, 217-219.
-
 # 6. References

### 2020-07-22 simulation_studies/simulation_plot_4_params.R
updated simulation studies

@@ -1,0 +1,70 @@
+# Since the alogrithm does not yet work properly for all possible models
+# this script analyses its functionality for 2-dim beta and gamma
+# also scale is regressed on the same predictor as location
+library(R6)
+library(asp20model)
+library(asp20boost)
+
+boost_true_model <- function(beta_0, beta_1, gamma_0, gamma_1){
+
+  # simulate data -------------------------
+  set.seed(1337)
+  n <- 500
+  x <- runif(n)
+  y <- beta_0 + beta_1 * x + rnorm(
+    n,
+    sd = exp(gamma_0 + gamma_1 * x)
+  )
+
+  # estimate data by boosting ------------
+  model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+  gradient_boost(model, stepsize = 0.001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
+
+
+  # plot results -------------------------
+  x_grid <- seq(0,1, length.out = 100)
+  y_grid <- beta_0 + beta_1 * x_grid
+
+  # limits for plot
+  ylim_upper <- model$beta[1] + model$beta[2] * 1 + 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
+  ylim_lower <- model$beta[1] + model$beta[2] * 1 - 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
+
+
+  # plot data points
+  plot(x, y, col = 2,
+       xlim = c(-0.3,1.3),
+       ylim = c(ylim_upper * 1.05, ylim_lower * 1.05))
+
+  # plot true model
+  lines(x_grid, y_grid)
+  lines(x_grid, y_grid + 1.96 * exp(gamma_0 + gamma_1 * x_grid))
+  lines(x_grid, y_grid - 1.96 * exp(gamma_0 + gamma_1 * x_grid))
+
+  #plot estimated effect
+  location_estimate <- model$beta[1] + model$beta[2] * x_grid
+  scale_estimate_sd <- exp(model$gamma[1] + model$gamma[2] * x_grid)
+  lines(x_grid, location_estimate, col = 3)
+  lines(x_grid, location_estimate + 1.96 * scale_estimate_sd, col = 3)
+  lines(x_grid, location_estimate - 1.96 * scale_estimate_sd, col = 3)
+
+
+
+  return(list(
[…]

### 2020-07-22 simulation_studies/simulation_study_hetsked.R

@@ -1,32 +1,0 @@
-library(R6)
-library(asp20model)
-library(asp20boost)
-
-set.seed(1337)
-n <- 500
-x <- matrix(runif(10 * n), ncol = 10) #10 covariates x_1 to x_10
-
-y <- 0 + x[, 1] + x[, 3] + x[, 5] + rnorm(n,
-    sd = 0 + x[, 2] + x[, 3] + x[, 4])
-
-start_time <- Sys.time()
-model <- asp20boost::LocationScaleRegressionBoost$new(
-  y ~ x[, 1] + x[, 2] + x[, 3] + x[, 4] + + x[, 5], #location
-  ~ x[, 1] + x[, 2] + x[, 3] + x[, 4] + + x[, 5] #scale
-)
-asp20boost::gradient_boost(model, stepsize = 0.0001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
-end_time <- Sys.time()
-
-
-
-
-# True Coefficients Location: 0, 1, 0, 1, 0, 1
-# Identified Coefficients Location:
-model$beta
-
-# True Coefficients Scale: 0, 0, 1, 1, 1, 0
-# Identified Coefficients Scale:
-model$gamma
-
-#Time Taken
-end_time - start_time

### 2020-07-22 simulation_studies/simulation_study_homosked.R

@@ -1,26 +1,0 @@
-library(R6)
-library(asp20model)
-library(asp20boost)
-
-set.seed(1337)
-n <- 500
-x <- matrix(runif(10 * n), ncol = 10) #10 covariates x_1 to x_10
-
-y <- 0 + x[, 1] + x[, 3] + x[, 5] + rnorm(n, sd = 1)
-# only half of them influence response
-
-start_time <- Sys.time()
-model <- asp20boost::LocationScaleRegressionBoost$new(
-  y ~ x[, 1] + x[, 2] + x[, 3] + x[, 4] + + x[, 5],
-  ~ 1 #homogenous scale
-  )
-asp20boost::gradient_boost(model, stepsize = 0.0001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
-end_time <- Sys.time()
-
-
-# True Coefficients Location: 0, 1, 0, 1, 0, 1
-# Identified Coefficients Location:
-model$beta
-
-#Time Taken
-end_time - start_time

### 2020-07-22 simulation_studies/simulation_study_simple.R

@@ -1,26 +1,0 @@
-library(R6)
-library(asp20model)
-library(asp20boost)
-
-set.seed(1337)
-n <- 500
-x <- runif(n)
-y <- x + rnorm(n, sd = exp(-3 + 2 * x))
-
-start_time <- Sys.time()
-model <- asp20boost::LocationScaleRegressionBoost$new(y ~ x, ~ x)
-asp20boost::gradient_boost(model, stepsize = 0.001, maxit = 2000, abstol = 0.0001, verbose = FALSE)
-end_time <- Sys.time()
-
-# True Coefficients Location: 0, 1
-# Identified Coefficients Location:
-model$beta[1]
-model$beta[2]
-
-# True Coefficients Scale: -3, 2
-# Identified Coefficients Scale:
-model$gamma[1]
-model$gamma[2]
-
-#Time Taken
-end_time - start_time

### 2020-07-22 simulation_studies/simulation_variable_selection.R

@@ -1,0 +1,66 @@
+# This script analyses if variable selection in the case of multidemnsional
+# beta and gamma vectors works
+# For this reason some coefficients are 0.
+library(R6)
+library(asp20model)
+library(asp20boost)
+
+boost_true_model <- function(beta_vec, gamma_vec){
+
+  # simulate data -------------------------
+  set.seed(1337)
+  n <- 500
+  X <- matrix(runif(n*5), ncol = 5)
+  y <- beta_vec[1] + beta_vec[2] * X[, 1] + beta_vec[3] * X[, 2] +  beta_vec[4] * X[, 3] + beta_vec[5] * X[, 4] + beta_vec[6] * X[, 5] + rnorm(
+    n,
+    sd = exp(gamma_vec[1] + gamma_vec[2] * X[, 1] + gamma_vec[3] * X[, 2] +  gamma_vec[4] * X[, 3] + gamma_vec[5] * X[, 4] + gamma_vec[6] * X[, 5])
+  )
+
+  # estimate data by boosting ------------
+  model <- asp20boost::LocationScaleRegressionBoost$new(
+    y ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5],
+    ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5])
+  asp20boost::gradient_boost(model, stepsize = 0.001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
+
+
+  return(list(
+    beta = model$beta,
+    gamma = model$gamma)
+  )
+}
+
+
+
+
+
+
+
+
+
+# beta_1 <- c(1,1,0,0,0,0)
+# gamma_1 <- c(1,1,0,0,0,0)
+# boost_true_model(beta_1, gamma_1)
+
+# beta_2 <- c(0,1,0,1,0,1)
+# gamma_2 <- c(0,0,1,0,1,0)
+# boost_true_model(beta_2, gamma_2)
+
+# beta_3 <- c(1,1,0,1,0,1)
+# gamma_3 <- c(1,0,1,0,1,0)
+# boost_true_model(beta_3, gamma_3)
+
+# homoskedastic model
+# beta_4 <- c(1,1,0,1,0,1)
+# gamma_4 <- c(1,0,0,0,0,0)
+# boost_true_model(beta_4, gamma_4)
+
+# intercept model
[…]

### 2020-07-22 tests/testthat/helper-boosting.R
added automated unit tests

@@ -1,0 +1,9 @@
+set.seed(1337)
+
+n <- 500
+x1 <- runif(n)
+x2 <- runif(n)
+x3 <- runif(n)
+y <- rnorm(n, x1 + x3, exp(-3 + x2 + x3))
+model <- LocationScaleRegressionBoost$new(y ~ x1 + x3, ~ x2 + x3)
+gradient_boost(model, stepsize = 0.001, maxit = 10000, abstol = 0.0001, verbose = FALSE)

### 2020-07-22 tests/testthat/helper_model_class_creation.R

@@ -1,0 +1,21 @@
+set.seed(1337)
+
+
+# create a model class - simplest case
+n <- 500
+x <- runif(n)
+z <- runif(n)
+y <- x + rnorm(n, sd = exp(z))
+model_simple <- LocationScaleRegressionBoost$new(y ~ x, ~ z)
+
+
+# create a model class - large model
+n <- 500
+x <- matrix(runif(5 * n), ncol = 5) # 5 location-covariates x_1 to x_10
+z <- matrix(runif(5 * n), ncol = 5) # 5 scale-covariates z_1 to z_10
+# formulate the true model - dependent on only some covariates
+y <- 21 + 22 * x[, 1] + 22 * x[, 3] + 23 * x[, 5] + rnorm(n,
+                                                          sd = 31 + 32 * z[, 2] + 33 * z[, 3] + 34 * z[, 4])
+model_large <- LocationScaleRegressionBoost$new(
+  y ~  x[, 1] + x[, 2] + x[, 3] + x[, 4] + x[, 5],
+  ~ z[, 1] + z[, 2] + z[, 3] + z[, 4] + z[, 5])

### 2020-07-22 tests/testthat/test_boosting_alg.R

@@ -1,0 +1,22 @@
+# context("functionality boosting algorithm")
+#
+# source("helper_model_class_creation.R")
+# # creates one simple and one large model
+#
+# # for ad-hoc loading use: source("tests/testthat/helper_model_class_creation.R")
+#
+#
+# # estimate models
+# gradient_boost(model_simple, stepsize = 0.01, maxit = 10000, abstol = 0.0001, verbose = TRUE)
+# gradient_boost(model_large, stepsize = 0.001, maxit = 2000, abstol = 0.0001, verbose = FALSE)
+#
+#
+# test_that("beta estimates are correct in simple model",{
+#   model_1 <- model_simple$clone()
+#   expect_equal(model_1$beta, c(0, 1))
+# })
+#
+# test_that("gamma estimates are correct in simple model",{
+#   model_1 <- model_simple$clone()
+#   expect_equal(model_1$gamma, c(0, 1))
+# })

### 2020-07-22 tests/testthat/test_class_creation.R

@@ -1,0 +1,35 @@
+context("model class creation")
+
+source("helper_model_class_creation.R")
+# for ad-hoc loading use: source("tests/testthat/helper_model_class_creation.R")
+#creates one simple model and one large model
+
+
+test_that("inheritance works", {
+  model_1 <- model_simple$clone()
+  model_2 <- model_large$clone()
+  expect_is(model_1, "LocationScaleRegression")
+  expect_is(model_1, "LocationScaleRegressionBoost")
+  expect_is(model_2, "LocationScaleRegression")
+  expect_is(model_2, "LocationScaleRegressionBoost")
+})
+
+
+test_that("beta vector initialized correctly", {
+  model_1 <- model_simple$clone()
+  model_2 <- model_large$clone()
+  expect_equal(length(model_1$beta), 2)
+  expect_equal(model_1$beta, c(0,0 ))
+  expect_equal(length(model_2$beta), 6)
+  expect_equal(model_2$beta, c(0,0,0,0,0,0))
+})
+
+
+test_that("gamma vector initialized correctly", {
+  model_1 <- model_simple$clone()
+  model_2 <- model_large$clone()
+  expect_equal(length(model_1$gamma), 2)
+  expect_equal(model_1$gamma, c(0,0 ))
+  expect_equal(length(model_2$gamma), 6)
+  expect_equal(model_2$gamma, c(0,0,0,0,0,0))
+})

### 2020-07-22 tests/testthat/test_gradients.R

@@ -1,3 +1,5 @@
+context("gradients")
+
 library(numDeriv)
 
 set.seed(1337)
@@ -29,9 +31,3 @@ f <- function(x) {
 test_that("gamma gradient works", {
   expect_equivalent(model$grad_gamma(), grad(f, model$gamma))
 })
-
-
-#test_that("gradient boost does not throw an error", {
-#  expect_error(gradient_boost(model, verbose = TRUE), NA)
-#})
-

### 2020-07-22 simulation_studies/not_working_simulation_plot_4_params.R
updated simulation studies

@@ -1,0 +1,83 @@
+# Since the alogrithm does not yet work properly for all possible models
+# this script analyses its functionality for 2-dim beta and gamma
+# also scale is regressed on the same predictor as location
+library(R6)
+library(asp20model)
+library(asp20boost)
+
+
+# in order to make this script work an extra step needs to be done:
+# the error is as follow: model creation inside the "boost_true_model"-function
+# below does not work, UNLESS a LocationScaleRegressionBoost-model class has been created before
+# for this reason a simple empty model call and estimation is performed here
+{ x <- runif(100)
+  y <- y <- x + rnorm(100, sd = exp(-3 + 2 * x))
+  model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+  gradient_boost(model, stepsize = 0.001, maxit = 1000, abstol = 0.0001, verbose = FALSE)
+}
+
+
+
+boost_true_model <- function(beta_0, beta_1, gamma_0, gamma_1){
+
+  # simulate data -------------------------
+  set.seed(1337)
+  n <- 500
+  x <- runif(n)
+  y <- beta_0 + beta_1 * x + rnorm(
+    n,
+    sd = exp(gamma_0 + gamma_1 * x)
+  )
+
+  # estimate data by boosting ------------
+  model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+  gradient_boost(model, stepsize = 0.001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
+
+
+  # plot results -------------------------
+  x_grid <- seq(0,1, length.out = 100)
+  y_grid <- beta_0 + beta_1 * x_grid
+
+  # limits for plot
+  ylim_upper <- model$beta[1] + model$beta[2] * 1 + 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
+  ylim_lower <- model$beta[1] + model$beta[2] * 1 - 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
+
+
+  # plot data points
+  plot(x, y, col = 2,
+       xlim = c(-0.3,1.3),
+       ylim = c(ylim_upper * 1.05, ylim_lower * 1.05))
+
+  # plot true model
+  lines(x_grid, y_grid)
[…]

### 2020-07-22 simulation_studies/now_working_simulation_variable_selection.R

@@ -1,0 +1,80 @@
+# This script analyses if variable selection in the case of multidemnsional
+# beta and gamma vectors works
+# For this reason some coefficients are 0.
+library(R6)
+library(asp20model)
+library(asp20boost)
+
+
+# in order to make this script work an extra step needs to be done:
+# the error is as follow: model creation inside the "boost_true_model"-function
+# below does not work, UNLESS a LocationScaleRegressionBoost-model class has been created before
+# for this reason asimple model call and estimation is performed here
+{ x <- runif(100)
+  y <- runif(100)
+  model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+  gradient_boost(model, stepsize = 0.001, maxit = 1000, abstol = 0.0001, verbose = FALSE)
+}
+
+
+
+
+boost_true_model <- function(beta_vec, gamma_vec){
+
+  # simulate data -------------------------
+  set.seed(1337)
+  n <- 500
+  X <- matrix(runif(n*5), ncol = 5)
+  y <- beta_vec[1] + beta_vec[2] * X[, 1] + beta_vec[3] * X[, 2] +  beta_vec[4] * X[, 3] + beta_vec[5] * X[, 4] + beta_vec[6] * X[, 5] + rnorm(
+    n,
+    sd = exp(gamma_vec[1] + gamma_vec[2] * X[, 1] + gamma_vec[3] * X[, 2] +  gamma_vec[4] * X[, 3] + gamma_vec[5] * X[, 4] + gamma_vec[6] * X[, 5])
+  )
+
+  # estimate data by boosting ------------
+  model <- asp20boost::LocationScaleRegressionBoost$new(
+    y ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5],
+    ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5])
+  asp20boost::gradient_boost(model, stepsize = 0.001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
+
+
+  return(list(
+    beta = model$beta,
+    gamma = model$gamma)
+  )
+}
+
+
+
+
+
+
+
+
+
+# beta_1 <- c(1,1,0,0,0,0)
+# gamma_1 <- c(1,1,0,0,0,0)
[…]

### 2020-07-22 simulation_studies/simulation_plot_4_params.R

@@ -1,70 +1,52 @@
-# Since the alogrithm does not yet work properly for all possible models
-# this script analyses its functionality for 2-dim beta and gamma
-# also scale is regressed on the same predictor as location
-library(R6)
-library(asp20model)
 library(asp20boost)
 
-boost_true_model <- function(beta_0, beta_1, gamma_0, gamma_1){
-
-  # simulate data -------------------------
-  set.seed(1337)
-  n <- 500
-  x <- runif(n)
-  y <- beta_0 + beta_1 * x + rnorm(
-    n,
-    sd = exp(gamma_0 + gamma_1 * x)
-  )
-
-  # estimate data by boosting ------------
-  model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
-  gradient_boost(model, stepsize = 0.001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
-
-
-  # plot results -------------------------
-  x_grid <- seq(0,1, length.out = 100)
-  y_grid <- beta_0 + beta_1 * x_grid
-
-  # limits for plot
-  ylim_upper <- model$beta[1] + model$beta[2] * 1 + 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
-  ylim_lower <- model$beta[1] + model$beta[2] * 1 - 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
-
-
-  # plot data points
-  plot(x, y, col = 2,
-       xlim = c(-0.3,1.3),
-       ylim = c(ylim_upper * 1.05, ylim_lower * 1.05))
-
-  # plot true model
-  lines(x_grid, y_grid)
-  lines(x_grid, y_grid + 1.96 * exp(gamma_0 + gamma_1 * x_grid))
-  lines(x_grid, y_grid - 1.96 * exp(gamma_0 + gamma_1 * x_grid))
-
-  #plot estimated effect
-  location_estimate <- model$beta[1] + model$beta[2] * x_grid
-  scale_estimate_sd <- exp(model$gamma[1] + model$gamma[2] * x_grid)
-  lines(x_grid, location_estimate, col = 3)
-  lines(x_grid, location_estimate + 1.96 * scale_estimate_sd, col = 3)
-  lines(x_grid, location_estimate - 1.96 * scale_estimate_sd, col = 3)
[…]

### 2020-07-22 simulation_studies/simulation_variable_selection.R

@@ -5,62 +5,45 @@ library(R6)
 library(asp20model)
 library(asp20boost)
 
-boost_true_model <- function(beta_vec, gamma_vec){
+beta_vec <- c(1,0,1,0,1,0)
+gamma_vec <- c(0,1,0,1,0,1)
 
-  # simulate data -------------------------
-  set.seed(1337)
-  n <- 500
-  X <- matrix(runif(n*5), ncol = 5)
-  y <- beta_vec[1] + beta_vec[2] * X[, 1] + beta_vec[3] * X[, 2] +  beta_vec[4] * X[, 3] + beta_vec[5] * X[, 4] + beta_vec[6] * X[, 5] + rnorm(
-    n,
-    sd = exp(gamma_vec[1] + gamma_vec[2] * X[, 1] + gamma_vec[3] * X[, 2] +  gamma_vec[4] * X[, 3] + gamma_vec[5] * X[, 4] + gamma_vec[6] * X[, 5])
-  )
-
-  # estimate data by boosting ------------
-  model <- asp20boost::LocationScaleRegressionBoost$new(
-    y ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5],
-    ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5])
-  asp20boost::gradient_boost(model, stepsize = 0.001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
-
-
-  return(list(
-    beta = model$beta,
-    gamma = model$gamma)
-  )
-}
-
-
-
-
-
-
-
-
-
-# beta_1 <- c(1,1,0,0,0,0)
-# gamma_1 <- c(1,1,0,0,0,0)
-# boost_true_model(beta_1, gamma_1)
 
 # beta_2 <- c(0,1,0,1,0,1)
 # gamma_2 <- c(0,0,1,0,1,0)
-# boost_true_model(beta_2, gamma_2)
 
 # beta_3 <- c(1,1,0,1,0,1)
 # gamma_3 <- c(1,0,1,0,1,0)
-# boost_true_model(beta_3, gamma_3)
 
 # homoskedastic model
 # beta_4 <- c(1,1,0,1,0,1)
 # gamma_4 <- c(1,0,0,0,0,0)
-# boost_true_model(beta_4, gamma_4)
 
-# intercept model
-beta_5 <- c(1,0,0,0,0,0)
[…]

### 2020-07-24 Reports/report_2/2_REPORT.Rmd
Report 2 rmd. Skeleton and bib.tex libraries.

--- 2020-06-05 Reports/report_1/1_REPORT.Rmd
@@
 ---
-title: "Report ASP20 Boost"
+title: "ASP20 Boost"
 author:
@@
 - packages.bib
-subtitle: First Report
+subtitle: Second Report
 biblio-style: apalike
@@
 ```
-
-# 1. Introduction:
-
-The course "Advanced Statistical Programming with R" consists in implementing a package to address location-scale regression by building upon an R6 class [@R-R6] given to us as part of the `asp20model` repository [@R-asp20model].
-The sample dataset consists of a response $y$, i.e, the expectation is in linear relationship to a set of predictors $x_i$. 
-Its variance, in contrast, is dependent on a different set of linear predictors $z_i$, however transformed by the response function $g(x) = exp(x)$. 
-This ensures positiveness of the variances. 
-The goal approach is to estimate the effects of both: $x_i$ on the expectation of $y$ (location) - this is captured by the estimates $\beta$, and $z_i$ on the variance of $y$ (scale) - this is captured by the estimates $\gamma$.
-Our student group `asp20boost` solves this task applying the concept of boosting.
-
-```{r data, echo=FALSE, cache=TRUE, fig.cap = "A visual example of Heteroskedasticity. Source: asp20model"}
-n <- 500
-x <- runif(n)
-y <- x + rnorm(n, sd = exp(-3 + 2 * x))
-plot(x, y)
-abline(0, 1, lwd = 2)
-curve(x + 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
-curve(x - 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
-```
-
-In *part 2* of this report we explain the concept of boosting and elaborate on how it suites to address the problem of location-scale regression. 
-In *part 3* we describe our implementation milestones, whereby *part 4* especially elaborates on componentwise boosting. 
-In *part 5* we present our thoughts on upcoming challenges and critical aspects of the package development.
-
-# 2. Concept: Boosting
-
-The first step in our progress was to understand the theoretical concept of boosting and which types of problems it is able to solve.
-To explain the concept and follow our process of understanding we first explain boosting for location parameters. 
-Boosting scale parameters required another way of thinking about it and constituted a first milestone in our understanding and implementation.
-In the following, the term "booster" refers to a boosting algorithm.
-
-## Location-Booster
-
-The following description of simple boosting algorithm is mainly based on @Fahrmeir.2013, 217-219.
-
-The boosting concept relies on the idea of iterative estimation. 
-To estimate the location parameters $\beta$ via boosting, one starts with initial estimates, which may be far from optimal. 
-However, with estimates and the response vector at hand, residuals can be calculated. 
-The boosting algorithm focuses on those calculated residuals.
-The effect of the predictors $x_i$ on these residuals is estimated using least squares and the resulting effect is added to the location-estimate $\beta$ adjusting it  by means of the learning rate $\nu$.
-This yields a better fit of the location-model, which results in smaller residuals.
-> These new residuals are estimated again in the next iteration, yielding smaller effect sizes, and hence a convergence of the boosting algorithm towards the OLS-estimate.
-
-A simple booster stops after a fixed number of iterations is reached.
-
-\begin{equation}
-u = y - X\hat{\beta}^{(t-1)}
-\end{equation}
-
-## Scale-Booster
-
-A boosting algorithm for scale parameters $\gamma$ consists in the same two repeating steps of first estimating a term, and second updating $\gamma$ by adding the estimated effect - adjusted by a learning rate.
-The term in the estimating step is, however, not the residual.
-For boosting $\gamma$ this term equals the score function of the model i.e  the derivative of the log likelihood of the normal distribution.
-
-\begin{equation}
-ll = \frac{1}{2\pi\exp(Z\Gamma)^2}\times\exp(-\frac{u^2}{2\exp(Z\Gamma)^2})
-\end{equation}
-
-# 3. Implementation
-
-To implement a boosting algorithm we constructed the function `gradient_boost` as a part of our R6-Class `LocationScaleRegressionBoost`.
-which  directly builds upon the `LocationScaleRegression` R6 class of the `asp20model`package set at our disposal for this course and this purpose specifically [@R-asp20model].
-
-In the current version of the `asp20boost`-package, the boosting algorithm is already implemented in order to allow for componentwise boosting, which will be explained in part 4.
-
-To further provide clarity, accountability and understanding during to our development process we describe briefly our first implementation of an ordinary boosting algorithm.
-This isn't visible in our code anymore.
-
-A simple booster for the location parameters $\beta$ works without extending the `asp20model`'s `LocationScaleRegression` class. 
-The current residuals may be extracted with the `resid()` command and then estimated.
-The location parameter is updated. 
-This causes the `LocationScaleRegression` class to calculate updated state-dependent objects, such as the log likelihood, the gradients and the residuals. 
-This objects are again used to repeatedly  estimate residuals and update the location parameter $\beta$.
-
-The simple booster for the scale parameters $\gamma$, as already mentioned, consists in repeating the same two steps of first, estimating a term and then updating the scale parameters $\gamma$ by adding the estimated effect - adjusted by the learning rate $\nu$.
-The estimated term here is the derivative of the log likelihood of the normal distribution.
-To calculate this, our code makes use of the `resid-function()`, but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
-This simple implementation of a boosting algorithm for location and scale served the code-basis or "skeleton" for our package and can, hence, be thought as the file milestone of this project.
-
-# 4. Functionality: Componentwise Boosting
-
-A useful extension of our simple boosting algorithm regards *componentwise boosting*. 
-The key idea here is to not update a whole parameter vector, but only one entry of it.
-The entry chosen for the update at this juncture is the one yielding the best improvement in terms of minimization of the loss function.
-> Corrected: Check again. The loss function in our case is the deviance calculated using $\gamma$ and our $\beta$.
-
-Hence, in each iteration only one component of the location- and one component of the scale-parameters are updated.
-
-We implement this in our package by the extending the `LocationScaleRegression` class to include the two active fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.
-
-These functions partition the design matrix $X$- respectively $Z$ - into its single columns and then estimate the residuals - respectively the scores - separately for each component, and determine loss functions for a hypothetical update with the respective component.
-Respective to the old loss function value, the highest loss-improvement is determined and the respective component is used to update the parameter-vector.
-
-We are in the process of reconsidering the design of this implementation.
-Other design possibilities are the following:
-
-* Create public fields for the heavily used score-function values.
-* Move the calculation of componentwise losses to an external function.
-* Harmonize boosting and componentwise boosting into one external function, determining the mode of operation by an argument `componentwise = TRUE`.
-
-Another conceptual questions that comes up is if the best loss-improvement may be indicated by the already existing gradients, and hence there is no need for extra calculation.
-
-# 5. Prospects
-
-## Further Functionalities
-
-The most important functionality we intend to implement in our package is the optimization of the number of iterations via cross validation. 
-To work out the theoretical concept behind this idea is, as well as to implementing it are our current next goals.
-One further, rather loose, idea is to find ways to optimize the learning rate in our boosting algorithm.
-
-## Stability
-
-We intend to implement further automated unit tests. 
-This will enable us to assess quickly the stability of our code given differing inputs.
-We also intend to implement measures for proneness reduction to input errors.
-
-## Performance
-
-* A major performance problem in our code is that small learning rates for the $\beta$-booster lead to a  sharp increase of our code's processing time. This remains to be solved.
-* A good choice of starting values for $\beta$ and $\gamma$ may enhance performance. A possible candidate is the mean of the response.
-
-## Further Aspects
-
-* Allow user-input of the LocationScaleRegressionBoost-model in form of a data frame, for example by allowing to pass an optional "data"-argument as known from the `lm-call`.
-This may reduce proneness to input-errors.
-
-* Visualize the results of the boosted estimates
-
-* Document the extension of the R6-Class properly and inherit documentation of the `LocationScaleRegression-Class` using roxygen2.
-
-## Application
-
-[Johannes: Munich Rent Data? - Identify Outliers]
-If you want to cite Kneib to include this copy the following and update the page number:  @Fahrmeir.2013, 217-219.
-
-# 6. References

### 2020-07-24 Reports/report_2/2_REPORT.Rmd
Levin's trasnscript. Spell Check. Formatting. First knitt.

@@
 ```
+
+# 1. Introduction
+
+Following up on the first report from June 5th 2020, this document describes the further progress of developing the R-package asp20boost.
+In part 2, we describe the changes in design and new features since completion of the first report, these include amongst others automated unit tests and the improvement of matrix calculations via Cholesky-decomposition.
+To assess the performance of the gradient-boost algorithm, we conducted several simulation studies.
+The results are described in part 3. Finally, in part 4, we outline which challenges remain.
+
+# 2. Code Updates
+
+## Cholesky Decomposition
+
+Following the issue raised by Hannes Riebl in our Gitlab-Repo, we implemented a Cholesky decomposition in the matrix calculations of the `LocationScaleRegressionBoost` class.
+As described in the first report the idea of boosting consists in the iterative estimation of
+* Residuals in the case of location-estimation
+* The gradient (wrt. scale parameters) of the loss-function in the case of scale estimation.
+
+In both cases this is done by the least squares procedure [insert formula: (X’X)^(-1) * X‘ * residuals_vector]. 
+If repeatedly executed, this calculation may be improved by decomposing the symmetric positive definite matrix X’X via Cholesky's procedure.
+The inverse then is calculated rapidly by the R-command `chol2inv()`.
+
+## Automated Unit Tests
+
+To automate the testing process of our package we began with the implementation of unit tests.
+Up to now, we automated the check for correct model class creation.
+In order to test the boosting algorithm with simulated data, external R-scripts seem more adequate.
+
+## API Design
+
+Componentwise boosting demands selection of the component, which is to be updated.
+Both this selection and the ensuing calculation of the update are done by the fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.
+We decided to move these calculations from active fields to public fields of the `LocationScaleRegressionBoost` class.
+This is motivated by the fact that a substantial amount of calculations happen in these fields, while active fields should be of low computational cost.
+Hence public fields are more accurate.
+
+# 3. Simulation Studies
+
+In order to assess the functionality of the `gradient_boost` function, we conducted several simulation studies which may be categorized into two sets.
+The first set is implemented in the file *simulation_plot_4_params.R*.
+It sets up a simple location-scale model with two parameters beta for location, as well as two parameters gamma for scale.
+The algorithm does not yet work properly for all possible parameter combinations of beta and gamma.
+The second set is implemented in the file *simulation_variable_selection.R*.
+It aims to check if variable selection works correctly for multidimensional (dim > 2) beta or gamma.
+Results also depend on the concrete parameter configuration.
+
+## Two-dimensional Case
+
+Setting up the model identically to the example in the *asp20model-Repo* lead to algorithm convergence towards the true values beta = (0, 1)‘ and gamma = (-3, 2)‘.
+
+Let the true model be characterized by the following parameter configuration: beta = (1, 1)‘ and gamma = (1, 1)‘. 
+After three iterations, the algorithm yields estimates of beta_hat = (0.01, 0.05)‘ and gamma_hat = (12.30, 0.00)‘.
+The algorithm detects no further change in gradients and stops prematurely.
+By lowering the `abstol` parameter functionality may be restored.
+After 3412 iterations estimates are beta_hat = (0.80, 1.44)‘ and gamma_hat = (0.96, 1.12)‘.
+These values are not yet good estimates but reasonably close to conclude basic functionality of the boosting algorithm.
+
+In the case of parameter configuration : beta = (2, 2)‘ and gamma = (2, 2)‘ early stopping remains as a problem. 
+Functionality can not be achieved by lowering the tolerance threshold `abstol` of the break condition.
+The break condition is a comparison of gradients of successive iterations.
+
+## Multidimensional Case
+
+In the simulation studies for multidimensional models, a total of six parameters for beta as well as six for gamma may be specified.
+
+### Simple Case
+
+A simple case consists in setting intercept and one slope for beta and gamma to one while the remaining 2x4 parameters are zero: beta = (1, 1, 0, 0, 0, 0)‘ and gamma = (1, 1, 0, 0, 0, 0)‘.
+The algorithm should depict the 2x2 parameters of interest and estimate them close to one while the others should remain zero.
+The resulting estimates displayed below indicate proper estimates for the scale parameter vector `gamma_hat` but not for the location estimates `beta_hat`.
+```
+model$beta
+[1]  0.15122097  1.72916260  0.61363927 -0.14620695  0.02756381  0.28560006
+> model$gamma
+[1]  0.95706770  1.15196564 -0.02756374  0.00000000 -0.06839582  0.00000000
+```
+```
+model$beta
+[1]  0.15122097  1.72916260  0.61363927 -0.14620695  0.02756381  0.28560006
+model$gamma
+[1]  0.95706770  1.15196564 -0.02756374  0.00000000 -0.06839582  0.00000000
+```
+### Example Case
+
+Setting up the example model from the asp20model repo, the algorithm, in turn, finds correct estimates – however being somewhat inaccurate for gamma_1_hat:
+```
+model$beta
+[1] -0.0148539032  1.0130172152  0.0108580428 -0.0026268607  0.0001592711
+[6]  0.0049089286
+model$gamma
+[1] -3.103392229  1.258857920  0.017474511 -0.005902031 -0.092450456
+[6]  0.000000000
+```
+### 1-0-Shifted Case
+
+Another model setup beta = (1,0,1,0,1,0)‘ and gamma = (0,1,0,1,0,1)‘ yields again correct estimates for gamma, but not for beta:
+
+```
+model$beta
+[1]  0.69898406  0.53163733  1.64776999 -0.92116479  1.10914562 -0.02877867
+model$gamma
+[1]  0.01368480  1.17157405 -0.03057068  0.94892465 -0.03256148  0.89579817
+```
+
+## Conclusions of the simulation studies
+
+From these findings we conclude that the evolution of gradients during iterations needs to be checked.
+We need to find out why gradients converge without yielding sufficiently adequate parameter estimates.
+At the moment component-wise boosting is hardwired into our code.
+A useful implementation would be to allow for non-componentwise boosting again.
+Being able to switch between overall and componentwise boosting may facilitate locating weak points in our code.
+
+# 4. Further Tasks
+
+## Ensure functionality of gradient-boost algorithm
+
+As depicted in part 3, functionality of our algorithm needs to be ensured independent of the concrete data-generating process.
+
+## Choice of initial parameter values
+
+The performance of the boosting algorithm may be enhanced if initial estimates are chosen sensibly.
+For the location parameters beta this may be the OLS-estimate or the mean response. For the scale parameters gamma this may be the (homoskedastic) sample variance.
+Another possibility is to ask other student groups for straight-forward and fast numerical procedures to determine good starting values.
+However, this may be costly in terms of computation time and hence slow down overall performance of the algorithm.
+Trade-offs need to be considered.

### 2020-07-24 Reports/report_2/2_REPORT.Rmd
Report 2 Final Knit: - Latex - Citation - Form, appereance. - Spelling. Removed irrelevant question file.

@@
 
-Following up on the first report from June 5th 2020, this document describes the further progress of developing the R-package asp20boost.
+Following up on the first report from June 5th 2020, this document describes the further progress of developing the R-package `asp20boost`.
 In part 2, we describe the changes in design and new features since completion of the first report, these include amongst others automated unit tests and the improvement of matrix calculations via Cholesky-decomposition.
 To assess the performance of the gradient-boost algorithm, we conducted several simulation studies.
-The results are described in part 3. Finally, in part 4, we outline which challenges remain.
+The results are described in part 3. Finally, in part 4, we outline remaining challenges.
 
@@
 
-## Cholesky Decomposition
+## Cholesky decomposition
 
-Following the issue raised by Hannes Riebl in our Gitlab-Repo, we implemented a Cholesky decomposition in the matrix calculations of the `LocationScaleRegressionBoost` class.
+Following the issue raised by Hannes Riebl in our Gitlab-repository, we implemented a Cholesky decomposition in the matrix calculations of the `LocationScaleRegressionBoost` class.
 As described in the first report the idea of boosting consists in the iterative estimation of
+
 * Residuals in the case of location-estimation
@@
 
-In both cases this is done by the least squares procedure [insert formula: (X’X)^(-1) * X‘ * residuals_vector]. 
-If repeatedly executed, this calculation may be improved by decomposing the symmetric positive definite matrix X’X via Cholesky's procedure.
+In both cases this is done by the least squares procedure $(X'X)^{-1} X'u$ [@Fahrmeir.2013, 226]. 
+If repeatedly executed, this calculation may be improved by decomposing the symmetric positive definite matrix $X'X$ via Cholesky's procedure.
 The inverse then is calculated rapidly by the R-command `chol2inv()`.
@@
 
-## API Design
+## API design
 
 Componentwise boosting demands selection of the component, which is to be updated.
-Both this selection and the ensuing calculation of the update are done by the fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.
+Both this selection and the ensuing calculation of the update are done by the fields `bestFittingVariablebeta` and `bestFittingVariableGamma`.
 We decided to move these calculations from active fields to public fields of the `LocationScaleRegressionBoost` class.
@@
 The first set is implemented in the file *simulation_plot_4_params.R*.
-It sets up a simple location-scale model with two parameters beta for location, as well as two parameters gamma for scale.
-The algorithm does not yet work properly for all possible parameter combinations of beta and gamma.
+It sets up a simple location-scale model with two parameters $\beta$ for location, as well as two parameters $\gamma$ for scale.
+The algorithm does not yet work properly for all possible parameter combinations of $\beta$ and $\gamma$.
 The second set is implemented in the file *simulation_variable_selection.R*.
-It aims to check if variable selection works correctly for multidimensional (dim > 2) beta or gamma.
+It aims to check if variable selection works correctly for multidimensional (dim $>$ 2) $\beta$ or $\gamma$.
 Results also depend on the concrete parameter configuration.
 
-## Two-dimensional Case
+## Two-dimensional case
 
-Setting up the model identically to the example in the *asp20model-Repo* lead to algorithm convergence towards the true values beta = (0, 1)‘ and gamma = (-3, 2)‘.
+Setting up the model identically to the example in the `asp20model` repository leads to algorithm convergence towards the true values $\beta = (0, 1)'$ and $\gamma = (-3, 2)'$ [@R-asp20model].
 
-Let the true model be characterized by the following parameter configuration: beta = (1, 1)‘ and gamma = (1, 1)‘. 
-After three iterations, the algorithm yields estimates of beta_hat = (0.01, 0.05)‘ and gamma_hat = (12.30, 0.00)‘.
+Let the true model be characterized by the following parameter configuration: $\beta = (1, 1)'$ and $\gamma = (1, 1)'$. 
+After three iterations, the algorithm yields estimates of $\hat{\beta} = (0.01, 0.05)'$ and $\hat{\gamma} = (12.30, 0.00)'$.
 The algorithm detects no further change in gradients and stops prematurely.
 By lowering the `abstol` parameter functionality may be restored.
-After 3412 iterations estimates are beta_hat = (0.80, 1.44)‘ and gamma_hat = (0.96, 1.12)‘.
+After 3412 iterations estimates are $\hat{\beta} = (0.80, 1.44)'$ and $\hat{\gamma} = (0.96, 1.12)'$.
 These values are not yet good estimates but reasonably close to conclude basic functionality of the boosting algorithm.
 
-In the case of parameter configuration : beta = (2, 2)‘ and gamma = (2, 2)‘ early stopping remains as a problem. 
+In the case of parameter configuration : $\beta = (2, 2)'$ and $\gamma = (2, 2)'$ early stopping remains as a problem. 
 Functionality can not be achieved by lowering the tolerance threshold `abstol` of the break condition.
@@
 
-## Multidimensional Case
+## Multidimensional case
 
-In the simulation studies for multidimensional models, a total of six parameters for beta as well as six for gamma may be specified.
+In the simulation studies for multidimensional models, a total of six parameters for $\beta$ as well as six for $\gamma$ may be specified.
 
@@
 
-A simple case consists in setting intercept and one slope for beta and gamma to one while the remaining 2x4 parameters are zero: beta = (1, 1, 0, 0, 0, 0)‘ and gamma = (1, 1, 0, 0, 0, 0)‘.
-The algorithm should depict the 2x2 parameters of interest and estimate them close to one while the others should remain zero.
-The resulting estimates displayed below indicate proper estimates for the scale parameter vector `gamma_hat` but not for the location estimates `beta_hat`.
+A simple case consists in setting intercept and one slope for $\beta$ and $\gamma$ to one while the remaining $2 \times 4$ parameters are zero: $\beta = (1, 1, 0, 0, 0, 0)'$ and $\gamma = (1, 1, 0, 0, 0, 0)'$.
+The algorithm should depict the $2 \times 2$ parameters of interest and estimate them close to one while the others should remain zero.
+The resulting estimates displayed below indicate proper estimates for the scale parameter vector $\hat{\gamma}$ but not for the location estimates $\hat{\beta}$.
+
 ```
@@
 
-Setting up the example model from the asp20model repo, the algorithm, in turn, finds correct estimates – however being somewhat inaccurate for gamma_1_hat:
+Setting up the example model from the `asp20model` repository, the algorithm, in turn, finds correct estimates being however somewhat inaccurate for $\hat{\gamma_{1}}$ [@R-asp20model]:
 ```
@@
 
-Another model setup beta = (1,0,1,0,1,0)‘ and gamma = (0,1,0,1,0,1)‘ yields again correct estimates for gamma, but not for beta:
+Another model setup $\beta = (1,0,1,0,1,0)'$ and $\gamma = (0,1,0,1,0,1)'$ yields again correct estimates for $\gamma$, but not for $\beta$:
 
@@
 
-## Conclusions of the simulation studies
+## Conclusions
 
@@
 The performance of the boosting algorithm may be enhanced if initial estimates are chosen sensibly.
-For the location parameters beta this may be the OLS-estimate or the mean response. For the scale parameters gamma this may be the (homoskedastic) sample variance.
+For the location parameters $\beta$ this may be the OLS-estimate or the mean response. For the scale parameters $\gamma$ this may be the (homoskedastic) sample variance.
 Another possibility is to ask other student groups for straight-forward and fast numerical procedures to determine good starting values.
@@
 Trade-offs need to be considered.
+
+# 5. References

### 2020-08-12 R/model.R
implement swtich for componentwise boosting and structural changes , graph for parameter change

@@ -17,7 +17,52 @@ LocationScaleRegressionBoost <- R6Class(
   "LocationScaleRegressionBoost",
   inherit = LocationScaleRegression,
 
-  public = list(
+
+
+
+  private = list(
+
+    componentwise = TRUE,
+
+    X_HAT = numeric(),
+    X_HAT_componentwise = numeric(),
+    X_number_of_columns = numeric(),
+
+    Z_HAT = numeric(),
+    Z_HAT_componentwise = numeric(),
+    Z_number_of_columns = numeric(),
+
+    #Extend existing update beta to calculate least square estimates for current parameters
+    update_beta = function(value) {
+      super$update_beta(value)
+      #update least squares estimate only when model is initialized
+      if(length(private$X_number_of_columns)>0){
+        if(private$componentwise)
+        {
+          self$lstSqrEstBeta_componentwise <- private$componentwiseLossBeta()
+        }
+        else
+        {
+
+          self$lstSqrEstBeta <- private$X_HAT %*% crossprod(private$X, self$resid())
+        }
+      }
+    },
+
+    #Extend existing update gamma to calculate least square estimates for current parameters
+    update_gamma = function(value) {
+      super$update_gamma(value)
+
+      #update least squares estimate only when model is initialized
+      if(length(private$X_number_of_columns)>0){
+        if(private$componentwise){
+          self$lstSqrEstGamma_componentwise <- private$componentwiseLossGamma()
+        }
+        else{
+          self$lstSqrEstGamma <- private$Z_HAT %*% crossprod(private$Z, self$resid("deviance"))
+        }
+      }
+    },
 
     #' @details
     #' Returns the loss - squared error of the partial deviance for each covariate of Beta
@@ -29,8 +74,8 @@ LocationScaleRegressionBoost <- R6Class(
       #Fit separate linear models for all covariates of beta
       loss <- numeric()
       resid_working <- (self$resid()) #  Our u_i for beta in page 226 step 2.
-      number_of_columns <- dim(private$X)[2]
-      for(n in 1:number_of_columns) {
+
+      for(n in 1:private$X_number_of_columns) {
         predictor_colwise <- private$X[,n]
 
         # old way to do it: estimate residuals with predictor componentwise (columnwise)
[…]

### 2020-08-12 R/model.R
comment testing

@@ -367,18 +367,18 @@ gradient_boost = function(model,
   invisible(model)
 }
 
-library(R6)
-library(asp20model)
-start_time <- Sys.time()
-library(asp20model)
-set.seed(1337)
-n <- 500
-x <- runif(n)
-y <- x + rnorm(n, sd = exp(-3 + 2 * x))
-model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
-gradient_boost(model,stepsize = 0.001, maxit = 1000, abstol = 0.0001,componentwise = FALSE, verbose = TRUE, plot=TRUE)
-end_time <- Sys.time()
-end_time - start_time
+# library(R6)
+# library(asp20model)
+# start_time <- Sys.time()
+# library(asp20model)
+# set.seed(1337)
+# n <- 500
+# x <- runif(n)
+# y <- x + rnorm(n, sd = exp(-3 + 2 * x))
+# model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+# gradient_boost(model,stepsize = 0.001, maxit = 1000, abstol = 0.0001,componentwise = TRUE, verbose = TRUE, plot=TRUE)
+# end_time <- Sys.time()
+# end_time - start_time
 
 
 

### 2020-08-12 R/model.R
small fix

@@ -39,12 +39,12 @@ LocationScaleRegressionBoost <- R6Class(
       if(length(private$X_number_of_columns)>0){
         if(private$componentwise)
         {
-          self$lstSqrEstBeta_componentwise <- private$componentwiseLossBeta()
+          self$lstSqrEstBetaLossLoss_componentwise <- private$componentwiseLossBeta()
         }
         else
         {
 
-          self$lstSqrEstBeta <- private$X_HAT %*% crossprod(private$X, self$resid())
+          self$lstSqrEstBetaLoss <- private$X_HAT %*% crossprod(private$X, self$resid())
         }
       }
     },
@@ -56,10 +56,10 @@ LocationScaleRegressionBoost <- R6Class(
       #update least squares estimate only when model is initialized
       if(length(private$X_number_of_columns)>0){
         if(private$componentwise){
-          self$lstSqrEstGamma_componentwise <- private$componentwiseLossGamma()
+          self$lstSqrEstGammaLossLoss_componentwise <- private$componentwiseLossGamma()
         }
         else{
-          self$lstSqrEstGamma <- private$Z_HAT %*% crossprod(private$Z, self$resid("deviance"))
+          self$lstSqrEstGammaLoss <- private$Z_HAT %*% crossprod(private$Z, self$resid("deviance"))
         }
       }
     },
@@ -138,10 +138,10 @@ LocationScaleRegressionBoost <- R6Class(
   ),
 
   public = list(
-    lstSqrEstBeta = numeric(),
-    lstSqrEstBeta_componentwise = numeric(),
-    lstSqrEstGamma = numeric(),
-    lstSqrEstGamma_componentwise = numeric(),
+    lstSqrEstBetaLoss = numeric(),
+    lstSqrEstBetaLossLoss_componentwise = numeric(),
+    lstSqrEstGammaLoss = numeric(),
+    lstSqrEstGammaLossLoss_componentwise = numeric(),
 
     par_log = list(),
 
@@ -172,8 +172,8 @@ LocationScaleRegressionBoost <- R6Class(
       }
 
       #Init least square estimates for components of beta and gamma
-      self$lstSqrEstBeta_componentwise <- private$componentwiseLossBeta()
-      self$lstSqrEstGamma_componentwise <- private$componentwiseLossGamma()
+      self$lstSqrEstBetaLossLoss_componentwise <- private$componentwiseLossBeta()
[…]

### 2020-08-15 Reports/report_2/2_REPORT.Rmd
- Fixed spelling errors across various documents. - New folder for term paper.

@@
 - Sebastian Aristizabal
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: 24.07.2020
 output:
@@
 [1]  0.15122097  1.72916260  0.61363927 -0.14620695  0.02756381  0.28560006
-> model$gamma
-[1]  0.95706770  1.15196564 -0.02756374  0.00000000 -0.06839582  0.00000000
-```
-```
-model$beta
-[1]  0.15122097  1.72916260  0.61363927 -0.14620695  0.02756381  0.28560006
 model$gamma

### 2020-08-15 R/model.R

@@ -93,7 +93,7 @@ LocationScaleRegressionBoost <- R6Class(
 
 
     #' @details
-    #' Determins the best variable for componentwise boosting
+    #' Determines the best variable for componentwise boosting
     #'
     #' @return
     #' New Beta Vector with updated values on the best fitted covariate
@@ -118,7 +118,7 @@ LocationScaleRegressionBoost <- R6Class(
 
 
     #' @details
-    #' Determins the best variable for componentwise boosting
+    #' Determines the best variable for componentwise boosting
     #' @return
     #' New Gamma Vector with updated values on the best fitted covariate
     #' @export

### 2020-08-15 simulation_studies/now_working_simulation_variable_selection.R

@@ -1,4 +1,4 @@
-# This script analyses if variable selection in the case of multidemnsional
+# This script analyses if variable selection in the case of multidimensional
 # beta and gamma vectors works
 # For this reason some coefficients are 0.
 library(R6)
@@ -9,7 +9,7 @@ library(asp20boost)
 # in order to make this script work an extra step needs to be done:
 # the error is as follow: model creation inside the "boost_true_model"-function
 # below does not work, UNLESS a LocationScaleRegressionBoost-model class has been created before
-# for this reason asimple model call and estimation is performed here
+# for this reason a simple model call and estimation is performed here
 { x <- runif(100)
   y <- runif(100)
   model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)

### 2020-08-15 Reports/term_paper/TERM_PAPER.Rmd
- Term Paper - Base structe - Cover page.

--- 2020-06-05 Reports/report_1/1_REPORT.Rmd
@@
 ---
-title: "Report ASP20 Boost"
-author:
-- Johannes Strauß
-- Levin Wiebelt
-- Sebastian Aristizabal
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
     toc_depth: 2
+    includes:
+        in_header: preamble.tex
+        before_body: frontpage.tex
 documentclass: report
@@
 - packages.bib
-subtitle: First Report
 biblio-style: apalike
 ---
-
 ```{r setup, include=FALSE}
@@
 
-In *part 2* of this report we explain the concept of boosting and elaborate on how it suites to address the problem of location-scale regression. 
-In *part 3* we describe our implementation milestones, whereby *part 4* especially elaborates on componentwise boosting. 
-In *part 5* we present our thoughts on upcoming challenges and critical aspects of the package development.
+In *part 2*  
+In *part 3*  
+In *part 5*
 
-# 2. Concept: Boosting
+# 2 Model and Methods
+
+## Concept: Boosting
 
@@
 \begin{equation}
-ll = \frac{1}{2\pi\exp(Z\Gamma)^2}\times\exp(-\frac{u^2}{2\exp(Z\Gamma)^2})
+ll = \frac{1}{2\pi\exp(Z\gamma)^2}\times\exp(-\frac{u^2}{2\exp(Z\gamma)^2})
 \end{equation}
 
-# 3. Implementation
+## Functionality: Componentwise Boosting
+
+A useful extension of our simple boosting algorithm regards *componentwise boosting*. 
+The key idea here is to not update a whole parameter vector, but only one entry of it.
+The entry chosen for the update at this juncture is the one yielding the best improvement in terms of minimization of the loss function.
+The loss function in our case is the deviance calculated using $\gamma$ and our $\beta$.
+
+Hence, in each iteration only one component of the location- and one component of the scale-parameters are updated.
+
+### 1. Initialize the regression coefficients
+
+\begin{equation} 
+\begin{aligned}
+\beta_{j}^{(0)} = 0, \text{ for} j = 1,..,k. \\
+\gamma_{j}^{(0)} = 0, \text{ for} j = 1,..,k. \\
+\end{aligned}
+\end{equation}
+
+### 2. Compute the negative gradients
+
+Beta
+\begin{equation} 
+\begin{aligned}
+u_i = y - X\hat{\beta}^{(t-1)} \\
+\hat{b}_j = ((\vec{x}^j)^T \space\vec{x}^j)^{-1}((\vec{x}^j)^T \vec{u} \\
+\end{aligned}
+\end{equation}
+
+Gamma
+\begin{equation} 
+\begin{aligned}
+u = y - X\hat{\beta}^{(t-1)} \\
+u_i = X\hat{\beta}^{(t-1)}/Z\hat{\gamma}^{(t-1)} \\
+\hat{b}_j = u^2(\vec{z}^j)^Te^{-2(Z\hat{\gamma}^{(t-1)})} \\
+\end{aligned}
+\end{equation}
+
+### 3. Fit separate models for all covariates
+
+Beta
+\begin{equation} 
+\hat{b}_j = ((\vec{x}^j)^T \space\vec{x}^j)^{-1}((\vec{x}^j)^T \vec{u}, \text{for} \space j = 1,..,k. 
+\end{equation}
+
+Gamma
+\begin{equation} 
+\hat{b}_j = ((\vec{z}^j)^T \space\vec{z}^j)^{-1}((\vec{z}^j)^T \vec{u}, \text{ for} \space j = 1,..,k. 
+\end{equation}
+
+And determine the best fitting variable via:
+
+Beta
+\begin{equation} 
+\begin{aligned}
+u_i = y-X\hat{\beta}^{(t-1)} \\
+j^* = \arg\min_{j = 0,..,k} \sum_{i=1}^{n}(u_i-x_{ij}\hat{b}_j)^2\\
+\end{aligned}
+\end{equation}
+
+Gamma
+\begin{equation} 
+\begin{aligned}
+u_i = X\hat{\beta}^{(t-1)}/Z\hat{\gamma}^{(t-1)} \\
+j^* = \arg\min_{j = 0,..,k} \sum_{i=1}^{n}(u_i-x_{ij}\hat{b}_j)^2
+\end{aligned}
+\end{equation}
+
+### 4. Update Coefficients.
+
+Beta
+\begin{equation} 
+\begin{aligned}
+\hat{\beta}_{j^*}^{(t)}  = \hat{\beta}_{j^*}^{(t-1)}+ \nu\hat{b}_{j^*} \\
+\hat{\beta}_{j}^{(t)}  = \hat{\beta}_{j}^{(t-1)}, j\neq j^*.
+\end{aligned}
+\end{equation}
+
+Gamma
+\begin{equation} 
+\begin{aligned}
+\hat{\gamma}_{j^*}^{(t)}  = \hat{\gamma}_{j^*}^{(t-1)}+ \nu\hat{b}_{j^*} \\
+\hat{\gamma}_{j}^{(t)}  = \hat{\gamma}_{j}^{(t-1)}, j\neq j^*.
+\end{aligned}
+\end{equation}
+
+We implement this in our package by extending the `LocationScaleRegression` class to include the two active fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.
+
+These functions partition the design matrix $X$- respectively $Z$ - into its single columns and then estimate the residuals - respectively the scores - separately for each component, and determine loss functions for a hypothetical update with the respective component.
+Respective to the old loss function value, the highest loss-improvement is determined and the respective component is used to update the parameter-vector.
+
+We are in the process of reconsidering the design of this implementation.
+Other design possibilities are the following:
+
+* Create public fields for the heavily used score-function values.
+* Move the calculation of componentwise losses to an external function.
+* Harmonize boosting and componentwise boosting into one external function, determining the mode of operation by an argument `componentwise = TRUE`.
+
+Another conceptual questions that comes up is if the best loss-improvement may be indicated by the already existing gradients, and hence there is no need for extra calculation.
+
+# 3. Software Implementation
 
@@
 
-# 4. Functionality: Componentwise Boosting
+# 4. Simulation Studies
 
-A useful extension of our simple boosting algorithm regards *componentwise boosting*. 
-The key idea here is to not update a whole parameter vector, but only one entry of it.
-The entry chosen for the update at this juncture is the one yielding the best improvement in terms of minimization of the loss function.
-The loss function in our case is the deviance calculated using $\gamma$ and our $\beta$.
-
-Hence, in each iteration only one component of the location- and one component of the scale-parameters are updated.
-
-### 1. Beta
-
-\begin{equation} 
-\begin{aligned}
-\beta_{j}^{(0)} = 0, \text{ for} j = 1,..,k. \\
-u_i = y - X\hat{\beta}^{(t-1)} \\
-\hat{b}_j = ((\vec{x}^j)^T \space\vec{x}^j)^{-1}((\vec{x}^j)^T \vec{u} \\
-\end{aligned}
-\end{equation}
-
-### 2. Gamma
-
-\begin{equation} 
-\begin{aligned}
-\gamma_{j}^{(0)} = 0, \text{ for} j = 1,..,k. \\
-u_i = \text{deviance} \\
-\hat{b}_j = U^2Z^2e^{-2(Z\Gamma)} \\
-\end{aligned}
-\end{equation}
-
-### 3. Same procedure.
-
-Fit separate models for all covariates i.e obtain
-
-\begin{equation} 
-\hat{b}_j = ((\vec{x}^j)^T \space\vec{x}^j)^{-1}((\vec{x}^j)^T \vec{u}, \text{ for} \space j = 1,..,k. 
-\end{equation}
-
-And determine the best fitting variable via
-
-\begin{equation} 
-j^* = \arg\min_{j = 0,..,k} \sum_{i=1}^{n}(u_i-x_{ij}\hat{b}_j)^2.
-\end{equation}
-
-### 4. Update Coefficients.
-
-\begin{equation} 
-\begin{aligned}
-\hat{\beta}_{j^*}^{(1)}  = \hat{\beta}_{j^*}^{(0)}+ \nu\hat{b}_{j^*} \\
-\hat{\beta}_{j}^{(1)}  = \hat{\beta}_{j}^{(0)}, j\neq j^*.
-\end{aligned}
-\end{equation}
-
-We implement this in our package by extending the `LocationScaleRegression` class to include the two active fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.
-
-These functions partition the design matrix $X$- respectively $Z$ - into its single columns and then estimate the residuals - respectively the scores - separately for each component, and determine loss functions for a hypothetical update with the respective component.
-Respective to the old loss function value, the highest loss-improvement is determined and the respective component is used to update the parameter-vector.
-
-We are in the process of reconsidering the design of this implementation.
-Other design possibilities are the following:
-
-* Create public fields for the heavily used score-function values.
-* Move the calculation of componentwise losses to an external function.
-* Harmonize boosting and componentwise boosting into one external function, determining the mode of operation by an argument `componentwise = TRUE`.
-
-Another conceptual questions that comes up is if the best loss-improvement may be indicated by the already existing gradients, and hence there is no need for extra calculation.
-
-# 5. Prospects
-
-## Further Functionalities
-
-The most important functionality we intend to implement in our package is the optimization of the number of iterations via cross validation. 
-To work out the theoretical concept behind this idea is, as well as to implementing it are our current next goals.
-One further, rather loose, idea is to find ways to optimize the learning rate in our boosting algorithm.
-
-## Stability
-
-We intend to implement further automated unit tests. 
-This will enable us to assess quickly the stability of our code given differing inputs.
-We also intend to implement measures for proneness reduction to input errors.
-
-## Performance
-
-* A major performance problem in our code is that small learning rates for the $\beta$-booster lead to a  sharp increase of our code's processing time. This remains to be solved.
-* A good choice of starting values for $\beta$ and $\gamma$ may enhance performance. A possible candidate is the mean of the response.
-
-## Further Aspects
-
-* Allow user-input of the LocationScaleRegressionBoost-model in form of a data frame, for example by allowing to pass an optional "data"-argument as known from the `lm-call`.
-This may reduce proneness to input-errors.
-
-* Visualize the results of the boosted estimates
-
-* Document the extension of the R6-Class properly and inherit documentation of the `LocationScaleRegression-Class` using roxygen2.
+# 5. Discussion/Outlook.
 

### 2020-08-18 Reports/term_paper/TERM_PAPER.Rmd
- Changes on final report. - First update of Roxygen comments.

@@
 
+# Abstract
+
+We describe version 1.0 of the R add-on package asp20boost. The package  implements boosting *Note: Expand*. 
+Keywords: component-wise boosting,  gradient descent, location and scale.
+
 # 1. Introduction:
@@
 
-## Concept: Boosting
+## Concept: Simple Boosting
 
@@
 
-## Location-Booster
+### Location-Booster
 
@@
 
+*Note to self: Add the description of the whole algorithm. Not just the residuals*
+
 \begin{equation}
@@
 
-## Scale-Booster
+### Scale-Booster
 
@@
 
-## Functionality: Componentwise Boosting
+## Componentwise Boosting
+
+*Note to self: Add rationale of CWB and motivation for prefering it over simple B*
 
@@
 
-We are in the process of reconsidering the design of this implementation.
-Other design possibilities are the following:
+# 3. Software Implementation
 
-* Create public fields for the heavily used score-function values.
-* Move the calculation of componentwise losses to an external function.
-* Harmonize boosting and componentwise boosting into one external function, determining the mode of operation by an argument `componentwise = TRUE`.
-
-Another conceptual questions that comes up is if the best loss-improvement may be indicated by the already existing gradients, and hence there is no need for extra calculation.
-
-# 3. Software Implementation
+*Note to self: I pooled all statements of report 1 regarding inplementation on this section. There is thus not coherent argument. Most of it is superfluous. Needs to be updated.*
 
@@
 
+We are in the process of reconsidering the design of this implementation.
+Other design possibilities are the following:
+
+* Create public fields for the heavily used score-function values.
+* Move the calculation of componentwise losses to an external function.
+* Harmonize boosting and componentwise boosting into one external function, determining the mode of operation by an argument `componentwise = TRUE`.
+
+Another conceptual questions that comes up is if the best loss-improvement may be indicated by the already existing gradients, and hence there is no need for extra calculation.
+
 # 4. Simulation Studies

### 2020-08-18 R/model.R
- Changes on final report. - First update of Roxygen comments.

@@ -1,13 +1,24 @@
-#' @title asp20boost
+#' @title R6 class for performing either simple and generic componentwise
+#'  boosting on location-scale regression models.
 #'
-#' This model class [...]
+#'
+#' This model class builts upon and thus inherits the basic structure
+#' from the `LocationScaleRegression` model class of the `asp20model` package.
+#' Consequently, it share similar syntax and structure. (Please  see the help
+#' page of that class *Note: It would be nice to just link the vignette*)
+#'
+#' It assumes a normally distributed response variable with
+#' one linear predictor for the location (i.e. the mean) and one for the
+#' scale (i.e. the standard deviation). The linear predictors for the
+#' location and the scale are called Xβ and Zγ respectively. The scale
+#' uses a log link.
 #'
 #' @field componentwiseLossGamma Calculates loss function for
-#'   every component in scale-dimension/gamma
+#'   every component in scale-dimension `gamma`
 #' @field componentwiseLossBeta Calculates loss function for
-#'   every component in location-dimension/beta
-#' @field bestFittingVariableGamma Calculates the actual update for gamma
-#' @field bestFittingVariableBeta Calculates the actual update for beta
+#'   every component in location-dimension `beta`
+#' @field bestFittingVariableGamma Calculates the actual update for `gamma`
+#' @field bestFittingVariableBeta Calculates the actual update for `beta`
 #'
 #' @import R6
 #' @import asp20model

### 2020-08-18 R/model.R
Roxygen, Documentation page.

@@ -1,24 +1,22 @@
-#' @title R6 class for performing either simple and generic componentwise
-#'  boosting on location-scale regression models.
+#' @title R6 class for performing either simple or generic componentwise
+#'  boosting (CWB) on location-scale regression models.
 #'
-#'
-#' This model class builts upon and thus inherits the basic structure
+#' @description This model class builds upon and thus inherits the basic structure
 #' from the `LocationScaleRegression` model class of the `asp20model` package.
-#' Consequently, it share similar syntax and structure. (Please  see the help
-#' page of that class *Note: It would be nice to just link the vignette*)
 #'
+#' Consequently, it shares a similar syntax and structure with the former#' #'
 #' It assumes a normally distributed response variable with
 #' one linear predictor for the location (i.e. the mean) and one for the
 #' scale (i.e. the standard deviation). The linear predictors for the
 #' location and the scale are called Xβ and Zγ respectively. The scale
 #' uses a log link.
 #'
-#' @field componentwiseLossGamma Calculates loss function for
-#'   every component in scale-dimension `gamma`
-#' @field componentwiseLossBeta Calculates loss function for
-#'   every component in location-dimension `beta`
-#' @field bestFittingVariableGamma Calculates the actual update for `gamma`
-#' @field bestFittingVariableBeta Calculates the actual update for `beta`
+#' @field `componentwiseLossGamma` Calculates loss function for
+#'   every component in scale-dimension `gamma` in case of CWB
+#' @field `componentwiseLossBeta` Calculates loss function for
+#'   every component in location-dimension `beta` in case of CWB.
+#' @field `bestFittingVariableGamma` Calculates the actual update for `gamma`
+#' @field `bestFittingVariableBeta` Calculates the actual update for `beta`
 #'
 #' @import R6
 #' @import asp20model
@@ -27,8 +25,6 @@
 LocationScaleRegressionBoost <- R6Class(
   "LocationScaleRegressionBoost",
   inherit = LocationScaleRegression,
-
-
 
 
   private = list(
@@ -106,7 +102,8 @@ LocationScaleRegressionBoost <- R6Class(
 
 
     #' @details
-    #' Returns the loss - squared error of the partial deviance for each covariate of Gamma
+    #' Calculates the loss i.e the squared error of the partial deviance for each
+    #' covariate of `gamma`
     #' @return
[…]

### 2020-09-07 R/model.R
deleted redundant gradient check

@@ -194,27 +194,12 @@ gradient_boost = function(model,
                           verbose = FALSE,
                           plot = FALSE) {
 
-  # store gradients from last iteration step to compare them with newly calculated ones
-  grad_old_mu <- model$gradients_loglik_mu()
-  grad_old_sigma <- model$gradients_loglik_sigma()
-
   model$par_log <- list()
   for(iter in 1:maxit) {
 
     if(componentwise == T) model$update_parameters_compwise()
     if(componentwise == F) model$update_parameters_conventional()
 
-    # check for substantial change in gradients of loglikelihood (wrt eta_mu, eta_sigma)
-    {
-    grad_change_mu <- abs(grad_old_mu - model$gradients_loglik_mu())
-    grad_change_sigma <- abs(grad_old_sigma - model$gradients_loglik_sigma())
-    grad_changes <- c(grad_change_mu, grad_change_sigma)
-    }
-
-    if(iter > 1 && all(grad_changes <= abstol)){
-      message("early stopping at iteration: ",iter)
-      break()
-    }
 
     model$par_log[[iter]]<-c(model$beta, model$gamma)
     if(plot == T) model$plot()
@@ -224,7 +209,7 @@ gradient_boost = function(model,
       par_msg <- format(par_msg, trim = TRUE, digits = 3)
       par_msg <- paste(par_msg, collapse = " ")
 
-      loglik_msg <- format(model$loglik(), digits = 3)
+      loglik_msg <- format(model$loglik(), digits = 6)
 
       message(
         "Iteration:      ", iter, "\n",

### 2020-09-07 README.Rmd
- Changes on final report. - First update of Roxygen comments.

@@
 ---
-
-<!-- README.md is generated from README.Rmd. Please edit that file -->
 
@@
 
+Welcome to *asp20boost*.
+We implement both simple boosting and component-wise boosting ensemble algorithms for location and scale settings.
+
  * Supervisor: Thomas Kneib
  * Students: Johannes, Levin, Sebastián
-* Topic: Parameter estimation and variable selection using the boosting ensemble learning algorithm
 
-## Tasks
+We have successfully implemented both boosting variants.
+Currently, we work on the publication of the package i.e. we perform unit tests and expand the documentation.
 
-### until ~ mid June
-
-* extend documentation of R6-Class
-  + inherit documentation from Hannes
-  + add own extensions
-  + both can be done via roxygen (see Wickham R-PAckages chp5)
-
-* maybe move component-selection mechanism from active field in R6-Class to gradient_boost-function
-  + intuition tells me R6-Class is not the right place to do central calculations of the algorithm
-  + maybe create new helper function component_selection() which gradien_boost() can call
-
-### until ~ 11 May
-
-* simple boost function for scale (estimate var(yi))
-* read Wickham "R-Packages"
-* try implementing function-testing
-* get used to roxygen2
-
-## Literature
+### Literature
 * Section **4.3** on "Boosting Linear Regression Models" in "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx
@@
 
-## Location Scale Regression Model
+This is a visual representation of a location scale Regression Model (taken from asp20model)
 
@@
 
-## Working with Location Scale Regression Boost Model
-First, you need to install the `asp20model` &  `asp20boost` package from GitLab and load it:
+To install and work with out package you need first to install the `asp20model` &  `asp20boost` package from GitLab and load it. 
+The installing of devtools requires attention.
 
@@
 #install.packages("devtools")
-#This throws an error for asp20boost
 #devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
 #devtools::install_gitlab("asp20/asp20boost", host = "gitlab.gwdg.de")
-
-model$gamma
 ```

### 2020-09-14 simulation_studies/draft_sim_study_13_sep.Rmd
Added simulation studies

---
title: "Simulation Study"
output: html_notebook
---

In this part we design and perform a simulation study in order to demonstrate functionality of the implemented boosting algorithm. For this, we first employ small models with two location and scale parameters each and ensure basic functionality. Second, validity of the variabel selection features is demonstrated.

```{r}
# setting the stage
rm(list=ls())
options(scipen=999)
source("helper_plot_model.R")
source("helper_init_model.R")
source("helper_calc_mse.R")
source("helper_init_large_model.R")
library(asp20boost)
library(gamboostLSS)
```

Helper functions for model initialization, plotting, etc. are called, as well as the asp20boost package itself and the package gamboostLSS containing comparison functions.

## ToDo
* add labels to beta gamma output
* befor every subchapter "Evaluation Config.*", put a formula with the current parameter configuration that is being evaluated.
* add substantial effect in large model to make clear how effect is visible

## Small Model - 4 Parameter Case

The goal here is to demonstrate, that the algorithm finds correct estimates for sensible choices of /beta and /gamma. By evaluating different parameter configurations it becomes clear what model properties they represent. We use the following four different parameter configurations to explore location-scale regression model characteristics:

```{r}
# config 1 ----------------------
beta_1 <- c(0,1)
gamma_1 <- c(-3,2)

# config 2 ----------------------
beta_2 <- c(2,2)
gamma_2 <- c(2,2)

# config 3 ----------------------
beta_3 <- c(0,7)
gamma_3 <- c(-3,3)

# config 4 ----------------------
beta_4 <- c(0,1)
gamma_4 <- c(3,-3)

# config 5 ----------------------
beta_5 <- c(0,7)
gamma_5 <- c(1,-1)
```

### Evaluation Configuration 1

We call the helper-function to create a model object of 1000 observations and plot it with a second helper function.

```{r}
mod_1 <- init_model(beta_1, gamma_1, 1000)
plot_model(beta_1, gamma_1, 1000)
```

We have already seen the basic plot structure, visualizing a location-scale regression model. What’s added is the depiction of two important quantities: Indicated by the blue arrow, one is the location slope /beta_1. The red arrow, in contrasts, indicates [modeled?] standard deviation of the observed responses at the point x=1. The ratio of those two quantities is essential for possibility of estimation. Here, we have a good signal-noise ratio, which enables the gradient_boost() function to estimate it.

```{r}
gradient_boost(mod_1)
round(mod_1$beta, 3)
round(mod_1$gamma, 3)
```

We see that the estimates are close to the true model parameters.

#### Evaluation Configuration 2

With all parameters set to 2, the algorithm yields NaNs.

```{r}
mod_2 <- init_model(beta_2, gamma_2, 1000)
gradient_boost(mod_2)
calc_MSE(beta_2, gamma_2, mod_2$beta, mod_2$gamma)
```

To understand the model characteristics of parameter configuration 2-2-2-2, we look at the model plot.

```{r}
plot_model(beta_2, gamma_2, 1000)
```

Here, the red arrow is large, while the blue arrow is hardly visible. This represents the fact, that gamma parameters lead to SD being immensly high at x=1 and hence Untergehen of location effect. The model is not estimable. This is due to the fact that sd(x=1) = exp(gamma_0 + gamma_1 * 1) = exp(4) = 54.60. The location slope is 2. It is impossible to estimate an effect of 2 (change of two unit if x increases by one), if the random noise of the response is already very likely to be +/- 55 at the point x=1. Non-linearity of SDs makes this comparison hard to do intuitively. However, with the exponential function in mind one can abschätzen where sds may lie given parameter configuration. 
In contrast, location predictions are easily imagined by simply seeing beta_0 and beta_1. For x=0 the vlue is equal to the intercept, for x=1 the value is the sum of both betas. X values restricted to 0,1-range make it even easier to imagine preditors since the highest SD of interest is at x=1, where the predictor may be easily calculated by adding gamma_0 and gamma_1.

#### Evaluation Configuration 3

```{r}
mod_3 <- init_model(beta_3, gamma_3, 1000)
plot_model(beta_3, gamma_3, 1000)
```

Configuration 3 yields a good signal-noise ratio.

```{r}
gradient_boost(mod_3)
round(mod_3$beta, 3)
round(mod_3$gamma, 3)
```

The algorithm performs well with estimates close to true parameters.To quantify how close estimates are to true parameters, we quantify this distance employing the mean squared error.

```{r}
calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
```

#### Evaluation Configuration 4

```{r}
mod_4 <- init_model(beta_4, gamma_4, 1000)
plot_model(beta_4, gamma_4, 1000)
```

Configuration 4 involves a declining SD, which leads to the other end of the extreme - extremely low sd, leading to NaN problems (numerical underflow).

```{r}
gradient_boost(mod_4)
round(mod_4$beta, 3)
round(mod_4$gamma, 3)
```

The alogirthm also yields NaNs.

#### Evaluation Configuration 5

```{r}
mod_5 <- init_model(beta_5, gamma_5, 1000)
plot_model(beta_5, gamma_5, 1000)
```

Configuration 5 involves a declining SD.

```{r}
gradient_boost(mod_5)
round(mod_5$beta, 3)
round(mod_5$gamma, 3)
```

This time, the algorithm behaves poorly. The estimates are rather far from the true parameters. This, in turn, leads to high MSE.

```{r}
calc_MSE(beta_5, gamma_5, mod_5$beta, mod_5$gamma)
```

### Conclusions

* The specified model must have a signal-noise ratio, that permits estimation.
* Gamma-slope has to be positive. (Here, the package may be further developed.)

## Large Model - Variable Selection

[...]

```{r}
# initialize model-----------------------------------------------
result <- init_large_model()
mod_1 <- result$model
X <- result$design
y <- result$response
fitted_loc <- result$fitted_response
fitted_scale <- result$fitted_sd

# check model via plotting --------------------------------------
par(mfrow = c(4,3))
for(i in 1:11) plot(X[,i], y, ylim = c(0,200))

```

No effect becomes clearly visible, since each individually is too small to have a substantial effect on the response.

```{r}
# check model via linear predictors ---------------------------------------------------------
fitted_loc
fitted_scale

```

Model estimation currently does not work.

```{r}
gradient_boost(mod_1, maxit = 1000, componentwise = TRUE, verbose = F)

```
Resulting parameter estimates are wrong. Almost no component become updated.
```{r}
mod_1$beta
mod_1$gamma
```

[...]

## Benchmark gamboostLSS

[...]

[...]

[...]

### 2020-09-14 simulation_studies/compare_gamboostlss.R

@@ -1,0 +1,41 @@
+library(asp20boost)
+
+# set true model parameters
+beta_0  <- 0
+beta_1  <- 1
+gamma_0 <- -3
+gamma_1 <- 2
+
+
+
+# simulate data -------------------------
+set.seed(1337)
+n <- 1000
+x <- runif(n)
+y <- beta_0 + beta_1 * x + rnorm(
+  n,
+  sd = exp(gamma_0 + gamma_1 * x)
+)
+
+xy_df <- cbind(x,y)
+head(xy_df)
+
+# estimate data by boosting ------------
+mod_1 <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+gradient_boost(model, maxit = 1000)
+
+
+# estimate with external package --------
+ctrl <- boost_control(trace = FALSE,
+                      mstop = c(mu = 1000, sigma = 1000),
+                      nu = c(mu = 0.01, sigma = 0.001)
+                      )
+mod_2 <- gamboostLSS(formula = y ~ x,
+                     control = ctrl)
+eta_mu <- fitted(mod_2)$mu
+eta_sigma <- fitted(mod_2)$sigma
+
+
+lm(eta_mu ~ x)
+lm(eta_sigma ~ x)
+

### 2020-09-14 simulation_studies/estimate_model.R

@@ -1,0 +1,58 @@
+rm(list=ls())
+source("simulation_studies/helper_plot_model.R")
+source("simulation_studies/helper_init_model.R")
+source("simulation_studies/helper_calc_mse.R")
+library(asp20boost)
+library(gamboostLSS)
+
+beta_1 <- c(0,1)
+gamma_1 <- c(-3,2)
+beta_2 <- c(0,1)
+gamma_2 <- c(-0.5,0)
+beta_3 <- c(0,1)
+gamma_3 <- c(-3,3)
+beta_4 <- c(0,1)
+gamma_4 <- c(3,-3)
+
+
+
+mod_1 <- init_model(beta_1, gamma_1, 1000)
+plot_model(beta_1, gamma_1, 1000)
+gradient_boost(mod_1)
+calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
+
+
+# mod_2 <- init_model(beta_2, gamma_2, 1000)
+# plot_model(beta_2, gamma_2, 1000)
+# gradient_boost(mod_2)
+# calc_MSE(beta_2, gamma_2, mod_2$beta, mod_2$gamma)
+
+
+# mod_3 <- init_model(beta_3, gamma_3, 1000)
+# plot_model(beta_3, gamma_3, 1000)
+# gradient_boost(mod_3)
+# mod_3$beta
+# mod_3$gamma
+
+
+# mod_4 <- init_model(beta_4, gamma_4, 1000)
+# plot_model(beta_4, gamma_4, 1000)
+# gradient_boost(mod_4)
+# mod_4$beta
+# mod_4$gamma
+
+
+
+# ausprobieren --------------------------------------
+
+
+
+
+
+
+
+
+
+
+
+

### 2020-09-14 simulation_studies/helper_calc_mse.R

@@ -1,0 +1,13 @@
+calc_MSE <- function(true_beta, true_gamma, est_beta, est_gamma) {
+  result_list <- list()
+  result_list$beta <- 0
+  result_list$gamma <- 0
+  result_list$pooled <- 0
+
+  for(i in 1:2){
+    result_list$beta <- result_list$beta + (true_beta[i] - est_beta[i])^2
+    result_list$gamma <- result_list$gamma + (true_gamma[i] - est_gamma[i])^2
+  }
+  result_list$pooled <- result_list$beta + result_list$gamma
+  return(result_list)
+}

### 2020-09-14 simulation_studies/helper_init_large_model.R

@@ -1,0 +1,64 @@
+init_large_model <- function(beta_range = c(10, 20),
+                             gamma_range = c(0.1, 0.3)
+){
+
+
+  # sample coefficients ------------------------------------------------------------
+  beta <- round(runif(11, min = 10, max = 20), digits = 0)
+  gamma <- round(runif(11, min = 0.1, max = 0.3), digits = 1)
+
+  beta[seq(3,11,3)] <- 0 # set every third beta coef to 0
+  gamma[seq(2,10,2)] <- 0 # set every second gamma coef to 0
+  beta[1] <- 0 # set intercept to 0 for convenience
+  gamma[1] <- -3 # set gamma intercept to keep sds in an estimable range
+  #beta[2] <- 280 # set one very high effect for checking]
+
+
+  # sample covariates --------------------------------------------------------------
+  # we sample 100 covariates. intercept is automatically added by model class
+  covariates_X <- runif(100 * 11) # generate 100 obs each for 100 covariates
+  covariates_Z <- runif(100 * 11)
+
+  X <- matrix(covariates_X, nrow = 100, byrow = TRUE)
+  X[, 1] <- 1
+  Z <- matrix(covariates_Z, nrow = 100, byrow = TRUE)
+  Z[, 1] <- 1
+
+
+  # sample response ----------------------------------------------------------------
+  y <- rep(NA, times = 100)
+  eta_mu <- rep(NA, 100)
+  eta_sigma <- rep(NA, 100)
+
+  for(i in 1:100) {
+    x_i <- X[i, ]
+    eta_mu[i] <- x_i %*% beta
+    eta_sigma[i] <- x_i %*% gamma
+    y[i] <- rnorm(n = 1,
+                  mean = x_i %*% beta,
+                  sd = exp(x_i %*% gamma)
+                  #sd = exp(z_i) %*% gamma
+                  #sd = 0.01
+    )
+  }
+
+
+  # create model call --------------------------------------------------------------
+  xnam <- paste("x", 1:10, sep="")
+  for(i in 1:10) assign(xnam[i], X[, i+1]) # i+1 since first term is no covariate, but intercept-1-vector
+  fmla_loc <- as.formula(paste("y ~ ", paste(xnam, collapse= "+")))
+  fmla_scale <- as.formula(paste("~ ", paste(xnam, collapse= "+")))
+  mod_1 <- LocationScaleRegressionBoost$new(fmla_loc, fmla_scale)
+
+
[…]

### 2020-09-14 simulation_studies/helper_init_model.R

@@ -1,0 +1,15 @@
+# set up main plotting function ----------------------------------
+init_model <- function(beta, gamma, n) {
+
+  # set up linear predictor functions -------------------------------
+  eta_mu <- function(x) beta[1] + beta[2] * x
+  eta_sigma <- function(x) gamma[1] + gamma[2] * x
+
+  # set up random variables -----------------------------------------
+  x <- runif(n)
+  y <- eta_mu(x) + rnorm(n, sd = exp(eta_sigma(x)))
+
+  # create model object ---------------------------------------------
+  LocationScaleRegressionBoost$new(y ~ x, ~ x)
+
+}

### 2020-09-14 simulation_studies/helper_plot_model.R

@@ -1,0 +1,31 @@
+# set up main plotting function ----------------------------------
+plot_model <- function(beta, gamma, n) {
+
+  # set up linear predictor functions -------------------------------
+  eta_mu <- function(x) beta[1] + beta[2] * x
+  eta_sigma <- function(x) gamma[1] + gamma[2] * x
+
+  # set up random variables -----------------------------------------
+  x <- runif(n)
+  y <- eta_mu(x) + rnorm(n, sd = exp(eta_sigma(x)))
+
+  # configure plot --------------------------------------------------
+  plot(x, y, col = "grey", cex = 0.5, pch = 16)
+  abline(beta[1], beta[2], lwd = 2, col = 1)
+  curve(eta_mu(x) + 1 * exp(eta_sigma(x)), -0.1, 1.1, add = TRUE, col = 1, lty = 3, lwd = 2)
+  curve(eta_mu(x) - 1 * exp(eta_sigma(x)), -0.1, 1.1, add = TRUE, col = 1, lty = 3, lwd = 2)
+  abline(h = beta[1], lty = 2, col = 4)
+
+  arrows(x0 = 1, x1 = 1,
+         y0 = eta_mu(1),
+         y1 = eta_mu(1) + 1 * exp(eta_sigma(1)),
+         length = 0.05, code = 3, col = 2, lty = 1, lwd = 2)
+  arrows(x0 = 0.985, x1 = 0.985,
+         y0 = eta_mu(1),
+         y1 = beta[1],
+         length = 0.05, code = 3, col = 4, lty = 1, lwd = 2)
+
+  legend("topleft", legend=c("location slope", "standard deviation at x=1"),
+         col=c(4, 2), lty=c(1,1), cex=0.8)
+
+}

### 2020-09-14 simulation_studies/helper_plot_model_high_p.r

@@ -1,0 +1,13 @@
+helper_plot_model_high_p <- function(beta, gamma, X, index){
+  
+  # set up linear predictor functions -------------------------------
+  eta_mu <- function(x) beta[1] + beta[index] * x
+  eta_sigma <- function(x) gamma[1] + gamma[index] * x
+  
+  x <- X[ , index]
+  
+  plot(x, y) # plot y against column vectors of X
+  curve(mean(y) + eta_mu(x) + 1 * exp(eta_sigma(x)), -0.1, 1.1, add = TRUE, col = 1, lty = 3, lwd = 2)
+  curve(mean(y) + eta_mu(x) - 1 * exp(eta_sigma(x)), -0.1, 1.1, add = TRUE, col = 1, lty = 3, lwd = 2)
+  
+}

### 2020-09-14 simulation_studies/large_variable_selection.r

@@ -1,0 +1,36 @@
+rm(list = ls())
+library(asp20boost)
+options(scipen=999)
+source("simulation_studies/helper_init_large_model.R")
+windows()
+
+
+# initialize model
+result <- init_large_model()
+mod_1 <- result$model
+X <- result$design
+y <- result$response
+fitted_loc <- result$fitted_response
+fitted_scale <- result$fitted_sd
+
+# check model via plotting ------------------------------------------------------------------
+par(mfrow = c(4,3))
+for(i in 1:11) plot(X[,i], y, ylim = c(0,200))
+# only one effect is clearly indicated, the others are too small compared to the very high response
+# the effect is also indicated if heteroskedastic sds are added. Hence, signal-noise ratio is OK.
+# ALSO FOR OTHER effects? or only for the exceptionally high one?
+
+
+
+# check model via linear predictors ---------------------------------------------------------
+fitted_loc
+fitted_scale
+
+# estimate huge model -----------------------------------------------------------------------
+#gradient_boost(mod_1, maxit = 1000, componentwise = TRUE)
+
+
+
+# ausprobieren ------------------------------------------------------------------------------
+mod_1$beta
+mod_1$gamma

### 2020-09-14 R/model.R
add caching of cholsky,  improved plot performance, add stepsize vector

@@ -35,22 +35,33 @@ LocationScaleRegressionBoost <- R6Class(
 
   private = list(
 
+    X_HAT = numeric(),
+    X_HAT_componentwise = numeric(),
+    X_number_of_columns = numeric(),
+
+    Z_HAT = numeric(),
+    Z_HAT_componentwise = numeric(),
+    Z_number_of_columns = numeric(),
+
+
     # estimators (bj_hat) for location and scale,
     # conventional <-> componentwise --------------------------------------------------------
 
     gradient_estimators_mu = function() {
-      chol2inv(chol(crossprod(private$X, private$X))) %*% crossprod(private$X, self$gradients_loglik_mu())
+      private$X_HAT %*% crossprod(private$X, self$gradients_loglik_mu())
     },
     gradient_estimators_sigma = function() {
-      chol2inv(chol(crossprod(private$Z, private$Z))) %*% crossprod(private$Z, self$gradients_loglik_sigma())
+      private$Z_HAT %*% crossprod(private$Z, self$gradients_loglik_sigma())
     },
 
     gradient_estimators_mu_compwise = function() {
       number_of_cols <- ncol(private$X)
       result_list <- rep(NA, number_of_cols)
       for(i in 1:number_of_cols){
-        result_list[i] <- chol2inv(chol(crossprod(private$X[, i], private$X[, i]))) %*% crossprod(private$X[, i], self$gradients_loglik_mu())
-      }
+        #result_list[i] <- chol2inv(chol(crossprod(private$X[, i], private$X[, i]))) %*% crossprod(private$X[, i], self$gradients_loglik_mu())
+        result_list[i] <- private$X_HAT_componentwise[i] %*% crossprod(private$X[, i], self$gradients_loglik_mu())
+
+       }
       return(result_list)
     },
 
@@ -58,7 +69,9 @@ LocationScaleRegressionBoost <- R6Class(
       number_of_cols <- ncol(private$Z)
       result_list <- rep(NA, number_of_cols)
       for(i in 1:number_of_cols){
-        result_list[i] <- chol2inv(chol(crossprod(private$Z[, i], private$Z[, i]))) %*% crossprod(private$Z[, i], self$gradients_loglik_sigma())
+        #result_list[i] <- chol2inv(chol(crossprod(private$Z[, i], private$Z[, i]))) %*% crossprod(private$Z[, i], self$gradients_loglik_sigma())
[…]

### 2020-09-14 simulation_studies/helper_calc_mse.R
updating simulation studies

@@ -2,12 +2,12 @@ calc_MSE <- function(true_beta, true_gamma, est_beta, est_gamma) {
   result_list <- list()
   result_list$beta <- 0
   result_list$gamma <- 0
-  result_list$pooled <- 0
+  result_list$total <- 0
 
   for(i in 1:2){
     result_list$beta <- result_list$beta + (true_beta[i] - est_beta[i])^2
     result_list$gamma <- result_list$gamma + (true_gamma[i] - est_gamma[i])^2
   }
-  result_list$pooled <- result_list$beta + result_list$gamma
+  result_list$total <- result_list$beta + result_list$gamma
   return(result_list)
 }

### 2020-09-14 simulation_studies/helper_init_model.R

@@ -10,6 +10,14 @@ init_model <- function(beta, gamma, n) {
   y <- eta_mu(x) + rnorm(n, sd = exp(eta_sigma(x)))
 
   # create model object ---------------------------------------------
-  LocationScaleRegressionBoost$new(y ~ x, ~ x)
+  model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
+
+  result_list <- list()
+  result_list$model <- model
+  result_list$covariate <- x
+  result_list$response <- y
+  result_list$fitted_response <- eta_mu
+  #result_list$fitted_sd <- exp(eta_sigma)
+  return(result_list)
 
 }

### 2020-09-14 simulation_studies/draft_sim_study_13_sep.Rmd

@@
 In this part we design and perform a simulation study in order to demonstrate functionality of the implemented boosting algorithm.
-For this, we first employ small models with two location and scale parameters each and ensure basic functionality.
-Second, validity of the variabel selection features is demonstrated.
+For this, we first employ small models with two location and scale parameters each, and ensure basic functionality.
+Second, validity of the variable selection features is demonstrated.
+Lastly, we compare our algorithm with the one implemented in the gamboostLSS package.
+
+In order ensure legibility we employ some helper functions for model initialization, plotting, etc. These and the packages are loaded in the follwing chunk.
 
@@
 library(gamboostLSS)
+library(microbenchmark)
 ```
-
-Helper functions for model initialization, plotting, etc. are called, as well as the asp20boost package itself and the package gamboostLSS containing comparison functions.
 
@@
 * add substantial effect in large model to make clear how effect is visible
+* try different stepsizes and mstops for other algorithm
+* MSE erklären
+
+In case of higher dimensional parameter vectors or idiosyncratic regressors \zi for scale, a look at the distribution of the linear location predictor \etamu, as well as the exponentiated scale predictor \etasigma is of help.
+
+```{r}
+# check model via linear predictors ---------------------------------------------------------
+# eta_mu_2 <- init$fitted_loc
+# eta_sigma_2 <- init$fitted_scale
+# 
+# hist(eta_mu_2, breaks = 25)
+# hist(exp(eta_sigma_2), breaks = 25)
+```
 
@@
 
-The goal here is to demonstrate, that the algorithm finds correct estimates for sensible choices of /beta and /gamma.
-By evaluating different parameter configurations it becomes clear what model properties they represent.
-We use the following four different parameter configurations to explore location-scale regression model characteristics:
+The goal in this first part is to demonstrate, that the algorithm finds correct estimates for sensible choices of \beta and \gamma.
+We use the following 5 differing parameter configurations to explore location-scale regression model characteristics.
 
@@
 # config 5 ----------------------
-beta_5 <- c(0,7)
-gamma_5 <- c(1,-1)
+beta_5 <- c(0,1)
+gamma_5 <- c(0.3,-0.3)
+
+# config 6 ----------------------
+beta_6 <- c(0,15)
+gamma_6 <- c(1,-1)
 ```
@@
 
-We call the helper-function to create a model object of 1000 observations and plot it with a second helper function.
+[insert model config by formula]
+
+This parameter configuration is equivalent to the example case of the asp20model package.
+A model object of 1000 observations is created and the model is plotted by employing the helper functions mentioned above.
 
 ```{r}
-mod_1 <- init_model(beta_1, gamma_1, 1000)
+mod_1 <- init_model(beta_1, gamma_1, 1000)$model
 plot_model(beta_1, gamma_1, 1000)
@@
 
-We have already seen the basic plot structure, visualizing a location-scale regression model.
-What’s added is the depiction of two important quantities: Indicated by the blue arrow, one is the location slope /beta_1.
-The red arrow, in contrasts, indicates [modeled?] standard deviation of the observed responses at the point x=1.
-The ratio of those two quantities is essential for possibility of estimation.
-Here, we have a good signal-noise ratio, which enables the gradient_boost() function to estimate it.
+As seen above the range of response values, as well as the mean increase as the single covariate x increases.
+The black dashed line here, however, does not represent the 95% confidence interval as is usually the case, but rather the standard deviation of the reponse at the respective covariate value.
+The lower dashed line represents one standard deviation departing from the expected response in negative direction.
+The upper dashed line, in turn, in positive direction.
+Hence the interval of the dashed lines span over two standard deviations.
+
+The red arrow indicates the standard deviation at the point x = 1 - that is the upper limit of covariate values.
+We designed covariate values to be on the unit interval for convenience.
+For cases with increasing scale, the red arrow hence indicates maximum standard deviation of the response.
+
+The blue arrow depicts the slope of the location parameter vector \beta1
+
+The ratio of those two quantities is essential for estimability of the model.
+Here the blue arrow is dominant, hence the model may well be estimated.
+We will see the opposite case in the following subpart.
 
@@
 
-We see that the estimates are close to the true model parameters.
+We see that the estimates are close to the true model parameters - for \beta as well as for \gamma.
 
@@
 
-With all parameters set to 2, the algorithm yields NaNs.
+Configuration 2 sets all parameters set to value 2. Estimating this model with the gradient_boost() function does not yield valid results.
 
 ```{r}
-mod_2 <- init_model(beta_2, gamma_2, 1000)
+init <- init_model(beta_2, gamma_2, 1000)
+mod_2 <- init$model
 gradient_boost(mod_2)
-calc_MSE(beta_2, gamma_2, mod_2$beta, mod_2$gamma)
+mod_2$beta
+mod_2$gamma
 ```
 
-To understand the model characteristics of parameter configuration 2-2-2-2, we look at the model plot.
+To understand why this happens we need to understand the model characteristics resulting from parameter configuration 2-2-2-2.
 
@@
 
-Here, the red arrow is large, while the blue arrow is hardly visible.
-This represents the fact, that gamma parameters lead to SD being immensly high at x=1 and hence Untergehen of location effect.
-The model is not estimable.
-This is due to the fact that sd(x=1) = exp(gamma_0 + gamma_1 * 1) = exp(4) = 54.60.
-The location slope is 2.
-It is impossible to estimate an effect of 2 (change of two unit if x increases by one), if the random noise of the response is already very likely to be +/- 55 at the point x=1.
-Non-linearity of SDs makes this comparison hard to do intuitively.
-However, with the exponential function in mind one can abschätzen where sds may lie given parameter configuration.
-In contrast, location predictions are easily imagined by simply seeing beta_0 and beta_1.
-For x=0 the vlue is equal to the intercept, for x=1 the value is the sum of both betas.
-X values restricted to 0,1-range make it even easier to imagine preditors since the highest SD of interest is at x=1, where the predictor may be easily calculated by adding gamma_0 and gamma_1.
+Here, the red arrow is of large size, while the blue arrow is hardly visible.
+This is due to the presence of huge standard deviations in response \y, concealing possible changes in expectation of \y.
+In fact, the standard deviation at x = 1 is:
+
+```{r}
+# sd(x = 1) = exp(2 + 2 * 1) = 
+exp(gamma_2[1] + gamma_2[2] * 1)
+```
+The blue arrow indicates change of expected response y due to a one unit increase in the regressor x, which is \beta1 = 2.
+The estimable signal perishes, due to the high amount of noise caused by the high gamma-parameters.
+Resulting is a signal-noise ration, that is impossible to estimate.
+Values for parameters gamma need to be chosen sensibly for simulating data.
+This is no straight-forward task, since the effect of \gamma on the response standard deviation is not linear, but exponential.
+To check in the simple case, the model plot containing arrows are of help.
 
@@
 
+[insert parameter configuration as formula]
+
 ```{r}
-mod_3 <- init_model(beta_3, gamma_3, 1000)
+mod_3 <- init_model(beta_3, gamma_3, 1000)$model
 plot_model(beta_3, gamma_3, 1000)
@@
 
-The algorithm performs well with estimates close to true parameters.To quantify how close estimates are to true parameters, we quantify this distance employing the mean squared error.
+The algorithm performs well with estimates close to true parameters.
+In order quantify proximity of estimated model to true data generating closeness we employ the mean squared error.
+This enables a comparison of model fit for configuration 1 and 3.
 
 ```{r}
-calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
+mse_1 <- calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
+mse_1
+```
+
+```{r}
+mse_3 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
+mse_3
+```
+
+Model 3 yields a slightly better fit in terms of MSE as model 1, differing only at the 4. decimal place. Realtively, it improved by about 9 percent.
+
+```{r}
+round((mse_3$total - mse_1$total) / mse_1$total * 100, digits = 2) #percent
 ```
@@
 
-```{r}
-mod_4 <- init_model(beta_4, gamma_4, 1000)
-plot_model(beta_4, gamma_4, 1000)
-```
-
-Configuration 4 involves a declining SD, which leads to the other end of the extreme - extremely low sd, leading to NaN problems (numerical underflow).
+[include configuration as formulas]
 
 ```{r}
+mod_4 <- init_model(beta_4, gamma_4, 1000)$model
 gradient_boost(mod_4)
@@
 
-The alogirthm also yields NaNs.
+Configuration 4, again, is not estimable by the gradient_boost() function. A look at the model plot helps also in this case.
+
+```{r}
+plot_model(beta_4, gamma_4, 1000)
+```
+
+Here SDs are considerably high for x = 0, namely exp(3) = 20.1 however decreasing fast until reaching a value of exp(3 - 3) = 1 at x = 1.
+With a \beta slope of 1, the algorithm runs into numerical problems yielding NaNs as results.
 
@@
 
+[insert model configuration as formula]
+
+Changing the gamma parameters by factor 0.1 changes model characteristics considerably.
+Signal-noise ratio is well balanced in this case.
+The algorithm is able to detect the true parameter configuration.
+
 ```{r}
-mod_5 <- init_model(beta_5, gamma_5, 1000)
+mod_5 <- init_model(beta_5, gamma_5, 1000)$model
 plot_model(beta_5, gamma_5, 1000)
 ```
-
-Configuration 5 involves a declining SD.
 
@@
 
+Estimation errors are present, especially for slope estimates \gamma.
 This time, the algorithm behaves poorly. The estimates are rather far from the true parameters. This, in turn, leads to high MSE.
@@
 
+### Evaluation Configuration 6
+
+```{r}
+mod_6 <- init_model(beta_6, gamma_6, 1000)$model
+plot_model(beta_6, gamma_6, 1000)
+```
+
+```{r}
+gradient_boost(mod_6)
+round(mod_6$beta, 3)
+round(mod_6$gamma, 3)
+```
+
+For very high signal-noise ratios, the algorithm performs poorly if negative scale slopes a involved in the model.
+
 ### Conclusions
@@
 * The specified model must have a signal-noise ratio, that permits estimation.
-* Gamma-slope has to be positive. (Here, the package may be further developed.)
+* Gamma-slopes need to be chosen sensibly, especially if multiple slopes are used.
+* Negative gamma slopes lead to poorly behaving algortihm.
 
@@
 # initialize model-----------------------------------------------
-result <- init_large_model()
-mod_1 <- result$model
-X <- result$design
-y <- result$response
-fitted_loc <- result$fitted_response
-fitted_scale <- result$fitted_sd
+init <- init_large_model()
+mod_1 <- init$model
+X <- init$design
+y <- init$response
+fitted_loc <- init$fitted_response
+fitted_scale <- init$fitted_sd
 
@@
 # check model via linear predictors ---------------------------------------------------------
-fitted_loc
-fitted_scale
-
+par(mfrow = c(2,1))
+hist(fitted_loc, breaks = 25)
+hist(fitted_scale, breaks = 25)
 ```
@@
 ```{r}
-gradient_boost(mod_1, maxit = 1000, componentwise = TRUE, verbose = F)
+#gradient_boost(mod_1, maxit = 1000, componentwise = TRUE, verbose = F)
 
@@
 
-[...]
+In order to assess our algorithm we compare it to the existing package gamboostLSS. It is the common way in R to adress location-scale regression models by the boosting method.
+
+### gamboostLSS- package
 
@@
 
+Description of package
+
+In this package we do not first create a model and then boost it.
+Creation and boosting is performed in one call, already yielding estimates.
+These estimates are, however not beta and gamma, but rather eta_mu and eta_gamma.
+So these unit-wise linear predictors need to be regressed on the covariate in order to arrive at the estimates of interest.
+
+As the package contains more functionalities than ours, we have to specify a control variable, that matches stepsize, and mstop to our package to allow for comparison.
+
+```{r}
+# set controls of package in order to make it comparable with ours
+ctrl <- boost_control(trace = FALSE,
+                      mstop = c(mu = 1000, sigma = 1000),
+                      nu = c(mu = 0.01, sigma = 0.001)
+)
+```
+
+### Evaluation Configuration 1
+
+We recap the first model configuration:
+
+```{r}
+init <- init_model(beta_1, gamma_1, 1000)
+mod_1 <- init$model
+response_1 <- init$response
+covariate_1 <- init$covariate
+plot_model(beta_1, gamma_1, 1000)
+```
+
+```{r}
+gradient_boost(mod_1)
+mod_1$beta
+mod_1$gamma
+```
+
+```{r}
+# estimate with external package --------
+
+mod_11 <- gamboostLSS(formula = response_1 ~ covariate_1,
+                      control = ctrl)
+eta_mu <- fitted(mod_11)$mu
+eta_sigma <- fitted(mod_11)$sigma
+
+lm(eta_mu ~ covariate_1)
+lm(eta_sigma ~ covariate_1)
+
+```
+
 [...]
+
+### Evaluation Configuration 3
+
+```{r}
+init <- init_model(beta_3, gamma_3, 1000)
+mod_3 <- init$model
+response_3 <- init$response
+covariate_3 <- init$covariate
+plot_model(beta_3, gamma_3, 1000)
+```
+[...]
+
+```{r}
+# estimate with external package --------
+
+mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3,
+                      control = ctrl)
+eta_mu <- fitted(mod_33)$mu
+eta_sigma <- fitted(mod_33)$sigma
+
+lm(eta_mu ~ covariate_3)
+lm(eta_sigma ~ covariate_3)
+
+```
+
+[...]
+```{r}
+gradient_boost(mod_3)
+round(mod_3$beta, 3)
+round(mod_3$gamma, 3)
+```
+
+```{r}
+microbenchmark(
+  gradient_boost(mod_3)
+)
+```
+
+```{r}
+microbenchmark(
+  mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3, control = ctrl),
+  eta_mu <- fitted(mod_33)$mu,
+  eta_sigma <- fitted(mod_33)$sigma,
+  lm(eta_mu ~ covariate_3),
+  lm(eta_sigma ~ covariate_3)
+)
+```
+
+Ensuring comparability is a difficult task, since the two packages are implemented in rather different ways.
+Running gradient_boost takes about 657000 microseconds (0.66 seconds).
+In contrasts, running the external algorithm gamboostLSS takes 2.829.493 microseconds (2.8 seconds), which is a multiple of our algorithm.
+Additionally, beta and gamma parameters are not yielded directly, but need to be determined by regression as well.
+In contrats, this happens directly in our algorithm.
+
+To be fair we evaluate construction of the model object as well.
+
+```{r}
+microbenchmark(
+  mod_1 <- LocationScaleRegressionBoost$new(response_3 ~ covariate_3, ~ covariate_3)
+)
+```
+
+This changes the situation only marginally, as this is a quite fast step.

### 2020-09-14 R/asp20boost.R
add some unit tests and minor fixes to roxygen

@@ -4,8 +4,8 @@
 #' specification, and a function `gradient_boost` for boosting the specified model object.
 #'
 #' @usage
-#' example_model <- LocationScaleRegressionBoost(formula, ...)
-#' gradient_boost(example_model, ...)
+#' model <- LocationScaleRegressionBoost(formula, ...)
+#' gradient_boost(model, ...)
 #'
 #' @details
 #' This package resulted from a series of student projects. Thus, it is part of a

### 2020-09-14 R/model.R

@@ -8,7 +8,7 @@
 #' page of that class *Note: It would be nice to just link the vignette*)
 #'
 #' @usage
-#' example_model <- LocationScaleRegression$new(formula, ...)
+#' example_model <- LocationScaleRegressionBoost$new(formula, ...)
 #'
 #' @details
 #' It assumes a normally distributed response variable with
@@ -132,7 +132,7 @@ LocationScaleRegressionBoost <- R6Class(
     #' @return a numeric vector nx1
     #' @examples
     #' y <- rnorm(30)
-    #' model <- LocationScaleRegression$new(y ~ 1)
+    #' model <- LocationScaleRegressionBoost$new(y ~ 1)
     #' model$gradients_loglik_mu()
 
     gradients_loglik_mu = function() {
@@ -150,13 +150,13 @@ LocationScaleRegressionBoost <- R6Class(
 
 
     # helper functions for boosting conventional <-> compwise ----------------------------------
-    #' a function using the gradient estimators to update parameters beta and gamma
+    # a function using the gradient estimators to update parameters beta and gamma
     #' @return a numeric vector nx1
     update_parameters_conventional = function() {
       self$beta <- self$beta + self$stepsize_beta * private$gradient_estimators_mu()
       self$gamma <- self$gamma + self$stepsize_gamma * private$gradient_estimators_sigma()
     }, #possible naming: update parameters conventional
-    #' a function using the gradient estimators to update parameters beta and gamma
+    # a function using the gradient estimators to update parameters beta and gamma
     #' @return a numeric vector nx1
     update_parameters_compwise = function() {
 

### 2020-09-14 tests/testthat/test_boosting_alg.R

@@ -1,22 +1,38 @@
-# context("functionality boosting algorithm")
-#
-# source("helper_model_class_creation.R")
-# # creates one simple and one large model
-#
-# # for ad-hoc loading use: source("tests/testthat/helper_model_class_creation.R")
-#
-#
-# # estimate models
-# gradient_boost(model_simple, stepsize = 0.01, maxit = 10000, abstol = 0.0001, verbose = TRUE)
-# gradient_boost(model_large, stepsize = 0.001, maxit = 2000, abstol = 0.0001, verbose = FALSE)
-#
-#
-# test_that("beta estimates are correct in simple model",{
-#   model_1 <- model_simple$clone()
-#   expect_equal(model_1$beta, c(0, 1))
-# })
-#
-# test_that("gamma estimates are correct in simple model",{
-#   model_1 <- model_simple$clone()
-#   expect_equal(model_1$gamma, c(0, 1))
-# })
+context("functionality boosting algorithm")
+
+source("helper_test_boosting.R")
+# creates one simple boosting model
+
+# for ad-hoc loading use: source("tests/testthat/helper_model_class_creation.R")
+
+
+test_that("length of covariates equals beta parameters",{
+  model_test <- model$clone()
+  expect_equal(length(model_test$beta), 3)
+})
+
+test_that("length of covariates equals gamma parameters",{
+  model_test <- model$clone()
+  expect_equal(length(model_test$gamma), 3)
+})
+
+test_that("model beta step size equals boosting parameter",{
+  model_test <- model$clone()
+  expect_equal(model_test$stepsize_beta, 0.01)
+})
+
+test_that("model gamma step size equals boosting parameter",{
+  model_test <- model$clone()
+  expect_equal(model_test$stepsize_gamma, 0.1)
+})
+
+test_that("beta values are finite values",{
+  model_test <- model$clone()
+  expect_equal(all(is.finite(model_test$beta)==TRUE), TRUE)
+})
+test_that("gama values are finite values",{
+  model_test <- model$clone()
+  expect_equal(all(is.finite(model_test$gamma)==TRUE), TRUE)
+})
+
+

### 2020-09-15 simulation_studies/draft_sim_study_13_sep.Rmd
further developed sim studies

@@
 library(microbenchmark)
+set.seed(1337)
 ```
@@
 
-Here SDs are considerably high for x = 0, namely exp(3) = 20.1 however decreasing fast until reaching a value of exp(3 - 3) = 1 at x = 1.
-With a \beta slope of 1, the algorithm runs into numerical problems yielding NaNs as results.
+Here SDs are considerably high for x = 0, namely exp(3) = 20.1, however decreasing fast until reaching a value of exp(3 - 3) = 1 at x = 1.
+With a beta slope of 1, the algorithm runs into numerical problems yielding NaNs as results.
 
@@
 Signal-noise ratio is well balanced in this case.
-The algorithm is able to detect the true parameter configuration.
+The algorithm is able to detect the true parameters considerably close.
 
@@
 
-Estimation errors are present, especially for slope estimates \gamma.
-This time, the algorithm behaves poorly. The estimates are rather far from the true parameters. This, in turn, leads to high MSE.
+### Evaluation Configuration 6
 
-```{r}
-calc_MSE(beta_5, gamma_5, mod_5$beta, mod_5$gamma)
-```
-
-### Evaluation Configuration 6
+For robustness, we check a configuration with high location slope and comparably low gamma parameters, resulting in a very good signal-noise ratio.
+This induces the expectation of good estimates, as seen in configuration 3.
 
@@
 
-For very high signal-noise ratios, the algorithm performs poorly if negative scale slopes a involved in the model.
+It turns out that the algorithm performs very poorly. Is does not run into numerical problems, but it yields estimate far away from the true parameter values, leading to high MSE.
+
+```{r}
+calc_MSE(beta_6, gamma_6, mod_6$beta, mod_6$gamma)
+```
+
+This poor behaviour remains without explanation for now.
 
@@
 
-* The specified model must have a signal-noise ratio, that permits estimation.
-* Gamma-slopes need to be chosen sensibly, especially if multiple slopes are used.
-* Negative gamma slopes lead to poorly behaving algortihm.
+The observed 4 parameter cases let us conclude that not every arbitrary parameter configuration is estimable by the implemented gradient_boost() algorithm.
+A first insight is that signal-noise ration needs to be substantially high in order to prevent that location effect gets concealed by inflated standard deviations.
+Second, we observe poor behaviour if negative scale slopes are involved.
+Since we have no explanation for this behaviour we exclude these in our further assessment.
 
@@
 
-[...]
+For models with many regressors componentwise boosting allows to perform variable selection.
+Plotting such a model yields only limited overview, since each regressor needs to be plotted seperately.
+However, the behaviour of response location dependent on the first four components may be observed in the following.
 
@@
 # initialize model-----------------------------------------------
-init <- init_large_model()
+init <- init_large_model(beta_range = c(20,50), gamma_range = c(2, 3))
 mod_1 <- init$model
@@
 y <- init$response
-fitted_loc <- init$fitted_response
-fitted_scale <- init$fitted_sd
 
 # check model via plotting --------------------------------------
-par(mfrow = c(4,3))
-for(i in 1:11) plot(X[,i], y, ylim = c(0,200))
+par(mfrow = c(2,2))
+for(i in 1:4) plot(X[,i], y, ylim = c(0,200))
 
@@
 
-No effect becomes clearly visible, since each individually is too small to have a substantial effect on the response.
+No effect is clearly visible, since each individually is too small to have a substantial effect on the response. However, a slight upwards slope may be observed.
+
+A more suitable method to check for signal-noise balance is to observe the linear location predictors, as well as the exponentiated scale predictors, determining the response standard deviation.
 
@@
 par(mfrow = c(2,1))
+fitted_loc <- init$fitted_response
+fitted_scale <- init$fitted_sd
 hist(fitted_loc, breaks = 25)
@@
 
-Model estimation currently does not work.
+A balanced signal-noise ratio  is assured, as expected response spread at about 60, while response SD spreads at 0.05. This yields an estimable model.
+
+However, lastly there occured errors in the functionality of componentwise boosting. The alogrithm happens to update only intercepts, leading to no variable being chosen. No components become updated.
 
 ```{r}
-#gradient_boost(mod_1, maxit = 1000, componentwise = TRUE, verbose = F)
+gradient_boost(mod_1, maxit = 1000, componentwise = TRUE, verbose = F)
 
 ```
-Resulting parameter estimates are wrong. Almost no component become updated.
+
 ```{r}
@@
 ```
-
-[...]
 

### 2020-09-15 simulation_studies/#simulation_variable_selection.R

@@ -1,12 +1,10 @@
 # This script analyses if variable selection in the case of multidemnsional
 # beta and gamma vectors works
 # For this reason some coefficients are 0.
-library(R6)
-library(asp20model)
 library(asp20boost)
 
-beta_vec <- c(1,0,1,0,1,0)
-gamma_vec <- c(0,1,0,1,0,1)
+beta_vec <- c(1,0,10,0,10,0)
+gamma_vec <- c(0,0.1,0,0.1,0,0.1)
 
 
 # beta_2 <- c(0,1,0,1,0,1)
@@ -41,9 +39,11 @@ y <- beta_vec[1] + beta_vec[2] * X[, 1] + beta_vec[3] * X[, 2] +  beta_vec[4] * 
 model <- LocationScaleRegressionBoost$new(
   y ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5],
   ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5])
-gradient_boost(model, stepsize = 0.001, maxit = 1000, abstol = 0.0001, verbose = FALSE)
+gradient_boost(model, stepsize = c(0.1, 0.01), maxit = 1000, verbose = TRUE, componentwise=T)
 
 
 model$beta
 model$gamma
 
+beta_vec
+gamma_vec

### 2020-09-15 simulation_studies/helper_init_large_model.R

@@ -1,23 +1,26 @@
 init_large_model <- function(beta_range = c(10, 20),
-                             gamma_range = c(0.1, 0.3)
+                             gamma_range = c(0.1, 0.3),
+                             dim_param = 10,
+                             int_gamma = -3
 ){
 
 
   # sample coefficients ------------------------------------------------------------
-  beta <- round(runif(11, min = 10, max = 20), digits = 0)
-  gamma <- round(runif(11, min = 0.1, max = 0.3), digits = 1)
+  beta <- round(runif(dim_param + 1, min = beta_range[1], max = beta_range[2]), digits = 0)
+  gamma <- round(runif(dim_param + 1, min = gamma_range[1], max = gamma_range[2]), digits = 1)
 
-  beta[seq(3,11,3)] <- 0 # set every third beta coef to 0
-  gamma[seq(2,10,2)] <- 0 # set every second gamma coef to 0
-  beta[1] <- 0 # set intercept to 0 for convenience
-  gamma[1] <- -3 # set gamma intercept to keep sds in an estimable range
+  beta[seq(2, dim_param + 1, 3)] <- 0 # set every third beta coef to 0
+  beta[seq(3, dim_param + 1, 6)] <- beta[seq(2, dim_param + 1, 6)] * (-1) # set every third beta coef to 0
+  gamma[seq(2,dim_param + 1, 2)] <- 0 # set every second gamma coef to 0
+  #beta[1] <- 0 # set intercept to 0 for convenience
+  gamma[1] <- int_gamma # set gamma intercept to keep sds in an estimable range
   #beta[2] <- 280 # set one very high effect for checking]
 
 
   # sample covariates --------------------------------------------------------------
   # we sample 100 covariates. intercept is automatically added by model class
-  covariates_X <- runif(100 * 11) # generate 100 obs each for 100 covariates
-  covariates_Z <- runif(100 * 11)
+  covariates_X <- runif(100 * (dim_param+1))
+  covariates_Z <- runif(100 * (dim_param+1))
 
   X <- matrix(covariates_X, nrow = 100, byrow = TRUE)
   X[, 1] <- 1
@@ -34,8 +37,8 @@ init_large_model <- function(beta_range = c(10, 20),
     x_i <- X[i, ]
     eta_mu[i] <- x_i %*% beta
     eta_sigma[i] <- x_i %*% gamma
-    y[i] <- rnorm(n = 1,
[…]

### 2020-09-15 simulation_studies/variable_selection_2.R

@@ -1,0 +1,24 @@
+source("simulation_studies/helper_init_large_model.R")
+
+
+# init_1 <- init_large_model(c(5,10), c(0.5,1), 5, -0.2) #works quite
+# init_2 <- init_large_model(c(5,10), c(0.5,1), 10, -1) #bad performance
+# init_3 <- init_large_model(c(3,6), c(0.2,0.8), 10, -0.2) #works quite
+init <- init_large_model(c(5,10), c(0.01,0.08), 50, 0)
+
+mod_1 <- init$model
+init$fitted_response
+init$fitted_sd
+
+ init$beta
+init$gamma
+
+
+
+gradient_boost(mod_1, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE, verbose = TRUE)
+
+init$beta
+mod_1$beta
+
+init$gamma
+mod_1$gamma

### 2020-09-15 simulation_studies/draft_sim_study_13_sep.Rmd
updated simulation studies

@@
 source("helper_init_large_model.R")
-library(asp20boost)
+library(asp20boost)K
 library(gamboostLSS)
@@
 
-As seen above the range of response values, as well as the mean increase as the single covariate x increases.
-The black dashed line here, however, does not represent the 95% confidence interval as is usually the case, but rather the standard deviation of the reponse at the respective covariate value.
-The lower dashed line represents one standard deviation departing from the expected response in negative direction.
-The upper dashed line, in turn, in positive direction.
+As seen above the range of response values, as well as the mean increase as the covariate x increases.
+The black dashed line here, however, does not represent the 95% confidence interval as is usually the case, but rather the standard deviation of the response at the respective covariate value.
+The lower dashed line represents one standard deviation departing from the expected response in negative direction - the upper dashed line, in turn, in positive direction.
 Hence the interval of the dashed lines span over two standard deviations.
 
-The red arrow indicates the standard deviation at the point x = 1 - that is the upper limit of covariate values.
+The red arrow indicates the standard deviation at the point x = 1,  that is the upper limit of covariate values.
 We designed covariate values to be on the unit interval for convenience.
@@
 
-The blue arrow depicts the slope of the location parameter vector \beta1
+The blue arrow depicts the slope of the location parameter vector *beta_1.*
 
@@
 
-We see that the estimates are close to the true model parameters - for \beta as well as for \gamma.
+We see that the estimates are close to the true model parameters - for *beta* as well as for *gamma.*
 
@@
 Signal-noise ratio is well balanced in this case.
-The algorithm is able to detect the true parameters considerably close.
+The algorithm is able to detect the true parameters considerably close, however not satisfactorily.
 
@@
 The observed 4 parameter cases let us conclude that not every arbitrary parameter configuration is estimable by the implemented gradient_boost() algorithm.
-A first insight is that signal-noise ration needs to be substantially high in order to prevent that location effect gets concealed by inflated standard deviations.
+A first insight is that signal-noise ratio needs to be substantially high in order to prevent that location effect gets concealed by inflated standard deviations.
 Second, we observe poor behaviour if negative scale slopes are involved.
@@
 
+### Setup 1
+
 For models with many regressors componentwise boosting allows to perform variable selection.
-Plotting such a model yields only limited overview, since each regressor needs to be plotted seperately.
-However, the behaviour of response location dependent on the first four components may be observed in the following.
 
@@
 # initialize model-----------------------------------------------
-init <- init_large_model(beta_range = c(20,50), gamma_range = c(2, 3))
-mod_1 <- init$model
-X <- init$design
-y <- init$response
+init_1 <- init_large_model(beta_range = c(5,10), 
+                           gamma_range = c(0.5,1), 
+                           dim_param = 5, 
+                           int_gamma = -0.2,
+                           n = 100)
+mod_1 <- init_1$model
 
 # check model via plotting --------------------------------------
-par(mfrow = c(2,2))
-for(i in 1:4) plot(X[,i], y, ylim = c(0,200))
+# par(mfrow = c(2,2))
+# for(i in 1:4) plot(X[,i], y, ylim = c(0,200))
 
@@
 
-No effect is clearly visible, since each individually is too small to have a substantial effect on the response. However, a slight upwards slope may be observed.
-
-A more suitable method to check for signal-noise balance is to observe the linear location predictors, as well as the exponentiated scale predictors, determining the response standard deviation.
+Checking the constructed response is crucial: is the signal-noise ratio appropriate?
+Unfortunately, this can not be done easily by plotting - as in the 2-dimensional case.
+In the multidimensional case it helps to look at the n = 100 unit-wise location and scale predictors.
+Mean and sd of these 100 values help getting an overview.
 
 ```{r}
-# check model via linear predictors ---------------------------------------------------------
-par(mfrow = c(2,1))
-fitted_loc <- init$fitted_response
-fitted_scale <- init$fitted_sd
-hist(fitted_loc, breaks = 25)
-hist(fitted_scale, breaks = 25)
+mean(init_1$fitted_response)
+sd(init_1$fitted_response)
+mean(init_1$fitted_sd)
+sd(init_1$fitted_sd)
 ```
 
-A balanced signal-noise ratio  is assured, as expected response spread at about 60, while response SD spreads at 0.05. This yields an estimable model.
-
-However, lastly there occured errors in the functionality of componentwise boosting. The alogrithm happens to update only intercepts, leading to no variable being chosen. No components become updated.
+Histograms also may help.
 
 ```{r}
-gradient_boost(mod_1, maxit = 1000, componentwise = TRUE, verbose = F)
+par(mfrow=c(2,1))
+hist(init_1$fitted_response, breaks = 25)
+hist(init_1$fitted_sd, breaks = 25)
+```
 
+A balanced signal-noise ratio  is assured, as expected response values span the range of 8 to 22 (varaiteion 14), while response SD cover values of 1 to 4.
+A variantion amount of 14 is towards a mixumu sd of 4, while way more sds are at about 2.5 and lower.
+This should result in estimable effect sizes.
+We have a look at the randomly drawn parameter values for beta and gamma:
+```{r}
+init_1$beta
+init_1$gamma
+```
+
+And now boost and find the non-zeros.
+
+```{r}
+gradient_boost(mod_1, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+
+mod_1$beta
+mod_1$gamma
+```
+
+For location, the right parameters are chosen, however not at the proper values.
+For scale too many parameters become chosen.
+This may be improved by setting less maxit.
+However this may lead to move away from the yet quite optimal choice for *beta*.
+It could also be improved by lowering stepsize for gamma.
+This results in a pattern where the latter entries of gamma are chosen exactly wrong.
+We will observe this pattern also in the next examples and have no explanation.
+
+```{r}
+mod_1$beta <- rep(0,6)
+mod_1$gamma <- rep(0,6)
+gradient_boost(mod_1, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
+
+mod_1$beta
+mod_1$gamma
+```
+
+### Setup 2
+
+We scale dimension of the parameter vectors up to 10.
+
+```{r}
+init_2 <- init_large_model(c(3,6), c(0.2,0.8), 10, -0.2, n = 100)
+mod_2 <- init_2$model
 ```
@@
 ```{r}
-mod_1$beta
-mod_1$gamma
+par(mfrow=c(2,1))
+hist(init_2$fitted_response, breaks = 25)
+hist(init_2$fitted_sd, breaks = 25)
 ```
+
+```{r}
+gradient_boost(mod_2, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+
+cbind(true = init_2$beta, estimate = round(mod_2$beta,2))
+cbind(true = init_2$gamma, estimate = round(mod_2$gamma,2))
+```
+
+For beta the algo chose 1 too few.
+For gamma entirs 1-3 are chosen correctly. Then the algorithm yields errors. Entry 4 is chosen, while in fact it is zero. Entry 5 is not chosen, while in the true model an effect exists.
+
+### Setup 2
+
+Scale dimensions further up
+
+```{r}
+# init_3 <- init_large_model(c(5,10), c(0.01,0.08), 50, 0, n = 100)
+# mod_3 <- init_3$model
+```
+
+```{r}
+par(mfrow=c(2,1))
+hist(init_3$fitted_response, breaks = 25)
+hist(init_3$fitted_sd, breaks = 25)
+```
+
+```{r}
+#gradient_boost(mod_3, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+
+cbind(true_beta = init_3$beta, estimate_beta = round(mod_3$beta,2), 
+      true_gamma = init_3$gamma, estimate_gamma = round(mod_3$gamma,2))
+```
+
+This problem occurs if we scale parameter dimensions up: only intercepts become updated.
+No covariates are chosen to have an effect on the response.
+If algorithm progress is checked with verbose=TRUE, one sees that intercepts inflate in the very first step, and then the proceedsing iterations work on lowering the intercepts, which takes place very slowly due to stepsize parameters.
+
+Possibly a sensible choice of initial parameter values may help. However we did not try this.
 

### 2020-09-15 simulation_studies/helper_init_large_model.R

@@ -1,7 +1,8 @@
 init_large_model <- function(beta_range = c(10, 20),
                              gamma_range = c(0.1, 0.3),
                              dim_param = 10,
-                             int_gamma = -3
+                             int_gamma = -3,
+                             n = 100
 ){
 
 
@@ -19,21 +20,21 @@ init_large_model <- function(beta_range = c(10, 20),
 
   # sample covariates --------------------------------------------------------------
   # we sample 100 covariates. intercept is automatically added by model class
-  covariates_X <- runif(100 * (dim_param+1))
-  covariates_Z <- runif(100 * (dim_param+1))
+  covariates_X <- runif(n * (dim_param+1))
+  covariates_Z <- runif(n * (dim_param+1))
 
-  X <- matrix(covariates_X, nrow = 100, byrow = TRUE)
+  X <- matrix(covariates_X, nrow = n, byrow = TRUE)
   X[, 1] <- 1
-  Z <- matrix(covariates_Z, nrow = 100, byrow = TRUE)
+  Z <- matrix(covariates_Z, nrow = n, byrow = TRUE)
   Z[, 1] <- 1
 
 
   # sample response ----------------------------------------------------------------
-  y <- rep(NA, times = 100)
-  eta_mu <- rep(NA, 100)
-  eta_sigma <- rep(NA, 100)
+  y <- rep(NA, times = n)
+  eta_mu <- rep(NA, n)
+  eta_sigma <- rep(NA, n)
 
-  for(i in 1:100) {
+  for(i in 1:n) {
     x_i <- X[i, ]
     eta_mu[i] <- x_i %*% beta
     eta_sigma[i] <- x_i %*% gamma

### 2020-09-15 simulation_studies/variable_selection_2.R

@@ -4,13 +4,13 @@ source("simulation_studies/helper_init_large_model.R")
 # init_1 <- init_large_model(c(5,10), c(0.5,1), 5, -0.2) #works quite
 # init_2 <- init_large_model(c(5,10), c(0.5,1), 10, -1) #bad performance
 # init_3 <- init_large_model(c(3,6), c(0.2,0.8), 10, -0.2) #works quite
-init <- init_large_model(c(5,10), c(0.01,0.08), 50, 0)
+init <- init_large_model(c(5,10), c(0.01,0.08), 50, 0, n = 1000)
 
 mod_1 <- init$model
 init$fitted_response
 init$fitted_sd
 
- init$beta
+init$beta
 init$gamma
 
 

### 2020-09-15 R/model.R
add comments and update software implementation in term paper

@@ -35,6 +35,7 @@ LocationScaleRegressionBoost <- R6Class(
 
   private = list(
 
+    #helper variables to improve computation time
     X_HAT = numeric(),
     X_HAT_componentwise = numeric(),
     X_number_of_columns = numeric(),
@@ -45,41 +46,31 @@ LocationScaleRegressionBoost <- R6Class(
 
 
     # estimators (bj_hat) for location and scale,
-    # conventional <-> componentwise --------------------------------------------------------
-
+    # conventional boosting --------------------------------------------------------
     gradient_estimators_mu = function() {
       private$X_HAT %*% crossprod(private$X, self$gradients_loglik_mu())
     },
     gradient_estimators_sigma = function() {
       private$Z_HAT %*% crossprod(private$Z, self$gradients_loglik_sigma())
     },
-
+    # componentwise boosting --------------------------------------------------------
     gradient_estimators_mu_compwise = function() {
       number_of_cols <- ncol(private$X)
       result_list <- rep(NA, number_of_cols)
       for(i in 1:number_of_cols){
-        #result_list[i] <- chol2inv(chol(crossprod(private$X[, i], private$X[, i]))) %*% crossprod(private$X[, i], self$gradients_loglik_mu())
         result_list[i] <- private$X_HAT_componentwise[i] %*% crossprod(private$X[, i], self$gradients_loglik_mu())
-
        }
       return(result_list)
     },
-
     gradient_estimators_sigma_compwise = function() {
       number_of_cols <- ncol(private$Z)
       result_list <- rep(NA, number_of_cols)
       for(i in 1:number_of_cols){
-        #result_list[i] <- chol2inv(chol(crossprod(private$Z[, i], private$Z[, i]))) %*% crossprod(private$Z[, i], self$gradients_loglik_sigma())
         result_list[i] <- private$Z_HAT_componentwise[i] %*% crossprod(private$Z[, i], self$gradients_loglik_sigma())
-
       }
       return(result_list)
     }
-
-
-
-
-  ),
+),
 
   public = list(
 
@@ -126,7 +117,7 @@ LocationScaleRegressionBoost <- R6Class(
       private$Z %*% self$gamma
     },
 
-    # gradients of the log-likelihood wrt eta_mu / eta_sigma.
+    # gradients of the log-likelihood wrt eta_mu
     # these gradients are estimated and then stepwisely added
     #' @details Returns the unit-wise gradients of the loss function, which is the negative loglikelihood.
[…]

### 2020-09-15 R/asp20boost.R
Hilfe Page

@@ -1,21 +1,25 @@
-#' Gradient Boosting for Location-Scale Regression Models
+#' Gradient Boosting for Location-Scale Regression Models ANO ANO
 #'
 #' The package provides an R6-class `LocationScaleRegressionBoost` for model
-#' specification, and a function `gradient_boost` for boosting the specified model object.
+#' specification, and a function `gradient_boost` for boosting the specified
+#' model object.
 #'
 #' @usage
 #' model <- LocationScaleRegressionBoost(formula, ...)
 #' gradient_boost(model, ...)
 #'
 #' @details
-#' This package resulted from a series of student projects. Thus, it is part of a
-#' bundle of packages all starting with `asp20...`. The overall goal of these packages
-#' is to estimate location-scale regression models, while each project/package employs
-#' different methods. All packages, including `asp20boost` build on the central
-#' package `asp20model`. Here, an R6-class is specified, which is inherited and expanded
-#' in the `asp20boost`-package.
+#' This package resulted from a series of student projects in the context of the
+#' seminar *Advanced Statistical Programming 2020* of the University of
+#' Göttingen. Thus, it is part of a bundle of the `asp20__` packages. The main
+#' goal of these projects is implement a method to estimate location-scale
+#' regression models. Each project's package implements a different method. All
+#' packages, including `asp20boost`, build upon the `asp20model` package.
+#' Specifically, an R6-class with inherited properties from `asp20model` is
+#' `asp20model` expanded in the `asp20boost`-package.
 #'
-#' @author Johannes Strauß, Levin Wiebelt, Sebastiàn Aristizabal
+#'
+#' @author Johannes Strauß, Levin Wiebelt, Sebastián Aristizábal
 #'
 #' @import R6
 #' @import asp20model

### 2020-09-15 R/asp20boost.R
Chapter 4 (or 5? IDK anymore)

@@ -1,4 +1,4 @@
-#' Gradient Boosting for Location-Scale Regression Models ANO ANO
+#' Gradient Boosting for Location-Scale Regression Models HAI HAI
 #'
 #' The package provides an R6-class `LocationScaleRegressionBoost` for model
 #' specification, and a function `gradient_boost` for boosting the specified

### 2020-09-15 simulation_studies/draft_sim_study_13_sep.Rmd
2-dim case sim study done

@@
 
-In order ensure legibility we employ some helper functions for model initialization, plotting, etc. These and the packages are loaded in the follwing chunk.
+In order ensure legibility we employ some helper functions for model initialization, plotting, etc. These and the packages are loaded in the following chunk.
 
@@
 source("helper_init_large_model.R")
-library(asp20boost)K
+library(asp20boost)
 library(gamboostLSS)
@@
 
+Second, we observe poor behaviour if negative scale slopes are involved. Since we have no explanation for this behaviour we exclude these in our further assessment.
+
 ```{r}
@@
 
-## Small Model - 4 Parameter Case
+## Small Model - Two-dimnesional Case
 
 The goal in this first part is to demonstrate, that the algorithm finds correct estimates for sensible choices of \beta and \gamma.
-We use the following 5 differing parameter configurations to explore location-scale regression model characteristics.
+We use the following 4 different parameter configurations to explore location-scale regression model characteristics.
 
@@
 # config 3 ----------------------
-beta_3 <- c(0,7)
-gamma_3 <- c(-3,3)
+beta_3 <- c(0,1)
+gamma_3 <- c(-1,-0.3)
 
 # config 4 ----------------------
-beta_4 <- c(0,1)
-gamma_4 <- c(3,-3)
+beta_4 <- c(0,15)
+gamma_4 <- c(1,-1)
 
-# config 5 ----------------------
-beta_5 <- c(0,1)
-gamma_5 <- c(0.3,-0.3)
+# config 7 ----------------------
+# beta_7 <- c(0,1)
+# gamma_7 <- c(3,-3)
+```
 
-# config 6 ----------------------
-beta_6 <- c(0,15)
-gamma_6 <- c(1,-1)
+```{r}
+mod_1 <- init_model(beta_1, gamma_1, 1000)$model
+gradient_boost(mod_1)
+mod_2 <- init_model(beta_2, gamma_2, 1000)$model
+gradient_boost(mod_2)
+mod_3 <- init_model(beta_3, gamma_3, 1000)$model
+gradient_boost(mod_3)
+mod_4 <- init_model(beta_4, gamma_4, 1000)$model
+gradient_boost(mod_4)
+```
+
+```{r, cache=TRUE, fig.height = 8, fig.width = 11}
+par(mfrow = c(2,2))
+plot_model(beta_1, gamma_1, 1000, "Config 1")
+plot_model(beta_2, gamma_2, 1000, "Config 2")
+plot_model(beta_3, gamma_3, 1000, "Config 3")
+plot_model(beta_4, gamma_4, 1000, "Config 4")
+```
+```{r}
+result_df <- data.frame("Config" = c(1,1,2,2,3,3,4,4),
+                        "beta_true" = c(beta_1, beta_2, beta_3, beta_4),
+                        "beta est." = c(round(mod_1$beta, 3), round(mod_2$beta, 3), round(mod_3$beta, 3), round(mod_4$beta, 3)),
+                        "gamma_true" = c(gamma_1, gamma_2, gamma_3, gamma_4),
+                        "gamma est." = c(round(mod_1$gamma, 3), round(mod_2$gamma, 3), round(mod_3$gamma, 3), round(mod_4$gamma, 3))
+                        )
+result_df
 ```
@@
 
-This parameter configuration is equivalent to the example case of the asp20model package.
-A model object of 1000 observations is created and the model is plotted by employing the helper functions mentioned above.
-
-```{r}
-mod_1 <- init_model(beta_1, gamma_1, 1000)$model
-plot_model(beta_1, gamma_1, 1000)
-```
-
-As seen above the range of response values, as well as the mean increase as the covariate x increases.
-The black dashed line here, however, does not represent the 95% confidence interval as is usually the case, but rather the standard deviation of the response at the respective covariate value.
-The lower dashed line represents one standard deviation departing from the expected response in negative direction - the upper dashed line, in turn, in positive direction.
-Hence the interval of the dashed lines span over two standard deviations.
-
-The red arrow indicates the standard deviation at the point x = 1,  that is the upper limit of covariate values.
-We designed covariate values to be on the unit interval for convenience.
+This parameter configuration is equivalent to the example case of the *asp20model* package. 
+As already mentioned, heteroskedasticity is present and indicated by the funnel shape of the point cloud.
+The solid black line represents the true response mean, while the dashed lines represent the true upwards and downwards standard deviations from this mean.
+The red arrow indicates the standard deviation at the point x = 1,  that is the upper limit of covariate values, since covariates are design to be on the unit interval.
 For cases with increasing scale, the red arrow hence indicates maximum standard deviation of the response.
-
 The blue arrow depicts the slope of the location parameter vector *beta_1.*
-
 The ratio of those two quantities is essential for estimability of the model. 
 Here the blue arrow is dominant, hence the model may well be estimated. 
-We will see the opposite case in the following subpart.
-
-```{r}
-gradient_boost(mod_1)
-round(mod_1$beta, 3)
-round(mod_1$gamma, 3)
-```
-
-We see that the estimates are close to the true model parameters - for *beta* as well as for *gamma.*
+Estimates are displayed in *table[...]* above. We see that both *betahat* and *gammahat* estimates are close to the true parameter vectors *beta* and *gamma.*
 
@@
 
-Configuration 2 sets all parameters set to value 2. Estimating this model with the gradient_boost() function does not yield valid results.
-
-```{r}
-init <- init_model(beta_2, gamma_2, 1000)
-mod_2 <- init$model
-gradient_boost(mod_2)
-mod_2$beta
-mod_2$gamma
-```
-
+Configuration 2 is a somewhat naive parameter configuration, where all parameters are set to value 2. 
+Estimating this model does not yield valid results.
+The gradient_boost() function runs into numerical problems.
 To understand why this happens we need to understand the model characteristics resulting from parameter configuration 2-2-2-2.
-
-```{r}
-plot_model(beta_2, gamma_2, 1000)
-```
-
 Here, the red arrow is of large size, while the blue arrow is hardly visible. 
-This is due to the presence of huge standard deviations in response \y, concealing possible changes in expectation of \y.
+This is due to the presence of huge standard deviations in response *y*, concealing changes in expectation of *y.* 
 In fact, the standard deviation at x = 1 is:
@@
 ```{r}
-# sd(x = 1) = exp(2 + 2 * 1) = 
 exp(gamma_2[1] + gamma_2[2] * 1)
 ```
-The blue arrow indicates change of expected response y due to a one unit increase in the regressor x, which is \beta1 = 2.
+In contrast the response expectation values range from 0 to 2.
+The blue arrow indicates change of expected response *y* due to a one unit increase in the regressor *x*, which is *beta1* = 2. 
 The estimable signal perishes, due to the high amount of noise caused by the high gamma-parameters. 
@@
 Values for parameters gamma need to be chosen sensibly for simulating data. 
-This is no straight-forward task, since the effect of \gamma on the response standard deviation is not linear, but exponential.
-To check in the simple case, the model plot containing arrows are of help.
+This is no straight-forward task, since the effect of *gamma* on the response standard deviation is not linear, but exponential. 
+To check this in the two-dimensional case, the arrows in the model plot are of help.
+In the following part handling multidimensional cases, we'll introduce another form of checking.
 
@@
 
+Configuration 3, in turn, yields reasonable estimates.
+It contains a negative gamma slope, leading to decreasing response standard deviations.
+This can be observed by the funnel shape pointing in the opposite direction as before.
+
+It may be of interest to check if estimates are fit better (if they lie closer to the true parameters) than e.g. in configuration 1.
+To do this we measure the gap between true values and estimates by the mean squared error.
+This enables comparison of configurations 1 and 3.
+
 ```{r}
-mod_3 <- init_model(beta_3, gamma_3, 1000)$model
-plot_model(beta_3, gamma_3, 1000)
+mse_1 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
+mse_3 <- calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3)
 ```
 
-Configuration 3 yields a good signal-noise ratio.
-
-```{r}
-gradient_boost(mod_3)
-round(mod_3$beta, 3)
-round(mod_3$gamma, 3)
-```
-
-The algorithm performs well with estimates close to true parameters.
-In order quantify proximity of estimated model to true data generating closeness we employ the mean squared error.
-This enables a comparison of model fit for configuration 1 and 3.
-
-```{r}
-mse_1 <- calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
-mse_1
-```
-
-```{r}
-mse_3 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
-mse_3
-```
-
-Model 3 yields a slightly better fit in terms of MSE as model 1, differing only at the 4. decimal place. Realtively, it improved by about 9 percent.
-
-```{r}
-round((mse_3$total - mse_1$total) / mse_1$total * 100, digits = 2) #percent
-```
+Configuration 3 leads to smaller MSE, especially for scale estimates *gamma_hat*.
+We conclude that our alogrithm performs better in configuration 3.
 
@@
 
+In configuration 4, we again included a negative gamma slope, as well as a very high beta slope.
+The idea is to ensure a very good signal noise ratio, in order to ensure good model fit, indicated by low MSE.
+Contrary to our expectations, the algorithm performs poorly for this parameter configuration.
+
 ```{r}
-mod_4 <- init_model(beta_4, gamma_4, 1000)$model
-gradient_boost(mod_4)
-round(mod_4$beta, 3)
-round(mod_4$gamma, 3)
+mse_4 <- calc_MSE(beta_4, gamma_4, mod_4$beta, mod_4$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3,
+      "config 4" = mse_4)
 ```
 
-Configuration 4, again, is not estimable by the gradient_boost() function. A look at the model plot helps also in this case.
-
-```{r}
-plot_model(beta_4, gamma_4, 1000)
-```
-
-Here SDs are considerably high for x = 0, namely exp(3) = 20.1, however decreasing fast until reaching a value of exp(3 - 3) = 1 at x = 1.
-With a beta slope of 1, the algorithm runs into numerical problems yielding NaNs as results.
-
-#### Evaluation Configuration 5
-
-[insert model configuration as formula]
-
-Changing the gamma parameters by factor 0.1 changes model characteristics considerably.
-Signal-noise ratio is well balanced in this case.
-The algorithm is able to detect the true parameters considerably close, however not satisfactorily.
-
-```{r}
-mod_5 <- init_model(beta_5, gamma_5, 1000)$model
-plot_model(beta_5, gamma_5, 1000)
-```
-
-```{r}
-gradient_boost(mod_5)
-round(mod_5$beta, 3)
-round(mod_5$gamma, 3)
-```
-
-### Evaluation Configuration 6
-
-For robustness, we check a configuration with high location slope and comparably low gamma parameters, resulting in a very good signal-noise ratio.
-This induces the expectation of good estimates, as seen in configuration 3.
-
-```{r}
-mod_6 <- init_model(beta_6, gamma_6, 1000)$model
-plot_model(beta_6, gamma_6, 1000)
-```
-
-```{r}
-gradient_boost(mod_6)
-round(mod_6$beta, 3)
-round(mod_6$gamma, 3)
-```
-
-It turns out that the algorithm performs very poorly. Is does not run into numerical problems, but it yields estimate far away from the true parameter values, leading to high MSE.
-
-```{r}
-calc_MSE(beta_6, gamma_6, mod_6$beta, mod_6$gamma)
-```
-
-This poor behaviour remains without explanation for now.
+The bad model fit is due to the fact, that the location slope is exceptionally underestimated.
 
@@
 
-The observed 4 parameter cases let us conclude that not every arbitrary parameter configuration is estimable by the implemented gradient_boost() algorithm.
-A first insight is that signal-noise ratio needs to be substantially high in order to prevent that location effect gets concealed by inflated standard deviations.
-Second, we observe poor behaviour if negative scale slopes are involved.
-Since we have no explanation for this behaviour we exclude these in our further assessment.
+The observed two-dimensional cases let us conclude that not every arbitrary parameter configuration is estimable by the implemented gradient_boost() algorithm.
+A first insight is that signal-noise ratio needs to be sufficiently high in order to prevent that location effect gets concealed by inflated standard deviations.
 
-## Large Model - Variable Selection
+## Large Model - Variable Selection in the multidimensional case
+
+### ToDo
+
+This may be improved by setting less maxit.
+However this may lead to move away from the yet quite optimal choice for *beta*.
+It could also be improved by lowering stepsize for gamma.
+This results in a pattern where the latter entries of gamma are chosen exactly wrong.
+We will observe this pattern also in the next examples and have no explanation.
 
@@
 For models with many regressors componentwise boosting allows to perform variable selection.
+The helper function *init_large_model* creates sample data.
+The parameter vectors beta and gamma are of size *dim_param* and contain some zeros.
+The non-zero elements are assigned random values in the specified ranges.
 
@@
 mod_1 <- init_1$model
-
-# check model via plotting --------------------------------------
-# par(mfrow = c(2,2))
-# for(i in 1:4) plot(X[,i], y, ylim = c(0,200))
-
+cbind(beta_true = init_1$beta, gamma_true = init_1$gamma)
 ```
 
-Checking the constructed response is crucial: is the signal-noise ratio appropriate?
-Unfortunately, this can not be done easily by plotting - as in the 2-dimensional case.
-In the multidimensional case it helps to look at the n = 100 unit-wise location and scale predictors.
-Mean and sd of these 100 values help getting an overview.
+In *mod_1* beta and gamma ranges are chosen in order to create a model that is estimable.
+In the multidimensional it is again important to create data with a balanced signal-noise ratio.
+If gamma parameters add up to high values, the resulting simulated response has huge standard deviation, which conceals the location effects.
+
+In the twodimensional case we used plots to check this ratio.
+Now, we have a look at the unit-wise location predictors *eta_mu*, as well as the exponentiated scale predictors *exp(eta_sigma)*, which determine the response sd.
 
 ```{r}
-mean(init_1$fitted_response)
-sd(init_1$fitted_response)
-mean(init_1$fitted_sd)
-sd(init_1$fitted_sd)
+cbind("eta_mu" = rbind(mean = mean(init_1$fitted_response), var = var(init_1$fitted_response)), 
+      "exp_eta_sigma" = rbind(mean = mean(init_1$fitted_sd), var = var(init_1$fitted_sd)))
 ```
 
-Histograms also may help.
+Plotting the empirical distributions of *eta_mu* and *exp_eta_sigma* as histograms gives a visual and more detailed overview of the location-scale regression model characteristics given the assigned parameters *beta* and *gamma*
 
@@
 par(mfrow=c(2,1))
-hist(init_1$fitted_response, breaks = 25)
-hist(init_1$fitted_sd, breaks = 25)
+hist(init_1$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_1$fitted_sd, breaks = 25, main = "exp_eta_sigma")
 ```
 
-A balanced signal-noise ratio  is assured, as expected response values span the range of 8 to 22 (varaiteion 14), while response SD cover values of 1 to 4.
-A variantion amount of 14 is towards a mixumu sd of 4, while way more sds are at about 2.5 and lower.
-This should result in estimable effect sizes.
-We have a look at the randomly drawn parameter values for beta and gamma:
+Location predictors take values between 8 and 22, peaking at 15.
+In contrast, the highest response-SD is 4, while most values are at about 2.5 and lower.
+The resulting response variation should be small enough to not conceal the location effects, as well as big enough to be estimated.
+
+Performing the boosting alogrithm by calling the gradient_boost() function leads to the follwing results.
+
 ```{r}
-init_1$beta
-init_1$gamma
+gradient_boost(mod_1, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
+
+cbind(true = init_1$beta, estimate = round(mod_1$beta,2))
+cbind(true = init_1$gamma, estimate = round(mod_1$gamma,2))
 ```
 
-And now boost and find the non-zeros.
+The algorithm depicts correctly which effects on response location exists, that is it finds the non-zero elements of *beta*.
+For scale estimate *gamma_hat* too many parameters become chosen, since two zero-elements of true parameter *gamma* are assigned non-zero values by the estimate *gamma_hat*.
+
+The values estimated differ considerably from the true parameter values. 
+The resulting MSE is higher than in the four parameter case.
 
 ```{r}
-gradient_boost(mod_1, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
-
-mod_1$beta
-mod_1$gamma
-```
-
-For location, the right parameters are chosen, however not at the proper values.
-For scale too many parameters become chosen.
-This may be improved by setting less maxit.
-However this may lead to move away from the yet quite optimal choice for *beta*.
-It could also be improved by lowering stepsize for gamma.
-This results in a pattern where the latter entries of gamma are chosen exactly wrong.
-We will observe this pattern also in the next examples and have no explanation.
-
-```{r}
-mod_1$beta <- rep(0,6)
-mod_1$gamma <- rep(0,6)
-gradient_boost(mod_1, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
-
-mod_1$beta
-mod_1$gamma
+calc_MSE(init_1$beta, init_1$gamma, mod_1$beta, mod_1$gamma)
 ```
@@
 mod_2 <- init_2$model
+cbind(beta_true = init_2$beta, gamma_true = init_2$gamma)
 ```
@@
 par(mfrow=c(2,1))
-hist(init_2$fitted_response, breaks = 25)
-hist(init_2$fitted_sd, breaks = 25)
+hist(init_2$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_2$fitted_sd, breaks = 25, main = "exp_eta_sigma")
 ```
@@
 
-### Setup 2
+### Setup 3
 
@@
 
-In order to assess our algorithm we compare it to the existing package gamboostLSS. It is the common way in R to adress location-scale regression models by the boosting method.
+In order to assess our algorithm we compare it to the existing package *gamboostLSS.* Therefore we use configurations one and three of the two-dimensional cases
 
 ### gamboostLSS- package
+
+In the *gamboostLSS* package boosting a location-scale regression model works as follows.
+
+```{r}
+init <- init_model(beta_1, gamma_1, 1000)
+mod_1 <- init$model
+response_1 <- init$response
+covariate_1 <- init$covariate
+
+```
 
@@
 
-### Evaluation Configuration 1
+### Recap Configuration 1
 
 We recap the first model configuration:
-
-```{r}
-init <- init_model(beta_1, gamma_1, 1000)
-mod_1 <- init$model
-response_1 <- init$response
-covariate_1 <- init$covariate
-plot_model(beta_1, gamma_1, 1000)
-```
 
@@
 
-### Evaluation Configuration 3
+### Recap Configuration 3
 

### 2020-09-15 simulation_studies/helper_plot_model.R

@@ -1,5 +1,5 @@
 # set up main plotting function ----------------------------------
-plot_model <- function(beta, gamma, n) {
+plot_model <- function(beta, gamma, n, plot_title) {
 
   # set up linear predictor functions -------------------------------
   eta_mu <- function(x) beta[1] + beta[2] * x
@@ -10,7 +10,7 @@ plot_model <- function(beta, gamma, n) {
   y <- eta_mu(x) + rnorm(n, sd = exp(eta_sigma(x)))
 
   # configure plot --------------------------------------------------
-  plot(x, y, col = "grey", cex = 0.5, pch = 16)
+  plot(x, y, col = "grey", cex = 0.5, pch = 16, main = plot_title)
   abline(beta[1], beta[2], lwd = 2, col = 1)
   curve(eta_mu(x) + 1 * exp(eta_sigma(x)), -0.1, 1.1, add = TRUE, col = 1, lty = 3, lwd = 2)
   curve(eta_mu(x) - 1 * exp(eta_sigma(x)), -0.1, 1.1, add = TRUE, col = 1, lty = 3, lwd = 2)
@@ -25,7 +25,7 @@ plot_model <- function(beta, gamma, n) {
          y1 = beta[1],
          length = 0.05, code = 3, col = 4, lty = 1, lwd = 2)
 
-  legend("topleft", legend=c("location slope", "standard deviation at x=1"),
-         col=c(4, 2), lty=c(1,1), cex=0.8)
+  #legend("topleft", legend=c("location slope", "standard deviation at x=1"),
+  #       col=c(4, 2), lty=c(1,1), cex=0.8)
 
 }

### 2020-09-15 R/model.R
- Work on Roxygen Comments - First draft Chapter 6

@@ -1,11 +1,10 @@
 #' R6 Class for Specifying Location-Scale Regression Models
 #'
 #' @description
-#' Model obejcts created by this class may be estimated by the `gradient_boost` function.
+#' Model objects created by this class may be estimated using the `gradient_boost` function.
 #' This model class builds upon and thus inherits the basic structure
-#' from the `LocationScaleRegression` model class of the `asp20model` package.
-#' Consequently, it shares similar syntax and structure. (Please  see the help
-#' page of that class *Note: It would be nice to just link the vignette*)
+#' from the `LocationScaleRegression` model class of the \pkg{asp20model} package
+#' thus sharing a similar syntax and structure.
 #'
 #' @usage
 #' example_model <- LocationScaleRegressionBoost$new(formula, ...)

### 2020-09-15 R/model.R
update to help page

@@ -20,10 +20,10 @@
 #' which is the negative loglikelihood.
 #' @field gradients_loglik_sigma Caclulates the unit-wise gradients of the loss-functions,
 #' which is the negative loglikelihood.
- #' @field update_parameters_conventional ccc
-#' @field update_parameters_compwise ddd
-#' @field par_log eee
-#' @field plot fff
+#' @field update_parameters_conventional Adds the gradients estimators to the active fields $\bm{\beta}$ and $\bm{\gamma}$ multiplied by the defined stepsize for $\bm{\beta} $and $\bm{\gamma}$.
+#' @field update_parameters_compwise Adds update vector to the active fields $\bm{\beta}$ and $\bm{\gamma by first performing the loss improvement calculation and then building an update vector  consisting of zeros with exception of the best fitting component gradient multiplied by the defined stepsize.
+#' @field par_log Stores parameter values for each iteration to use it in the plotting function
+#' @field plot Plots the stored parameter values of each iteration by accessing the par_log public field
 #'
 # #' @inheritSection asp20model::LocationScaleRegression Section description
 #'

### 2020-09-15 simulation_studies/sim_study_large_model.Rmd
updated sim_studies

--- 2020-09-15 simulation_studies/draft_sim_study_13_sep.Rmd
@@
 library(microbenchmark)
-set.seed(1337)
 ```
-
-## ToDo
-* add labels to beta gamma output
-* befor every subchapter "Evaluation Config.*", put a formula with the current parameter configuration that is being evaluated.
-* add substantial effect in large model to make clear how effect is visible
-* try different stepsizes and mstops for other algorithm
-* MSE erklären
-
-In case of higher dimensional parameter vectors or idiosyncratic regressors \zi for scale, a look at the distribution of the linear location predictor \etamu, as well as the exponentiated scale predictor \etasigma is of help.
-
-Second, we observe poor behaviour if negative scale slopes are involved. Since we have no explanation for this behaviour we exclude these in our further assessment.
-
-```{r}
-# check model via linear predictors ---------------------------------------------------------
-# eta_mu_2 <- init$fitted_loc
-# eta_sigma_2 <- init$fitted_scale
-# 
-# hist(eta_mu_2, breaks = 25)
-# hist(exp(eta_sigma_2), breaks = 25)
-```
-
-## Small Model - Two-dimnesional Case
-
-The goal in this first part is to demonstrate, that the algorithm finds correct estimates for sensible choices of \beta and \gamma.
-We use the following 4 different parameter configurations to explore location-scale regression model characteristics.
-
-```{r}
-# config 1 ----------------------
-beta_1 <- c(0,1)
-gamma_1 <- c(-3,2)
-
-# config 2 ----------------------
-beta_2 <- c(2,2)
-gamma_2 <- c(2,2)
-
-# config 3 ----------------------
-beta_3 <- c(0,1)
-gamma_3 <- c(-1,-0.3)
-
-# config 4 ----------------------
-beta_4 <- c(0,15)
-gamma_4 <- c(1,-1)
-
-# config 7 ----------------------
-# beta_7 <- c(0,1)
-# gamma_7 <- c(3,-3)
-```
-
-```{r}
-mod_1 <- init_model(beta_1, gamma_1, 1000)$model
-gradient_boost(mod_1)
-mod_2 <- init_model(beta_2, gamma_2, 1000)$model
-gradient_boost(mod_2)
-mod_3 <- init_model(beta_3, gamma_3, 1000)$model
-gradient_boost(mod_3)
-mod_4 <- init_model(beta_4, gamma_4, 1000)$model
-gradient_boost(mod_4)
-```
-
-```{r, cache=TRUE, fig.height = 8, fig.width = 11}
-par(mfrow = c(2,2))
-plot_model(beta_1, gamma_1, 1000, "Config 1")
-plot_model(beta_2, gamma_2, 1000, "Config 2")
-plot_model(beta_3, gamma_3, 1000, "Config 3")
-plot_model(beta_4, gamma_4, 1000, "Config 4")
-```
-```{r}
-result_df <- data.frame("Config" = c(1,1,2,2,3,3,4,4),
-                        "beta_true" = c(beta_1, beta_2, beta_3, beta_4),
-                        "beta est." = c(round(mod_1$beta, 3), round(mod_2$beta, 3), round(mod_3$beta, 3), round(mod_4$beta, 3)),
-                        "gamma_true" = c(gamma_1, gamma_2, gamma_3, gamma_4),
-                        "gamma est." = c(round(mod_1$gamma, 3), round(mod_2$gamma, 3), round(mod_3$gamma, 3), round(mod_4$gamma, 3))
-                        )
-result_df
-```
-
-### Evaluation Configuration 1
-
-[insert model config by formula]
-
-This parameter configuration is equivalent to the example case of the *asp20model* package. 
-As already mentioned, heteroskedasticity is present and indicated by the funnel shape of the point cloud.
-The solid black line represents the true response mean, while the dashed lines represent the true upwards and downwards standard deviations from this mean.
-The red arrow indicates the standard deviation at the point x = 1,  that is the upper limit of covariate values, since covariates are design to be on the unit interval.
-For cases with increasing scale, the red arrow hence indicates maximum standard deviation of the response.
-The blue arrow depicts the slope of the location parameter vector *beta_1.*
-The ratio of those two quantities is essential for estimability of the model. 
-Here the blue arrow is dominant, hence the model may well be estimated. 
-Estimates are displayed in *table[...]* above. We see that both *betahat* and *gammahat* estimates are close to the true parameter vectors *beta* and *gamma.*
-
-#### Evaluation Configuration 2
-
-Configuration 2 is a somewhat naive parameter configuration, where all parameters are set to value 2. 
-Estimating this model does not yield valid results.
-The gradient_boost() function runs into numerical problems.
-To understand why this happens we need to understand the model characteristics resulting from parameter configuration 2-2-2-2.
-Here, the red arrow is of large size, while the blue arrow is hardly visible. 
-This is due to the presence of huge standard deviations in response *y*, concealing changes in expectation of *y.* 
-In fact, the standard deviation at x = 1 is:
-
-```{r}
-exp(gamma_2[1] + gamma_2[2] * 1)
-```
-In contrast the response expectation values range from 0 to 2.
-The blue arrow indicates change of expected response *y* due to a one unit increase in the regressor *x*, which is *beta1* = 2. 
-The estimable signal perishes, due to the high amount of noise caused by the high gamma-parameters. 
-Resulting is a signal-noise ration, that is impossible to estimate. 
-Values for parameters gamma need to be chosen sensibly for simulating data. 
-This is no straight-forward task, since the effect of *gamma* on the response standard deviation is not linear, but exponential. 
-To check this in the two-dimensional case, the arrows in the model plot are of help.
-In the following part handling multidimensional cases, we'll introduce another form of checking.
-
-#### Evaluation Configuration 3
-
-[insert parameter configuration as formula]
-
-Configuration 3, in turn, yields reasonable estimates.
-It contains a negative gamma slope, leading to decreasing response standard deviations.
-This can be observed by the funnel shape pointing in the opposite direction as before.
-
-It may be of interest to check if estimates are fit better (if they lie closer to the true parameters) than e.g. in configuration 1.
-To do this we measure the gap between true values and estimates by the mean squared error.
-This enables comparison of configurations 1 and 3.
-
-```{r}
-mse_1 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
-mse_3 <- calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
-cbind("config 1" = mse_1,
-      "config_3" = mse_3)
-```
-
-Configuration 3 leads to smaller MSE, especially for scale estimates *gamma_hat*.
-We conclude that our alogrithm performs better in configuration 3.
-
-#### Evaluation Configuration 4
-
-[include configuration as formulas]
-
-In configuration 4, we again included a negative gamma slope, as well as a very high beta slope.
-The idea is to ensure a very good signal noise ratio, in order to ensure good model fit, indicated by low MSE.
-Contrary to our expectations, the algorithm performs poorly for this parameter configuration.
-
-```{r}
-mse_4 <- calc_MSE(beta_4, gamma_4, mod_4$beta, mod_4$gamma)
-cbind("config 1" = mse_1,
-      "config_3" = mse_3,
-      "config 4" = mse_4)
-```
-
-The bad model fit is due to the fact, that the location slope is exceptionally underestimated.
-
-### Conclusions
-
-The observed two-dimensional cases let us conclude that not every arbitrary parameter configuration is estimable by the implemented gradient_boost() algorithm.
-A first insight is that signal-noise ratio needs to be sufficiently high in order to prevent that location effect gets concealed by inflated standard deviations.
-
-## Large Model - Variable Selection in the multidimensional case
 
@@
 
-### Setup 1
+## Large Model - Variable Selection in the multidimensional case
 
-For models with many regressors componentwise boosting allows to perform variable selection.
-The helper function *init_large_model* creates sample data.
-The parameter vectors beta and gamma are of size *dim_param* and contain some zeros.
-The non-zero elements are assigned random values in the specified ranges.
+In this part we demonstrate how the gradient_boost() function performs variable selection.
+We have a look at three large model configurations.
+Configuration 5 contains 6-dimensional parameter vectors, oncluding the intercept terms.
+In configuration 6 *beta* and *gamma* are of length 11, while in configuration 7 they contain 51 components.
+The parameter vectors are created randomly, then some components are set to 0, resulting in the configurations displayed below.
+In order to ensure a good signal-noise ratio, there are restrictions to random parameter creation, which are explained below.
 
 ```{r}
-# initialize model-----------------------------------------------
-init_1 <- init_large_model(beta_range = c(5,10), 
+set.seed(1337)
+# initialize models --------------------------------------------------------------
+init_5 <- init_large_model(beta_range = c(5,10),
                            gamma_range = c(0.5,1),
@@
                            n = 100)
-mod_1 <- init_1$model
-cbind(beta_true = init_1$beta, gamma_true = init_1$gamma)
+init_6 <- init_large_model(beta_range = c(3,6),
+                           gamma_range = c(0.2,0.8),
+                           dim_param = 10,
+                           int_gamma = -0.2,
+                           n = 100)
+init_7 <- init_large_model(beta_range = c(5,10),
+                           gamma_range = c(0.01,0.08),
+                           dim_param = 50,
+                           int_gamma = 0,
+                           n = 100)
+mod_5 <- init_5$model
+mod_6 <- init_6$model
+mod_7 <- init_7$model
+
+# estimate models -----------------------------------------------------------------
+gradient_boost(mod_5, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
+gradient_boost(mod_6, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+gradient_boost(mod_7, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+
 ```
 
-In *mod_1* beta and gamma ranges are chosen in order to create a model that is estimable.
-In the multidimensional it is again important to create data with a balanced signal-noise ratio.
-If gamma parameters add up to high values, the resulting simulated response has huge standard deviation, which conceals the location effects.
-
-In the twodimensional case we used plots to check this ratio.
-Now, we have a look at the unit-wise location predictors *eta_mu*, as well as the exponentiated scale predictors *exp(eta_sigma)*, which determine the response sd.
+```{r}
+# Configuration 5
+cbind(beta_true = init_5$beta, 
+      beta_est = round(mod_5$beta, 3),
+      gamma_true = init_5$gamma,
+      gamma_est = round(mod_5$gamma, 3)
+)
+```
 
 ```{r}
-cbind("eta_mu" = rbind(mean = mean(init_1$fitted_response), var = var(init_1$fitted_response)), 
-      "exp_eta_sigma" = rbind(mean = mean(init_1$fitted_sd), var = var(init_1$fitted_sd)))
+# Configuration 6
+cbind(beta_true = init_6$beta, 
+      beta_est = round(mod_6$beta, 3),
+      gamma_true = init_6$gamma,
+      gamma_est = round(mod_6$gamma, 3)
+)
 ```
 
-Plotting the empirical distributions of *eta_mu* and *exp_eta_sigma* as histograms gives a visual and more detailed overview of the location-scale regression model characteristics given the assigned parameters *beta* and *gamma*
+### Configuration 5
+As mentioned above there are restrictions on parameter creation, since not every possible parameter configuration leads to estimable models, as we have seen in the two-dimensional case.
+In the multidimensional case, the problem remains: 
+If gamma parameters add up to high values, the resulting simulated response has huge standard deviation, which conceals the location effects.
+Checking for signal-noise ratio remains to be done in the multidimensional case as well.
+Plotting is hardly an option as dimensions increase.
+Therefore we check by another method, that is looking at the distributions of true response means and true response SDs.
 
@@
 par(mfrow=c(2,1))
-hist(init_1$fitted_response, breaks = 25, main = "eta_mu")
-hist(init_1$fitted_sd, breaks = 25, main = "exp_eta_sigma")
+hist(init_5$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_5$fitted_sd, breaks = 25, main = "exp_eta_sigma")
 ```
@@
 
-Performing the boosting alogrithm by calling the gradient_boost() function leads to the follwing results.
-
-```{r}
-gradient_boost(mod_1, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
-
-cbind(true = init_1$beta, estimate = round(mod_1$beta,2))
-cbind(true = init_1$gamma, estimate = round(mod_1$gamma,2))
-```
+In configuration 5 beta components are correctly selected.
+Zero-components are also zero in the estimate vector, while non-zero components are estimated as non-zero effects.
+The effect sizes differ considerably more than we observed in the two-dimensional case.
 
@@
 ```{r}
-calc_MSE(init_1$beta, init_1$gamma, mod_1$beta, mod_1$gamma)
+mse_5 <- calc_MSE(init_5$beta, init_5$gamma, mod_5$beta, mod_5$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3,
+      "config 4" = mse_4,
+      "config_5" = mse_5)
 ```
 
-### Setup 2
+### Configuration 6
 
-We scale dimension of the parameter vectors up to 10.
-
-```{r}
-init_2 <- init_large_model(c(3,6), c(0.2,0.8), 10, -0.2, n = 100)
-mod_2 <- init_2$model
-cbind(beta_true = init_2$beta, gamma_true = init_2$gamma)
-```
+For 10-dimensional configuration 6 we get the following mean and SD distributions.
 
@@
 par(mfrow=c(2,1))
-hist(init_2$fitted_response, breaks = 25, main = "eta_mu")
-hist(init_2$fitted_sd, breaks = 25, main = "exp_eta_sigma")
-```
-
-```{r}
-gradient_boost(mod_2, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
-
-cbind(true = init_2$beta, estimate = round(mod_2$beta,2))
-cbind(true = init_2$gamma, estimate = round(mod_2$gamma,2))
+hist(init_6$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_6$fitted_sd, breaks = 25, main = "exp_eta_sigma")
 ```
@@
 
-### Setup 3
+### Configuration 7
 
-Scale dimensions further up
-
-```{r}
-# init_3 <- init_large_model(c(5,10), c(0.01,0.08), 50, 0, n = 100)
-# mod_3 <- init_3$model
-```
+Scaling parameter dimension up to 51, the algortihm runs into problems.
+A solution may be to properly choose initial values.
 
@@
 par(mfrow=c(2,1))
-hist(init_3$fitted_response, breaks = 25)
-hist(init_3$fitted_sd, breaks = 25)
+hist(init_7$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_7$fitted_sd, breaks = 25, main = "exp_eta_sigma")
 ```
 
-```{r}
-#gradient_boost(mod_3, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
-
-cbind(true_beta = init_3$beta, estimate_beta = round(mod_3$beta,2), 
-      true_gamma = init_3$gamma, estimate_gamma = round(mod_3$gamma,2))
-```
-
+Even with a well-balanced signal-noise ratio as indicated by the histograms, problems occur.
 This problem occurs if we scale parameter dimensions up: only intercepts become updated. 
@@
 If algorithm progress is checked with verbose=TRUE, one sees that intercepts inflate in the very first step, and then the proceedsing iterations work on lowering the intercepts, which takes place very slowly due to stepsize parameters.
+Possibly a sensible choice of initial parameter values may help. However we did not try this.
 
-Possibly a sensible choice of initial parameter values may help. However we did not try this.
+```{r}
+cbind(beta_true = init_7$beta, 
+      beta_est = round(mod_7$beta, 3),
+      gamma_true = init_7$gamma,
+      gamma_est = round(mod_7$gamma, 3)
+)
+```
 
@@
 
-In order to assess our algorithm we compare it to the existing package *gamboostLSS.* Therefore we use configurations one and three of the two-dimensional cases
+In order to assess performance of our algorithm further we compare it to the existing package *gamboostLSS.* 
+Therefore we use configurations 1 and 4 the two-dimensional cases.
+In the former case our algorithm performed well, in the latter rather bad.
 
@@
 In the *gamboostLSS* package boosting a location-scale regression model works as follows.
-
-```{r}
-init <- init_model(beta_1, gamma_1, 1000)
-mod_1 <- init$model
-response_1 <- init$response
-covariate_1 <- init$covariate
-
-```
-
-[...]
-
-Description of package
-
-In this package we do not first create a model and then boost it.
-Creation and boosting is performed in one call, already yielding estimates.
-These estimates are, however not beta and gamma, but rather eta_mu and eta_gamma.
-So these unit-wise linear predictors need to be regressed on the covariate in order to arrive at the estimates of interest.
-
+First we need to define the vector of controls.
 As the package contains more functionalities than ours, we have to specify a control variable, that matches stepsize, and mstop to our package to allow for comparison.
@@
 ```{r}
-# set controls of package in order to make it comparable with ours
 ctrl <- boost_control(trace = FALSE,
@@
 
-We recap the first model configuration:
+```{r}
+set.seed(1337)
+init_1 <- init_model(beta_1, gamma_1, 1000)$model
+mod_1 <- init_1$model
+response_1 <- init_1$response
+covariate_1 <- init_1$covariate
+gamboostLSS(formula = response_1 ~ covariate_1,
+                      control = ctrl)
+```
+
+What we get as result are not yet parameter estimates, but linear predictors.
+These must be regressed on the parameters to arrive at the desired estimates.
+In this package we do not first create a model and then boost it. Creation and boosting is performed in one call, already yielding estimates. 
+These estimates are, however not beta and gamma, but rather eta_mu and eta_gamma. 
+So these unit-wise linear predictors need to be regressed on the covariate in order to arrive at the estimates of interest.
 
 ```{r}
-gradient_boost(mod_1)
-mod_1$beta
-mod_1$gamma
-```
+eta_mu <- fitted(est_1)$mu
+eta_sigma <- fitted(est_1)$sigma
 
-```{r}
-# estimate with external package --------
-
-mod_11 <- gamboostLSS(formula = response_1 ~ covariate_1,
-                      control = ctrl)
-eta_mu <- fitted(mod_11)$mu
-eta_sigma <- fitted(mod_11)$sigma
-
-lm(eta_mu ~ covariate_1)
-lm(eta_sigma ~ covariate_1)
+est_beta <- lm(eta_mu ~ covariate_1)
+est_gamma <- lm(eta_sigma ~ covariate_1)
 
 ```
-
-[...]
 

### 2020-09-15 simulation_studies/sim_study_small_model.Rmd
updated sim_studies

--- 2020-09-15 simulation_studies/draft_sim_study_13_sep.Rmd
@@
 ---
-title: "Simulation Study"
+title: "R Notebook"
 output: html_notebook
 ---
-
-In this part we design and perform a simulation study in order to demonstrate functionality of the implemented boosting algorithm.
-For this, we first employ small models with two location and scale parameters each, and ensure basic functionality.
-Second, validity of the variable selection features is demonstrated.
-Lastly, we compare our algorithm with the one implemented in the gamboostLSS package.
-
-In order ensure legibility we employ some helper functions for model initialization, plotting, etc. These and the packages are loaded in the following chunk.
 
@@
 library(microbenchmark)
-set.seed(1337)
 ```
 
-## ToDo
-* add labels to beta gamma output
-* befor every subchapter "Evaluation Config.*", put a formula with the current parameter configuration that is being evaluated.
-* add substantial effect in large model to make clear how effect is visible
-* try different stepsizes and mstops for other algorithm
-* MSE erklären
-
-In case of higher dimensional parameter vectors or idiosyncratic regressors \zi for scale, a look at the distribution of the linear location predictor \etamu, as well as the exponentiated scale predictor \etasigma is of help.
-
-Second, we observe poor behaviour if negative scale slopes are involved. Since we have no explanation for this behaviour we exclude these in our further assessment.
-
-```{r}
-# check model via linear predictors ---------------------------------------------------------
-# eta_mu_2 <- init$fitted_loc
-# eta_sigma_2 <- init$fitted_scale
-# 
-# hist(eta_mu_2, breaks = 25)
-# hist(exp(eta_sigma_2), breaks = 25)
-```
+### ToDo Sebastián
+* Formula in front of every 'Configuration' paragraph
+* Make Data Frame look like a nice table
 
@@
 
-The goal in this first part is to demonstrate, that the algorithm finds correct estimates for sensible choices of \beta and \gamma.
-We use the following 4 different parameter configurations to explore location-scale regression model characteristics.
+The goal in this first part is to demonstrate, that the algorithm finds correct estimates for sensible choices of *beta* and *gamma.* We use the following 4 different parameter configurations to explore location-scale regression model characteristics.
 
@@
 gamma_4 <- c(1,-1)
-
-# config 7 ----------------------
-# beta_7 <- c(0,1)
-# gamma_7 <- c(3,-3)
 ```
@@
 ```{r}
+set.seed(1337)
 mod_1 <- init_model(beta_1, gamma_1, 1000)$model
@@
 Here the blue arrow is dominant, hence the model may well be estimated. 
-Estimates are displayed in *table[...]* above. We see that both *betahat* and *gammahat* estimates are close to the true parameter vectors *beta* and *gamma.*
+Estimates are displayed in the table above. 
+We see that both *betahat* and *gammahat* estimates are close to the true parameter vectors *beta* and *gamma.*
 
@@
 Here, the red arrow is of large size, while the blue arrow is hardly visible. 
-This is due to the presence of huge standard deviations in response *y*, concealing changes in expectation of *y.* 
+This is due to the presence of huge standard deviations in response *y*, concealing changes in mean of *y.* 
 In fact, the standard deviation at x = 1 is:
@@
 ```
-In contrast the response expectation values range from 0 to 2.
-The blue arrow indicates change of expected response *y* due to a one unit increase in the regressor *x*, which is *beta1* = 2. 
-The estimable signal perishes, due to the high amount of noise caused by the high gamma-parameters. 
-Resulting is a signal-noise ration, that is impossible to estimate. 
+In contrast the true response mean values range from 0 to 2.
+The change in response location (the signal), indicated by the blue arrow, perishes due to the high amount of noise caused by the high gamma-parameters. 
+Resulting is a signal-noise ratio, that is impossible to estimate. 
 Values for parameters gamma need to be chosen sensibly for simulating data. 
@@
 To check this in the two-dimensional case, the arrows in the model plot are of help.
-In the following part handling multidimensional cases, we'll introduce another form of checking.
 
@@
 
-It may be of interest to check if estimates are fit better (if they lie closer to the true parameters) than e.g. in configuration 1.
+It may be of interest to check if estimates lie closer to the true parameters than e.g. in configuration 1.
 To do this we measure the gap between true values and estimates by the mean squared error.
@@
 
-Configuration 3 leads to smaller MSE, especially for scale estimates *gamma_hat*.
-We conclude that our alogrithm performs better in configuration 3.
+Configuration 1 leads to smaller MSE, hance our algorithm performs better in configuration 1.
 
@@
 
-The bad model fit is due to the fact, that the location slope is exceptionally underestimated.
+The bad model fit is due to the fact, that the location slope is highly underestimated.
 
@@
 A first insight is that signal-noise ratio needs to be sufficiently high in order to prevent that location effect gets concealed by inflated standard deviations.
-
-## Large Model - Variable Selection in the multidimensional case
-
-### ToDo
-
-This may be improved by setting less maxit.
-However this may lead to move away from the yet quite optimal choice for *beta*.
-It could also be improved by lowering stepsize for gamma.
-This results in a pattern where the latter entries of gamma are chosen exactly wrong.
-We will observe this pattern also in the next examples and have no explanation.
-
-### Setup 1
-
-For models with many regressors componentwise boosting allows to perform variable selection.
-The helper function *init_large_model* creates sample data.
-The parameter vectors beta and gamma are of size *dim_param* and contain some zeros.
-The non-zero elements are assigned random values in the specified ranges.
-
-```{r}
-# initialize model-----------------------------------------------
-init_1 <- init_large_model(beta_range = c(5,10), 
-                           gamma_range = c(0.5,1), 
-                           dim_param = 5, 
-                           int_gamma = -0.2,
-                           n = 100)
-mod_1 <- init_1$model
-cbind(beta_true = init_1$beta, gamma_true = init_1$gamma)
-```
-
-In *mod_1* beta and gamma ranges are chosen in order to create a model that is estimable.
-In the multidimensional it is again important to create data with a balanced signal-noise ratio.
-If gamma parameters add up to high values, the resulting simulated response has huge standard deviation, which conceals the location effects.
-
-In the twodimensional case we used plots to check this ratio.
-Now, we have a look at the unit-wise location predictors *eta_mu*, as well as the exponentiated scale predictors *exp(eta_sigma)*, which determine the response sd.
-
-```{r}
-cbind("eta_mu" = rbind(mean = mean(init_1$fitted_response), var = var(init_1$fitted_response)), 
-      "exp_eta_sigma" = rbind(mean = mean(init_1$fitted_sd), var = var(init_1$fitted_sd)))
-```
-
-Plotting the empirical distributions of *eta_mu* and *exp_eta_sigma* as histograms gives a visual and more detailed overview of the location-scale regression model characteristics given the assigned parameters *beta* and *gamma*
-
-```{r}
-par(mfrow=c(2,1))
-hist(init_1$fitted_response, breaks = 25, main = "eta_mu")
-hist(init_1$fitted_sd, breaks = 25, main = "exp_eta_sigma")
-```
-
-Location predictors take values between 8 and 22, peaking at 15.
-In contrast, the highest response-SD is 4, while most values are at about 2.5 and lower.
-The resulting response variation should be small enough to not conceal the location effects, as well as big enough to be estimated.
-
-Performing the boosting alogrithm by calling the gradient_boost() function leads to the follwing results.
-
-```{r}
-gradient_boost(mod_1, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
-
-cbind(true = init_1$beta, estimate = round(mod_1$beta,2))
-cbind(true = init_1$gamma, estimate = round(mod_1$gamma,2))
-```
-
-The algorithm depicts correctly which effects on response location exists, that is it finds the non-zero elements of *beta*.
-For scale estimate *gamma_hat* too many parameters become chosen, since two zero-elements of true parameter *gamma* are assigned non-zero values by the estimate *gamma_hat*.
-
-The values estimated differ considerably from the true parameter values. 
-The resulting MSE is higher than in the four parameter case.
-
-```{r}
-calc_MSE(init_1$beta, init_1$gamma, mod_1$beta, mod_1$gamma)
-```
-
-### Setup 2
-
-We scale dimension of the parameter vectors up to 10.
-
-```{r}
-init_2 <- init_large_model(c(3,6), c(0.2,0.8), 10, -0.2, n = 100)
-mod_2 <- init_2$model
-cbind(beta_true = init_2$beta, gamma_true = init_2$gamma)
-```
-
-```{r}
-par(mfrow=c(2,1))
-hist(init_2$fitted_response, breaks = 25, main = "eta_mu")
-hist(init_2$fitted_sd, breaks = 25, main = "exp_eta_sigma")
-```
-
-```{r}
-gradient_boost(mod_2, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
-
-cbind(true = init_2$beta, estimate = round(mod_2$beta,2))
-cbind(true = init_2$gamma, estimate = round(mod_2$gamma,2))
-```
-
-For beta the algo chose 1 too few.
-For gamma entirs 1-3 are chosen correctly. Then the algorithm yields errors. Entry 4 is chosen, while in fact it is zero. Entry 5 is not chosen, while in the true model an effect exists.
-
-### Setup 3
-
-Scale dimensions further up
-
-```{r}
-# init_3 <- init_large_model(c(5,10), c(0.01,0.08), 50, 0, n = 100)
-# mod_3 <- init_3$model
-```
-
-```{r}
-par(mfrow=c(2,1))
-hist(init_3$fitted_response, breaks = 25)
-hist(init_3$fitted_sd, breaks = 25)
-```
-
-```{r}
-#gradient_boost(mod_3, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
-
-cbind(true_beta = init_3$beta, estimate_beta = round(mod_3$beta,2), 
-      true_gamma = init_3$gamma, estimate_gamma = round(mod_3$gamma,2))
-```
-
-This problem occurs if we scale parameter dimensions up: only intercepts become updated.
-No covariates are chosen to have an effect on the response.
-If algorithm progress is checked with verbose=TRUE, one sees that intercepts inflate in the very first step, and then the proceedsing iterations work on lowering the intercepts, which takes place very slowly due to stepsize parameters.
-
-Possibly a sensible choice of initial parameter values may help. However we did not try this.
-
-## Benchmark gamboostLSS
-
-In order to assess our algorithm we compare it to the existing package *gamboostLSS.* Therefore we use configurations one and three of the two-dimensional cases
-
-### gamboostLSS- package
-
-In the *gamboostLSS* package boosting a location-scale regression model works as follows.
-
-```{r}
-init <- init_model(beta_1, gamma_1, 1000)
-mod_1 <- init$model
-response_1 <- init$response
-covariate_1 <- init$covariate
-
-```
-
-[...]
-
-Description of package
-
-In this package we do not first create a model and then boost it.
-Creation and boosting is performed in one call, already yielding estimates.
-These estimates are, however not beta and gamma, but rather eta_mu and eta_gamma.
-So these unit-wise linear predictors need to be regressed on the covariate in order to arrive at the estimates of interest.
-
-As the package contains more functionalities than ours, we have to specify a control variable, that matches stepsize, and mstop to our package to allow for comparison.
-
-```{r}
-# set controls of package in order to make it comparable with ours
-ctrl <- boost_control(trace = FALSE,
-                      mstop = c(mu = 1000, sigma = 1000),
-                      nu = c(mu = 0.01, sigma = 0.001)
-)
-```
-
-### Recap Configuration 1
-
-We recap the first model configuration:
-
-```{r}
-gradient_boost(mod_1)
-mod_1$beta
-mod_1$gamma
-```
-
-```{r}
-# estimate with external package --------
-
-mod_11 <- gamboostLSS(formula = response_1 ~ covariate_1,
-                      control = ctrl)
-eta_mu <- fitted(mod_11)$mu
-eta_sigma <- fitted(mod_11)$sigma
-
-lm(eta_mu ~ covariate_1)
-lm(eta_sigma ~ covariate_1)
-
-```
-
-[...]
-
-### Recap Configuration 3
-
-```{r}
-init <- init_model(beta_3, gamma_3, 1000)
-mod_3 <- init$model
-response_3 <- init$response
-covariate_3 <- init$covariate
-plot_model(beta_3, gamma_3, 1000)
-```
-[...]
-
-```{r}
-# estimate with external package --------
-
-mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3,
-                      control = ctrl)
-eta_mu <- fitted(mod_33)$mu
-eta_sigma <- fitted(mod_33)$sigma
-
-lm(eta_mu ~ covariate_3)
-lm(eta_sigma ~ covariate_3)
-
-```
-
-[...]
-```{r}
-gradient_boost(mod_3)
-round(mod_3$beta, 3)
-round(mod_3$gamma, 3)
-```
-
-```{r}
-microbenchmark(
-  gradient_boost(mod_3)
-)
-```
-
-```{r}
-microbenchmark(
-  mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3, control = ctrl),
-  eta_mu <- fitted(mod_33)$mu,
-  eta_sigma <- fitted(mod_33)$sigma,
-  lm(eta_mu ~ covariate_3),
-  lm(eta_sigma ~ covariate_3)
-)
-```
-
-Ensuring comparability is a difficult task, since the two packages are implemented in rather different ways.
-Running gradient_boost takes about 657000 microseconds (0.66 seconds).
-In contrasts, running the external algorithm gamboostLSS takes 2.829.493 microseconds (2.8 seconds), which is a multiple of our algorithm.
-Additionally, beta and gamma parameters are not yielded directly, but need to be determined by regression as well.
-In contrats, this happens directly in our algorithm.
-
-To be fair we evaluate construction of the model object as well.
-
-```{r}
-microbenchmark(
-  mod_1 <- LocationScaleRegressionBoost$new(response_3 ~ covariate_3, ~ covariate_3)
-)
-```
-
-This changes the situation only marginally, as this is a quite fast step.
+For the finding of configuration 4 we have several explanation approaches:
+* There are further constraints on simulating data meaningfully which we were not yet able to work out.
+* Our algorithm does not reliably yield robust results if negative gamma slopes are included.
+* Our algortihm runs into problems if the signal is too strong compared to noise.

### 2020-09-15 simulation_studies/#not_working_simulation_plot_4_params.R

@@ -1,83 +1,0 @@
-# Since the alogrithm does not yet work properly for all possible models
-# this script analyses its functionality for 2-dim beta and gamma
-# also scale is regressed on the same predictor as location
-library(R6)
-library(asp20model)
-library(asp20boost)
-
-
-# in order to make this script work an extra step needs to be done:
-# the error is as follow: model creation inside the "boost_true_model"-function
-# below does not work, UNLESS a LocationScaleRegressionBoost-model class has been created before
-# for this reason a simple empty model call and estimation is performed here
-{ x <- runif(100)
-  y <- y <- x + rnorm(100, sd = exp(-3 + 2 * x))
-  model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
-  gradient_boost(model, stepsize = 0.001, maxit = 1000, abstol = 0.0001, verbose = FALSE)
-}
-
-
-
-boost_true_model <- function(beta_0, beta_1, gamma_0, gamma_1){
-
-  # simulate data -------------------------
-  set.seed(1337)
-  n <- 500
-  x <- runif(n)
-  y <- beta_0 + beta_1 * x + rnorm(
-    n,
-    sd = exp(gamma_0 + gamma_1 * x)
-  )
-
-  # estimate data by boosting ------------
-  model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
-  gradient_boost(model, stepsize = 0.001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
-
-
-  # plot results -------------------------
-  x_grid <- seq(0,1, length.out = 100)
-  y_grid <- beta_0 + beta_1 * x_grid
-
-  # limits for plot
-  ylim_upper <- model$beta[1] + model$beta[2] * 1 + 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
-  ylim_lower <- model$beta[1] + model$beta[2] * 1 - 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
-
-
-  # plot data points
-  plot(x, y, col = 2,
-       xlim = c(-0.3,1.3),
-       ylim = c(ylim_upper * 1.05, ylim_lower * 1.05))
-
-  # plot true model
-  lines(x_grid, y_grid)
[…]

### 2020-09-15 simulation_studies/#now_working_simulation_variable_selection.R

@@ -1,80 +1,0 @@
-# This script analyses if variable selection in the case of multidimensional
-# beta and gamma vectors works
-# For this reason some coefficients are 0.
-library(R6)
-library(asp20model)
-library(asp20boost)
-
-
-# in order to make this script work an extra step needs to be done:
-# the error is as follow: model creation inside the "boost_true_model"-function
-# below does not work, UNLESS a LocationScaleRegressionBoost-model class has been created before
-# for this reason a simple model call and estimation is performed here
-{ x <- runif(100)
-  y <- runif(100)
-  model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
-  gradient_boost(model, stepsize = 0.001, maxit = 1000, abstol = 0.0001, verbose = FALSE)
-}
-
-
-
-
-boost_true_model <- function(beta_vec, gamma_vec){
-
-  # simulate data -------------------------
-  set.seed(1337)
-  n <- 500
-  X <- matrix(runif(n*5), ncol = 5)
-  y <- beta_vec[1] + beta_vec[2] * X[, 1] + beta_vec[3] * X[, 2] +  beta_vec[4] * X[, 3] + beta_vec[5] * X[, 4] + beta_vec[6] * X[, 5] + rnorm(
-    n,
-    sd = exp(gamma_vec[1] + gamma_vec[2] * X[, 1] + gamma_vec[3] * X[, 2] +  gamma_vec[4] * X[, 3] + gamma_vec[5] * X[, 4] + gamma_vec[6] * X[, 5])
-  )
-
-  # estimate data by boosting ------------
-  model <- asp20boost::LocationScaleRegressionBoost$new(
-    y ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5],
-    ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5])
-  asp20boost::gradient_boost(model, stepsize = 0.001, maxit = 5000, abstol = 0.0001, verbose = FALSE)
-
-
-  return(list(
-    beta = model$beta,
-    gamma = model$gamma)
-  )
-}
-
-
-
-
-
-
-
-
-
-# beta_1 <- c(1,1,0,0,0,0)
-# gamma_1 <- c(1,1,0,0,0,0)
[…]

### 2020-09-15 simulation_studies/#simulation_plot_4_params.R

@@ -1,52 +1,0 @@
-library(asp20boost)
-
-# set true model parameters
-beta_0  <- 2
-beta_1  <- 2
-gamma_0 <- 2
-gamma_1 <- 2
-
-
-
-# simulate data -------------------------
-set.seed(1337)
-n <- 500
-x <- runif(n)
-y <- beta_0 + beta_1 * x + rnorm(
-  n,
-  sd = exp(gamma_0 + gamma_1 * x)
-)
-
-# estimate data by boosting ------------
-model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
-gradient_boost(model, stepsize = 0.001, maxit = 5000, abstol = 0.00000000000000000000000000000000000000000001, verbose = TRUE)
-
-
-# plot results -------------------------
-x_grid <- seq(0,1, length.out = 100)
-y_grid <- beta_0 + beta_1 * x_grid
-
-# limits for plot
-ylim_upper <- model$beta[1] + model$beta[2] * 1 + 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
-ylim_lower <- model$beta[1] + model$beta[2] * 1 - 1.96 * exp(model$gamma[1] + model$gamma[2] * 1)
-
-
-# plot data points
-plot(x, y, col = 2,
-     xlim = c(-0.3,1.3),
-     ylim = c(ylim_upper * 1.05, ylim_lower * 1.05))
-
-# plot true model
-lines(x_grid, y_grid)
-lines(x_grid, y_grid + 1.96 * exp(gamma_0 + gamma_1 * x_grid))
-lines(x_grid, y_grid - 1.96 * exp(gamma_0 + gamma_1 * x_grid))
-
-#plot estimated effect
-location_estimate <- model$beta[1] + model$beta[2] * x_grid
-scale_estimate_sd <- exp(model$gamma[1] + model$gamma[2] * x_grid)
-lines(x_grid, location_estimate, col = 3)
-lines(x_grid, location_estimate + 1.96 * scale_estimate_sd, col = 3)
-lines(x_grid, location_estimate - 1.96 * scale_estimate_sd, col = 3)
-
-model$beta
-model$gamma

### 2020-09-15 simulation_studies/#simulation_variable_selection.R

@@ -1,49 +1,0 @@
-# This script analyses if variable selection in the case of multidemnsional
-# beta and gamma vectors works
-# For this reason some coefficients are 0.
-library(asp20boost)
-
-beta_vec <- c(1,0,10,0,10,0)
-gamma_vec <- c(0,0.1,0,0.1,0,0.1)
-
-
-# beta_2 <- c(0,1,0,1,0,1)
-# gamma_2 <- c(0,0,1,0,1,0)
-
-# beta_3 <- c(1,1,0,1,0,1)
-# gamma_3 <- c(1,0,1,0,1,0)
-
-# homoskedastic model
-# beta_4 <- c(1,1,0,1,0,1)
-# gamma_4 <- c(1,0,0,0,0,0)
-
-# intercept model (heteroskedastic)
-# beta_5 <- c(1,0,0,0,0,0)
-# gamma_5 <- c(1,0,1,0,1,0)
-
-
-
-
-
-
-# simulate data -------------------------
-set.seed(1337)
-n <- 500
-X <- matrix(runif(n*5), ncol = 5)
-y <- beta_vec[1] + beta_vec[2] * X[, 1] + beta_vec[3] * X[, 2] +  beta_vec[4] * X[, 3] + beta_vec[5] * X[, 4] + beta_vec[6] * X[, 5] + rnorm(
-  n,
-  sd = exp(gamma_vec[1] + gamma_vec[2] * X[, 1] + gamma_vec[3] * X[, 2] +  gamma_vec[4] * X[, 3] + gamma_vec[5] * X[, 4] + gamma_vec[6] * X[, 5])
-)
-
-# estimate data by boosting ------------
-model <- LocationScaleRegressionBoost$new(
-  y ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5],
-  ~ X[, 1] + X[, 2] + X[, 3] + X[, 4] + X[, 5])
-gradient_boost(model, stepsize = c(0.1, 0.01), maxit = 1000, verbose = TRUE, componentwise=T)
-
-
-model$beta
-model$gamma
-
-beta_vec
-gamma_vec

### 2020-09-15 simulation_studies/compare_gamboostlss.R

@@ -1,42 +1,0 @@
-library(asp20boost)
-library(gamboostLSS)
-
-# set true model parameters
-beta_0  <- 0
-beta_1  <- 1
-gamma_0 <- -3
-gamma_1 <- 2
-
-
-
-# simulate data -------------------------
-set.seed(1337)
-n <- 1000
-x <- runif(n)
-y <- beta_0 + beta_1 * x + rnorm(
-  n,
-  sd = exp(gamma_0 + gamma_1 * x)
-)
-
-xy_df <- cbind(x,y)
-head(xy_df)
-
-# estimate data by boosting ------------
-mod_1 <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
-gradient_boost(mod_1, maxit = 1000)
-
-
-# estimate with external package --------
-ctrl <- boost_control(trace = FALSE,
-                      mstop = c(mu = 1000, sigma = 1000),
-                      nu = c(mu = 0.01, sigma = 0.001)
-                      )
-mod_2 <- gamboostLSS(formula = y ~ x,
-                     control = ctrl)
-eta_mu <- fitted(mod_2)$mu
-eta_sigma <- fitted(mod_2)$sigma
-
-
-lm(eta_mu ~ x)
-lm(eta_sigma ~ x)
-

### 2020-09-15 simulation_studies/estimate_model.R

@@ -1,58 +1,0 @@
-rm(list=ls())
-source("simulation_studies/helper_plot_model.R")
-source("simulation_studies/helper_init_model.R")
-source("simulation_studies/helper_calc_mse.R")
-library(asp20boost)
-library(gamboostLSS)
-
-beta_1 <- c(0,1)
-gamma_1 <- c(-3,2)
-beta_2 <- c(0,1)
-gamma_2 <- c(-0.5,0)
-beta_3 <- c(0,1)
-gamma_3 <- c(-0.3,0.3)
-beta_4 <- c(0,1)
-gamma_4 <- c(0.3,-0.3)
-
-
-#
-# mod_1 <- init_model(beta_1, gamma_1, 1000)
-# plot_model(beta_1, gamma_1, 1000)
-# gradient_boost(mod_1)
-# calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
-
-
-# mod_2 <- init_model(beta_2, gamma_2, 1000)
-# plot_model(beta_2, gamma_2, 1000)
-# gradient_boost(mod_2)
-# calc_MSE(beta_2, gamma_2, mod_2$beta, mod_2$gamma)
-
-
-# mod_3 <- init_model(beta_3, gamma_3, 1000)
-# plot_model(beta_3, gamma_3, 1000)
-# gradient_boost(mod_3)
-# mod_3$beta
-# mod_3$gamma
-
-
-mod_4 <- init_model(beta_4, gamma_4, 1000)$model
-plot_model(beta_4, gamma_4, 1000)
-gradient_boost(mod_4)
-mod_4$beta
-mod_4$gamma
-
-
-
-# ausprobieren --------------------------------------
-
-
-
-
-
-
-
-
-
-
-
-

### 2020-09-15 simulation_studies/helper_plot_model_high_p.r

@@ -1,13 +1,0 @@
-helper_plot_model_high_p <- function(beta, gamma, X, index){
-  
-  # set up linear predictor functions -------------------------------
-  eta_mu <- function(x) beta[1] + beta[index] * x
-  eta_sigma <- function(x) gamma[1] + gamma[index] * x
-  
-  x <- X[ , index]
-  
-  plot(x, y) # plot y against column vectors of X
-  curve(mean(y) + eta_mu(x) + 1 * exp(eta_sigma(x)), -0.1, 1.1, add = TRUE, col = 1, lty = 3, lwd = 2)
-  curve(mean(y) + eta_mu(x) - 1 * exp(eta_sigma(x)), -0.1, 1.1, add = TRUE, col = 1, lty = 3, lwd = 2)
-  
-}

### 2020-09-15 simulation_studies/large_variable_selection.r

@@ -1,36 +1,0 @@
-rm(list = ls())
-library(asp20boost)
-options(scipen=999)
-source("simulation_studies/helper_init_large_model.R")
-windows()
-
-
-# initialize model
-result <- init_large_model()
-mod_1 <- result$model
-X <- result$design
-y <- result$response
-fitted_loc <- result$fitted_response
-fitted_scale <- result$fitted_sd
-
-# check model via plotting ------------------------------------------------------------------
-par(mfrow = c(4,3))
-for(i in 1:11) plot(X[,i], y, ylim = c(0,200))
-# only one effect is clearly indicated, the others are too small compared to the very high response
-# the effect is also indicated if heteroskedastic sds are added. Hence, signal-noise ratio is OK.
-# ALSO FOR OTHER effects? or only for the exceptionally high one?
-
-
-
-# check model via linear predictors ---------------------------------------------------------
-fitted_loc
-fitted_scale
-
-# estimate huge model -----------------------------------------------------------------------
-#gradient_boost(mod_1, maxit = 1000, componentwise = TRUE)
-
-
-
-# ausprobieren ------------------------------------------------------------------------------
-mod_1$beta
-mod_1$gamma

### 2020-09-15 simulation_studies/variable_selection_2.R

@@ -1,24 +1,0 @@
-source("simulation_studies/helper_init_large_model.R")
-
-
-# init_1 <- init_large_model(c(5,10), c(0.5,1), 5, -0.2) #works quite
-# init_2 <- init_large_model(c(5,10), c(0.5,1), 10, -1) #bad performance
-# init_3 <- init_large_model(c(3,6), c(0.2,0.8), 10, -0.2) #works quite
-init <- init_large_model(c(5,10), c(0.01,0.08), 50, 0, n = 1000)
-
-mod_1 <- init$model
-init$fitted_response
-init$fitted_sd
-
-init$beta
-init$gamma
-
-
-
-gradient_boost(mod_1, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE, verbose = TRUE)
-
-init$beta
-mod_1$beta
-
-init$gamma
-mod_1$gamma

### 2020-09-15 simulation_studies/sim_study_small_model.Rmd
updated sim_studies

@@
 ```{r}
-# setting the stage
+# run all code chunks in this document
+# do not display output of this codechunk
+# DO display code
 rm(list=ls())
@@
 ```{r}
+# dont display code
+# dont display output
 # config 1 ----------------------
@@
 ```{r}
+# display code
+# dont display output
 set.seed(1337)
@@
 ```{r, cache=TRUE, fig.height = 8, fig.width = 11}
+# display code
+# display output
 par(mfrow = c(2,2))
@@
 ```{r}
+# dont display code
+# display output
 result_df <- data.frame("Config" = c(1,1,2,2,3,3,4,4),
@@
 This is due to the presence of huge standard deviations in response *y*, concealing changes in mean of *y.* 
-In fact, the standard deviation at x = 1 is:
-
-```{r}
-exp(gamma_2[1] + gamma_2[2] * 1)
-```
+In fact, the standard deviation at x = 1 is exp(2+2) = 54.60.
 In contrast the true response mean values range from 0 to 2.
@@
 ```{r}
+# display code
+# display output
 mse_1 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
@@
 ```{r}
+# display code
+# display output
 mse_4 <- calc_MSE(beta_4, gamma_4, mod_4$beta, mod_4$gamma)

### 2020-09-15 simulation_studies/sim_study.Rmd
update simulation studies

--- 2020-09-15 simulation_studies/sim_study_large_model.Rmd
@@
 ---
-title: "Simulation Study"
+title: "R Notebook"
 output: html_notebook
@@
 
-In this part we design and perform a simulation study in order to demonstrate functionality of the implemented boosting algorithm.
-For this, we first employ small models with two location and scale parameters each, and ensure basic functionality.
-Second, validity of the variable selection features is demonstrated.
-Lastly, we compare our algorithm with the one implemented in the gamboostLSS package.
-
-In order ensure legibility we employ some helper functions for model initialization, plotting, etc. These and the packages are loaded in the following chunk.
-
 ```{r}
-# setting the stage
+# run all code chunks in this document
+# do not display output of this codechunk
+# DO display code
 rm(list=ls())
@@
 library(microbenchmark)
-```
-
-### ToDo
-
-This may be improved by setting less maxit.
-However this may lead to move away from the yet quite optimal choice for *beta*.
-It could also be improved by lowering stepsize for gamma.
-This results in a pattern where the latter entries of gamma are chosen exactly wrong.
-We will observe this pattern also in the next examples and have no explanation.
-
-## Large Model - Variable Selection in the multidimensional case
-
-In this part we demonstrate how the gradient_boost() function performs variable selection.
-We have a look at three large model configurations.
-Configuration 5 contains 6-dimensional parameter vectors, oncluding the intercept terms.
-In configuration 6 *beta* and *gamma* are of length 11, while in configuration 7 they contain 51 components.
-The parameter vectors are created randomly, then some components are set to 0, resulting in the configurations displayed below.
-In order to ensure a good signal-noise ratio, there are restrictions to random parameter creation, which are explained below.
-
-```{r}
-set.seed(1337)
-# initialize models --------------------------------------------------------------
-init_5 <- init_large_model(beta_range = c(5,10),
-                           gamma_range = c(0.5,1),
-                           dim_param = 5,
-                           int_gamma = -0.2,
-                           n = 100)
-init_6 <- init_large_model(beta_range = c(3,6),
-                           gamma_range = c(0.2,0.8),
-                           dim_param = 10,
-                           int_gamma = -0.2,
-                           n = 100)
-init_7 <- init_large_model(beta_range = c(5,10),
-                           gamma_range = c(0.01,0.08),
-                           dim_param = 50,
-                           int_gamma = 0,
-                           n = 100)
-mod_5 <- init_5$model
-mod_6 <- init_6$model
-mod_7 <- init_7$model
-
-# estimate models -----------------------------------------------------------------
-gradient_boost(mod_5, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
-gradient_boost(mod_6, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
-gradient_boost(mod_7, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
-
-```
-
-```{r}
-# Configuration 5
-cbind(beta_true = init_5$beta, 
-      beta_est = round(mod_5$beta, 3),
-      gamma_true = init_5$gamma,
-      gamma_est = round(mod_5$gamma, 3)
-)
-```
-
-```{r}
-# Configuration 6
-cbind(beta_true = init_6$beta, 
-      beta_est = round(mod_6$beta, 3),
-      gamma_true = init_6$gamma,
-      gamma_est = round(mod_6$gamma, 3)
-)
-```
-
-### Configuration 5
-As mentioned above there are restrictions on parameter creation, since not every possible parameter configuration leads to estimable models, as we have seen in the two-dimensional case.
-In the multidimensional case, the problem remains: 
-If gamma parameters add up to high values, the resulting simulated response has huge standard deviation, which conceals the location effects.
-Checking for signal-noise ratio remains to be done in the multidimensional case as well.
-Plotting is hardly an option as dimensions increase.
-Therefore we check by another method, that is looking at the distributions of true response means and true response SDs.
-
-```{r}
-par(mfrow=c(2,1))
-hist(init_5$fitted_response, breaks = 25, main = "eta_mu")
-hist(init_5$fitted_sd, breaks = 25, main = "exp_eta_sigma")
-```
-
-Location predictors take values between 8 and 22, peaking at 15.
-In contrast, the highest response-SD is 4, while most values are at about 2.5 and lower.
-The resulting response variation should be small enough to not conceal the location effects, as well as big enough to be estimated.
-
-In configuration 5 beta components are correctly selected.
-Zero-components are also zero in the estimate vector, while non-zero components are estimated as non-zero effects.
-The effect sizes differ considerably more than we observed in the two-dimensional case.
-
-The algorithm depicts correctly which effects on response location exists, that is it finds the non-zero elements of *beta*.
-For scale estimate *gamma_hat* too many parameters become chosen, since two zero-elements of true parameter *gamma* are assigned non-zero values by the estimate *gamma_hat*.
-
-The values estimated differ considerably from the true parameter values. 
-The resulting MSE is higher than in the four parameter case.
-
-```{r}
-mse_5 <- calc_MSE(init_5$beta, init_5$gamma, mod_5$beta, mod_5$gamma)
-cbind("config 1" = mse_1,
-      "config_3" = mse_3,
-      "config 4" = mse_4,
-      "config_5" = mse_5)
-```
-
-### Configuration 6
-
-For 10-dimensional configuration 6 we get the following mean and SD distributions.
-
-```{r}
-par(mfrow=c(2,1))
-hist(init_6$fitted_response, breaks = 25, main = "eta_mu")
-hist(init_6$fitted_sd, breaks = 25, main = "exp_eta_sigma")
-```
-
-For beta the algo chose 1 too few.
-For gamma entirs 1-3 are chosen correctly. Then the algorithm yields errors. Entry 4 is chosen, while in fact it is zero. Entry 5 is not chosen, while in the true model an effect exists.
-
-### Configuration 7
-
-Scaling parameter dimension up to 51, the algortihm runs into problems.
-A solution may be to properly choose initial values.
-
-```{r}
-par(mfrow=c(2,1))
-hist(init_7$fitted_response, breaks = 25, main = "eta_mu")
-hist(init_7$fitted_sd, breaks = 25, main = "exp_eta_sigma")
-```
-
-Even with a well-balanced signal-noise ratio as indicated by the histograms, problems occur.
-This problem occurs if we scale parameter dimensions up: only intercepts become updated. 
-No covariates are chosen to have an effect on the response. 
-If algorithm progress is checked with verbose=TRUE, one sees that intercepts inflate in the very first step, and then the proceedsing iterations work on lowering the intercepts, which takes place very slowly due to stepsize parameters.
-Possibly a sensible choice of initial parameter values may help. However we did not try this.
-
-```{r}
-cbind(beta_true = init_7$beta, 
-      beta_est = round(mod_7$beta, 3),
-      gamma_true = init_7$gamma,
-      gamma_est = round(mod_7$gamma, 3)
-)
 ```

### 2020-09-15 simulation_studies/sim_study_benchmark.Rmd
update simulation studies

--- 2020-09-15 simulation_studies/sim_study_small_model.Rmd
@@
 ---
-title: "R Notebook"
+title: "Simulation Study"
 output: html_notebook
@@
 
+In this part we design and perform a simulation study in order to demonstrate functionality of the implemented boosting algorithm. 
+For this, we first employ models with two location and scale parameters each, and show basic functionality. 
+Second, validity of the variable selection features is demonstrated, by using location and scale parameters with up to 50 components.
+Lastly, we compare our algorithm with the one implemented in the gamboostLSS package. *really?*
+
 ```{r}
-# run all code chunks in this document
-# do not display output of this codechunk
-# DO display code
+# setting the stage
 rm(list=ls())
@@
 ### ToDo Sebastián
-* Formula in front of every 'Configuration' paragraph
-* Make Data Frame look like a nice table
+* Formula in front of every 'Configuration' paragraph in "Small Model"
+* Make Data Frame look like a nice table (optional, maybe just chill)
 
-## Small Model - Two-dimnesional Case
+## Two-dimesional Case
 
@@
 * Our algortihm runs into problems if the signal is too strong compared to noise.
+
+## Large Model - Variable Selection in the multidimensional case
+
+In this part we demonstrate how the gradient_boost() function performs variable selection.
+We have a look at three large model configurations.
+Configuration 5 contains 6-dimensional parameter vectors.
+In configuration 6 *beta* and *gamma* are of length 11, while in configuration 7 they contain 51 components.
+The parameter vectors are created randomly, then some components are set to 0, resulting in the configurations displayed below.
+In order to ensure a good signal-noise ratio, there are restrictions to random parameter creation, which are explained below.
+
+```{r}
+# show code
+# dont show output
+set.seed(1337)
+# initialize models --------------------------------------------------------------
+init_5 <- init_large_model(beta_range = c(5,10),
+                           gamma_range = c(0.5,1),
+                           dim_param = 5,
+                           int_gamma = -0.2,
+                           n = 100)
+init_6 <- init_large_model(beta_range = c(3,6),
+                           gamma_range = c(0.2,0.8),
+                           dim_param = 10,
+                           int_gamma = -0.2,
+                           n = 100)
+init_7 <- init_large_model(beta_range = c(5,10),
+                           gamma_range = c(0.01,0.08),
+                           dim_param = 50,
+                           int_gamma = 0,
+                           n = 100)
+mod_5 <- init_5$model
+mod_6 <- init_6$model
+mod_7 <- init_7$model
+
+# estimate models -----------------------------------------------------------------
+gradient_boost(mod_5, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
+gradient_boost(mod_6, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+gradient_boost(mod_7, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+
+```
+
+```{r}
+#dont show code
+#show output
+# Configuration 5
+cbind(beta_true = init_5$beta, 
+      beta_est = round(mod_5$beta, 3),
+      gamma_true = init_5$gamma,
+      gamma_est = round(mod_5$gamma, 3)
+)
+```
+
+```{r}
+#dont show code
+#show output
+# Configuration 6
+cbind(beta_true = init_6$beta, 
+      beta_est = round(mod_6$beta, 3),
+      gamma_true = init_6$gamma,
+      gamma_est = round(mod_6$gamma, 3)
+)
+```
+
+### Configuration 5
+As mentioned above there are restrictions on parameter creation, since not every possible parameter configuration leads to estimable models, as we have seen in the two-dimensional case.
+In the multidimensional case, the problem remains: 
+If gamma parameters add up to high values, the resulting simulated response has huge standard deviation, which conceals the location effects.
+Checking for signal-noise ratio remains to be done in the multidimensional case as well.
+Plotting is hardly an option as dimensions increase.
+Therefore we check by another method, that is looking at the distributions of true response means and true response SDs.
+
+```{r}
+#show code
+#show output
+par(mfrow=c(2,1))
+hist(init_5$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_5$fitted_sd, breaks = 25, main = "exp_eta_sigma")
+```
+
+Location predictors take values between 8 and 22, peaking at 15. 
+In contrast, the highest response-SD is 4, while most values are at about 2.5 and lower. 
+The resulting response variation should be small enough to not conceal the location effects, as well as big enough to be estimated.
+
+In configuration 5, the algorithm correctly depicts which effects on response location exists, that is it finds the non-zero elements of *beta*.
+For scale estimate *gamma_hat* too many parameters become chosen, since two zero-elements of true parameter *gamma* are assigned non-zero values by the estimate *gamma_hat*.
+The estimated effect sizes differ considerably more from the true values than we observed in the two-dimensional case, resulting in higher MSE
+
+```{r}
+#show code
+#show output
+mse_5 <- calc_MSE(init_5$beta, init_5$gamma, mod_5$beta, mod_5$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3,
+      "config 4" = mse_4,
+      "config_5" = mse_5)
+```
+
+### Configuration 6
+
+For 10-dimensional configuration 6 we get the following mean and SD distributions.
+
+```{r}
+#show code
+#show output
+par(mfrow=c(2,1))
+hist(init_6$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_6$fitted_sd, breaks = 25, main = "exp_eta_sigma")
+```
+
+For beta the algo chose 1 too few.
+For gamma entirs 1-3 are chosen correctly. Then the algorithm yields errors. Entry 4 is chosen, while in fact it is zero. Entry 5 is not chosen, while in the true model an effect exists.
+
+### Configuration 7
+
+Scaling parameter dimension up to 51, the algortihm runs into problems: all slope estimates are zero.
+
+```{r}
+#show code
+#show output
+round(mod_7$beta, 3)
+round(mod_7$gamma, 3)
+```
+
+Histograms indicate a well-balanced signal-noise ratio, so the reason lies somewhere else.
+
+```{r}
+#show code
+#show output
+par(mfrow=c(2,1))
+hist(init_7$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_7$fitted_sd, breaks = 25, main = "exp_eta_sigma")
+```
+
+The problem of slope estimates all being zero is due to the fact, that the algorithm fits only intercepts in the first step, resulting in very high values.
+The remaining iterations are used to lower intercepts, this may be checked by setting verbose = TRUE.
+For this reason no other components of the parameter estimates become updated.
+Possibly a sensible choice of initial parameter values may address this problem.

### 2020-09-15 simulation_studies/sim_study.Rmd
update sim_studies

@@
 
-```{r}
-# run all code chunks in this document
-# do not display output of this codechunk
-# DO display code
+In this part we design and perform a simulation study in order to demonstrate functionality of the implemented boosting algorithm. 
+For this, we first employ models with two location and scale parameters each, and show basic functionality. 
+Second, validity of the variable selection features is demonstrated, by using location and scale parameters with up to 50 components.
+Lastly, we compare our algorithm with the one implemented in the gamboostLSS package. *really?*
+
+```{r setting, cache=TRUE}
+# setting the stage
 rm(list=ls())
@@
 
-## Benchmark gamboostLSS
+### ToDo Sebastián
+* Formula in front of every 'Configuration' paragraph in "Small Model"
+* Make Data Frame look like a nice table (optional, maybe just chill)
 
-In order to assess performance of our algorithm further we compare it to the existing package *gamboostLSS.* 
-Therefore we use configurations 1 and 4 the two-dimensional cases.
-In the former case our algorithm performed well, in the latter rather bad.
+## Two-dimesional Case
 
-### gamboostLSS- package
+The goal in this first part is to demonstrate, that the algorithm finds correct estimates for sensible choices of *beta* and *gamma.* We use the following 4 different parameter configurations to explore location-scale regression model characteristics.
 
-In the *gamboostLSS* package boosting a location-scale regression model works as follows.
-First we need to define the vector of controls.
-As the package contains more functionalities than ours, we have to specify a control variable, that matches stepsize, and mstop to our package to allow for comparison.
+```{r include=FALSE}
+# dont display code
+# dont display output
+# config 1 ----------------------
+beta_1 <- c(0,1)
+gamma_1 <- c(-3,2)
+
+# config 2 ----------------------
+beta_2 <- c(2,2)
+gamma_2 <- c(2,2)
+
+# config 3 ----------------------
+beta_3 <- c(0,1)
+gamma_3 <- c(-1,-0.3)
+
+# config 4 ----------------------
+beta_4 <- c(0,15)
+gamma_4 <- c(1,-1)
+```
 
 ```{r}
-ctrl <- boost_control(trace = FALSE,
-                      mstop = c(mu = 1000, sigma = 1000),
-                      nu = c(mu = 0.01, sigma = 0.001)
-)
+# display code
+# dont display output
+set.seed(1337)
+mod_1 <- init_model(beta_1, gamma_1, 1000)$model
+gradient_boost(mod_1)
+mod_2 <- init_model(beta_2, gamma_2, 1000)$model
+gradient_boost(mod_2)
+mod_3 <- init_model(beta_3, gamma_3, 1000)$model
+gradient_boost(mod_3)
+mod_4 <- init_model(beta_4, gamma_4, 1000)$model
+gradient_boost(mod_4)
 ```
 
-### Recap Configuration 1
+```{r, cache=TRUE, fig.height = 8, fig.width = 11}
+# display code
+# display output
+par(mfrow = c(2,2))
+plot_model(beta_1, gamma_1, 1000, "Config 1")
+plot_model(beta_2, gamma_2, 1000, "Config 2")
+plot_model(beta_3, gamma_3, 1000, "Config 3")
+plot_model(beta_4, gamma_4, 1000, "Config 4")
+```
+```{r}
+# dont display code
+# display output
+result_df <- data.frame("Config" = c(1,1,2,2,3,3,4,4),
+                        "beta_true" = c(beta_1, beta_2, beta_3, beta_4),
+                        "beta est." = c(round(mod_1$beta, 3), round(mod_2$beta, 3), round(mod_3$beta, 3), round(mod_4$beta, 3)),
+                        "gamma_true" = c(gamma_1, gamma_2, gamma_3, gamma_4),
+                        "gamma est." = c(round(mod_1$gamma, 3), round(mod_2$gamma, 3), round(mod_3$gamma, 3), round(mod_4$gamma, 3))
+                        )
+result_df
+```
+
+### Evaluation Configuration 1
+
+[insert model config by formula]
+
+This parameter configuration is equivalent to the example case of the *asp20model* package. 
+As already mentioned, heteroskedasticity is present and indicated by the funnel shape of the point cloud.
+The solid black line represents the true response mean, while the dashed lines represent the true upwards and downwards standard deviations from this mean.
+The red arrow indicates the standard deviation at the point x = 1,  that is the upper limit of covariate values, since covariates are design to be on the unit interval.
+For cases with increasing scale, the red arrow hence indicates maximum standard deviation of the response.
+The blue arrow depicts the slope of the location parameter vector *beta_1.*
+The ratio of those two quantities is essential for estimability of the model. 
+Here the blue arrow is dominant, hence the model may well be estimated. 
+Estimates are displayed in the table above. 
+We see that both *betahat* and *gammahat* estimates are close to the true parameter vectors *beta* and *gamma.*
+
+#### Evaluation Configuration 2
+
+Configuration 2 is a somewhat naive parameter configuration, where all parameters are set to value 2. 
+Estimating this model does not yield valid results.
+The gradient_boost() function runs into numerical problems.
+To understand why this happens we need to understand the model characteristics resulting from parameter configuration 2-2-2-2.
+Here, the red arrow is of large size, while the blue arrow is hardly visible. 
+This is due to the presence of huge standard deviations in response *y*, concealing changes in mean of *y.* 
+In fact, the standard deviation at x = 1 is exp(2+2) = 54.60.
+In contrast the true response mean values range from 0 to 2.
+The change in response location (the signal), indicated by the blue arrow, perishes due to the high amount of noise caused by the high gamma-parameters. 
+Resulting is a signal-noise ratio, that is impossible to estimate. 
+Values for parameters gamma need to be chosen sensibly for simulating data. 
+This is no straight-forward task, since the effect of *gamma* on the response standard deviation is not linear, but exponential. 
+To check this in the two-dimensional case, the arrows in the model plot are of help.
+
+#### Evaluation Configuration 3
+
+[insert parameter configuration as formula]
+
+Configuration 3, in turn, yields reasonable estimates.
+It contains a negative gamma slope, leading to decreasing response standard deviations.
+This can be observed by the funnel shape pointing in the opposite direction as before.
+
+It may be of interest to check if estimates lie closer to the true parameters than e.g. in configuration 1.
+To do this we measure the gap between true values and estimates by the mean squared error.
+This enables comparison of configurations 1 and 3.
 
 ```{r}
-set.seed(1337)
-init_1 <- init_model(beta_1, gamma_1, 1000)$model
-mod_1 <- init_1$model
-response_1 <- init_1$response
-covariate_1 <- init_1$covariate
-gamboostLSS(formula = response_1 ~ covariate_1,
-                      control = ctrl)
+# display code
+# display output
+mse_1 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
+mse_3 <- calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3)
 ```
 
-What we get as result are not yet parameter estimates, but linear predictors.
-These must be regressed on the parameters to arrive at the desired estimates.
-In this package we do not first create a model and then boost it. Creation and boosting is performed in one call, already yielding estimates. 
-These estimates are, however not beta and gamma, but rather eta_mu and eta_gamma. 
-So these unit-wise linear predictors need to be regressed on the covariate in order to arrive at the estimates of interest.
+Configuration 1 leads to smaller MSE, hance our algorithm performs better in configuration 1.
+
+#### Evaluation Configuration 4
+
+[include configuration as formulas]
+
+In configuration 4, we again included a negative gamma slope, as well as a very high beta slope.
+The idea is to ensure a very good signal noise ratio, in order to ensure good model fit, indicated by low MSE.
+Contrary to our expectations, the algorithm performs poorly for this parameter configuration.
 
 ```{r}
-eta_mu <- fitted(est_1)$mu
-eta_sigma <- fitted(est_1)$sigma
+# display code
+# display output
+mse_4 <- calc_MSE(beta_4, gamma_4, mod_4$beta, mod_4$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3,
+      "config 4" = mse_4)
+```
 
-est_beta <- lm(eta_mu ~ covariate_1)
-est_gamma <- lm(eta_sigma ~ covariate_1)
+The bad model fit is due to the fact, that the location slope is highly underestimated.
+
+### Conclusions
+
+The observed two-dimensional cases let us conclude that not every arbitrary parameter configuration is estimable by the implemented gradient_boost() algorithm.
+A first insight is that signal-noise ratio needs to be sufficiently high in order to prevent that location effect gets concealed by inflated standard deviations.
+For the finding of configuration 4 we have several explanation approaches:
+* There are further constraints on simulating data meaningfully which we were not yet able to work out.
+* Our algorithm does not reliably yield robust results if negative gamma slopes are included.
+* Our algortihm runs into problems if the signal is too strong compared to noise.
+
+## Large Model - Variable Selection in the multidimensional case
+
+In this part we demonstrate how the gradient_boost() function performs variable selection.
+We have a look at three large model configurations.
+Configuration 5 contains 6-dimensional parameter vectors.
+In configuration 6 *beta* and *gamma* are of length 11, while in configuration 7 they contain 51 components.
+The parameter vectors are created randomly, then some components are set to 0, resulting in the configurations displayed below.
+In order to ensure a good signal-noise ratio, there are restrictions to random parameter creation, which are explained below.
+
+```{r}
+# show code
+# dont show output
+set.seed(1337)
+# initialize models --------------------------------------------------------------
+init_5 <- init_large_model(beta_range = c(5,10),
+                           gamma_range = c(0.5,1),
+                           dim_param = 5,
+                           int_gamma = -0.2,
+                           n = 100)
+init_6 <- init_large_model(beta_range = c(3,6),
+                           gamma_range = c(0.2,0.8),
+                           dim_param = 10,
+                           int_gamma = -0.2,
+                           n = 100)
+init_7 <- init_large_model(beta_range = c(5,10),
+                           gamma_range = c(0.01,0.08),
+                           dim_param = 50,
+                           int_gamma = 0,
+                           n = 100)
+mod_5 <- init_5$model
+mod_6 <- init_6$model
+mod_7 <- init_7$model
+
+# estimate models -----------------------------------------------------------------
+gradient_boost(mod_5, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
+gradient_boost(mod_6, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+gradient_boost(mod_7, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
 
@@
 
-### Recap Configuration 3
-
 ```{r}
-init <- init_model(beta_3, gamma_3, 1000)
-mod_3 <- init$model
-response_3 <- init$response
-covariate_3 <- init$covariate
-plot_model(beta_3, gamma_3, 1000)
-```
-[...]
-
-```{r}
-# estimate with external package --------
-
-mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3,
-                      control = ctrl)
-eta_mu <- fitted(mod_33)$mu
-eta_sigma <- fitted(mod_33)$sigma
-
-lm(eta_mu ~ covariate_3)
-lm(eta_sigma ~ covariate_3)
-
-```
-
-[...]
-```{r}
-gradient_boost(mod_3)
-round(mod_3$beta, 3)
-round(mod_3$gamma, 3)
-```
-
-```{r}
-microbenchmark(
-  gradient_boost(mod_3)
+#dont show code
+#show output
+# Configuration 5
+cbind(beta_true = init_5$beta, 
+      beta_est = round(mod_5$beta, 3),
+      gamma_true = init_5$gamma,
+      gamma_est = round(mod_5$gamma, 3)
 )
@@
 ```{r}
-microbenchmark(
-  mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3, control = ctrl),
-  eta_mu <- fitted(mod_33)$mu,
-  eta_sigma <- fitted(mod_33)$sigma,
-  lm(eta_mu ~ covariate_3),
-  lm(eta_sigma ~ covariate_3)
+#dont show code
+#show output
+# Configuration 6
+cbind(beta_true = init_6$beta, 
+      beta_est = round(mod_6$beta, 3),
+      gamma_true = init_6$gamma,
+      gamma_est = round(mod_6$gamma, 3)
 )
@@
 
-Ensuring comparability is a difficult task, since the two packages are implemented in rather different ways.
-Running gradient_boost takes about 657000 microseconds (0.66 seconds).
-In contrasts, running the external algorithm gamboostLSS takes 2.829.493 microseconds (2.8 seconds), which is a multiple of our algorithm.
-Additionally, beta and gamma parameters are not yielded directly, but need to be determined by regression as well.
-In contrats, this happens directly in our algorithm.
-
-To be fair we evaluate construction of the model object as well.
+### Configuration 5
+As mentioned above there are restrictions on parameter creation, since not every possible parameter configuration leads to estimable models, as we have seen in the two-dimensional case.
+In the multidimensional case, the problem remains: 
+If gamma parameters add up to high values, the resulting simulated response has huge standard deviation, which conceals the location effects.
+Checking for signal-noise ratio remains to be done in the multidimensional case as well.
+Plotting is hardly an option as dimensions increase.
+Therefore we check by another method, that is looking at the distributions of true response means and true response SDs.
 
 ```{r}
-microbenchmark(
-  mod_1 <- LocationScaleRegressionBoost$new(response_3 ~ covariate_3, ~ covariate_3)
-)
+#show code
+#show output
+par(mfrow=c(2,1))
+hist(init_5$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_5$fitted_sd, breaks = 25, main = "exp_eta_sigma")
 ```
 
-This changes the situation only marginally, as this is a quite fast step.
+Location predictors take values between 8 and 22, peaking at 15. 
+In contrast, the highest response-SD is 4, while most values are at about 2.5 and lower. 
+The resulting response variation should be small enough to not conceal the location effects, as well as big enough to be estimated.
+
+In configuration 5, the algorithm correctly depicts which effects on response location exists, that is it finds the non-zero elements of *beta*.
+For scale estimate *gamma_hat* too many parameters become chosen, since two zero-elements of true parameter *gamma* are assigned non-zero values by the estimate *gamma_hat*.
+The estimated effect sizes differ considerably more from the true values than we observed in the two-dimensional case, resulting in higher MSE
+
+```{r}
+#show code
+#show output
+mse_5 <- calc_MSE(init_5$beta, init_5$gamma, mod_5$beta, mod_5$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3,
+      "config 4" = mse_4,
+      "config_5" = mse_5)
+```
+
+### Configuration 6
+
+For 10-dimensional configuration 6 we get the following mean and SD distributions.
+
+```{r}
+#show code
+#show output
+par(mfrow=c(2,1))
+hist(init_6$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_6$fitted_sd, breaks = 25, main = "exp_eta_sigma")
+```
+
+For beta the algo chose 1 too few.
+For gamma entirs 1-3 are chosen correctly. Then the algorithm yields errors. Entry 4 is chosen, while in fact it is zero. Entry 5 is not chosen, while in the true model an effect exists.
+
+### Configuration 7
+
+Scaling parameter dimension up to 51, the algortihm runs into problems: all slope estimates are zero.
+
+```{r}
+#show code
+#show output
+round(mod_7$beta, 3)
+round(mod_7$gamma, 3)
+```
+
+Histograms indicate a well-balanced signal-noise ratio, so the reason lies somewhere else.
+
+```{r}
+#show code
+#show output
+par(mfrow=c(2,1))
+hist(init_7$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_7$fitted_sd, breaks = 25, main = "exp_eta_sigma")
+```
+
+The problem of slope estimates all being zero is due to the fact, that the algorithm fits only intercepts in the first step, resulting in very high values.
+The remaining iterations are used to lower intercepts, this may be checked by setting verbose = TRUE.
+For this reason no other components of the parameter estimates become updated.
+Possibly a sensible choice of initial parameter values may address this problem.

### 2020-09-15 simulation_studies/sim_study.Rmd
vignette almost finished + PDF. Missing abstract and conclusion

--- 2020-09-15 simulation_studies/sim_study.Rmd
@@
 
-```{r}
-# run all code chunks in this document
-# do not display output of this codechunk
-# DO display code
+In this part we design and perform a simulation study in order to demonstrate functionality of the implemented boosting algorithm. 
+For this, we first employ models with two location and scale parameters each, and show basic functionality. 
+Second, validity of the variable selection features is demonstrated by using location and scale parameters with up to 50 components.
+Lastly, we compare our algorithm with the one implemented in the `gamboostLSS` package. *really?*
+
+```{r setting, cache=TRUE}
 rm(list=ls())
 options(scipen=999)
-source("helper_plot_model.R")
-source("helper_init_model.R")
-source("helper_calc_mse.R")
-source("helper_init_large_model.R")
+source("../simulation_studies/helper_plot_model.R")
+source("../simulation_studies/helper_init_model.R")
+source("../simulation_studies/helper_calc_mse.R")
+source("../simulation_studies/helper_init_large_model.R")
 library(asp20boost)
@@
 
-## Benchmark gamboostLSS
+## Two-dimesional Case
 
-In order to assess performance of our algorithm further we compare it to the existing package *gamboostLSS.* 
-Therefore we use configurations 1 and 4 the two-dimensional cases.
-In the former case our algorithm performed well, in the latter rather bad.
+The goal in this first part is to demonstrate that the algorithm finds correct estimates for sensible choices of $\bm{\beta}$ and $\bm{\gamma}$.
+We use the following 4 different parameter configurations to explore location-scale regression model characteristics.
 
-### gamboostLSS- package
+```{r ,include=FALSE, echo=FALSE, cache=TRUE}
+# config 1 ----------------------
+beta_1 <- c(0,1)
+gamma_1 <- c(-3,2)
 
-In the *gamboostLSS* package boosting a location-scale regression model works as follows.
-First we need to define the vector of controls.
-As the package contains more functionalities than ours, we have to specify a control variable, that matches stepsize, and mstop to our package to allow for comparison.
+# config 2 ----------------------
+beta_2 <- c(2,2)
+gamma_2 <- c(2,2)
+
+# config 3 ----------------------
+beta_3 <- c(0,1)
+gamma_3 <- c(-1,-0.3)
+
+# config 4 ----------------------
+beta_4 <- c(0,15)
+gamma_4 <- c(1,-1)
+```
+
+```{r, cache=TRUE}
+set.seed(1337)
+mod_1 <- init_model(beta_1, gamma_1, 1000)$model
+gradient_boost(mod_1)
+mod_2 <- init_model(beta_2, gamma_2, 1000)$model
+gradient_boost(mod_2)
+mod_3 <- init_model(beta_3, gamma_3, 1000)$model
+gradient_boost(mod_3)
+mod_4 <- init_model(beta_4, gamma_4, 1000)$model
+gradient_boost(mod_4)
+```
+
+```{r, cache=TRUE, fig.height = 8, fig.width = 11}
+# display code
+# display output
+par(mfrow = c(2,2))
+plot_model(beta_1, gamma_1, 1000, "Config 1")
+plot_model(beta_2, gamma_2, 1000, "Config 2")
+plot_model(beta_3, gamma_3, 1000, "Config 3")
+plot_model(beta_4, gamma_4, 1000, "Config 4")
+```
+
+```{r, include=FALSE, cache=TRUE}
+# We need output. IMPORTANT
+
+result_df <- data.frame("Config" = c(1,1,2,2,3,3,4,4),
+                        "beta_true" = c(beta_1, beta_2, beta_3, beta_4),
+                        "beta est." = c(round(mod_1$beta, 3), round(mod_2$beta, 3), round(mod_3$beta, 3), round(mod_4$beta, 3)),
+                        "gamma_true" = c(gamma_1, gamma_2, gamma_3, gamma_4),
+                        "gamma est." = c(round(mod_1$gamma, 3), round(mod_2$gamma, 3), round(mod_3$gamma, 3), round(mod_4$gamma, 3))
+                        )
+result_df
+```
+
+### Evaluation Configuration 1
+
+Values: $\bm{\beta} = (0,1)'$ and $\bm{\gamma} = (-3,2)'$
+
+This parameter configuration is equivalent to the base example case of the `asp20model` package. 
+As already mentioned, heteroskedasticity is present, as indicated visually by the funnel shape of the scatter plot.
+The solid black line represents the true response mean, while the dashed lines represent the true upwards and downwards standard deviations from this mean.
+The red arrow indicates the standard deviation at the point $x = 1$ i.e. the upper limit of covariate values, since covariates are designed to be on the unit interval.
+For cases with increasing scale, the red arrow hence indicates maximum standard deviation of the response.
+The blue arrow depicts the slope of the location parameter vector $\bm{\beta_1}$. 
+The ratio of those two quantities is essential for the estimability of the model. 
+Here, the blue arrow is dominant indicating that the model may be well estimated. 
+Estimates are displayed in the table above. 
+We see that both $\bm{\beta}hat$ and $\hat{\bm{\gamma}}$ estimates are close to the true parameter vectors $\bm{\beta}$ and $\bm{\gamma}$.
+
+### Evaluation Configuration 2
+
+Values: $\bm{\beta} = (2,2)'$ and $\bm{\gamma} = (2,2)'$
+
+Configuration 2 is a somewhat naive parameter configuration where all parameters are set to value $2$. 
+Estimating this model does not yield valid results.
+The `gradient_boost()` function runs into numerical problems.
+To understand why this happens, we need to understand the model characteristics resulting from parameter configuration `2-2-2-2`.
+Here, the red arrow shows a large size, while the blue arrow is hardly visible. 
+This happens because of the huge standard deviations present in response $y$ that concealing the changes in the mean $\bar{y}$. 
+In fact, the standard deviation at $x = 1$ is $exp(2+2) = 54.60$.
+In contrast, the true response mean values range from $0$ to $2$.
+The change in response location (the signal), indicated by the blue arrow, perishes *stirbt?* due to the high amount of noise caused by the high $\bm{\gamma}$ parameters. 
+Thus, a signal-noise ratio results that is impossible to estimate. 
+Values for parameters $\bm{\gamma}$ need to be chosen sensibly for simulating data. 
+Nevertheless, this is no straight-forward task, since the effect of $\bm{\gamma}$ on the response standard deviation is not linear but exponential. 
+To check this *check the expoentiality?* in the two-dimensional case, the arrows in the model plot may be of help.
+
+#### Evaluation Configuration 3
+
+Values: $\bm{\beta} = (0,1)'$ and $\bm{\gamma} = (-1,-0.3)'$
+
+Configuration 3, in turn, yields reasonable estimates.
+It contains a negative $\bm{\gamma}$ slope leading to decreasing response standard deviations.
+This can be observed by the funnel shape of the plot pointing in the opposite direction as before.
+
+It may be of interest to check if estimates lie closer to the true parameters relative to  e.g. configuration 1.
+To do this, we measure the gap between true values and estimates by the mean squared error *why not MSE*.
+This enables comparison of configurations 1 and 3.
+
+```{r, cache=TRUE}
+
+mse_1 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
+mse_3 <- calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3)
+```
+
+Configuration 1 leads to smaller MSE. 
+This indicates a better performance in this  configuration.
+
+#### Evaluation Configuration 4
+
+Values: $\bm{\beta} = (0,15)'$ and $\bm{\gamma} = (1,-1)'$
+
+In configuration 4, we again included both a negative slope for $\bm{\gamma}$  and a very high slope $\bm{\beta}$.
+The idea is to ensure a very good signal-noise ratio to guarantee a  good model fit measured by a low MSE.
+Contrary to our expectations, the algorithm performs poorly for this parameter configuration.
+
+```{r, cache=TRUE}
+mse_4 <- calc_MSE(beta_4, gamma_4, mod_4$beta, mod_4$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3,
+      "config 4" = mse_4)
+```
+
+The bad model fit is due to the fact that the location slope is highly underestimated.
+
+### Conclusions
+
+The observed two-dimensional cases let us conclude that not every arbitrary parameter configuration is estimable by the implemented `gradient_boost()` algorithm. 
+A first insight is that signal-noise ratio needs to be sufficiently high in order to prevent that the location effect gets concealed by inflated standard deviations. 
+Regarding the finding of configuration 4, we offer several explanations:
+
+  * Further considerations to  simulating data sensibly need to be taken,  which we were not yet able to work out.
+  * Our algorithm does not reliably yield robust results if negative $\bm{\gamma}$ slopes are included.
+  * Our algorithm runs into problems if the signal is too strong compared to noise.
+
+## Large Model - Variable Selection in the multidimensional case
+
+In this part we demonstrate how the `gradient_boost()` function performs variable selection.
+We have a look at three large model configurations.
+Configuration 5 present 6-dimensional parameter setting.
+In configuration 6 $\bm{\beta}$ and $\bm{\gamma}$ are of length 11, while in configuration 7 they contain 51 components.
+The parameter vectors are created randomly. 
+Then some components are set to 0. 
+This results in the configurations displayed below.
+In order to ensure a good signal-noise ratio, to creation of random parameter is restricted, as explained below.
+
+```{r, echo=FALSE, cache=TRUE}
+set.seed(1337)
+# initialize models --------------------------------------------------------------
+init_5 <- init_large_model(beta_range = c(5,10),
+                           gamma_range = c(0.5,1),
+                           dim_param = 5,
+                           int_gamma = -0.2,
+                           n = 100)
+init_6 <- init_large_model(beta_range = c(3,6),
+                           gamma_range = c(0.2,0.8),
+                           dim_param = 10,
+                           int_gamma = -0.2,
+                           n = 100)
+init_7 <- init_large_model(beta_range = c(5,10),
+                           gamma_range = c(0.01,0.08),
+                           dim_param = 50,
+                           int_gamma = 0,
+                           n = 100)
+mod_5 <- init_5$model
+mod_6 <- init_6$model
+mod_7 <- init_7$model
+
+# estimate models -----------------------------------------------------------------
+gradient_boost(mod_5, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
+gradient_boost(mod_6, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+gradient_boost(mod_7, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
+
+```
 
 ```{r}
-ctrl <- boost_control(trace = FALSE,
-                      mstop = c(mu = 1000, sigma = 1000),
-                      nu = c(mu = 0.01, sigma = 0.001)
+#dont show code
+#show output
+# Configuration 5
+cbind(beta_true = init_5$beta, 
+      beta_est = round(mod_5$beta, 3),
+      gamma_true = init_5$gamma,
+      gamma_est = round(mod_5$gamma, 3)
 )
@@
 
-### Recap Configuration 1
-
-```{r}
-set.seed(1337)
-init_1 <- init_model(beta_1, gamma_1, 1000)$model
-mod_1 <- init_1$model
-response_1 <- init_1$response
-covariate_1 <- init_1$covariate
-gamboostLSS(formula = response_1 ~ covariate_1,
-                      control = ctrl)
-```
-
-What we get as result are not yet parameter estimates, but linear predictors.
-These must be regressed on the parameters to arrive at the desired estimates.
-In this package we do not first create a model and then boost it. Creation and boosting is performed in one call, already yielding estimates. 
-These estimates are, however not beta and gamma, but rather eta_mu and eta_gamma. 
-So these unit-wise linear predictors need to be regressed on the covariate in order to arrive at the estimates of interest.
-
-```{r}
-eta_mu <- fitted(est_1)$mu
-eta_sigma <- fitted(est_1)$sigma
-
-est_beta <- lm(eta_mu ~ covariate_1)
-est_gamma <- lm(eta_sigma ~ covariate_1)
-
-```
-
-### Recap Configuration 3
-
-```{r}
-init <- init_model(beta_3, gamma_3, 1000)
-mod_3 <- init$model
-response_3 <- init$response
-covariate_3 <- init$covariate
-plot_model(beta_3, gamma_3, 1000)
-```
-[...]
-
-```{r}
-# estimate with external package --------
-
-mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3,
-                      control = ctrl)
-eta_mu <- fitted(mod_33)$mu
-eta_sigma <- fitted(mod_33)$sigma
-
-lm(eta_mu ~ covariate_3)
-lm(eta_sigma ~ covariate_3)
-
-```
-
-[...]
-```{r}
-gradient_boost(mod_3)
-round(mod_3$beta, 3)
-round(mod_3$gamma, 3)
-```
-
-```{r}
-microbenchmark(
-  gradient_boost(mod_3)
+```{r echo=FALSE, cache=TRUE}
+#dont show code
+#show output
+# Configuration 6
+cbind(beta_true = init_6$beta, 
+      beta_est = round(mod_6$beta, 3),
+      gamma_true = init_6$gamma,
+      gamma_est = round(mod_6$gamma, 3)
 )
@@
 
-```{r}
-microbenchmark(
-  mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3, control = ctrl),
-  eta_mu <- fitted(mod_33)$mu,
-  eta_sigma <- fitted(mod_33)$sigma,
-  lm(eta_mu ~ covariate_3),
-  lm(eta_sigma ~ covariate_3)
-)
+### Configuration 5
+
+As mentioned above there are restrictions on parameter creation since not every possible parameter configuration leads to estimable models, as we have seen in the two-dimensional case.
+In the multidimensional case, the following  problem remains: if $\bm{\gamma}$ parameters add up to high values, the resulting simulated response has huge standard deviation. 
+This leads to the concealing of the location effects.
+Checking for well balanced signal-noise ratio remains to be done in the multidimensional case as well.
+Furthermore, plotting is hardly an option, as dimensions increase.
+Therefore we check using another method i.e. we look at the distributions of true response means and true response standard deviations.
+
+```{r, cache=TRUE}
+par(mfrow=c(2,1))
+hist(init_5$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_5$fitted_sd, breaks = 25, main = "exp_eta_sigma")
 ```
 
-Ensuring comparability is a difficult task, since the two packages are implemented in rather different ways.
-Running gradient_boost takes about 657000 microseconds (0.66 seconds).
-In contrasts, running the external algorithm gamboostLSS takes 2.829.493 microseconds (2.8 seconds), which is a multiple of our algorithm.
-Additionally, beta and gamma parameters are not yielded directly, but need to be determined by regression as well.
-In contrats, this happens directly in our algorithm.
+Location predictors take values between 8 and 22 with a maximum of 15. 
+In contrast, the highest response- standard deviation is 4, while most values are at about 2.5 and lower. 
+The resulting response variation should be both small enough to not conceal the location effects and big enough to be estimated.
 
-To be fair we evaluate construction of the model object as well.
+In configuration 5, the algorithm correctly depicts which effects exists on the location response *location response oder response location?* i.e. it finds the non-zero elements of $\bm{\beta}$.
+For the scale estimate $\bm{\gamma}_hat$ too many parameters are chosen since two zero elements of the true parameter $\bm{\gamma}$ are assigned non-zero values by the estimate $\hat{\bm{\gamma}}$.
+The estimated effect sizes differ considerably higher from the true values relative the two-dimensional case already presented.  This results in a higher MSE.
 
-```{r}
-microbenchmark(
-  mod_1 <- LocationScaleRegressionBoost$new(response_3 ~ covariate_3, ~ covariate_3)
-)
+```{r, cache=TRUE}
+mse_5 <- calc_MSE(init_5$beta, init_5$gamma, mod_5$beta, mod_5$gamma)
+cbind("config 1" = mse_1,
+      "config_3" = mse_3,
+      "config 4" = mse_4,
+      "config_5" = mse_5)
 ```
 
-This changes the situation only marginally, as this is a quite fast step.
+### Configuration 6
+
+In this case, we face a 10-dimensional configuration.
+We get the following distributions for the mean and the standard deviation respectively.
+
+```{r, cache = TRUE}
+par(mfrow=c(2,1))
+hist(init_6$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_6$fitted_sd, breaks = 25, main = "exp_eta_sigma")
+```
+
+For $\bm{\beta}$ the algorithm chose 1 *what?* too few.
+For $\bm{\gamma}$ entries 1-3 are chosen correctly.
+Then the algorithm yields errors.
+Entry 4 is chosen, while in fact being zero.
+Entry 5 is not chosen even when an effect in the true model exists.
+
+### Configuration 7
+
+When scaling the parameter dimension up to 51 variables, the algorithm runs into a critical problem: all slope estimates are zero.
+
+```{r, cache = TRUE}
+round(mod_7$beta, 3)
+round(mod_7$gamma, 3)
+```
+
+Histograms indicate a well-balanced signal-noise ratio. 
+The problem's cause lies thus somewhere else.
+
+```{r, cache = TRUE}
+par(mfrow=c(2,1))
+hist(init_7$fitted_response, breaks = 25, main = "eta_mu")
+hist(init_7$fitted_sd, breaks = 25, main = "exp_eta_sigma")
+```
+
+One explanation for this problem is that the algorithm fits only intercepts in the first step resulting in very high values.
+The remaining iterations are used to lower intercepts *Um den intercepts to reducing? oder was meints du?*. 
+This may be checked by setting `verbose = TRUE`.
+For this reason, no other components of the parameter estimates are updated.
+Possibly, a sensible choice of initial parameter values may address this problem.
